### 2023-05-26 1.Introduction/1.motivation.tex
Frist commit : transfer from overleaf

\section{Motivation}
\label{sec:int_motivation}

Nowadays, it's impossible not to have heard about \acp{AI}, they are everywhere and everybody talks about them. The most recent example being \textit{ChatGPT} from \textbf{OpenAI} and is getting very popular even outside of the "computer community"(TODO : Find better term). Everybody is using those tools. It's clear that \ac{AI} is becoming more and more important in the current world.

\ac{AI} is wide topic, usually when we talk about \acp{AI}, we refer to the most common type, Machine Learning. Especially, a sub part of Machine Learning called Deep Learning. Deep Learning has seen a surge in research the last decade, which lead to great advancement in this field and the numerous \ac{AI} tools that popped up in the last years.
Deep Learning works using what is called \acp{NN}.

Those \acp{NN} require matrix multiplication. This kind of operation is very time and energy consuming to run on a classic computer. There are options to better run those operations. GPUs can be better suited than a CPU for a \ac{NN} application as it is made for 3D graphics that also mainly use Matrix Multiplication. Other specialized hardware such as FPGAs and ASICs can improve both execution time and energy consumption. All those are digital computers, using analog computing could highly improve speed, power consumption and even smaller chip area. All this using a memristive crossbar array \cite{Xbar}. An analog computer 
The point of this thesis is to lay out the ground work to create a LSTM capable chip to use in embedded systems. It could be used detection and prediction of any kind of sequential events.
Using those memristors \cite{TheoMemristor}, which work like resistance with a memory component to them, that means that they can be set to a desired value.

\subsection{Neural Networks}\label{sec:nn}

\acfp{NN} are a network of several layers of artificial neurons. Figure \ref{fig:snn}, shows a simple representation of a \ac{NN}, the artificial neurons are the represented by the colored circles. On figure \ref{fig:snn} each arrow represent a synapse.

\begin{figure}[h!]
    \centering
    \includesvg[height=8cm]{Figures/Colored_neural_network.svg}
    \caption{Simple \acl{NN}}
    \label{fig:snn}
\end{figure}

They are several layers to a \ac{NN} : 
\begin{itemize}
    \item Input layer : This layer is simply the different inputs.
    \item Hidden layer : This layer can be (and usually is) wider than the one in figure \ref{fig:snn}, it is there to add more layers and thus more precision to the result
    \item Output layer : This layer is where you can find the result from the \ac{NN}.
\end{itemize}

In a computational \ac{NN}, synapses are represented by weights. Those weights are to be multiplied by the last neuron and then added to each other to produce the next stage. Using the names defined in figure \ref{fig:nn_explained}, the equation linking each layer is the following matrix multiplication : 
\begin{equation}
    \begin{bmatrix}
    H1\\ H2\\ H3\\ H4\\
    \end{bmatrix}
    =
    \begin{bmatrix}
        W1,1 & W1,2 & W1,3\\
        W2,1 & W2,2 & W2,3\\
        W3,1 & W3,2 & W3,3\\
        W4,1 & W4,2 & W4,3\\
    \end{bmatrix}
    \cdot
    \begin{bmatrix}
        I1\\ I2\\ I3\\
    \end{bmatrix}
\end{equation}

And it works the same way to go to the next layer.

\begin{figure}[h!]
    \centering
    \includesvg[height=8cm]{Figures/NN_explained.svg}
    \caption{\aclp{NN} with names}
    \label{fig:nn_explained}
\end{figure}

Those matrix multiplication are very power-hungry and their computation time scales up with the size of the \ac{NN} using classical computing.

Analog computation enables those same calculations almost instantly and being far more power-efficient. This can be done using the physical properties of electrical devices.
We want to reproduce one of those neural network using electrical circuits to perform the same computation that a computer would in much faster manner.
Using \hyperref[subsec:memristors]{memristors} as weights. This removes the need to copy the weights from the main memory

### 2023-05-26 1.Introduction/2.state.tex
Frist commit : transfer from overleaf

\section{State of The Art}
\label{sec:int_state}

\subsection{Memristors}\label{subsec:memristors}

Memristors are the lesser known fourth fundamental passive component of electronics, among resistors, capacitors and inductor.
It was first theorized in 1971 by L. Chua from UC Berkeley, in \cite{TheoMemristor}. The name comes from the blend of \textit{memory} and \textit{resistance}.
The theory behind the component was extracted from a missing component to link the fundamental circuit variables. In figure \ref{fig:fundComp}, the four fundamental variables are on each side of the square, with the ones on opposite sides being linked as shown in the following equations :
\begin{equation}
    d\phi = v\cdot dt
\end{equation}

\begin{equation}
    dq = i\cdot dt
\end{equation}
Resistors, capacitors and inductors were already discovered and well known, so it was theorized that a fourth device should then exists to physically link flux ($\phi$) and charge ($q$).  The flux in this case is not a magnetic flux and is defined as such : $ d\phi=V\cdot dt \implies \phi =  \int V(t) \,dt  $.\\
The component stayed theoretical until 2008 when it's been implemented in a physical device for the first time \cite{Strukov2008}. It took 37 years to actually have a working device.\\
There is then an extention of this theoretical device to another, the memristive device. It was theorized in 1976 by L. Chua and S. M. Kang \cite{memrestiveDev}. The difference between the memristor and the memristive device is its internal behavior. Memristive device are commonly referred to as memristor as well.

\begin{figure}[H]
    \centering
    \includegraphics{Figures/Memristor.png}
    \caption{Fundamental passive components}
    \label{fig:fundComp}
\end{figure}

\subsubsection{Equations}
A memristor links the flux ($\phi$) and charge ($q$) and creates the memristance.
This memristance is defined with the following equation : 
\begin{equation}
    M(q)=\frac{d\phi}{dq}
\end{equation}
It can be compared with the other fundamental components like resistor ($R(i)=\frac{dv}{di}$), capacitor ($\frac{1}{C(q)}=\frac{dv}{dq}$) and inductor ($L(i)=\frac{d\phi}{di}$).
We can then extract a more useful equation in an actual circuit :
\begin{equation}
    v(t)=M(q(t))\cdot i(t)
\end{equation}
Similarly, a memductance can be defined as such :
\begin{equation}
    W(\phi)=\frac{dq}{d\phi}
\end{equation}
A memristive device is slightly differently defined, it still uses a memristance, but here the memristance also depends on an internal state called $x$. This gives us this equation :
\begin{equation}
    v(t)=M(x,i)\cdot i(t)
\end{equation}
The internal state ($x$) is not linked to flux or charge in the case of a memristive device.\\
Once again, we can also define the memristive device using a memductance :
\begin{equation}
    i(t)=W(x,v)\cdot v(t)
\end{equation}
In all of the previous equations, $v$ is the voltage in Volt ($V$), $i$ is the current in Ampere ($A$), $\phi$ is the flux in Weber ($Wb$), $q$ is the charge in Coulomb ($C$), $M$ is the memristance in Ohm ($\Omega$) and $W$ is the memductance in Siesmens ($S$ or $\Omega^{-1}$).

\subsubsection{Behavior}
A memristor is defined as a non-linear two-terminal fundamental electrical component. It behaves as a resistance with memory (hence its name), meaning that it changes its resistance based on how much charge went through it. This enables us to manipulate the resistance of the component.
The huge benefit of memristors is the memory component, when you power it, it will have the resistance it had before.

Memristive devices have a similar behavior, the memristive device's resistance will change depending on the internal state ($x$). That internal state changes based on how much and how long voltage signals or currents are applied to the memristive device.

\subsubsection{Usage}
The main research for memristor usage is using them as ReRAM. The idea behind ReRAM is to use memristors as Non-Volatile memory. It uses two states of the device with known resistances ($R_{on}$ and $R_{off}$), giving it binary property. Reading the memory simply requires setting a voltage and reading the output current. It is better than current solutions (HDD, SSD) as it has a much lower latency. It is better than traditionnal DRAM because it keeps the information even when turned off. This makes ReRAM a good replacement for both RAM and HDDs/SSDs, thus eliminating Von Neumann bottleneck due to the Von Neumann architecture that is used in all modern computers.

They can also be used to set a resistance to be able to perform analog multiplication. Setting them in a crossbar array makes them a very strong candidate to be used in neuromorphic computing.

\subsection{Memristors Crossbar Array}\label{subsec:crossbar}

Setting memristors in a crossbar array to perform analog matrix multiplication typically called Multiply and Accumulate because of the nature of Matrix Multiplication. Figure \ref{fig:crossbar} shows what a typical crossbar array looks like.

\begin{figure}[H]
\centering
\begin{subfigure}{0.5\textwidth}
\centering
    %\includegraphics[width=.9\linewidth]{Figures/crossbar2D.jpg}
    \includesvg[width=.9\linewidth]{Figures/crossbar.svg}
    \caption{Schematics}
\end{subfigure}%
\hfill
\begin{subfigure}{0.5\textwidth}
\centering
    \includegraphics[width=.9\linewidth]{Figures/crossbar3D.jpg}
    \caption{3 dimensional representation}
\end{subfigure}
\caption{Memristor crossbar array}
\label{fig:crossbar}
\end{figure}

It uses physical properties of electrical systems to perform analog computing. Lets use the circuit node in figure \ref{fig:crossNode} to explain the theory behind the memristor crossbar array.
\begin{figure}[H]
    \centering
    %\includegraphics{Figures/crossbar-node.jpg}
    \includesvg[height=3.5cm]{Figures/crossbarNode.svg}
    \caption{Memristor crossbar node of the $i^{th}$ line and $j^{th}$ column}
    \label{fig:crossNode}
\end{figure}

First of all, a voltage is applied on the $i^{th}$ line, because every column is virtually grounded, the voltage applied to the memristor, of a set resistance of $R$, is $V_i$. As such and using Ohm's law, we know the the current flowing into the memristor is bound by the following equation :
\begin{equation}
    V_i = R\cdot I_{mem} \implies I_{mem} = V_i\cdot (\frac{1}{R})=I_{mem} = V_i\cdot G
\end{equation}
With $G$ being the conductance of memristor.
This line then joins the column where a current of $I_{i,j-1}$ is flowing, then according to kirchhoff's law the resulting current is :
\begin{equation}
    I_{j,i} = I_{j,i-1}+I_{mem} = I_{j,i-1} + V_i\cdot G
\end{equation}
By applying this process to the whole system we find that the current as the bottom of one column is :
\begin{equation}
    I_1=G_1\cdot V_1 + G_2\cdot V_2 + G_3\cdot V_3
\end{equation}
With $G_1$, $G_2$ and $G_3$ being the conductance of the 3 memristors in the first column.

This analog circuit removes the need to copy data from a main memory as the memristors themselves hold the value and the weights are then stored where the computation happens.

\subsection{memristor's model}
TODO
\subsection{LSTM}
\acp{LSTM} are a type of \ac{NN} used to analyze sequence of data. They are capable of predict data based of the previous results. \acp{LSTM} are part of \acp{RNN}. \acp{RNN} are different than feedfoward \acl{NN} because of their feedback connections. Meaning that the results from the last time step have an impact on the next time step.

\acp{LSTM} are able to forget information passed. This ability gave the uncommon name of \acl{LSTM} as it has both long and short term memory. This what gives \acp{LSTM} their use for sequence data. They can analyze the data and keep the information from the last time step to make a better decision afterwards. The most comprehensible example is considering a sentence. (TODO : find example)

An \ac{LSTM} is more complicated than just a simple feedforward \acl{NN}, they have several gates, which are all technically a \ac{NN} themselves. There is also a cell state which job is to hold a value for the next step.

\begin{figure}[H]
    \centering
    \includesvg[width=\textwidth]{Figures/lstmCellLatex.svg}
    \caption{LSTM Cell}
    \label{fig:lstmCell}
\end{figure}

Figure \ref{fig:lstmCell} shows the complexity of the \ac{LSTM} architecture. In an \ac{LSTM}, each gate is a different \ac{NN}.

An \ac{LSTM} works using different gates
\subsection{Dummy Subsection B}
\label{subsec:subsectionb}

State of Art Subsection B

### 2023-05-26 Figures/TimeDiagram/wavedromtikz.py

@@ -1,0 +1,877 @@
+#!/usr/bin/env python
+
+import yaml
+import argparse
+import sys
+import os.path
+from string import Template
+from collections import namedtuple
+
+STANDALONE_TEX_HEADER = r"""
+\documentclass[crop,tikz]{standalone}
+\usepackage{tikz}
+\usepackage{ifthen}
+\usepackage{listings}
+
+\usetikzlibrary{ calc, positioning, decorations.markings, patterns}
+% all other packages and stuff you need for the picture
+%
+\begin{document}
+\begin{tikzpicture}[thick]
+"""
+
+STANDALONE_TEX_FOOTER = r"""
+\end{tikzpicture}
+\end{document}
+"""
+
+# A TikZ (LaTeX) header to be included before a waveform diagram. Defines
+# various primitives for drawing parts of a waveform.
+TIKZ_HEADER = Template(r"""
+% Seperation between lines
+\pgfmathsetlengthmacro{\wavesep}{2.0em}
+
+% Height of a waveform
+\pgfmathsetlengthmacro{\waveheight}{1.2em}
+
+% Width of a brick (half a cycle)
+\pgfmathsetlengthmacro{\wavewidth}{${size}em}
+
+% Width of the slant on slanted signal changes
+\pgfmathsetlengthmacro{\transitionwidth}{0.3em}
+
+% Width of the curve on slow signal changes (e.g. to z)
+\pgfmathsetlengthmacro{\curvedtransitionwidth}{1.0em}
+
+% Special signal styles
+\tikzset{wave x/.style={pattern=north east lines}}
+\tikzset{wave bus/.style={fill=white}}
+\tikzset{wave busyellow/.style={fill=yellow!25!white}}
+\tikzset{wave busorange/.style={fill=orange!25!white}}
+\tikzset{wave busblue/.style={fill=blue!25!white}}
+\tikzset{wave pulled/.style={dotted}}
+
+% Initialise pointer
+\coordinate (last waveform);
+
+% Label for a signal. Arguments:
+%  #1: The human-readable label string
+\newcommand{\signallabel}[1]{
+   \node [ anchor=east
+         , left=0 of last waveform
+         , minimum height=\waveheight
+         ]
+         {#1};
+}
+
+% Advance the "last brick" coordinate to the next position
+%  #1: Width of a brick
[…]

### 2023-05-29 2.State/1.memristors.tex
Changed the file organization

--- 2023-05-26 1.Introduction/2.state.tex
@@
-\section{State of The Art}
-\label{sec:int_state}
-
-\subsection{Memristors}\label{subsec:memristors}
+\section{Memristors}
+\label{sec:memristors}
 
@@
 
-\subsubsection{Equations}
+\subsection{Equations}
 A memristor links the flux ($\phi$) and charge ($q$) and creates the memristance.
@@
 
-\subsubsection{Behavior}
+\subsection{Behavior}
 A memristor is defined as a non-linear two-terminal fundamental electrical component.
@@
 
-\subsubsection{Usage}
+\subsection{Usage}
 The main research for memristor usage is using them as ReRAM.
@@
 They can also be used to set a resistance to be able to perform analog multiplication. Setting them in a crossbar array makes them a very strong candidate to be used in neuromorphic computing.
-
-\subsection{Memristors Crossbar Array}\label{subsec:crossbar}
-
-Setting memristors in a crossbar array to perform analog matrix multiplication typically called Multiply and Accumulate because of the nature of Matrix Multiplication.
-Figure \ref{fig:crossbar} shows what a typical crossbar array looks like.
-
-\begin{figure}[H]
-\centering
-\begin{subfigure}{0.5\textwidth}
-\centering
-    %\includegraphics[width=.9\linewidth]{Figures/crossbar2D.jpg}
-    \includesvg[width=.9\linewidth]{Figures/crossbar.svg}
-    \caption{Schematics}
-\end{subfigure}%
-\hfill
-\begin{subfigure}{0.5\textwidth}
-\centering
-    \includegraphics[width=.9\linewidth]{Figures/crossbar3D.jpg}
-    \caption{3 dimensional representation}
-\end{subfigure}
-\caption{Memristor crossbar array}
-\label{fig:crossbar}
-\end{figure}
-
-It uses physical properties of electrical systems to perform analog computing. Lets use the circuit node in figure \ref{fig:crossNode} to explain the theory behind the memristor crossbar array.
-\begin{figure}[H]
-    \centering
-    %\includegraphics{Figures/crossbar-node.jpg}
-    \includesvg[height=3.5cm]{Figures/crossbarNode.svg}
-    \caption{Memristor crossbar node of the $i^{th}$ line and $j^{th}$ column}
-    \label{fig:crossNode}
-\end{figure}
-
-First of all, a voltage is applied on the $i^{th}$ line, because every column is virtually grounded, the voltage applied to the memristor, of a set resistance of $R$, is $V_i$.
-As such and using Ohm's law, we know the the current flowing into the memristor is bound by the following equation :
-\begin{equation}
-    V_i = R\cdot I_{mem} \implies I_{mem} = V_i\cdot (\frac{1}{R})=I_{mem} = V_i\cdot G
-\end{equation}
-With $G$ being the conductance of memristor.
-This line then joins the column where a current of $I_{i,j-1}$ is flowing, then according to kirchhoff's law the resulting current is :
-\begin{equation}
-    I_{j,i} = I_{j,i-1}+I_{mem} = I_{j,i-1} + V_i\cdot G
-\end{equation}
-By applying this process to the whole system we find that the current as the bottom of one column is :
-\begin{equation}
-    I_1=G_1\cdot V_1 + G_2\cdot V_2 + G_3\cdot V_3
-\end{equation}
-With $G_1$, $G_2$ and $G_3$ being the conductance of the 3 memristors in the first column.
-
-This analog circuit removes the need to copy data from a main memory as the memristors themselves hold the value and the weights are then stored where the computation happens.
-
-\subsection{memristor's model}
-TODO
-\subsection{LSTM}
-\acp{LSTM} are a type of \ac{NN} used to analyze sequence of data.
-They are capable of predict data based of the previous results.
-\acp{LSTM} are part of \acp{RNN}.
-\acp{RNN} are different than feedfoward \acl{NN} because of their feedback connections.
-Meaning that the results from the last time step have an impact on the next time step.
-
-\acp{LSTM} are able to forget information passed.
-This ability gave the uncommon name of \acl{LSTM} as it has both long and short term memory.
-This what gives \acp{LSTM} their use for sequence data.
-They can analyze the data and keep the information from the last time step to make a better decision afterwards.
-The most comprehensible example is considering a sentence.
-(TODO : find example)
-
-An \ac{LSTM} is more complicated than just a simple feedforward \acl{NN}, they have several gates, which are all technically a \ac{NN} themselves.
-There is also a cell state which job is to hold a value for the next step.
-
-\begin{figure}[H]
-    \centering
-    \includesvg[width=\textwidth]{Figures/lstmCellLatex.svg}
-    \caption{LSTM Cell}
-    \label{fig:lstmCell}
-\end{figure}
-
-Figure \ref{fig:lstmCell} shows the complexity of the \ac{LSTM} architecture. In an \ac{LSTM}, each gate is a different \ac{NN}.
-
-An \ac{LSTM} works using different gates
-\subsection{Dummy Subsection B}
-\label{subsec:subsectionb}
-
-State of Art Subsection B

### 2023-05-29 2.State/2.crossbar.tex
Changed the file organization

--- 2023-05-26 1.Introduction/2.state.tex
@@
-\section{State of The Art}
-\label{sec:int_state}
-
-\subsection{Memristors}\label{subsec:memristors}
-
-Memristors are the lesser known fourth fundamental passive component of electronics, among resistors, capacitors and inductor.
-It was first theorized in 1971 by L. Chua from UC Berkeley, in \cite{TheoMemristor}. The name comes from the blend of \textit{memory} and \textit{resistance}.
-The theory behind the component was extracted from a missing component to link the fundamental circuit variables.
-In figure \ref{fig:fundComp}, the four fundamental variables are on each side of the square, with the ones on opposite sides being linked as shown in the following equations :
-\begin{equation}
-    d\phi = v\cdot dt
-\end{equation}
-
-\begin{equation}
-    dq = i\cdot dt
-\end{equation}
-Resistors, capacitors and inductors were already discovered and well known, so it was theorized that a fourth device should then exists to physically link flux ($\phi$) and charge ($q$).
-The flux in this case is not a magnetic flux and is defined as such : $ d\phi=V\cdot dt \implies \phi =  \int V(t) \,dt  $.\\
-The component stayed theoretical until 2008 when it's been implemented in a physical device for the first time \cite{Strukov2008}. It took 37 years to actually have a working device.\\
-There is then an extention of this theoretical device to another, the memristive device.
-It was theorized in 1976 by L.
-Chua and S.
-M.
-Kang \cite{memrestiveDev}.
-The difference between the memristor and the memristive device is its internal behavior.
-Memristive device are commonly referred to as memristor as well.
-
-\begin{figure}[H]
-    \centering
-    \includegraphics{Figures/Memristor.png}
-    \caption{Fundamental passive components}
-    \label{fig:fundComp}
-\end{figure}
-
-\subsubsection{Equations}
-A memristor links the flux ($\phi$) and charge ($q$) and creates the memristance.
-This memristance is defined with the following equation : 
-\begin{equation}
-    M(q)=\frac{d\phi}{dq}
-\end{equation}
-It can be compared with the other fundamental components like resistor ($R(i)=\frac{dv}{di}$), capacitor ($\frac{1}{C(q)}=\frac{dv}{dq}$) and inductor ($L(i)=\frac{d\phi}{di}$).
-We can then extract a more useful equation in an actual circuit :
-\begin{equation}
-    v(t)=M(q(t))\cdot i(t)
-\end{equation}
-Similarly, a memductance can be defined as such :
-\begin{equation}
-    W(\phi)=\frac{dq}{d\phi}
-\end{equation}
-A memristive device is slightly differently defined, it still uses a memristance, but here the memristance also depends on an internal state called $x$. This gives us this equation :
-\begin{equation}
-    v(t)=M(x,i)\cdot i(t)
-\end{equation}
-The internal state ($x$) is not linked to flux or charge in the case of a memristive device.\\
-Once again, we can also define the memristive device using a memductance :
-\begin{equation}
-    i(t)=W(x,v)\cdot v(t)
-\end{equation}
-In all of the previous equations, $v$ is the voltage in Volt ($V$), $i$ is the current in Ampere ($A$), $\phi$ is the flux in Weber ($Wb$), $q$ is the charge in Coulomb ($C$), $M$ is the memristance in Ohm ($\Omega$) and $W$ is the memductance in Siesmens ($S$ or $\Omega^{-1}$).
-
-\subsubsection{Behavior}
-A memristor is defined as a non-linear two-terminal fundamental electrical component.
-It behaves as a resistance with memory (hence its name), meaning that it changes its resistance based on how much charge went through it.
-This enables us to manipulate the resistance of the component.
-The huge benefit of memristors is the memory component, when you power it, it will have the resistance it had before.
-
-Memristive devices have a similar behavior, the memristive device's resistance will change depending on the internal state ($x$).
-That internal state changes based on how much and how long voltage signals or currents are applied to the memristive device.
-
-\subsubsection{Usage}
-The main research for memristor usage is using them as ReRAM.
-The idea behind ReRAM is to use memristors as Non-Volatile memory.
-It uses two states of the device with known resistances ($R_{on}$ and $R_{off}$), giving it binary property.
-Reading the memory simply requires setting a voltage and reading the output current.
-It is better than current solutions (HDD, SSD) as it has a much lower latency.
-It is better than traditionnal DRAM because it keeps the information even when turned off.
-This makes ReRAM a good replacement for both RAM and HDDs/SSDs, thus eliminating Von Neumann bottleneck due to the Von Neumann architecture that is used in all modern computers.
-
-They can also be used to set a resistance to be able to perform analog multiplication. Setting them in a crossbar array makes them a very strong candidate to be used in neuromorphic computing.
-
-\subsection{Memristors Crossbar Array}\label{subsec:crossbar}
+\section{Memristors Crossbar Array}\label{sec:crossbar}
 
@@
   \centering
-\begin{subfigure}{0.5\textwidth}
-\centering
-    %\includegraphics[width=.9\linewidth]{Figures/crossbar2D.jpg}
-    \includesvg[width=.9\linewidth]{Figures/crossbar.svg}
-    \caption{Schematics}
-\end{subfigure}%
+  %\subfloat[Schematics]{\includesvg[width=.45\linewidth]{Figures/crossbar.svg}}\qquad
+  \subfloat[Schematics]{\includegraphics[width=.45\linewidth]{Figures/crossbar.eps}}%\qquad
   \hfill
-\begin{subfigure}{0.5\textwidth}
-\centering
-    \includegraphics[width=.9\linewidth]{Figures/crossbar3D.jpg}
-    \caption{3 dimensional representation}
-\end{subfigure}
+  \subfloat[3 dimensional representation]{\includegraphics[width=.45\linewidth]{Figures/crossbar3D.jpg}}\\
   \caption{Memristor crossbar array}
@@
 This analog circuit removes the need to copy data from a main memory as the memristors themselves hold the value and the weights are then stored where the computation happens.
-
-\subsection{memristor's model}
-TODO
-\subsection{LSTM}
-\acp{LSTM} are a type of \ac{NN} used to analyze sequence of data.
-They are capable of predict data based of the previous results.
-\acp{LSTM} are part of \acp{RNN}.
-\acp{RNN} are different than feedfoward \acl{NN} because of their feedback connections.
-Meaning that the results from the last time step have an impact on the next time step.
-
-\acp{LSTM} are able to forget information passed.
-This ability gave the uncommon name of \acl{LSTM} as it has both long and short term memory.
-This what gives \acp{LSTM} their use for sequence data.
-They can analyze the data and keep the information from the last time step to make a better decision afterwards.
-The most comprehensible example is considering a sentence.
-(TODO : find example)
-
-An \ac{LSTM} is more complicated than just a simple feedforward \acl{NN}, they have several gates, which are all technically a \ac{NN} themselves.
-There is also a cell state which job is to hold a value for the next step.
-
-\begin{figure}[H]
-    \centering
-    \includesvg[width=\textwidth]{Figures/lstmCellLatex.svg}
-    \caption{LSTM Cell}
-    \label{fig:lstmCell}
-\end{figure}
-
-Figure \ref{fig:lstmCell} shows the complexity of the \ac{LSTM} architecture. In an \ac{LSTM}, each gate is a different \ac{NN}.
-
-An \ac{LSTM} works using different gates
-\subsection{Dummy Subsection B}
-\label{subsec:subsectionb}
-
-State of Art Subsection B

### 2023-05-29 2.State/4.lstm.tex
Changed the file organization

--- 2023-05-26 1.Introduction/2.state.tex
@@
-\section{State of The Art}
-\label{sec:int_state}
-
-\subsection{Memristors}\label{subsec:memristors}
-
-Memristors are the lesser known fourth fundamental passive component of electronics, among resistors, capacitors and inductor.
-It was first theorized in 1971 by L. Chua from UC Berkeley, in \cite{TheoMemristor}. The name comes from the blend of \textit{memory} and \textit{resistance}.
-The theory behind the component was extracted from a missing component to link the fundamental circuit variables.
-In figure \ref{fig:fundComp}, the four fundamental variables are on each side of the square, with the ones on opposite sides being linked as shown in the following equations :
-\begin{equation}
-    d\phi = v\cdot dt
-\end{equation}
-
-\begin{equation}
-    dq = i\cdot dt
-\end{equation}
-Resistors, capacitors and inductors were already discovered and well known, so it was theorized that a fourth device should then exists to physically link flux ($\phi$) and charge ($q$).
-The flux in this case is not a magnetic flux and is defined as such : $ d\phi=V\cdot dt \implies \phi =  \int V(t) \,dt  $.\\
-The component stayed theoretical until 2008 when it's been implemented in a physical device for the first time \cite{Strukov2008}. It took 37 years to actually have a working device.\\
-There is then an extention of this theoretical device to another, the memristive device.
-It was theorized in 1976 by L.
-Chua and S.
-M.
-Kang \cite{memrestiveDev}.
-The difference between the memristor and the memristive device is its internal behavior.
-Memristive device are commonly referred to as memristor as well.
-
-\begin{figure}[H]
-    \centering
-    \includegraphics{Figures/Memristor.png}
-    \caption{Fundamental passive components}
-    \label{fig:fundComp}
-\end{figure}
-
-\subsubsection{Equations}
-A memristor links the flux ($\phi$) and charge ($q$) and creates the memristance.
-This memristance is defined with the following equation : 
-\begin{equation}
-    M(q)=\frac{d\phi}{dq}
-\end{equation}
-It can be compared with the other fundamental components like resistor ($R(i)=\frac{dv}{di}$), capacitor ($\frac{1}{C(q)}=\frac{dv}{dq}$) and inductor ($L(i)=\frac{d\phi}{di}$).
-We can then extract a more useful equation in an actual circuit :
-\begin{equation}
-    v(t)=M(q(t))\cdot i(t)
-\end{equation}
-Similarly, a memductance can be defined as such :
-\begin{equation}
-    W(\phi)=\frac{dq}{d\phi}
-\end{equation}
-A memristive device is slightly differently defined, it still uses a memristance, but here the memristance also depends on an internal state called $x$. This gives us this equation :
-\begin{equation}
-    v(t)=M(x,i)\cdot i(t)
-\end{equation}
-The internal state ($x$) is not linked to flux or charge in the case of a memristive device.\\
-Once again, we can also define the memristive device using a memductance :
-\begin{equation}
-    i(t)=W(x,v)\cdot v(t)
-\end{equation}
-In all of the previous equations, $v$ is the voltage in Volt ($V$), $i$ is the current in Ampere ($A$), $\phi$ is the flux in Weber ($Wb$), $q$ is the charge in Coulomb ($C$), $M$ is the memristance in Ohm ($\Omega$) and $W$ is the memductance in Siesmens ($S$ or $\Omega^{-1}$).
-
-\subsubsection{Behavior}
-A memristor is defined as a non-linear two-terminal fundamental electrical component.
-It behaves as a resistance with memory (hence its name), meaning that it changes its resistance based on how much charge went through it.
-This enables us to manipulate the resistance of the component.
-The huge benefit of memristors is the memory component, when you power it, it will have the resistance it had before.
-
-Memristive devices have a similar behavior, the memristive device's resistance will change depending on the internal state ($x$).
-That internal state changes based on how much and how long voltage signals or currents are applied to the memristive device.
-
-\subsubsection{Usage}
-The main research for memristor usage is using them as ReRAM.
-The idea behind ReRAM is to use memristors as Non-Volatile memory.
-It uses two states of the device with known resistances ($R_{on}$ and $R_{off}$), giving it binary property.
-Reading the memory simply requires setting a voltage and reading the output current.
-It is better than current solutions (HDD, SSD) as it has a much lower latency.
-It is better than traditionnal DRAM because it keeps the information even when turned off.
-This makes ReRAM a good replacement for both RAM and HDDs/SSDs, thus eliminating Von Neumann bottleneck due to the Von Neumann architecture that is used in all modern computers.
-
-They can also be used to set a resistance to be able to perform analog multiplication. Setting them in a crossbar array makes them a very strong candidate to be used in neuromorphic computing.
-
-\subsection{Memristors Crossbar Array}\label{subsec:crossbar}
-
-Setting memristors in a crossbar array to perform analog matrix multiplication typically called Multiply and Accumulate because of the nature of Matrix Multiplication.
-Figure \ref{fig:crossbar} shows what a typical crossbar array looks like.
-
-\begin{figure}[H]
-\centering
-\begin{subfigure}{0.5\textwidth}
-\centering
-    %\includegraphics[width=.9\linewidth]{Figures/crossbar2D.jpg}
-    \includesvg[width=.9\linewidth]{Figures/crossbar.svg}
-    \caption{Schematics}
-\end{subfigure}%
-\hfill
-\begin{subfigure}{0.5\textwidth}
-\centering
-    \includegraphics[width=.9\linewidth]{Figures/crossbar3D.jpg}
-    \caption{3 dimensional representation}
-\end{subfigure}
-\caption{Memristor crossbar array}
-\label{fig:crossbar}
-\end{figure}
-
-It uses physical properties of electrical systems to perform analog computing. Lets use the circuit node in figure \ref{fig:crossNode} to explain the theory behind the memristor crossbar array.
-\begin{figure}[H]
-    \centering
-    %\includegraphics{Figures/crossbar-node.jpg}
-    \includesvg[height=3.5cm]{Figures/crossbarNode.svg}
-    \caption{Memristor crossbar node of the $i^{th}$ line and $j^{th}$ column}
-    \label{fig:crossNode}
-\end{figure}
-
-First of all, a voltage is applied on the $i^{th}$ line, because every column is virtually grounded, the voltage applied to the memristor, of a set resistance of $R$, is $V_i$.
-As such and using Ohm's law, we know the the current flowing into the memristor is bound by the following equation :
-\begin{equation}
-    V_i = R\cdot I_{mem} \implies I_{mem} = V_i\cdot (\frac{1}{R})=I_{mem} = V_i\cdot G
-\end{equation}
-With $G$ being the conductance of memristor.
-This line then joins the column where a current of $I_{i,j-1}$ is flowing, then according to kirchhoff's law the resulting current is :
-\begin{equation}
-    I_{j,i} = I_{j,i-1}+I_{mem} = I_{j,i-1} + V_i\cdot G
-\end{equation}
-By applying this process to the whole system we find that the current as the bottom of one column is :
-\begin{equation}
-    I_1=G_1\cdot V_1 + G_2\cdot V_2 + G_3\cdot V_3
-\end{equation}
-With $G_1$, $G_2$ and $G_3$ being the conductance of the 3 memristors in the first column.
-
-This analog circuit removes the need to copy data from a main memory as the memristors themselves hold the value and the weights are then stored where the computation happens.
-
-\subsection{memristor's model}
-TODO
-\subsection{LSTM}
+\section{LSTM}\label{sec:lstm}
 \acp{LSTM} are a type of \ac{NN} used to analyze sequence of data.
@@
 An \ac{LSTM} works using different gates
-\subsection{Dummy Subsection B}
-\label{subsec:subsectionb}
-
-State of Art Subsection B

### 2023-05-29 2.State/main.tex
Changed the file organization

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% The State:
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\fancychapter{State of the art}
\label{cap:state}

\input{2.State/1.memristors.tex}
\input{2.State/2.crossbar.tex}
\input{2.State/3.model.tex}
\input{2.State/4.lstm.tex}

\cleardoublepage

### 2023-05-30 diapo/diapo2.tex
Updated makefile (clean) Added diapo

\documentclass{beamer}
\usetheme{Boadilla}
\usecolortheme{whale}

\usepackage{wrapfig}
\usepackage{pgf}

\newcommand{\includepgf}[3]{\resizebox{#1}{#2}{\input{#3}}}

\AtBeginSection[]{
  \begin{frame}
  \vfill
  \centering
  \begin{beamercolorbox}[sep=8pt,center,shadow=true,rounded=true]{title}
    \usebeamerfont{title}\insertsectionhead\par%
  \end{beamercolorbox}
  \vfill
  \end{frame}
}

\title[MSc Thesis 2]
{Second MSc thesis presentation}

\author[V. BARBAZA]{Valentin BARBAZA}

\date[2023] % (optional)
{May 2023}

\logo{\includegraphics[height=0.7cm]{figures/istnewlogo.pdf}}

\definecolor{istblue}{RGB}{0,102,153}
\setbeamercolor{titlelike}{bg=istblue}
\setbeamerfont{title}{series=\bfseries}

\begin{document}

\frame{\titlepage}

\begin{frame}
\frametitle{Table of Contents}
\tableofcontents
\end{frame}

\section{Sub circuits}
\subsection{Activation functions}

\begin{frame}
\frametitle{Sigmoid}
I reproduced the sigmoid circuit described in the base paper, adjusted the parameters to optain the following results :
\begin{figure}
    \centering
      \includegraphics[height=0.5\textheight]{figures/sigmoid.png}
\end{figure}
\end{frame}

\begin{frame}
\frametitle{Hyperbolic Tangent}
The second activation function used is the tanh function. Using the same circuit with a slight change in the parameters we get the following :
\begin{figure}    
    \centering
    \includegraphics[height=0.5\textheight]{figures/tanh.png}
\end{figure}
They both use ideal verilog-A model for their OpAmp
\end{frame}

\subsection{Memory cells}
\begin{frame}{The memory cell}
    In order to store an analog value until the next LSTM step.
    \begin{figure}    
        \centering
        \includegraphics[height=0.5\textheight]{figures/memcell-circuit.png}
    \end{figure}
    The memory cell has 2 CMOS switches to avoid the memory to be influenced by the voltage in the input.
\end{frame}
\begin{frame}{Memory leaks}
    Here is the value stored in the unit over time with one and two CMOS switches.
    \begin{figure}
        \centering
        \resizebox{!}{0.4\textheight}{\input{figures/memcell-graph.pgf}}
    \end{figure}
    This shows that with 2 CMOS switches the value is still $99\%$ of the original value after $120 \mu s$ while it's only $52\%$ with 1 CMOS switch. 
\end{frame}
\begin{frame}{Memory leaks}
    Here is the value stored in the unit over time with one and two CMOS switches.
    \begin{figure}
        \centering
        \includepgf{!}{.4\textheight}{figures/memcell-graph.pgf}
    \end{figure}
    This shows that with 2 CMOS switches the value is still $99\%$ of the original value after $120 \mu s$ while it's only $52\%$ with 1 CMOS switch. 
\end{frame}
% explain control of memcells

\section{Full LSTM circuit}

\begin{frame}{LSTM circuit}
    Then I joined all the elements into the final circuit :
    \begin{figure}
        \centering
        \includegraphics[height=0.5\textheight]{figures/LSTM-NP.png}
    \end{figure}
    Here the opamps and voltage multipliers are verilog-A models.
\end{frame}

% prepare question sigmoid 0 to 1

\section{Improvements to the model}
\begin{frame}{Serial vs parallel}
    The base paper uses serial computation.
    \begin{figure}
        \centering
        \includegraphics[height=0.5\textheight]{figures/parallel.png}
    \end{figure}
    It does so to save area on the chip.
\end{frame}
\begin{frame}{Serial vs parallel}
    That it is only useful for small LSTM networks.\\
    The area of the crossbar increase with $O(n_h^2)$.\\
    When the system is in parallel, the rest of the circuit is increasing with $O(n_h)$.\\
    When the system is in serial, the time it takes for one LSTM step increase with $O(n_h)$.\\
    At one point one LSTM step will become too long.\\
    We can fix that by either using a fully serialized system.\\
    The second would be to find out the optimal number of serial channel and use a mix of both.
\end{frame}

\section{Double OpAmps}
\begin{frame}{Double OpAmps}
    \begin{figure}
        \centering
        \includegraphics[height=0.35\textheight]{figures/doubleOpAmps.png}
    \end{figure}
    The final equation of this system is : $$y_j = R_f\cdot [A\cdot (\sigma_{A+}-\sigma_{A-})+ ... + \beta \cdot (\sigma_{\beta+}-\sigma_{\beta-})]$$
    That means that the weight is positive if $\sigma_+>\sigma_-$, and the inverse for the negative.
\end{frame}
\begin{frame}{Single OpAmp}
        There is a possibility of having only one OpAmp, that means having a threshold value to separate positive from negative. This also reduces the resolution of weights we can achieve.
\end{frame}

\section{Next step}
\begin{frame}{Matrix generation}
    I've been trying to generate an LSTM circuit from verilog, which doesn't work as we can't set the resistors values using that method. I'm going to use another netlisting tool (CDL, spice, other) to generate the system with preset resistances.
    Once I'll know how to do that, I'll be able to create a python script to create any LSTM wanted, the options can be changed very easily.
\end{frame}
\section{Question}
\begin{frame}{Normalized values}
    For the input the values have to be normalized. Usually using ranges like $[0.8,1]$(in V). Meaning that $1V$ is the maximum input value. The issue is that the sigmoid function goes from 0  to 1, but my design goes from $0.9V$ to $1V$, which with the conversion back to numbers goes from 0 to the max value.\\
    Have I missed somehting?
\end{frame}
\end{document}

### 2023-05-30 diapo/template1.tex
Updated makefile (clean) Added diapo

\documentclass{beamer}
\usetheme{Boadilla}
\usecolortheme{whale}

\usepackage{wrapfig}

\title[MSc Thesis]
{Memristors-based Recurrent Modules for Neural Computing}
%\subtitle{Using }

\institute[IST] % (optional)
{%
  Diogo Caetano\\
  INESC-MN
  \and%
  Ruxandra Barlulescu\\
  INESC-ID
}

\author[V. BARBAZA]{Valentin BARBAZA (MEEC)}

\date[2023] % (optional)
{May 2023}

\logo{\includegraphics[height=0.7cm]{figures/istnewlogo.pdf}}

\definecolor{istblue}{RGB}{0,102,153}
\setbeamercolor{titlelike}{bg=istblue}
\setbeamerfont{title}{series=\bfseries}

\begin{document}

\frame{\titlepage}

%\begin{frame}
%\frametitle{Table of Contents}
%\tableofcontents
%\end{frame}

\begin{frame}
\frametitle{Application/Motivation}
The goal of the thesis is to know whether an analog LSTM network would work. This would allow running a Neural Network in real-time in embedded systems. The analog implementation (using memristors crossbar array) allows for much greater computational speed and very small on-chip area.
\end{frame}
\begin{frame}
\frametitle{Technical Goals}
My point is to run simulations using cadence. Running a working simulation is the first step before fabricating the system.
\end{frame}
\begin{frame}
\frametitle{Progress}
So far, I have designed some smaller elements required for the system to work such as the activation function for example.\\
I've designed a full size LSTM (with 4 hidden state).
\begin{figure}
    \centering
    \includegraphics[height=0.35\textheight]{figures/LSTM.png}
\end{figure}
I am now working on a way to generate LSTM network of any size.
\end{frame}

\begin{frame}
\frametitle{What is missing}
I now need to run the simulations and compare the results of my simulation with the results a digital LSTM (running with Keras for example).
\end{frame}

\end{document}

### 2023-05-30 diapo/diapo1/main.tex
Worked out the slide organization

\documentclass{beamer}
\usetheme{Boadilla}
\usecolortheme{whale}

\usepackage{wrapfig}

\title[Thesis 1]
{First thesis presentation}

\subtitle{How far have I gone}

\author[V. BARBAZA]{Valentin BARBAZA}

\date[2023] % (optional)
{March 2023}

\logo{\includegraphics[height=0.7cm]{figures/istnewlogo.pdf}}

\definecolor{istblue}{RGB}{0,102,153}
\setbeamercolor{titlelike}{bg=istblue}
\setbeamerfont{title}{series=\bfseries}

\begin{document}

\frame{\titlepage}

\begin{frame}
\frametitle{Table of Contents}
\tableofcontents
\end{frame}

\section{What I have done}

\begin{frame}{What I have done}
Here are the things that I have done to this day.
    \begin{itemize}
        \item Implement Memristors in cadence
        \item Understand most of LSTMs
        \item Read and understand back-propagation
        \item Started writing thesis (Very beginning)
    \end{itemize}
\end{frame}

%\subsection{The Memristors}
\begin{frame}{Memristor implementation}
Here are the resulting curves I obtained. The curves are the exact same ones as in the paper. Nevertheless, this type of memristor (changing by applying large voltage) probably isn't suitable for out system. It would require too long of a time to update the weights.
\begin{figure}
    \includegraphics[height=0.5\textheight]{figures/Resistances-100us.png}
\end{figure}
\end{frame}

\section{LSTM}
\begin{frame}{LSTM}
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/LSTM.png}
\end{figure}
\end{frame}
\begin{frame}{LSTM}
\begin{figure}
    \centering
    \includegraphics[width=0.9\textwidth]{figures/Analog-LSTM.png}
\end{figure}

\end{frame}

\section{What I can replicate}
\begin{frame}{What I can replicate}
    \begin{itemize}
        \item The circuit using resistances
        \item The circuit using pre-charged memristors
        \item Same for all LSTMs architectures
        \item Try to implement back-propagation (HARD)
        \item Compare power usage ?
    \end{itemize}
\end{frame}

%\begin{frame}{Back-propagation}
 %   I think it would require a whole other system on the side, to calculate the new weights. 
%\end{frame}

\section{Questions}
\begin{frame}{Questions}
    \begin{itemize}
         \item How to feed the inputs to the simulation ?
         \item How are the values encoded ?
    \end{itemize}
\end{frame}

\begin{frame}{Questions}
    For training :
    \begin{itemize}
        \item How long does it take to back-propagate ?
        \item Is the expected value, the value for next time step?
        \item Does that mean that the known values are for training ? Constant training in real time ?
    \end{itemize}
\end{frame}

\end{document}

### 2023-05-31 diapo/groupTemplate/main.tex
Working on a new template
Changed a bit of the template, not done yet

--- 2023-05-30 diapo/template1.tex
@@
-\documentclass{beamer}
+\documentclass[table]{beamer}
 \usetheme{Boadilla}
@@
 
-\usepackage{wrapfig}
+%\usepackage[table]{xcolor}
 
-\title[MSc Thesis]
-{Memristors-based Recurrent Modules for Neural Computing}
-%\subtitle{Using }
+\title[Group Meeting]
+{Weekly group meetings}
 
-\institute[IST] % (optional)
-{%
-  Diogo Caetano\\
-  INESC-MN
-  \and%
-  Ruxandra Barlulescu\\
-  INESC-ID
-}
+\subtitle{Memristors-based recurent modules for neural computing}
 
-\author[V. BARBAZA]{Valentin BARBAZA (MEEC)}
+\author[V. BARBAZA]{Valentin BARBAZA}
 
-\date[2023] % (optional)
-{May 2023}
+\date[26 May] % (optional)
+{26 May 2023}
 
-\logo{\includegraphics[height=0.7cm]{figures/istnewlogo.pdf}}
+\logo{\includegraphics[height=0.9cm]{figures/logos.png}}
 
+\definecolor{text}{RGB}{25,116,146}
+\definecolor{firstRow}{HTML}{219BC3}
+\definecolor{evenRow}{HTML}{CCDEE9}
+\definecolor{oddRow}{HTML}{E7EFF4}
 \definecolor{istblue}{RGB}{0,102,153}
@@
 
-%\begin{frame}
-%\frametitle{Table of Contents}
-%\tableofcontents
-%\end{frame}
+\begin{frame}
+  \frametitle{Introduction}
+  Quick introduction to your topic.
 
-\begin{frame}
-\frametitle{Application/Motivation}
-The goal of the thesis is to know whether an analog LSTM network would work.
-This would allow running a Neural Network in real-time in embedded systems.
-The analog implementation (using memristors crossbar array) allows for much greater computational speed and very small on-chip area.
-\end{frame}
-\begin{frame}
-\frametitle{Technical Goals}
-My point is to run simulations using cadence. Running a working simulation is the first step before fabricating the system.
-\end{frame}
-\begin{frame}
-\frametitle{Progress}
-So far, I have designed some smaller elements required for the system to work such as the activation function for example.\\
-I've designed a full size LSTM (with 4 hidden state).
-\begin{figure}
-    \centering
-    \includegraphics[height=0.35\textheight]{figures/LSTM.png}
-\end{figure}
-I am now working on a way to generate LSTM network of any size.
+  \begin{itemize}
+      \color{text}
+    \item Briefly summarized your project. Some of the attendants may need some refresh.
+    \item Explain in which project stage you are, thus we all know how close you are to achieve your goals.
+  \end{itemize}
+
+  ex: I have to design a PCB to control a magnetic sensor. I have already done the PCB and now I am testing it.
 \end{frame}
@@
 \begin{frame}
-\frametitle{What is missing}
-I now need to run the simulations and compare the results of my simulation with the results a digital LSTM (running with Keras for example).
+  \frametitle{Last weeks...}
+
+  Last meeting, I promised to do…
+
+  \centering
+  \rowcolors{2}{evenRow}{oddRow}
+  \begin{tabular}{ c m{3cm} c c c }
+    \rowcolor{firstRow}
+    & \centering\color{white}Task & \color{white}Done? & \color{white}Why not? & \color{white}Help? \\
+    1 & test2 & test23 & test3 & test4\\
+    2 & test2 & test23 & test3 & test4\\
+    3 & test2 & test23 & test3 & test4\\
+    4 & test2 & test23 & test3 & test4\\
+    5 & test2 & test23 & test3 & test4\\
+  \end{tabular}
+
+\end{frame}
+
+\begin{frame}
+  \frametitle{}
 \end{frame}

### 2023-06-05 diapo/groupTemplate/main.tex
Almost completed the new template, needs theme now

@@
 \usecolortheme{whale}
+
+\usepackage{tikz}
+%\input{shape.tex}
 
@@
 
-\begin{frame}
-  \frametitle{}
+\begin{frame}{What's new}
+  \begin{itemize}
+    \item Results obtianed :
+      \begin{itemize}
+          \color{text}
+        \item
+      \end{itemize}
+    \item Improvements :
+      \begin{itemize}
+          \color{text}
+        \item
+      \end{itemize}
+    \item Relevant information :
+      \begin{itemize}
+          \color{text}
+        \item test
+      \end{itemize}
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Problems/doubts}
+  \begin{itemize}
+    \item Things that must me sorted out in order to continue with your work.\\ Explain the problem.
+    \item Things that you do not understand.
+  \end{itemize}
+\end{frame}
+
+\begin{frame}{Next weeks}
+
+  List of what I plan to do in the next weeks :
+
+  \centering
+  \rowcolors{2}{evenRow}{oddRow}
+  \begin{tabular}{ c m{4cm} }
+    \rowcolor{firstRow}
+    & \centering\color{white}Task \cr
+    1 & test23 \\
+    2 & test23 \\
+    3 & test23 \\
+    4 & test23 \\
+    5 & test23 \\
+  \end{tabular}
+\end{frame}
+
+\begin{frame}{Networking and sharing}
+
+  \begin{tikzpicture}
+    \node (rect) at (0,0) [rounded corners,draw=text,very thick,minimum width=5.7cm,minimum height=6cm] {Help/collaboration};
+    \node (rect) at (6.2,0) [rounded corners,draw=text,very thick,minimum width=5.7cm,minimum height=6cm] {\raggedleft autre chose };
+  \end{tikzpicture}
 \end{frame}

### 2023-06-05 diapo/groupTemplate/main.tex
Created theme for group meeting template

@@
 \documentclass[table]{beamer}
-\usetheme{Boadilla}
-\usecolortheme{whale}
+\usetheme{gpmeet}
+%\usecolortheme{whale}
 
@@
 
-\title[Group Meeting]
-{Weekly group meetings}
+\title{Group meetings}
 
@@
 
-\date[26 May] % (optional)
-{26 May 2023}
+\date{26/05/2023}
 
-\logo{\includegraphics[height=0.9cm]{figures/logos.png}}
-
-\definecolor{text}{RGB}{25,116,146}
-\definecolor{firstRow}{HTML}{219BC3}
-\definecolor{evenRow}{HTML}{CCDEE9}
-\definecolor{oddRow}{HTML}{E7EFF4}
-\definecolor{istblue}{RGB}{0,102,153}
-\setbeamercolor{titlelike}{bg=istblue}
-\setbeamerfont{title}{series=\bfseries}
+\logo{
+  \begin{tikzpicture}[overlay,remember picture]
+    \node[left=0.2cm] at (current page.34){
+      \includegraphics[height=1.2cm]{figures/logos}
+    };
+  \end{tikzpicture}
+}
 
@@
     \rowcolor{firstRow}
-    & \centering\color{white}Task & \color{white}Done? & \color{white}Why not? & \color{white}Help? \\
+    \color{white}\textbf{\#} & \centering\color{white}Task & \color{white}Done? & \color{white}Why not? & \color{white}Help? \\
     1 & test2 & test23 & test3 & test4\\
@@
     \rowcolor{firstRow}
-    & \centering\color{white}Task \cr
+    \color{white}\textbf{\#} & \centering\color{white}Task \cr
     1 & test23 \\
@@
   \begin{tikzpicture}
-    \node (rect) at (0,0) [rounded corners,draw=text,very thick,minimum width=5.7cm,minimum height=6cm] {Help/collaboration};
-    \node (rect) at (6.2,0) [rounded corners,draw=text,very thick,minimum width=5.7cm,minimum height=6cm] {\raggedleft autre chose };
+    \node (rect) at (0,0) [rounded corners,draw=text,very thick,minimum width=5.0cm,minimum height=6cm] {Help/collaboration};
+    \node (rect) at (6,0) [rounded corners,draw=text,very thick,minimum width=5.0cm,minimum height=6cm] {\raggedleft autre chose };
   \end{tikzpicture}

### 2023-06-05 diapo/05-06-23_meeting/main.tex
Finished today's presentation

\documentclass[table]{beamer}
\usetheme{gpmeet}
%\usecolortheme{whale}

\usepackage{tikz}
%\input{shape.tex}

%\usepackage[table]{xcolor}

\title{Group meetings}

\subtitle{Memristors-based recurent modules for neural computing}

\author[V. BARBAZA]{Valentin BARBAZA}

\date{05/06/2023}

\logo{
  \begin{tikzpicture}[overlay,remember picture]
    \node[left=0.2cm] at (current page.34){
      \includegraphics[height=1.2cm]{figures/logos}
    };
  \end{tikzpicture}
}

\begin{document}

\frame{\titlepage}

\begin{frame}
  \frametitle{Introduction}

  \begin{itemize}
      \color{text}
    \item I am working on an analog implementation of a LSTM Neural Networks.
    \item The final objective of my thesis is to have a successful simulation of such a circuit.
  \end{itemize}

\end{frame}

\begin{frame}
  \frametitle{Introduction}
  \centering\includegraphics[width=\textwidth]{figures/LSTM-NP}
\end{frame}

\begin{frame}
  \frametitle{Last weeks...}

  Last meeting, I promised to do :

  \centering
  \rowcolors{2}{evenRow}{oddRow}
  \begin{tabular}{ c m{3cm} c m{3cm} c }
    \rowcolor{firstRow}
    \color{white}\textbf{\#} & \centering\color{white}Task & \color{white}Done? & \color{white}Why not? & \color{white}Help? \\
    1 & Try to use LTSpice & No & Found out I could do what I want with Cadence & No\\
    2 & Find another\newline problem to work on & No & Lack of time & Yes\\
    3 & Work on the thesis intro & on going &  & No\\
  \end{tabular}

\end{frame}

\begin{frame}{What's new}
  \begin{itemize}
    \item Results obtianed :
      \begin{itemize}
          \color{text}
        \item With the help of Fabio, I can now simulate all the steps automatically
        \item I successfully ran the simulation
        \item The result are not correct
      \end{itemize}
    \item Improvements :
      \begin{itemize}
          \color{text}
        \item Debug the whole thing
      \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{What's new}
  \includegraphics[width=\textwidth]{figures/fulltrial1-all-curves}
\end{frame}

\begin{frame}<1>[label=prob]
  \frametitle{Problems/doubts}
  \begin{itemize}
    \item<1-> The memory doesn't output the right values
    \item<2-> The memory input seems to be active for too long
  \item<3-> The output result isn't what is expected    \end{itemize}
\end{frame}

\begin{frame}{Problems/doubts}
  The memory doesn't output the right values.

  \centering
  \includegraphics[width=0.7\textwidth]{figures/fulltrial1-memory-prob}
\end{frame}

\againframe<2>{prob}

\begin{frame}{Problems/doubts}
  \centering
  \includegraphics[width=1\textwidth]{figures/fulltrial1-mem-act-late}
\end{frame}

\againframe<3>{prob}
\begin{frame}{Problems/doubts}
  The output is ~0.91 V. This is $-1.1$ and should be $0.015$

  \centering
  \includegraphics[width=1\textwidth]{figures/fulltrial1-final-res}
\end{frame}

\begin{frame}{Next weeks}

  List of what I plan to do in the next weeks :

  \centering
  \rowcolors{2}{evenRow}{oddRow}
  \begin{tabular}{ c m{4cm} }
    \rowcolor{firstRow}
    \color{white}\textbf{\#} & \centering\color{white}Task \cr
    1 & Find another LSTM problem \\
    2 & "Debug" the circuit \\
    2.1 & Run the entire simulation \\
    3 & Work on thesis \\
  \end{tabular}
\end{frame}

\begin{frame}{Networking and sharing}
  \begin{tikzpicture}
    %\node (rect) at (0,0) [rounded corners,draw=text,very thick,minimum width=5.0cm,minimum height=6cm] {};
    %\node (rect) at (6,0) [rounded corners,draw=text,very thick,minimum width=5.0cm,minimum height=6cm] {\raggedleft autre chose };
    \node[draw=text,very thick,rounded corners, text width=0.46\textwidth,minimum height=6cm] at (0,0) {\centering\textbf{Help/Collaboration :}\\ \raggedright \color{text}- If anybody has any idea of an interesting LSTM/NN problem to work on.\\-Also if any phenomen happening are familiar to you, I'll take the help};

    %\end{tikzpicture}
    %\begin{tikzpicture}
    \node[draw=text,very thick,rounded corners, text width=0.46\textwidth,minimum height=6cm] at (0.54\textwidth,0) {\centering\textbf{Recommendations :}\\ \color{text} I worked on a latex/beamer version of the template};
    \end{tikzpicture}
\end{frame}

\end{document}

### 2023-06-05 diapo/05-06-23_meeting/main.tex
wrong number was typed
Fixed logos

@@
 \usetheme{gpmeet}
-%\usecolortheme{whale}
+
+\newcommand{\leftRect}[2]{\node[draw=text,very thick,rounded corners, text width=0.46\textwidth,minimum height=6cm] at (0,0) {\centering\textbf{#1}\\ \raggedright \color{text}#2};}
+\newcommand{\rightRect}[2]{\node[draw=text,very thick,rounded corners, text width=0.46\textwidth,minimum height=6cm] at (0.54\textwidth,0) {\centering\textbf{#1}\\ \raggedright \color{text}#2};}
 
 \usepackage{tikz}
-%\input{shape.tex}
-
-%\usepackage[table]{xcolor}
 
@@
   \begin{tikzpicture}[overlay,remember picture]
-    \node[left=0.2cm] at (current page.34){
-      \includegraphics[height=1.2cm]{figures/logos}
+    \node[left=0cm] at (current page.34){
+      \includegraphics[height=1cm]{../figures/ist.eps}
+      \hspace{1pt}
+      \includegraphics[height=1cm]{figures/inesc-mn.png}
+      \includegraphics[height=1cm]{../figures/inesc-id.eps}
     };
@@
 \begin{frame}{Problems/doubts}
-  The output is ~0.91 V. This is $-1.1$ and should be $0.015$
+  The output is ~0.791 V. This is $-1.09$ and should be $0.015$
 
@@
   \begin{tikzpicture}
-    %\node (rect) at (0,0) [rounded corners,draw=text,very thick,minimum width=5.0cm,minimum height=6cm] {};
-    %\node (rect) at (6,0) [rounded corners,draw=text,very thick,minimum width=5.0cm,minimum height=6cm] {\raggedleft autre chose };
-    \node[draw=text,very thick,rounded corners, text width=0.46\textwidth,minimum height=6cm] at (0,0) {\centering\textbf{Help/Collaboration :}\\ \raggedright \color{text}- If anybody has any idea of an interesting LSTM/NN problem to work on.\\-Also if any phenomen happening are familiar to you, I'll take the help};
+    \leftRect{Help/Collaboration :}{- If anybody has any idea of an interesting LSTM/NN problem to work on.\\-Also if any phenomen happening are familiar to you, I'll take the help}
 
-    %\end{tikzpicture}
-    %\begin{tikzpicture}
-    \node[draw=text,very thick,rounded corners, text width=0.46\textwidth,minimum height=6cm] at (0.54\textwidth,0) {\centering\textbf{Recommendations :}\\ \color{text} I worked on a latex/beamer version of the template};
+    \rightRect{Recommendations :}{I worked on a latex/beamer version of the template}
   \end{tikzpicture}

### 2023-06-05 diapo/groupTemplate/main.tex
Template finished
Fixed logos

@@
 \begin{frame}{Networking and sharing}
-
   \begin{tikzpicture}
-    \node (rect) at (0,0) [rounded corners,draw=text,very thick,minimum width=5.0cm,minimum height=6cm] {Help/collaboration};
-    \node (rect) at (6,0) [rounded corners,draw=text,very thick,minimum width=5.0cm,minimum height=6cm] {\raggedleft autre chose };
+    \leftRect{Help/Collaboration :}{- If anybody has any idea of an interesting LSTM/NN problem to work on.\\-Also if any phenomen happening are familiar to you, I'll take the help}
+    \rightRect{Recommendations :}{I worked on a latex/beamer version of the template}
   \end{tikzpicture}

### 2023-06-05 diapo/05-06-23_meeting/main.tex
Lowered logos
Changed slides' path

@@
 \usetheme{gpmeet}
+\graphicspath{{../../figures/}}
 
@@
   \begin{tikzpicture}[overlay,remember picture]
-    \node[left=0cm] at (current page.34){
-      \includegraphics[height=1cm]{../figures/ist.eps}
+    \node[below=1pt] at (current page.55){
+      \includegraphics[height=1cm]{logos/ist}
       \hspace{1pt}
-      \includegraphics[height=1cm]{figures/inesc-mn.png}
-      \includegraphics[height=1cm]{../figures/inesc-id.eps}
+      \includegraphics[height=1cm]{logos/inesc-mn.png}
+      \includegraphics[height=1cm]{logos/inesc-id.eps}
     };
@@
     \frametitle{Introduction}
-  \centering\includegraphics[width=\textwidth]{figures/LSTM-NP}
+    \centering\includegraphics[width=\textwidth]{lstm/lstm-np}
   \end{frame}
@@
   \begin{frame}{What's new}
-  \includegraphics[width=\textwidth]{figures/fulltrial1-all-curves}
+    \includegraphics[width=\textwidth]{fulltrial1/fulltrial1-all-curves}
   \end{frame}
@@
     \centering
-  \includegraphics[width=0.7\textwidth]{figures/fulltrial1-memory-prob}
+    \includegraphics[width=0.7\textwidth]{fulltrial1/fulltrial1-memory-prob}
   \end{frame}
@@
     \centering
-  \includegraphics[width=1\textwidth]{figures/fulltrial1-mem-act-late}
+    \includegraphics[width=1\textwidth]{fulltrial1/fulltrial1-mem-act-late}
   \end{frame}
@@
     \centering
-  \includegraphics[width=1\textwidth]{figures/fulltrial1-final-res}
+    \includegraphics[width=1\textwidth]{fulltrial1/fulltrial1-final-res}
   \end{frame}

### 2023-06-06 diapo/05-06-23_meeting/main.tex
New file, plus pre change to slides
New figure + updated diapo
New version of the slides
One more help i need

@@
     \item<2-> The memory input seems to be active for too long
-    \item<3-> The output result isn't what is expected    \end{itemize}
+    \item<3-> The output of the sigmoid doesn't look right
+    \item<4-> The output result isn't what is expected
+  \end{itemize}
 \end{frame}
@@
 \begin{frame}{Problems/doubts}
+  Here we can see that the output of the sigmoid isn't right
+  \begin{figure}[!tbp]
+    \centering
+    \begin{minipage}[b]{0.4\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{fulltrial1/sigmoid-prob.png}
+    \end{minipage}
+    \hspace{20pt}
+    \begin{minipage}[b]{0.4\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{activation/sigmoid-prob}
+    \end{minipage}
+  \end{figure}
+\end{frame}
+
+\againframe<4>{prob}
+\begin{frame}{Problems/doubts}
   The output is ~0.791 V. This is $-1.09$ and should be $0.015$
@@
   \centering
-    \includegraphics[width=1\textwidth]{fulltrial1/fulltrial1-final-res}
+  \includegraphics[width=\textwidth]{fulltrial1/fulltrial1-final-res}
 \end{frame}
@@
   \begin{tikzpicture}
-      \leftRect{Help/Collaboration :}{- If anybody has any idea of an interesting LSTM/NN problem to work on.\\-Also if any phenomen happening are familiar to you, I'll take the help}
+    \leftRect{Help/Collaboration :}{- If anybody has any idea of an interesting LSTM/NN problem to work on.\\-Also if any phenomen happening are familiar to you, I'll take the help\\- I would need the feature size of a typical memristor}
 

### 2023-06-26 diapo/26-06-23_meeting/main.tex
New group meeting

--- 2023-06-06 diapo/05-06-23_meeting/main.tex
@@
 
-\date{05/06/2023}
+\date{12/06/2023}
 
@@
     \item<2-> The memory input seems to be active for too long
-    \item<3-> The output of the sigmoid doesn't look right
-    \item<4-> The output result isn't what is expected
+    \item<3-> The output of the sigmoid doesn't look right (Now fixed)
+    \item<4-> The output result isn't what is expected (Scale issue)
   \end{itemize}
@@
 
+\begin{frame}{Problems/doubts}
+  To fix that I had to re-adjust the reference voltages (sigmoid):
+  \begin{figure}[!tbp]
+    \centering
+    \begin{minipage}[b]{0.4\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{activation/sigmoid}
+      \caption{before}
+    \end{minipage}
+    \hspace{20pt}
+    \begin{minipage}[b]{0.4\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{activation/realSigmoid}
+      \caption{after}
+    \end{minipage}
+  \end{figure}
+\end{frame}
+
+\begin{frame}{Problems/doubts}
+  To fix that I had to re-adjust the reference voltages (tanh):
+  \begin{figure}[!tbp]
+    \centering
+    \begin{minipage}[b]{0.4\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{activation/tanh}
+      \caption{before}
+    \end{minipage}
+    \hspace{20pt}
+    \begin{minipage}[b]{0.4\textwidth}
+      \centering
+      \includegraphics[width=\textwidth]{activation/realTanh}
+      \caption{after}
+    \end{minipage}
+  \end{figure}
+\end{frame}
+
 \againframe<4>{prob}
 \begin{frame}{Problems/doubts}
-  The output is ~0.791 V. This is $-1.09$ and should be $0.015$
+  The output is weirdly on a different scale but gives out the right overall values. (Scale factor ~9022)
 
   \centering
-  \includegraphics[width=\textwidth]{fulltrial1/fulltrial1-final-res}
+  \includegraphics[width=.8\textwidth]{output/results0}
 \end{frame}
@@
   \rowcolors{2}{evenRow}{oddRow}
-  \begin{tabular}{ c m{4cm} }
+  \begin{tabular}{ c m{6cm} }
     \rowcolor{firstRow}
@@
     1 & Find another LSTM problem \\
-    2 & "Debug" the circuit \\
-    2.1 & Run the entire simulation \\
-    3 & Work on thesis \\
+    2 & Find out what is causing the issue \\
+    3 & Find formula for approximate onChip area \\
+    4 & Work on thesis \\
   \end{tabular}

### 2023-06-26 diapo/26-06-23_meeting/main.tex
Finished juin 26 pres

--- 2023-06-05 diapo/05-06-23_meeting/main.tex
@@
 \usetheme{gpmeet}
+\setbeamertemplate{caption}{\raggedright\insertcaption\par}
 \graphicspath{{../../figures/}}
@@
     \rowcolors{2}{evenRow}{oddRow}
-    \begin{tabular}{ c m{3cm} c m{3cm} c }
+    \begin{tabular}{ c m{4cm} c m{2cm} c }
       \rowcolor{firstRow}
       \color{white}\textbf{\#} & \centering\color{white}Task & \color{white}Done? & \color{white}Why not? & \color{white}Help? \\
-      1 & Try to use LTSpice & No & Found out I could do what I want with Cadence & No\\
-      2 & Find another\newline problem to work on & No & Lack of time & Yes\\
-      3 & Work on the thesis intro & on going &  & No\\
+      1 & Find another LSTM problem & On going & Didn't find & Yes\\
+      2 & Find out what is causing the issue & Partly & I had a scale of 10 000 and reduced it to 1.5 & No\\
+      3 & Find formula for approximate onChip area & No & Can be done later & No\\
+      4 & Work on thesis & On going & & No\\
     \end{tabular}
@@
             \color{text}
-          \item With the help of Fabio, I can now simulate all the steps automatically
-          \item I successfully ran the simulation
-          \item The result are not correct
+          \item I reduced the factor of the results
+          \item The factor is now tolerable, will have to see how it behaves with other problems
         \end{itemize}
@@
             \color{text}
-          \item Debug the whole thing
+          \item Reduce the still existing factor
+        \end{itemize}
+      \item Relevant information :
+        \begin{itemize}
+            \color{text}
+          \item I reduced the factor to 16 first, and then to 1.5.
         \end{itemize}
@@
   \begin{frame}{What's new}
-    \includegraphics[width=\textwidth]{fulltrial1/fulltrial1-all-curves}
+    Here the scale factor is almost the same with the serialized and parallel version ( 1.50 vs 1.58 ).
+    \includegraphics[width=\textwidth]{output/factor1.5.png}
   \end{frame}
 
-  \begin{frame}<1>[label=prob]
+  \begin{frame}
     \frametitle{Problems/doubts}
     \begin{itemize}
-      \item<1-> The memory doesn't output the right values
-      \item<2-> The memory input seems to be active for too long
-    \item<3-> The output result isn't what is expected    \end{itemize}
+      \item The results are still not quite right.
+      \item I am wondering whether it is still due to the activation function circuit.
+    \end{itemize}
   \end{frame}
@@
   \begin{frame}{Problems/doubts}
-    The memory doesn't output the right values.
-
+    I had to train using custom activation function but still not perfect.
+    \begin{figure}[!tbp]
       \centering
-    \includegraphics[width=0.7\textwidth]{fulltrial1/fulltrial1-memory-prob}
+      \begin{minipage}[b]{0.4\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{activation/sigmoid}
+        \caption{ $Sigmoid(x)$ - RSME : 0.0051}
+      \end{minipage}
+      \hspace{20pt}
+      \begin{minipage}[b]{0.4\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{activation/cSigmoid}
+        \caption{ $ Sigmoid(0.67\cdot x) $ - RSME : 0.0014}
+      \end{minipage}
+      \caption{Sigmoid}
+    \end{figure}
   \end{frame}
 
-  \againframe<2>{prob}
-
   \begin{frame}{Problems/doubts}
+    I had to train using custom activation function but still not perfect.
+    \begin{figure}[!tbp]
       \centering
-    \includegraphics[width=1\textwidth]{fulltrial1/fulltrial1-mem-act-late}
-  \end{frame}
-
-  \againframe<3>{prob}
-  \begin{frame}{Problems/doubts}
-    The output is ~0.791 V. This is $-1.09$ and should be $0.015$
-
+      \begin{minipage}[b]{0.4\textwidth}
         \centering
-    \includegraphics[width=1\textwidth]{fulltrial1/fulltrial1-final-res}
+        \includegraphics[width=\textwidth]{activation/tanh}
+        \caption{ $ Tanh(x) $ - RSME : 0.022}
+      \end{minipage}
+      % \rightarrow
+      \hspace{20pt}
+      \begin{minipage}[b]{0.4\textwidth}
+        \centering
+        \includegraphics[width=\textwidth]{activation/cTanh}
+        \caption{ $2\cdot Sigmoid(0.67\cdot x)-1$ - RSME : 0.0029}
+      \end{minipage}
+      \caption{Tanh}
+    \end{figure}
   \end{frame}
@@
     \rowcolors{2}{evenRow}{oddRow}
-    \begin{tabular}{ c m{4cm} }
+    \begin{tabular}{ c m{6cm} }
       \rowcolor{firstRow}
       \color{white}\textbf{\#} & \centering\color{white}Task \cr
+      0 & Work on thesis \\
+      0.1 & Formula onChip Area \\
       1 & Find another LSTM problem \\
-      2 & "Debug" the circuit \\
-      2.1 & Run the entire simulation \\
-      3 & Work on thesis \\
+      2 & Try to reduce the scale factor again \\
+      2.1 & Tensorflow custom function \\
     \end{tabular}
@@
     \begin{tikzpicture}
-      \leftRect{Help/Collaboration :}{- If anybody has any idea of an interesting LSTM/NN problem to work on.\\-Also if any phenomen happening are familiar to you, I'll take the help}
+      \leftRect{Help/Collaboration :}{- If anybody has any idea of an interesting LSTM/NN problem to work on.}
 
-      \rightRect{Recommendations :}{I worked on a latex/beamer version of the template}
+      \rightRect{Recommendations :}{None}
     \end{tikzpicture}

### 2023-06-26 diapo/26-06-23_meeting/main.tex
Updated pres for tonight

@@
   \begin{frame}{What's new}
-    Here the scale factor is almost the same with the serialized and parallel version ( 1.50 vs 1.58 ).
+    Here the scale factor is almost the same with the serialized and parallel version (1.50 vs 1.58).
     \includegraphics[width=\textwidth]{output/factor1.5.png}
+  \end{frame}
+
+  \begin{frame}{What's new}
+    Today, I managed to reduce the scale factor even more. (1.21)
+    \includegraphics[width=\textwidth]{output/factor1.2.png}
   \end{frame}

### 2023-06-26 diapo/26-06-23_meeting/main.tex
Changed date
Small adjustments

@@
 
-\date{12/06/2023}
+\date{26/06/2023}
 
@@
     \rowcolors{2}{evenRow}{oddRow}
-    \begin{tabular}{ c m{4cm} c m{2cm} c }
+    \begin{tabular}{c p{0.3\textwidth} c p{0.3\textwidth} c}
       \rowcolor{firstRow}
@@
       1 & Find another LSTM problem & On going & Didn't find & Yes\\
-      2 & Find out what is causing the issue & Partly & I had a scale of 10 000 and reduced it to 1.5 & No\\
+      2 & Find out what is causing the issue & Partly & I had a scale of 10000 and reduced it to 1.5 & No\\
       3 & Find formula for approximate onChip area & No & Can be done later & No\\

### 2023-06-26 diapo/26-06-23_meeting/main.tex
Updated objectives

@@
       0.1 & Formula onChip Area \\
+      0.2 & power consumption \\
       1 & Find another LSTM problem \\
-      2 & Try to reduce the scale factor again \\
-      2.1 & Tensorflow custom function \\
+      1.1 & Mbedded Video stabilization \\
+      2 & Digital inference time \\
     \end{tabular}

### 2023-06-30 1.Introduction/1.motivation.tex
Changed paths in main doc
Worked on motivation

@@
 Nowadays, it's impossible not to have heard about \acp{AI}, they are everywhere and everybody talks about them.
-The most recent example being \textit{ChatGPT} from \textbf{OpenAI} and is getting very popular even outside of the "computer community"(TODO : Find better term).
+The most recent example being \textit{ChatGPT} from \textbf{OpenAI} and is getting very popular even among casual computer users.
 Everybody is using those tools.
 It's clear that \ac{AI} is becoming more and more important in the current world.
+
+\acp{AI} are very power hungry algorithms which limits its use in embedded systems and power efficient devices.
+Furthermore, the time an \ac{AI} takes to compute its output can become quite long.
+So much so that the most complex ones only run on online and very powerful servers, like the aforementioned \textit{ChatGPT}.
+
+There are several options to reduce execution speed and energy consumption such as running the algorithm on a \ac{GPU} rather than a \ac{CPU}.
+Using an \ac{FPGA} or \ac{ASIC} are other ways to improve power consumption and execution time.
+The latter is the most restricting (as it's name implies) but has the best results.
+
+Using an \ac{ASIC} allows the use of analog computation. Analog computation offers great advantages compared to digital computers as it's much faster while being very power-efficient.
+
+% For the State of the art
 
@@
   \centering
-    \includesvg[height=8cm]{Figures/Colored_neural_network.svg}
+  \includesvg[height=8cm]{Colored_neural_network.svg}
   \caption{Simple \acl{NN}}
@@
   \centering
-    \includesvg[height=8cm]{Figures/NN_explained.svg}
+  \includesvg[height=8cm]{NN_explained.svg}
   \caption{\aclp{NN} with names}

### 2023-07-04 1.Introduction/1.motivation.tex
Worked on motivation

@@
 
-There are several options to reduce execution speed and energy consumption such as running the algorithm on a \ac{GPU} rather than a \ac{CPU}.
+There are several options to reduce execution speed and energy consumption such as running the algorithm on a \ac{GPU} rather than on a \ac{CPU}.
 Using an \ac{FPGA} or \ac{ASIC} are other ways to improve power consumption and execution time.
@@
 
-% For the State of the art
+Making an analog chip with a very low power consumption allows to use chip like that in embedded systems such as video surveillance cameras.
+A small chip could be installed inside the system that could be able to do gait detection \cite{gaitDS,gaitDig,gait}.
+The video surveillance camera would then be able to know who is showing up on the camera and if it's not a known person send an alert.
 
-\ac{AI} is wide topic, usually when we talk about \acp{AI}, we refer to the most common type, Machine Learning.
-Especially, a sub part of Machine Learning called Deep Learning.
-Deep Learning has seen a surge in research the last decade, which lead to great advancement in this field and the numerous \ac{AI} tools that popped up in the last years.
-Deep Learning works using what is called \acp{NN}.
+Such a chip could also be used for video stabilization \cite{videoStab}.
+It would allow to directly store stabilized video, and not have to use online servers like described in \cite{videoStab}.
+Any action camera would benefit greatly from such an improvement since they're mostly used with a lot of movements.
 
-Those \acp{NN} require matrix multiplication.
-This kind of operation is very time and energy consuming to run on a classic computer.
-There are options to better run those operations.
-GPUs can be better suited than a CPU for a \ac{NN} application as it is made for 3D graphics that also mainly use Matrix Multiplication.
-Other specialized hardware such as FPGAs and ASICs can improve both execution time and energy consumption.
-All those are digital computers, using analog computing could highly improve speed, power consumption and even smaller chip area.
-All this using a memristive crossbar array \cite{Xbar}.
-An analog computer
-The point of this thesis is to lay out the ground work to create a LSTM capable chip to use in embedded systems. It could be used detection and prediction of any kind of sequential events.
-Using those memristors \cite{TheoMemristor}, which work like resistance with a memory component to them, that means that they can be set to a desired value.
-
-\subsection{Neural Networks}\label{sec:nn}
-
-\acfp{NN} are a network of several layers of artificial neurons.
-Figure \ref{fig:snn}, shows a simple representation of a \ac{NN}, the artificial neurons are the represented by the colored circles.
-On figure \ref{fig:snn} each arrow represent a synapse.
-
-\begin{figure}[h!]
-  \centering
-  \includesvg[height=8cm]{Colored_neural_network.svg}
-  \caption{Simple \acl{NN}}
-  \label{fig:snn}
-\end{figure}
-
-They are several layers to a \ac{NN} :
-\begin{itemize}
-  \item Input layer : This layer is simply the different inputs.
-  \item Hidden layer : This layer can be (and usually is) wider than the one in figure \ref{fig:snn}, it is there to add more layers and thus more precision to the result
-  \item Output layer : This layer is where you can find the result from the \ac{NN}.
-\end{itemize}
-
-In a computational \ac{NN}, synapses are represented by weights.
-Those weights are to be multiplied by the last neuron and then added to each other to produce the next stage.
-Using the names defined in figure \ref{fig:nn_explained}, the equation linking each layer is the following matrix multiplication :
-\begin{equation}
-  \begin{bmatrix}
-    H1\\ H2\\ H3\\ H4\\
-  \end{bmatrix}
-  =
-  \begin{bmatrix}
-    W1,1 & W1,2 & W1,3\\
-    W2,1 & W2,2 & W2,3\\
-    W3,1 & W3,2 & W3,3\\
-    W4,1 & W4,2 & W4,3\\
-  \end{bmatrix}
-  \cdot
-  \begin{bmatrix}
-    I1\\ I2\\ I3\\
-  \end{bmatrix}
-\end{equation}
-
-And it works the same way to go to the next layer.
-
-\begin{figure}[h!]
-  \centering
-  \includesvg[height=8cm]{NN_explained.svg}
-  \caption{\aclp{NN} with names}
-  \label{fig:nn_explained}
-\end{figure}
-
-Those matrix multiplication are very power-hungry and their computation time scales up with the size of the \ac{NN} using classical computing.
-
-Analog computation enables those same calculations almost instantly and being far more power-efficient. This can be done using the physical properties of electrical devices.
-We want to reproduce one of those neural network using electrical circuits to perform the same computation that a computer would in much faster manner.
-Using \hyperref[subsec:memristors]{memristors} as weights. This removes the need to copy the weights from the main memory
+Another use for \ac{AI} that works great in embedded use is for camera relocalization \cite{videoReloc}.
+That is useful especially for VR headsets that need to know their location in the room.
+Having an embedded chip to do all this computation.

### 2023-07-04 2.State/nn.tex
Worked on motivation

% For the State of the art

\ac{AI} is wide topic, usually when we talk about \acp{AI}, we refer to the most common type, Machine Learning. Especially, a sub part of Machine Learning called Deep Learning. Deep Learning has seen a surge in research the last decade, which lead to great advancement in this field and the numerous \ac{AI} tools that popped up in the last years.
Deep Learning works using what is called \acp{NN}.

Those \acp{NN} require matrix multiplication. This kind of operation is very time and energy consuming to run on a classic computer. There are options to better run those operations. GPUs can be better suited than a CPU for a \ac{NN} application as it is made for 3D graphics that also mainly use Matrix Multiplication. Other specialized hardware such as FPGAs and ASICs can improve both execution time and energy consumption. All those are digital computers, using analog computing could highly improve speed, power consumption and even smaller chip area. All this using a memristive crossbar array \cite{Xbar}. An analog computer
The point of this thesis is to lay out the ground work to create a LSTM capable chip to use in embedded systems. It could be used detection and prediction of any kind of sequential events.
Using those memristors \cite{TheoMemristor}, which work like resistance with a memory component to them, that means that they can be set to a desired value.

\section{Neural Networks}\label{sec:nn}

\acfp{NN} are a network of several layers of artificial neurons. Figure \ref{fig:snn}, shows a simple representation of a \ac{NN}, the artificial neurons are the represented by the colored circles. On figure \ref{fig:snn} each arrow represent a synapse.

\begin{figure}[h!]
  \centering
  \includesvg[height=8cm]{NN_explained.svg}
  \caption{Simple \acl{NN}}
  \label{fig:snn}
\end{figure}

They are several layers to a \ac{NN} :
\begin{itemize}
  \item Input layer : This layer is simply the different inputs.
  \item Hidden layer : This layer can be (and usually is) wider than the one in figure \ref{fig:snn}, it is there to add more layers and thus more precision to the result
  \item Output layer : This layer is where you can find the result from the \ac{NN}.
\end{itemize}

In a computational \ac{NN}, synapses are represented by weights. Those weights are to be multiplied by the last neuron and then added to each other to produce the next stage. Using the names defined in figure \ref{fig:snn}, the equation linking each layer is the following matrix multiplication :
\begin{equation}
  \begin{bmatrix}
    H1\\ H2\\ H3\\ H4\\
  \end{bmatrix}
  =
  \begin{bmatrix}
    W1,1 & W1,2 & W1,3\\
    W2,1 & W2,2 & W2,3\\
    W3,1 & W3,2 & W3,3\\
    W4,1 & W4,2 & W4,3\\
  \end{bmatrix}
  \cdot
  \begin{bmatrix}
    I1\\ I2\\ I3\\
  \end{bmatrix}
\end{equation}

And it works the same way to go to the next layer.

Those matrix multiplication are very power-hungry and their computation time scales up with the size of the \ac{NN} using classical computing.

Analog computation enables those same calculations almost instantly and being far more power-efficient. This can be done using the physical properties of electrical devices.
We want to reproduce one of those neural network using electrical circuits to perform the same computation that a computer would in much faster manner.
Using \hyperref[subsec:memristors]{memristors} as weights. This removes the need to copy the weights from the main memory

### 2023-07-05 1.Introduction/1.motivation.tex
Bunch of changes, mainly great advance in lstm state of the art

@@
 That is useful especially for VR headsets that need to know their location in the room.
-Having an embedded chip to do all this computation.
+The embedded chip would allow this computation to be much faster than other methods, and thus give more realistic results.
+
+The point of this thesis is to lay out the ground work to create a LSTM capable chip to use in embedded systems. It could be used detection and prediction of any kind of sequential events.

### 2023-07-05 2.State/lstm.tex
Working on lstm
Bunch of changes, mainly great advance in lstm state of the art

--- 2023-05-29 2.State/4.lstm.tex
@@
 
-\acp{LSTM} are able to forget information passed.
+\acp{LSTM} were born from the vanishing/exploding gradient problem.
+Traditionnal \ac{RNN} have the ability to model sequential events by propagating through time, for example forward and backward propagation.
+This is achieved by connecting these sequential events with the hidden state:
+$$h_t=f(h_{t-1},x_t)$$
+The hidden state $h_t$ carries all the past informations for the next time step.
+Even though \acp{RNN} work very well, they are subject to the vanishing/exploding gradient problem. This issue occurs when too many time steps are chained.
+
+\acp{LSTM} alleviate this issue by adding a cell state, this state gives it the ability to choose what input is important and which one isn't, thus giving it a long term memory.
 This ability gave the uncommon name of \acl{LSTM} as it has both long and short term memory.
@@
   \centering
-  \includesvg[width=\textwidth]{Figures/lstmCellLatex.svg}
+  \includesvg[width=\textwidth]{lstm/lstmCell.svg}
   \caption{LSTM Cell}
@@
 
-Figure \ref{fig:lstmCell} shows the complexity of the \ac{LSTM} architecture. In an \ac{LSTM}, each gate is a different \ac{NN}.
+Figure \ref{fig:lstmCell} shows the complexity of the \ac{LSTM} architecture.
+In an \ac{LSTM}, each gate is a different \ac{NN} and then activated with either a tanh or a sigmoid activation function.
+Each input to the cell is a vector.
+Those vector are of varying sizes depending on several factors.
+For example, both $h_t$ and $c_t$ are of the same size as the number of hidden state (sometimes refered to as cell state) for any $t\geq 0$.
+The input vector ($x_t$) is of size of the sample we want to feed for each time step.
 
-An \ac{LSTM} works using different gates
+\subsection{Equations}
+
+The equations of an LSTM are quite unusual.
+Let's start with the more classic gate equations. They are the ones that behave like the more classic \ac{NN}.
+The input (equation \ref{eq:inputG}), forget (equation \ref{eq:forgetG}) and ouput (equation \ref{eq:ouputG}) gates are described below.
+
+\begin{equation}\label{eq:inputG}
+  i_t=\sigma (w_i\cdot[x_{t_1},h_{t-1}] + b_i)
+\end{equation}
+\begin{equation}\label{eq:forgetG}
+  f_t=\sigma (w_f\cdot[x_{t_1},h_{t-1}] + b_f)
+\end{equation}
+\begin{equation}\label{eq:ouputG}
+  o_t=\sigma (w_o\cdot[x_{t_1},h_{t-1}] + b_o)
+\end{equation}
+
+Where ($w_i$,$b_i$), ($w_f$,$b_f$) and ($w_o$,$b_o$) are the pair of weights and biases for the input, forget and ouput gates respectively.
+$x_t$ is the input vector and $h_t$ is the hidden state vector.
+
+For the next equation describes the candidate cell state (equation \ref{eq:candCell}), that will next become the cell state (equation \ref{eq:cellS}).
+
+\begin{equation}\label{eq:candCell}
+  \tilde{c}_t=tanh(w_c\cdot[x_{t-1},h_{t-1}] + b_c)
+\end{equation}
+\begin{equation}\label{eq:cellS}
+  c_t=f_t\cdot c_{t-1} + i_t \cdot \tilde{c}_t
+\end{equation}
+
+Where $w_c$ and $b_c$ are the weights and bias for the candidate cell state.
+
+The final step of the \ac{LSTM} is to compute the hidden state (equation \ref{eq:hiddenS}).
+\begin{equation}\label{eq:hiddenS}
+  h_t=o_t\cdot tanh(c_t)
+\end{equation}
+
+We set $x_0$ as the first input and define $h_0$ as a zero only vector.

### 2023-07-05 2.State/nn.tex
Working on lstm
Bunch of changes, mainly great advance in lstm state of the art

@@
-% For the State of the art
+\section{Neural Networks}\label{sec:nn}
 
-\ac{AI} is wide topic, usually when we talk about \acp{AI}, we refer to the most common type, Machine Learning.
+\ac{AI} is wide topic, it refers to any computer system that can make a descision.
+It can take many forms but usually when we talk about \acp{AI}, we refer to the most common type, Machine Learning.
 Especially, a sub part of Machine Learning called Deep Learning.
@@
 Deep Learning works using what is called \acp{NN}.
-
-Those \acp{NN} require matrix multiplication.
-This kind of operation is very time and energy consuming to run on a classic computer.
-There are options to better run those operations.
-GPUs can be better suited than a CPU for a \ac{NN} application as it is made for 3D graphics that also mainly use Matrix Multiplication.
-Other specialized hardware such as FPGAs and ASICs can improve both execution time and energy consumption.
-All those are digital computers, using analog computing could highly improve speed, power consumption and even smaller chip area.
-All this using a memristive crossbar array \cite{Xbar}.
-An analog computer
-The point of this thesis is to lay out the ground work to create a LSTM capable chip to use in embedded systems. It could be used detection and prediction of any kind of sequential events.
-Using those memristors \cite{TheoMemristor}, which work like resistance with a memory component to them, that means that they can be set to a desired value.
-
-\section{Neural Networks}\label{sec:nn}
 
@@
   \begin{bmatrix}
-    H1\\ H2\\ H3\\ H4\\
+    H_1\\ H_2\\ H_3\\ H_4\\
   \end{bmatrix}
@@
   \begin{bmatrix}
-    W1,1 & W1,2 & W1,3\\
-    W2,1 & W2,2 & W2,3\\
-    W3,1 & W3,2 & W3,3\\
-    W4,1 & W4,2 & W4,3\\
+    W_{1,1} & W_{1,2} & W_{1,3}\\
+    W_{2,1} & W_{2,2} & W_{2,3}\\
+    W_{3,1} & W_{3,2} & W_{3,3}\\
+    W_{4,1} & W_{4,2} & W_{4,3}\\
   \end{bmatrix}
@@
   \begin{bmatrix}
-    I1\\ I2\\ I3\\
+    I_1\\ I_2\\ I_3\\
   \end{bmatrix}

### 2023-07-07 3.Chapter/2.sectionb.tex
Moved placeholder figure

\section{Section B}
\label{sec:sectionb}

\subsection{Subsection A}
\label{subsec:subasectionB}

The model described can also be represented as

\begin{equation}
  \dot{\mathbf{x}}(t) = \mathbf{T}\mathbf{z}(y),\  \mathbf{y}(0) = \mathbf{y}_0,\  z\geq 0 \\
  \label{eq:dummyeq1}
\end{equation}

\noindent where

\begin{equation}
  \mathbf{A} = \left[ \begin{array}{cc} -(a_{12} + a_{10}) & a_{21} \\ a_{12} & -(a_{21} + a_{20}) \end{array} \right],\ \mathbf{x} = \left[ \begin{array}{c} x_1 \\ x_2 \end{array} \right] \\
    \label{eq:dummyeq2}
\end{equation}

\subsection{Subsection B}
\label{subsec:subbsectionB}

\begin{table}[H]
  \centering
  \caption{Dummy Table.}
  \begin{tabular}{|c|c|c|c|} \hline
    \textbf{Vendor Name}        & \textbf{Short Name} & \textbf{Commercial Name}  & \textbf{Manufacturer} \\ \hline \hline
    \multirow{3}{*}{Text in Multiple Row}   & ABC       &  ABC\textreg        & ABC SA               \\ \cline{2-4}
    &        DEF        &  DEF\textreg        & DEF SA        \\ \cline{2-4}
    &        GHF      &  GHF\textreg        & GHF SA        \\ \hline
    Text in Single Row          &        IJK        & IJK\textreg       & IJK SA        \\ \hline
    Frescos SA            &        LMN      & LMN\textreg       & LMN SA        \\ \hline
    Carros Lda.           &    \multicolumn{3}{|c|}{Text in Multiple Column}              \\ \hline
  \end{tabular}
  \label{tab:dummytable}
\end{table}

\begin{figure}[H]
  \centering
  \input{figures/TimeDiagram/basic.tex}
  \caption{Caption}
  \label{fig:my_label}
\end{figure}

### 2023-07-07 3.Circuit/activFunc.tex
Added chapter about circuits

\section{The activation functions}
\label{sec:af}

Activation function play a great role in the results obtained \cite{af}. It is the reason why making good Activation function circuits is fundamental. Of course the easy way would be to simply design a hard sigmoid, the issue being that the hard sigmoid is much worse than the regular sigmoid. especially for regression problems \cite{hardSigm}.

\subsection{Subsection A}
\label{subsec:subasectionA}

This would be a citation \cite{dummy}.

\ac{acro}
% The first time you use this, the acronym will be written in full with the acronym in parentheses: supernova (SN). At later times it will just print the acronym: SN.

\acf{acro}
% written out form with acronym in parentheses, irrespective of previous use

\acs{acro}
% acronym form, irrespective of previous use

\acl{acro}
% written out form without following acronym

\acp{acro}
% plural form of acronym by adding an s. \acfp. \acsp, \aclp work as well.

As seen in \cite{wiki}. \emph{Enfatizar}

\subsection{Subsection B}
\label{subsec:subbsectiona}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\linewidth]{cover}
  \caption[Dummy Figure Caption for List of Figures.]{Dummy Figure Caption.}
  \label{fig:dummyfigure1}
\end{figure}

Remember you can change the reference style. Another dummy citation \cite{site}.

### 2023-07-10 3.Circuit/activFunc.tex
Removed duplicate example
AF

@@
 It is the reason why making good Activation function circuits is fundamental.
-Of course the easy way would be to simply design a hard sigmoid, the issue being that the hard sigmoid is much worse than the regular sigmoid.
-especially for regression problems \cite{hardSigm}.
-
-\subsection{Subsection A}
-\label{subsec:subasectionA}
-
-This would be a citation \cite{dummy}.
-
-\ac{acro}
-% The first time you use this, the acronym will be written in full with the acronym in parentheses: supernova (SN). At later times it will just print the acronym: SN.
-
-\acf{acro}
-% written out form with acronym in parentheses, irrespective of previous use
-
-\acs{acro}
-% acronym form, irrespective of previous use
-
-\acl{acro}
-% written out form without following acronym
-
-\acp{acro}
-% plural form of acronym by adding an s. \acfp. \acsp, \aclp work as well.
-
-As seen in \cite{wiki}. \emph{Enfatizar}
-
-\subsection{Subsection B}
-\label{subsec:subbsectiona}
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=0.5\linewidth]{cover}
-  \caption[Dummy Figure Caption for List of Figures.]{Dummy Figure Caption.}
-  \label{fig:dummyfigure1}
-\end{figure}
-
-Remember you can change the reference style. Another dummy citation \cite{site}.
+Of course the easy way would be to simply design a hard sigmoid, the issue being that the hard sigmoid is much worse than the regular sigmoid, especially for regression problems \cite{hardSigm}.
+Designing analog activation function as close to the original is very important for the final result's quality.

### 2023-07-10 diapo/10-07-23_meeting/main.tex
Prez de ce soir

--- 2023-06-26 diapo/26-06-23_meeting/main.tex
@@
       \color{white}\textbf{\#} & \centering\color{white}Task & \color{white}Done? & \color{white}Why not? & \color{white}Help? \\
-      1 & Find another LSTM problem & On going & Didn't find & Yes\\
-      2 & Find out what is causing the issue & Partly & I had a scale of 10000 and reduced it to 1.5 & No\\
-      3 & Find formula for approximate onChip area & No & Can be done later & No\\
-      4 & Work on thesis & On going & & No\\
+      0 & Work on thesis & onGoing & & No\\
+      0.1 & Formula onChip Area & Almost & & No\\
+      0.2 & power consumption & No & How would I approximate ? & Yes \\
+      1 & Find another LSTM problem & Yes and no & Found one but too complicated & No\\
+      1.1 & Mbedded Video stabilization & Yes & & No\\
+      2 & Digital inference time & No & & No\\
     \end{tabular}
@@
             \color{text}
-          \item I reduced the factor of the results
-          \item The factor is now tolerable, will have to see how it behaves with other problems
+          \item
         \end{itemize}
@@
             \color{text}
-          \item Reduce the still existing factor
+          \item
         \end{itemize}
@@
             \color{text}
-          \item I reduced the factor to 16 first, and then to 1.5.
+          \item I started really focusing on the writing of the thesis. I'll have few to no updates from now on. (No idea what else to do)
         \end{itemize}
@@
 
-  \begin{frame}{What's new}
-    Here the scale factor is almost the same with the serialized and parallel version (1.50 vs 1.58).
-    \includegraphics[width=\textwidth]{output/factor1.5.png}
-  \end{frame}
-
-  \begin{frame}{What's new}
-    Today, I managed to reduce the scale factor even more. (1.21)
-    \includegraphics[width=\textwidth]{output/factor1.2.png}
-  \end{frame}
-
   \begin{frame}
     \frametitle{Problems/doubts}
+    I am wondering what goes in the :
     \begin{itemize}
-      \item The results are still not quite right.
-      \item I am wondering whether it is still due to the activation function circuit.
+      \item abstract
+      \item thesis outline
     \end{itemize}
-  \end{frame}
-
-  \begin{frame}{Problems/doubts}
-    I had to train using custom activation function but still not perfect.
-    \begin{figure}[!tbp]
-      \centering
-      \begin{minipage}[b]{0.4\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{activation/sigmoid}
-        \caption{ $Sigmoid(x)$ - RSME : 0.0051}
-      \end{minipage}
-      \hspace{20pt}
-      \begin{minipage}[b]{0.4\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{activation/cSigmoid}
-        \caption{ $ Sigmoid(0.67\cdot x) $ - RSME : 0.0014}
-      \end{minipage}
-      \caption{Sigmoid}
-    \end{figure}
-  \end{frame}
-
-  \begin{frame}{Problems/doubts}
-    I had to train using custom activation function but still not perfect.
-    \begin{figure}[!tbp]
-      \centering
-      \begin{minipage}[b]{0.4\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{activation/tanh}
-        \caption{ $ Tanh(x) $ - RSME : 0.022}
-      \end{minipage}
-      % \rightarrow
-      \hspace{20pt}
-      \begin{minipage}[b]{0.4\textwidth}
-        \centering
-        \includegraphics[width=\textwidth]{activation/cTanh}
-        \caption{ $2\cdot Sigmoid(0.67\cdot x)-1$ - RSME : 0.0029}
-      \end{minipage}
-      \caption{Tanh}
-    \end{figure}
   \end{frame}
@@
       0.1 & Formula onChip Area \\
-      1 & Find another LSTM problem \\
-      2 & Try to reduce the scale factor again \\
-      2.1 & Tensorflow custom function \\
+      0.2 & power consumption \\
+      1 & Find another LSTM problem (optionnal) \\
+      2 & Digital inference time \\
     \end{tabular}

### 2023-07-10 2.State/memristors.tex
Changed memristor figures to svg, cleaner
Changed one bib name

--- 2023-05-29 2.State/1.memristors.tex
@@
 The flux in this case is not a magnetic flux and is defined as such : $ d\phi=V\cdot dt \implies \phi =  \int V(t) \,dt  $.\\
-The component stayed theoretical until 2008 when it's been implemented in a physical device for the first time \cite{Strukov2008}. It took 37 years to actually have a working device.\\
+The component stayed theoretical until 2008 when it's been implemented in a physical device for the first time \cite{memristorFab}. It took 37 years to actually have a working device.\\
 There is then an extention of this theoretical device to another, the memristive device.
@@
   \centering
-  \includegraphics{Figures/Memristor.png}
+  \includesvg[width=0.55\textwidth]{memristor/memristor}
   \caption{Fundamental passive components}

### 2023-07-11 3.Design/main.tex
New figures plus renamed 3

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Dummy Chapter:
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% The Introduction:
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\fancychapter{The circuits}
\label{cap:circuit}

In this chapter, I'm going to describe the different circuit designed and used in this thesis.
%\textit{Present the chapter content.}

TODO : Maybe add more ???
Explain 0->0.9V and 1->1V!!

\input{3.Design/activFunc.tex}
\input{3.Design/1.sectiona.tex}
\input{3.Design/2.sectionb.tex}
\cleardoublepage

### 2023-07-12 3.Design/main.tex
Forgot to change one name
chnaged file name
change to the main text chap 3

@@
 % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\fancychapter{The circuits}
-\label{cap:circuit}
+\fancychapter{The design}
+\label{cap:design}
 
-In this chapter, I'm going to describe the different circuit designed and used in this thesis.
-%\textit{Present the chapter content.}
+In this chapter, I'm going to describe the different design decisions taken during the study of the thesis.
 
-TODO : Maybe add more ???
-Explain 0->0.9V and 1->1V!!
+The system will be working with a $V_{dd}$ of $1.8V$. Such a value was chosen because this is a low power system.
 
-\input{3.Design/activFunc.tex}
+The way values are encoded in the analog system will be descibed here as it serves for the entire thesis.
+In order for the system to support negative numbers we're going to use a $V_{cm}$ set to $\frac{V_{dd}}{2}$.
+That means that $V_{cm}=\frac{V_{dd}}{2}=0.9V$.
+This $V_{cm}$ will then describe a zero.
+A step of one was chosen to be $0.1V$ in the analog circuit.
+Table \ref{tab:valConv} shows the conversions from a real number to it's voltage equivalent.
+
+\begin{table}[H]
+\centering
+\caption{Real/Voltage Conversion Table.}
+\begin{tabular}{|c|c|}
+\hline
+\rowcolor{gray}
+Real value & Voltage \\
+       \hline
+       0 & 0.9V \\
+         \hline
+         1 & 1.0V \\
+           \hline
+           $x$ & $\frac{x}{10}+V_{cm}$\\
+             \hline
+             $(V-0.9)\cdot 10$ & $V$\\
+               \hline
+               \end{tabular}
+               \label{tab:valConv}
+               \end{table}
+
+               Since the system cannot reach voltage outside of the operating range with the intended behavior, the voltage is then restricted to $V\in [0,1.8]$.
+This means that the range of real value that the systems can handle is $x\in [-9,9]$.
+
+               Any data inputed in the analog system should have been previously checked to make sure it stays in the range of accepted values.
+
+               \input{3.Design/activation.tex}
                \input{3.Design/1.sectiona.tex}

### 2023-07-13 2.State/lstm.tex
finishing lstms
New lstm figures and text
update to lstm, unfinished

@@
 We set $x_0$ as the first input and define $h_0$ as a zero only vector.
+
+\subsection{Usage}
+
+Using \ac{LSTM} can be a bit tricky.
+Due to it's sequential nature, it takes several time steps.
+Every hidden state ($h_t$) is passed to the next time step as an input.
+This hidden state can be used to compute an output at any time step $t$.
+
+\begin{figure}[H]
+  \centering
+  \includesvg[width=\textwidth]{lstm/lstmUse}
+  \caption{LSTM Cell}
+  \label{fig:lstmCell}
+\end{figure}
+
+\subsection{Variants}
+
+\acp{LSTM} come in reality in a few different flavor.
+Usually, when \acp{LSTN} are mentionned, the version used is the \ac{NP} version.
+This is the most common version because it gives the best results in most situations.
+There are other variants of the \ac{LSTM}, that have slight variations in the equations.

### 2023-07-14 2.State/lstm.tex
Small fix on figure + lstm state changes

@@
 
-\acp{LSTM} were born from the vanishing/exploding gradient problem.
+The first \ac{LSTM} was invented in 1997 by Hochreiter and Schmidhuber \cite{firstLSTM}. \acp{LSTM} changed a lot through the years to become what they are now.
+\acp{LSTM} fixed the vanishing/exploding gradient problem.
 Traditionnal \ac{RNN} have the ability to model sequential events by propagating through time, for example forward and backward propagation.
@@
   \includesvg[width=\textwidth]{lstm/lstmUse}
-  \caption{LSTM Cell}
-  \label{fig:lstmCell}
+  \caption{LSTM use case}
+  \label{fig:lstmUse}
 \end{figure}
@@
 \acp{LSTM} come in reality in a few different flavor.
-Usually, when \acp{LSTN} are mentionned, the version used is the \ac{NP} version.
-This is the most common version because it gives the best results in most situations.
-There are other variants of the \ac{LSTM}, that have slight variations in the equations.
+Usually, when \acp{LSTM} are mentionned, the version used is the \ac{NP} version.
+This is the most common version as those are the equations described on the wikipedia page \cite{wikiLSTM} and used in librairies like Keras \cite{Keras} or pyTorch.
+In fact is that there are total of 8 other variations of \acp{LSTM} \cite{nbLSTM}.
+The difference between each of them varies from one to the other, but they can all be described with new equations.
+
+\subsubsection{Vanilla}
+This variant of the \ac{LSTM} was origanally the only \ac{LSTM} network, hence it's name.
+It differs from the classic \ac{LSTM} with it's use of peephole weights, hence the name of the classic \ac{LSTM} being \acf{NP}.
+The Vanilla \ac{LSTM} is thus defined by those equations \cite{vanillaLSTM, nbLSTM} :
+
+\begin{equation}\label{eq:inputGVanilla}
+  i_t=\sigma (w_i\cdot[x_{t_1},h_{t-1}] + b_i+c_{t-1}\odot p_f)
+\end{equation}
+\begin{equation}\label{eq:forgetGVanilla}
+  f_t=\sigma (w_f\cdot[x_{t_1},h_{t-1}] + b_f+c_{t-1}\odot p_i)
+\end{equation}
+\begin{equation}\label{eq:ouputGVanilla}
+  o_t=\sigma (w_o\cdot[x_{t_1},h_{t-1}] + b_o+c_{t}\odot p_o)
+\end{equation}
+With $p_f$, $p_i$ and $p_o$ being the peephole weights vectors. Their size is the same as the size of the hidden state vector ($h_t$).
+
+The equations that haven't been rewritten simply stay the same.

### 2023-07-14 2.State/lstm.tex
lstm fixes

@@
 Usually, when \acp{LSTM} are mentionned, the version used is the \ac{NP} version.
-This is the most common version as those are the equations described on the wikipedia page \cite{wikiLSTM} and used in librairies like Keras \cite{Keras} or pyTorch.
+This is the most common version as those are the equations described on the wikipedia page \cite{wikiLSTM} and used in librairies like Keras \cite{Keras} or pyTorch \cite{PyTorch}.
 In fact is that there are total of 8 other variations of \acp{LSTM} \cite{nbLSTM}.
@@
 \end{equation}
-With $p_f$, $p_i$ and $p_o$ being the peephole weights vectors. Their size is the same as the size of the hidden state vector ($h_t$).
+With $p_f$, $p_i$ and $p_o$ being the peephole weights vectors. Their size is the same as the size of the hidden state vector ($h_t$). $\odot$ is the pointwise multiplication operator.
 

### 2023-07-14 2.State/crossbar.tex
Worked on motivation
Changed state

--- 2023-05-29 2.State/2.crossbar.tex
@@
 
-Setting memristors in a crossbar array to perform analog matrix multiplication typically called Multiply and Accumulate because of the nature of Matrix Multiplication.
+Setting memristors in a crossbar array allows to perform analog \ac{VMM} also called Multiply and Accumulate.
+This circuit works great for this use as all the computation is done almost instantly and using physics laws.
 Figure \ref{fig:crossbar} shows what a typical crossbar array looks like.
@@
   \centering
-  %\subfloat[Schematics]{\includesvg[width=.45\linewidth]{Figures/crossbar.svg}}\qquad
-  \subfloat[Schematics]{\includegraphics[width=.45\linewidth]{Figures/crossbar.eps}}%\qquad
+  \subfloat[Schematics]{\includesvg[width=.45\linewidth]{crossbar/crossbar}}%\qquad
   \hfill
-  \subfloat[3 dimensional representation]{\includegraphics[width=.45\linewidth]{Figures/crossbar3D.jpg}}\\
+  \subfloat[3 dimensional representation]{\includegraphics[width=.45\linewidth]{crossbar/crossbar3D.jpg}}\\
   \caption{Memristor crossbar array}
@@
   \centering
-  %\includegraphics{Figures/crossbar-node.jpg}
-  \includesvg[height=3.5cm]{Figures/crossbarNode.svg}
+  \includesvg[height=3.5cm]{crossbar/node}
+  %\def\svgheigth{3.5cm}
+  %\input{crossbar/node.pdf_tex}
   \caption{Memristor crossbar node of the $i^{th}$ line and $j^{th}$ column}

### 2023-07-14 2.State/lstm.tex
Changed state
Changed lstm

@@
 This is the most common version as those are the equations described on the wikipedia page \cite{wikiLSTM} and used in librairies like Keras \cite{Keras} or pyTorch \cite{PyTorch}.
-In fact is that there are total of 8 other variations of \acp{LSTM} \cite{nbLSTM}.
+In fact, there are total of 8 other variations of \acp{LSTM} \cite{nbLSTM}.
 The difference between each of them varies from one to the other, but they can all be described with new equations.
 
-\subsubsection{Vanilla}
+\subsubsection{Vanilla \ac{LSTM}}
 This variant of the \ac{LSTM} was origanally the only \ac{LSTM} network, hence it's name.
-It differs from the classic \ac{LSTM} with it's use of peephole weights, hence the name of the classic \ac{LSTM} being \acf{NP}.
+It differs from the classic \ac{LSTM} with it's use of peephole weights, hence the name of the classic \ac{LSTM} being \acl{NP}.
 The Vanilla \ac{LSTM} is thus defined by those equations \cite{vanillaLSTM, nbLSTM} :
@@
 The equations that haven't been rewritten simply stay the same.
+
+\subsubsection{\acf{FGR} \ac{LSTM}}
+This variant of the \ac{LSTM} is the most complex of all.
+This is due to the amount of feedback connections.
+This is basically a more complex version of the Vanilla \ac{LSTM}.
+The modified equations are :
+\begin{equation}\label{eq:inputGVanilla}
+  i_t=\sigma (w_i\cdot[x_{t_1},h_{t-1}]+ R_{ii}\cdot i_{t-1} + R_{if}\cdot f_{t-1} + R{io}\cdot o_{t-1} + b_i+c_{t-1}\odot p_f)
+\end{equation}
+\begin{equation}\label{eq:forgetGVanilla}
+  f_t=\sigma (w_f\cdot[x_{t_1},h_{t-1}]+ R_{fi}\cdot i_{t-1} + R_{ff}\cdot f_{t-1} + R{fo}\cdot o_{t-1} + b_f+c_{t-1}\odot p_i)
+\end{equation}
+\begin{equation}\label{eq:ouputGVanilla}
+  o_t=\sigma (w_o\cdot[x_{t_1},h_{t-1}]+ R_{oi}\cdot i_{t-1} + R_{of}\cdot f_{t-1} + R{oo}\cdot o_{t-1} + b_o+c_{t}\odot p_o)
+\end{equation}
+With $R_{fi}$ being the weight matrix of the input feedback connection to compute the forget gate, and the same goes for the other weight matrices.
+
+Once again, the equations that haven't been rewritten are unchanged.
+
+\subsubsection{\acf{CIFG} \ac{LSTM} or \acf{GRU}}
+The \ac{CIFG} \ac{LSTM} or more commonly known as \ac{GRU}. \cite{gru,gruKeras,gruPyTorch}

### 2023-07-14 2.State/lstm.tex
SMall fix

@@
 \begin{equation}\label{eq:ouputGVanilla}
-  o_t=\sigma (w_o\cdot[x_{t_1},h_{t-1}]+ R_{oi}\cdot i_{t-1} + R_{of}\cdot f_{t-1} + R{oo}\cdot o_{t-1} + b_o+c_{t}\odot p_o)
+  o_t=\sigma (w_o\cdot[x_{t_1},h_{t-1}]+ R_{oi}\cdot i_{t-1} + R_{of}\cdot f_{t-1} + R{oo}\cdot o_{t-1} + b_o+c_{t-1}\odot p_o)
 \end{equation}

### 2023-07-31 2.State/lstm.tex
Chnages, might not work anymore since latexmk...

--- 2023-07-14 2.State/lstm.tex
@@
 The modified equations are :
-\begin{equation}\label{eq:inputGVanilla}
+\begin{equation}\label{eq:inputGFGR}
   i_t=\sigma (w_i\cdot[x_{t_1},h_{t-1}]+ R_{ii}\cdot i_{t-1} + R_{if}\cdot f_{t-1} + R{io}\cdot o_{t-1} + b_i+c_{t-1}\odot p_f)
 \end{equation}
-\begin{equation}\label{eq:forgetGVanilla}
+\begin{equation}\label{eq:forgetGFGR}
   f_t=\sigma (w_f\cdot[x_{t_1},h_{t-1}]+ R_{fi}\cdot i_{t-1} + R_{ff}\cdot f_{t-1} + R{fo}\cdot o_{t-1} + b_f+c_{t-1}\odot p_i)
 \end{equation}
-\begin{equation}\label{eq:ouputGVanilla}
-  o_t=\sigma (w_o\cdot[x_{t_1},h_{t-1}]+ R_{oi}\cdot i_{t-1} + R_{of}\cdot f_{t-1} + R{oo}\cdot o_{t-1} + b_o+c_{t}\odot p_o)
+\begin{equation}\label{eq:ouputGFGR}
+  o_t=\sigma (w_o\cdot[x_{t_1},h_{t-1}]+ R_{oi}\cdot i_{t-1} + R_{of}\cdot f_{t-1} + R{oo}\cdot o_{t-1} + b_o+c_{t-1}\odot p_o)
 \end{equation}

### 2023-08-18 2.State/lstm.tex
Worked on lstm state in train

@@
 \subsubsection{\acf{CIFG} \ac{LSTM} or \acf{GRU}}
-The \ac{CIFG} \ac{LSTM} or more commonly known as \ac{GRU}. \cite{gru,gruKeras,gruPyTorch}
+The \ac{CIFG} \ac{LSTM} or more commonly known as \ac{GRU}.
+\cite{gru,gruKeras,gruPyTorch} As it's name implies, \acl{CIFG}, links the input and forget gates as such, $f_t=1-i_t$.
+Other differences are :
+\begin{itemize}
+  \item No peepholes and output activation function.
+  \item Combined hidden state and cell state.
+  \item The candidate cell state is point wise multiplied with the output gate before the activation function.
+\end{itemize}
+
+This is clearer with the math equations :
+\begin{equation}\label{eq:cellGGRU}
+  \tilde{c}_t = tanh(w_{cx}\cdot x_t + w_{ch}\cdot(o_t \odot h_{t-1}) + b_c)
+\end{equation}
+\begin{equation}\label{eq:hiddenGRU}
+  h_t=(1-i_t)\cdot h_{t-1} + i_t \cdot \tilde{c}_t
+\end{equation}

### 2023-09-02 2.State/lstm.tex
Finished all lstm variants

@@
 
+\subsubsection{\acf{CIFG} \ac{LSTM} or \acf{GRU}}
+The \ac{CIFG} \ac{LSTM} or more commonly known as \ac{GRU}.
+\cite{gru,gruKeras,gruPyTorch} As it's name implies, \acl{CIFG}, links the input and forget gates as such, $f_t=1-i_t$.
+Other differences are :
+\begin{itemize}
+  \item No peepholes and output activation function.
+  \item Combined hidden state and cell state.
+  \item The candidate cell state is point wise multiplied with the output gate before the activation function.
+\end{itemize}
+
+This is clearer with the math equations :
+\begin{equation}\label{eq:cellGGRU}
+  \tilde{c}_t = tanh(w_{cx}\cdot x_t + w_{ch}\cdot(o_t \odot h_{t-1}) + b_c)
+\end{equation}
+\begin{equation}\label{eq:hiddenGRU}
+  h_t=(1-i_t)\cdot h_{t-1} + i_t \cdot \tilde{c}_t
+\end{equation}
+
 \subsubsection{\acf{FGR} \ac{LSTM}}
@@
 
-\subsubsection{\acf{CIFG} \ac{LSTM} or \acf{GRU}}
-The \ac{CIFG} \ac{LSTM} or more commonly known as \ac{GRU}.
-\cite{gru,gruKeras,gruPyTorch} As it's name implies, \acl{CIFG}, links the input and forget gates as such, $f_t=1-i_t$.
-Other differences are :
+\subsubsection{Other small variants}
+
 \begin{itemize}
-  \item No peepholes and output activation function.
-  \item Combined hidden state and cell state.
-  \item The candidate cell state is point wise multiplied with the output gate before the activation function.
+  \item Removing the activation function from the Vanilla \ac{LSTM} for :
+    \begin{itemize}
+      \item the candidate cell and thus becomes $ \tilde{c}_t=(w_c\cdot[x_{t-1},h_{t-1}] + b_c) $, this is called the \ac{NIAF}
+      \item the output of the previous cell state and thus becomes $ h_t=o_t\cdot c_t $, and is called \ac{NOAF}
     \end{itemize}
-
-This is clearer with the math equations :
-\begin{equation}\label{eq:cellGGRU}
-  \tilde{c}_t = tanh(w_{cx}\cdot x_t + w_{ch}\cdot(o_t \odot h_{t-1}) + b_c)
-\end{equation}
-\begin{equation}\label{eq:hiddenGRU}
-  h_t=(1-i_t)\cdot h_{t-1} + i_t \cdot \tilde{c}_t
-\end{equation}
+  \item Using units instead of the respective gates :
+    \begin{itemize}
+      \item \ac{NIG} : $i_t=1$
+      \item \ac{NOG} : $o_t=1$
+      \item \ac{NFG} : $f_t=1$
+    \end{itemize}
+\end{itemize}

### 2023-09-04 3.Design/activation.tex
AF wasn't very productive today...

\section{The activation functions}
\label{sec:af}

Activation functions play a great role in the results obtained \cite{af}. It is the reason why making good Activation function circuits is fundamental. Of course the easy way would be to simply design a hard sigmoid ($h(x)=max(0,min(1,\frac{x+1}{2}))$), the issue being that the hard sigmoid is much worse than the regular sigmoid, especially for regression problems \cite{hardSigm}.

Designing an analog activation function as close to the original is very important for the final result's quality.

\subsection{The circuit}

The circuit used is the same as the one in \cite{thesisRef}, the circuit is the one shown in figure \ref{fig:afCircuit}. The technology used being different, all the parameters had to be determined empirically to best fit a sigmoid shape. The parameters can be found in ( TODO : show af parameters ).

Due to the nature of the functions we want to generate, we will use the same circuit for both a sigmoid and a hyperbolic tangent like functions. The two different functions are generated by changing two parameters.

The functions generated are the same and only differ by their output range.

\begin{figure}[H]
\centering
\includesvg[width=\textwidth]{activation/circuit}
\caption{Activation functions circuit}
\label{fig:afCircuit}
\end{figure}

\subsection{The results}

This circuit outputs a voltage that depends on the input voltage passed on. The relation between the two is shown in figure

### 2023-09-05 3.Design/activation.tex
added memcell part, to finsih

@@
 Of course the easy way would be to simply design a hard sigmoid ($h(x)=max(0,min(1,\frac{x+1}{2}))$), the issue being that the hard sigmoid is much worse than the regular sigmoid, especially for regression problems \cite{hardSigm}.
+The same goes for the \ac{tanh} and hard \ac{tanh} functions.
 
@@
 
-Due to the nature of the functions we want to generate, we will use the same circuit for both a sigmoid and a hyperbolic tangent like functions.
+Due to the nature of the functions we want to generate, we will use the same circuit for both a sigmoid and a \ac{tanh} like functions.
 The two different functions are generated by changing two parameters.

### 2023-09-05 3.Design/memcell.tex
added memcell part, to finsih

\section{Memory cells}
\label{sec:memcell}

The memory cell is a circuit that is able to store an analog value for a limited time. It works using capacitors that have the ability to store a voltage for a short time.

The circuit is shown in figure \ref{fig:memcellCircuit}. The value/voltage is trapped in the capacitor using CMOS switches.

I chose to use a two CMOS switches design to avoid voltage leakage through the swicthes. As figure \ref{fig:memcellLoss} shows that only using 1 CMOS swicth leads to a large memory leak due to the high voltage difference on both sides of the CMOS swicth. Using two CMOS switches allows for this difference to be mitigated.

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth]{memcell/data-loss}
  \caption{Memory conservation in a memory cell with 1 CMOS switch vs 2 CMOS swicthes}
  \label{fig:memcellLoss}
\end{figure}

\begin{figure}[H]
  \centering
  \includesvg[width=\textwidth]{memcell/circuit2}
  \caption{Memory cell circuit}
  \label{fig:memcellCircuit}
\end{figure}

### 2023-09-06 3.Design/activation.tex
Updated af part

@@
 It is the reason why making good Activation function circuits is fundamental.
-Of course the easy way would be to simply design a hard sigmoid ($h(x)=max(0,min(1,\frac{x+1}{2}))$), the issue being that the hard sigmoid is much worse than the regular sigmoid, especially for regression problems \cite{hardSigm}.
+Of course the easy way would be to simply design a hard sigmoid (TODO : Ref annex), the issue being that the hard sigmoid is much worse than the regular sigmoid, especially for regression problems \cite{hardSigm}.
 The same goes for the \ac{tanh} and hard \ac{tanh} functions.
@@
 
-\subsection{The circuit}
+\subsection{Circuit}
 
@@
 
-The functions generated are the same and only differ by their output range.
+The functions generated are the same shape and only differ by their output range.
 
@@
 
-\subsection{The results}
+The approximate onChip area of this circuit is :
 
-This circuit outputs a voltage that depends on the input voltage passed on. The relation between the two is shown in figure
+\begin{equation}
+  A_{af}=2\cdot A_{R_1} + 2\cdot A_{nmos} + 5\cdot A_{R_2} + A_{R_3} +2\cdot A_{opAmp}
+\end{equation}
+With $A_{nmos} = w\cdot l = 9\cdot 10^{-7} \cdot 6 \cdot 10^{-8} = 5.4 \cdot 10^{-14} m^2$
+TODO : Finish calc
+
+\subsection{Symbol}
+The symbol for both the sigmoid and \ac{tanh}
+
+\begin{figure}[H]
+  \centering
+  \hspace*{1.5cm}
+  \subfloat[Sigmoid symbol]{\includesvg[height=2.5cm]{activation/sigmoidSymbol}}%
+  \hfill
+  \subfloat[\ac{tanh} symbol]{\includesvg[height=2.5cm]{activation/tanhSymbol}}%
+  \hspace*{1.5cm}
+  \caption{Activation functions symbols}
+  \label{fig:afSymbol}
+\end{figure}
+
+\subsection{Results}
+
+This circuit outputs a voltage that depends on the input voltage passed on.
+The relation between the two is shown in figure \ref{fig:afGraph}.
+The graph also shows the \ac{RMSE} of each graph compared to the ideal result.
+
+The actual output of the circuit shown in figure \ref{fig:afCircuit} is shown in. TODO
+
+\begin{figure}[H]
+  \centering
+  \includesvg[width=\textwidth]{activation/activGraph}
+  \caption{Input/Output graph of the activation function circuit for both sigmoid and \ac{tanh} functions}
+  \label{fig:afGraph}
+\end{figure}

### 2023-09-06 3.Design/models.tex
Added a new section : models

\section{Verilog models}
\label{sec:models}

Due to a lack of time, some of the more common component were not designed by me but instead were simulated using a Verilog model. The only components that use a verilog model are the voltage multiplier and the \ac{opAmp}.

\subsection{Operational amplifier}

This component is the very famous \ac{opAmp}. I didn't design a specific \ac{opAmp} because of the current range of the system I'm trying to build.

This is the code (figure TODO : in annex) of a perfect \ac{opAmp}. TODO : more description.
TODO : add symbol.

\subsection{Voltage multiplier}

This component while a bit less popular than the latter, is just as useful for our specific use. It allows us to multiply, as its name implies, two voltages. It is used to compute the pointwise multiplications of the \ac{LSTM} (figure \ref{fig:lstmCell}).

It must follow the following equation :

\begin{equation}
V_{out}=(V_{in_1}-V_{cm})\cdot (V_{in_2}-V_{cm}) + V_{cm}
\end{equation}

Where $V_{out}$ is the output voltage of the circuit and $V_{in_1}$ and $V_{in_2}$ are the input voltages.

However, the output of the voltage multiplier needs to be ajusted after because of the voltage value of a unit. Indeed, we need to have $1\cdot 1=1$, but if we set $V_{in_1}=V_{in_2}=1V$, we would find $V_{out}=0.91V$. There is then a small circuit added following the output of the voltage multiplier. TODO : Annexe

### 2023-09-06 2.State/model.tex
Tried to write on this part

\section{Memristor's model}\label{sec:model}
They are several ways to model a memristor (TODO : cite).

One of them, works by applying voltage pulse to the device. The device then change it's internal resistance based on how far from the stable point (usually the initial resistance) it is.
TODO

### 2023-09-06 3.Design/activation.tex
Gros pc pas commits...

--- 2023-09-05 3.Design/activation.tex
@@
 
+\begin{table}[H]
+  \centering
+  \caption{Circuits parameters}
+  \begin{tabular}{|c|c|c|}
+    \hline
+    \rowcolor{gray}
+    Parameter & Sigmoid & \ac{tanh} \\
+    \hline
+    $V_1$ & \multicolumn{2}{c|}{$1.1V$}\\
+    \hline
+    $V_2$ & \multicolumn{2}{c|}{$635mV$}\\
+    \hline
+    $V_3$ & $0.8V$ & $550mV$\\
+    \hline
+    $i_{dc}$ & \multicolumn{2}{c|}{$150uA$}\\
+    \hline
+    $w$ & \multicolumn{2}{c|}{$900nm$}\\
+    \hline
+    $l$ & \multicolumn{2}{c|}{$60nm$}\\
+    \hline
+    $R_1$ & \multicolumn{2}{c|}{$5k\Omega$}\\
+    \hline
+    $R_2$ & \multicolumn{2}{c|}{$10k\Omega$}\\
+    \hline
+    $R_3$ & $2k\Omega$ & $4k\Omega$\\
+    \hline
+  \end{tabular}
+  \label{tab:afPar}
+\end{table}
+
 \subsection{The circuit}
@@
 The technology used being different, all the parameters had to be determined empirically to best fit a sigmoid shape.
-The parameters can be found in ( TODO : show af parameters ).
+The parameters can be found in table \ref{tab:afPar}.
 
@@
 
-This circuit outputs a voltage that depends on the input voltage passed on. The relation between the two is shown in figure
+This circuit outputs a voltage that depends on the input voltage passed on. The relation between the two is shown in figure. %TODO
+
+Note that the actual output of the circuit in figure \ref{fig:afCircuit} is inverted arround $V_{cm}$ a simple circuit such as the one shown in  (TODO show circuit in annex) does inverts the

### 2023-09-06 3.Design/activation.tex

--- 2023-09-06 3.Design/activation.tex
@@
 
+\begin{table}[H]
+  \centering
+  \caption{Circuits parameters}
+  \begin{tabular}{|c|c|c|}
+    \hline
+    \rowcolor{gray}
+    Parameter & Sigmoid & \ac{tanh} \\
+    \hline
+    $V_1$ & \multicolumn{2}{c|}{$1.1V$}\\
+    \hline
+    $V_2$ & \multicolumn{2}{c|}{$635mV$}\\
+    \hline
+    $V_3$ & $0.8V$ & $550mV$\\
+    \hline
+    $i_{dc}$ & \multicolumn{2}{c|}{$150uA$}\\
+    \hline
+    $w$ & \multicolumn{2}{c|}{$900nm$}\\
+    \hline
+    $l$ & \multicolumn{2}{c|}{$60nm$}\\
+    \hline
+    $R_1$ & \multicolumn{2}{c|}{$5k\Omega$}\\
+    \hline
+    $R_2$ & \multicolumn{2}{c|}{$10k\Omega$}\\
+    \hline
+    $R_3$ & $2k\Omega$ & $4k\Omega$\\
+    \hline
+  \end{tabular}
+  \label{tab:afPar}
+\end{table}
+
 \subsection{Circuit}
@@
 
-The actual output of the circuit shown in figure \ref{fig:afCircuit} is shown in. TODO
+Note that the actual output of the circuit in figure \ref{fig:afCircuit} is inverted arround $V_{cm}$ a simple circuit such as the one shown in  (TODO show circuit in annex) does inverts the
 

### 2023-09-06 3.Design/crossbar.tex

\section{Crossbar circuit}
\label{sec:xbarCircuit}

This section is here to show

### 2023-09-06 3.Design/inverter.tex

\section{Inverter}
\label{sec:inv}

This section simply describes a very well know electrical component, the inverter. There is nothing special about this specific inverter. When it a voltage of $v_{dd}$ is applied to the input, the output is grounded and the same goes for the other way around.

\begin{figure}[H]
  \centering
  \hspace*{1.5cm}
  \subfloat[Inverter's circuit]{\label{fig:invCircuit}\includesvg[width=0.2\textwidth]{inverter/invCircuit}}%
  \hfill
  \subfloat[Inverter's symbol]{\label{fig:invSymbol}\includesvg[height=2.5cm]{activation/tanhSymbol}}%
  \hspace*{1.5cm}
  \caption{}
  \label{fig:inv}
\end{figure}

Figure \ref{fig:inv} combines the circuit and the symbol of the inverter used in this thesis.

### 2023-09-07 3.Design/inverter.tex
Not so productive morning

@@
   \centering
-  \hspace*{1.5cm}
-  \subfloat[Inverter's circuit]{\label{fig:invCircuit}\includesvg[width=0.2\textwidth]{inverter/invCircuit}}%
+  \hspace*{2.5cm}
+  \subfloat[Inverter's circuit]{\label{fig:invCircuit}\includesvg[height=3cm]{inverter/invCircuit}}%
   \hfill
-  \subfloat[Inverter's symbol]{\label{fig:invSymbol}\includesvg[height=2.5cm]{activation/tanhSymbol}}%
+  \subfloat[Inverter's symbol]{\label{fig:invSymbol}\includesvg[height=3cm]{inverter/invSymbol}}%
   \hspace*{1.5cm}

### 2023-09-07 3.Design/activation.tex
Not so productive morning
worked on af

@@
   \centering
-  \includesvg[width=\textwidth]{activation/circuit}
+  \includesvg[width=\textwidth]{activation/afCircuit}
   \caption{Activation functions circuit}
@@
 
-\subsection{Symbol}
-The symbol for both the sigmoid and \ac{tanh}
+\subsection{Symbols}
+The symbols for the sigmoid and the \ac{tanh} are separated for better understanding.
 
@@
   \centering
-  \hspace*{1.5cm}
+  \hspace*{0.8cm}
   \subfloat[Sigmoid symbol]{\includesvg[height=2.5cm]{activation/sigmoidSymbol}}%
@@
   \subfloat[\ac{tanh} symbol]{\includesvg[height=2.5cm]{activation/tanhSymbol}}%
-  \hspace*{1.5cm}
-  \caption{Activation functions symbols}
+  \hspace*{0.8cm}
+  \caption{Activation functions symbols with the input pin on the left and the output pin on the right}
   \label{fig:afSymbol}
@@
 \end{figure}
+
+These functions are still a bit different from the original functions (especially for the \ac{tanh}).
+However that doesn't matter too much as the trainning will be happening within the final circuit, all weights will be set in the circuit.
+This is the reason why such a difference doesn't matter.
+As long as the curves have the similar shape, the result won't be drastically affected.

### 2023-09-07 3.Design/memcell.tex
Changed name in file
Not so productive morning
Changes to memcell

@@
 
-The memory cell is a circuit that is able to store an analog value for a limited time. It works using capacitors that have the ability to store a voltage for a short time.
+The memory cell is a circuit that is able to store an analog value for a limited time. It works using capacitors that have the ability to store a voltage for a short period of time.
+
+\subsection{Circuit}
 
@@
 
-I chose to use a two CMOS switches design to avoid voltage leakage through the swicthes.
-As figure \ref{fig:memcellLoss} shows that only using 1 CMOS swicth leads to a large memory leak due to the high voltage difference on both sides of the CMOS swicth.
-Using two CMOS switches allows for this difference to be mitigated.
+\begin{figure}[H]
+  \centering
+  \includesvg[width=\textwidth]{memcell/memCircuit}
+  \caption{Memory cell circuit}
+  \label{fig:memcellCircuit}
+\end{figure}
+
+The cirucit has a two CMOS switches design to avoid voltage leakage through the swicthes.
+The system has voltage leakage when only one CMOS swicth, and thus leads to a large memory leak.
+This is due to the high voltage difference between the two sides of the CMOS swicth.
+Using two CMOS switches allows for this difference to be mitigated (figure \ref{fig:memcellLoss}).
 
@@
 
+\subsection{Symbol}
+
+The symbol for this circuit is designed to show a capacitor because it is its memory mechanism.
+
 \begin{figure}[H]
   \centering
-  \includesvg[width=\textwidth]{memcell/circuit2}
-  \caption{Memory cell circuit}
+  \includesvg[height=2.5cm]{memcell/memcellSymbol}
+  \caption{Memory cell symbol with the input pin (left), the output pin (right), the input enable pin (top) and the output enable pin (bottom)}
   \label{fig:memcellCircuit}
 \end{figure}
+
+\subsection{Usage}
+
+This circuit is pretty straight forward, when the input enable pin is high (up to $V_{dd}$), then the first switch is open and the capacitor is being charged, if it is low, then the switch is closed and the capacitor keeps the voltage and thus the analog data.
+When the output enable is high, the buffer (TODO : Explain in annexe ?) forwards the voltage stored in the capactor to the output pin without emptying the capacitor.
+
+TODO : figure that explains the fonctionnement

### 2023-09-07 3.Design/inverter.tex
Minor details have been changed in the files

@@
 There is nothing special about this specific inverter.
-When it a voltage of $v_{dd}$ is applied to the input, the output is grounded and the same goes for the other way around.
+When it a voltage of $V_{dd}$ is applied to the input, the output is grounded and the same goes for the other way around.
 
@@
   \hfill
-  \subfloat[Inverter's symbol]{\label{fig:invSymbol}\includesvg[height=3cm]{inverter/invSymbol}}%
+  \subfloat[Inverter's symbol]{\label{fig:invSymbol}\includesvg[height=2.5cm]{inverter/invSymbol}}%
   \hspace*{1.5cm}

### 2023-09-07 3.Design/crossbar.tex
Added crossbar part, may be removed

@@
 
-This section is here to show
+The crossbar circuit has already been explained in the section \ref{sec:crossbar},
+
+    Since the analog system is only simulated in the inference phase of the \ac{NN}, the resistance doesn't need to change during this phase.
+The weights only change during the training phase, and the resistances are just a rendition of the weights (section TODO : ref weigths to res)

### 2023-09-07 3.Design/models.tex
Added explication whyneed to amplify aft voltmult, and added appendix A, need to find name

@@
 Indeed, we need to have $1\cdot 1=1$, but if we set $V_{in_1}=V_{in_2}=1V$, we would find $V_{out}=0.91V$.
-There is then a small circuit added following the output of the voltage multiplier.
-TODO : Annexe
+To fix this issue, a small circuit added following the output of the voltage multiplier, this circuit needs to do the following :
+
+\begin{equation}
+  V_{amp}=(V_{out}-V_{cm})\cdot10+V_{cm}
+\end{equation}
+
+Where $V_{amp}$ is the output of the amplificator and $V_{out}$ is the output of the voltage multiplier and the input of the amplifier circuit.
+This circuit is shown in appendix \ref{apsec:amp}.

### 2023-09-07 Appendices/appendixA.tex
Added explication whyneed to amplify aft voltmult, and added appendix A, need to find name

\fancychapter{TODO determine name}
\label{ap:a}

\section{Amplificator}\label{apsec:amp}

This is the circuit that comes after the voltage multiplier (section \ref{sec:models}), it simply consists of two inverting amplifiers with one amplifying the results by 10. Figure TODO shows this simple circuit.

A non-inverting amplifier was producing too much difference from the expected result (about a 20\% error) %TODO maybe show something

### 2023-09-07 3.Design/memcell.tex
experimenting with time diagram, not working yet
Now working

@@
 TODO : figure that explains the fonctionnement
+
+\begin{figure}[H]
+  \centering
+  \def\degr{${}^\circ$}
+  \begin{tikztimingtable}
+    Clock 128\,MHz 0\degr & H 2C N(A1) 8{2C} N(A5) 3{2C} G \\
+    Clock 128\,MHz 90\degr & [C] 2{2C} N(A2) 8{2C} N(A6) 2{2C} C \\
+    Clock 128\,MHz 180\degr & C 2{2C} N(A3) 8{2C} N(A7) 2{2C} G \\
+    Clock 128\,MHz 270\degr & 3{2C} N(A4) 8{2C} N(A8) 2C C \\
+    Coarse Pulse & 3L 16H 6L \\
+    Coarse Pulse - Delayed 1 & 4L N(B2) 16H N(B6) 5L \\
+    Coarse Pulse - Delayed 2 & 5L N(B3) 16H N(B7) 4L \\
+    Coarse Pulse - Delayed 3 & 6L 16H 3L \\
+    \\
+    Final Pulse Set & 3L 16H N(B5) 6L \\
+    Final Pulse $\overline{\mbox{Reset}}$ & 6L N(B4) 16H 3L \\
+    Final Pulse & 3L N(B1) 19H N(B8) 3L \\
+    \extracode
+    \tablerules
+    \begin{pgfonlayer}{background}
+      \foreach \n in {1,...,8}
+      \draw [help lines] (A\n) --56(B\n);
+    \end{pgfonlayer}
+  \end{tikztimingtable}
+  \caption{Memory cell symbol with the input pin (left), the output pin (right), the input enable pin (top) and the output enable pin (bottom)}
+  \label{fig:memcellCircuit}
+\end{figure}

### 2023-09-08 3.Design/memcell.tex
onchip area begin + timediagram for memcell

--- 2023-09-07 3.Design/memcell.tex
@@
 
+For this circuit the approximate onChip area :
+
+\begin{equation}
+  A_{memcell}=8\cdot A_{CMOS}+2\cdot A_{inv}+A_{opAmp}+A_{cap}
+\end{equation}
+
+With $A_{CMOS} = w\cdot l = 9\cdot 10^{-7} \cdot 6 \cdot 10^{-8} = 5.4 \cdot 10^{-14} m^2$ TODO check value w et l
+
 \subsection{Symbol}
@@
 This circuit is pretty straight forward, when the input enable pin is high (up to $V_{dd}$), then the first switch is open and the capacitor is being charged, if it is low, then the switch is closed and the capacitor keeps the voltage and thus the analog data.
-When the output enable is high, the buffer (TODO : Explain in annexe ?) forwards the voltage stored in the capactor to the output pin without emptying the capacitor.
+When the output enable is high, the buffer (appendix \ref{apsec:buffer}) forwards the voltage stored in the capactor to the output pin without emptying the capacitor.
 
-TODO : figure that explains the fonctionnement
+\begin{figure}[H]
+  \centering
+  \begin{tikztimingtable}
+    Input data & x6D{$x_1$}6D{$x_2$}6D{$x_3$}x \\
+    Input enable & xL2H2L3H7L2HLx\\
+    Output enable & x3L2H4L5H1L3Hx\\
+    Memory & xU4D{$x_1$}10D{$x_2$}3D{$x_3$}x \\
+    Output data & x3U2D{$x_1$}4U5D{$x_2$}U3D{$x_3$}x \\
+    \extracode
+    \tablerules
+    %\draw (0,0) circle (2pt); % Origin
+    \begin{pgfonlayer}{background}
+      \vertlines[help lines]{0.6,18.6}
+      %\vertlines[red]{1.6,5.6,15.6}
+      %\vertlines[blue]{3.6,9.6,15.6}
+    \end{pgfonlayer}
+  \end{tikztimingtable}
+  \caption{Time diagram that shows the logic of the memory cell}
+  \label{tim:memcell}
+\end{figure}

### 2023-09-11 3.Design/memcell.tex
changed CMOS to acro now

@@
 
-The circuit is shown in figure \ref{fig:memcellCircuit}. The value/voltage is trapped in the capacitor using CMOS switches.
+The circuit is shown in figure \ref{fig:memcellCircuit}. The value/voltage is trapped in the capacitor using \ac{CMOS} switches.
 
@@
 
-The cirucit has a two CMOS switches design to avoid voltage leakage through the swicthes.
-The system has voltage leakage when only one CMOS swicth, and thus leads to a large memory leak.
-This is due to the high voltage difference between the two sides of the CMOS swicth.
-Using two CMOS switches allows for this difference to be mitigated (figure \ref{fig:memcellLoss}).
+The cirucit has a two \ac{CMOS} switches design to avoid voltage leakage through the swicthes.
+The system has voltage leakage when only one \ac{CMOS} swicth, and thus leads to a large memory leak.
+This is due to the high voltage difference between the two sides of the \ac{CMOS} swicth.
+Using two \ac{CMOS} switches allows for this difference to be mitigated (figure \ref{fig:memcellLoss}).
 
@@
   \includesvg[width=\textwidth]{memcell/data-loss}
-  \caption{Memory conservation in a memory cell with 1 CMOS switch vs 2 CMOS swicthes}
+  \caption{Memory conservation in a memory cell with 1 \ac{CMOS} switch vs 2 \ac{CMOS} swicthes}
   \label{fig:memcellLoss}

### 2023-09-11 3.Design/crossbar.tex
CMOS is now acronym, and started writing crossbar part
Added two mem per synapses part

@@
 
-The crossbar circuit has already been explained in the section \ref{sec:crossbar},
+The crossbar circuit has already been explained in the section \ref{sec:crossbar}.
+This section describes how the crossbar circuit is actually implemented.
+The final circuit is the one in figure \ref{fig:xbarCircuit}.
+
+\begin{figure}[H]
+  \centering
+  \includesvg[width=\textwidth]{crossbar/crossbarUse}
+  \caption{Crossbar circuit used in the final system}
+  \label{fig:xbarCircuit}
+\end{figure}
+
+First of all, a design with two memristors per synapse has been used. This design is beneficial in two ways :
+
+\begin{itemize}
+  \item It doubles the range of weights
+  \item It allows to easily use negative weights
+\end{itemize}
+
+This design is the one that is used in \cite{doubleMem}.
+Let's assume that a given memristor has a resistance range of $R\in[R_{min},R_{max}]$, that means it's conductance range is $\sigma \in [\sigma_{min},\sigma_{max}]$ (with $\sigma_{min}= \frac{1}{R_{max}}$ and $\sigma_{max}= \frac{1}{R_{min}}$).
+The use of a design such as the one specifically displayed if figure TODO.
+It works with two \ac{opAmp}.
+
+\begin{equation}
+  V_{opAmp1}=-R\cdot i_+
+\end{equation}
+\begin{equation}
+  i_{R_f}=i_-+\frac{V_{opAmp1}}{R}=i_--i_+
+\end{equation}
+With $i_{opAmp}=0A$ because we assume an ideal \ac{opamp}.
+\begin{equation}
+  V_{opAmp2}=V_{out}=-R_f\cdot(i_--i_+)=R_f\cdot(i_+-i_-)=R_f\cdot\sum_{k=0}^n(\sigma_{k+}-\sigma_{k-})\cdot x_k
+\end{equation}
+With $i_+=\sum_{k=0}^n\sigma_{k+}\cdot x_k$ and $i_-=\sum_{k=0}^n\sigma_{k-}\cdot x_k$.
+
+In the previous equations, $x_k$ is the voltage for the input line $k$.
+
+\begin{table}[H]
+  \centering
+  \begin{tabular}{|c|c|c|}
+    \cline{2-3}
+    \rowcolor{gray}
+    \multicolumn{1}{c|}{\cellcolor[HTML]{FFFFFF}} & Two memristors per synapse & One memristor per synapse \\
+    \hline
+    Maximum weight & $\sigma_{max}-\sigma_{min}$ & $\sigma_{max} -\overline{\sigma}$\\
+    \hline
+    Minimum weight & $\sigma_{min}-\sigma_{max}$ & $\sigma_{min} -\overline{\sigma}$\\
+    \hline
+    Range & $2\cdot(\sigma_{max}-\sigma_{min})$&$\sigma_{max}-\sigma_{min}$\\
+    \hline
+  \end{tabular}
+  \caption{Synaptic weights precision (extracted from \cite{doubleMem})}
+  \label{tab:synapses}
+\end{table}
 

### 2023-09-11 3.Design/activation.tex
Minor details have been changed in the files
added  w and l explain for af
onchip area begin + timediagram for memcell
Moved table's caption to the bottom
Worked on xbar + new package for refs + cleaning repo

@@
   \centering
-  \caption{Circuits parameters}
   \begin{tabular}{|c|c|c|}
@@
   \end{tabular}
+  \caption{Circuits parameters}
   \label{tab:afPar}
@@
 
+Here, $w$ and $l$ are, respectively, the width and length of the two NMOS of the circuit.
+
 \subsection{Circuit}
 
-The circuit used is the same as the one in \cite{thesisRef}, the circuit is the one shown in figure \ref{fig:afCircuit}.
+The circuit used is the same as the one in \cite{thesisRef}, the circuit is the one shown in \cref{fig:afCircuit}.
 The technology used being different, all the parameters had to be determined empirically to best fit a sigmoid shape.
-The parameters can be found in table \ref{tab:afPar}.
+The parameters can be found in \cref{tab:afPar}.
 
@@
 \begin{equation}
-  A_{af}=2\cdot A_{R_1} + 2\cdot A_{nmos} + 5\cdot A_{R_2} + A_{R_3} +2\cdot A_{opAmp}
+  A_{af}=2\cdot A_{R_1} + 2\cdot A_{CMOS} + 5\cdot A_{R_2} + A_{R_3} +2\cdot A_{opAmp}
 \end{equation}
-With $A_{nmos} = w\cdot l = 9\cdot 10^{-7} \cdot 6 \cdot 10^{-8} = 5.4 \cdot 10^{-14} m^2$
+With $A_{CMOS} = w\cdot l = 9\cdot 10^{-7} \cdot 6 \cdot 10^{-8} = 5.4 \cdot 10^{-14} m^2$
 TODO : Finish calc
@@
 
-\subsection{Results}
+\subsection{Usage}
 
 This circuit outputs a voltage that depends on the input voltage passed on.
-The relation between the two is shown in figure \ref{fig:afGraph}.
+The relation between the two is shown in \cref{fig:afGraph}.
 The graph also shows the \ac{RMSE} of each graph compared to the ideal result.
 
-Note that the actual output of the circuit in figure \ref{fig:afCircuit} is inverted arround $V_{cm}$ a simple circuit such as the one shown in  (TODO show circuit in annex) does inverts the
+Note that the actual output of the circuit in \cref{fig:afCircuit} is inverted arround $V_{cm}$ a simple circuit such as the one shown in  (TODO show circuit in annex) does inverts the
 

### 2023-09-11 3.Design/crossbar.tex
Worked on xbar + new package for refs + cleaning repo

@@
 
-The crossbar circuit has already been explained in the section \ref{sec:crossbar}.
+The crossbar circuit theory has already been explained in the \cref{sec:crossbar}.
 This section describes how the crossbar circuit is actually implemented.
-The final circuit is the one in figure \ref{fig:xbarCircuit}.
+The final circuit is the one in \cref{circt:xbar}.
 
@@
   \caption{Crossbar circuit used in the final system}
-  \label{fig:xbarCircuit}
+  \label{circt:xbar}
 \end{figure}
@@
 \begin{itemize}
-  \item It doubles the range of weights
-  \item It allows to easily use negative weights
+  \item It doubles the range of weights (\cref{tab:synapses})
+  \item It allows to easily use negative weights, as this design simply allows the output values to be centered around $V_{cm}$ to be compliant with the standard set in \cref{tab:valConv}.
 \end{itemize}
@@
 Let's assume that a given memristor has a resistance range of $R\in[R_{min},R_{max}]$, that means it's conductance range is $\sigma \in [\sigma_{min},\sigma_{max}]$ (with $\sigma_{min}= \frac{1}{R_{max}}$ and $\sigma_{max}= \frac{1}{R_{min}}$).
-The use of a design such as the one specifically displayed if figure TODO.
-It works with two \ac{opAmp}.
+This design works using two \ac{opAmp} connected to $V_{cm}$ with the positive pin and the negative pin to the output of the crossbar array.
+\Cref{eq:doubleMem0,eq:doubleMem1,eq:doubleMem2} are describing how this architecture works.
 
 \begin{equation}
+  \label{eq:doubleMem0}
   V_{opAmp1}=-R\cdot i_+
@@
 \begin{equation}
+  \label{eq:doubleMem1}
   i_{R_f}=i_-+\frac{V_{opAmp1}}{R}=i_--i_+
@@
 \begin{equation}
+  \label{eq:doubleMem2}
   V_{opAmp2}=V_{out}=-R_f\cdot(i_--i_+)=R_f\cdot(i_+-i_-)=R_f\cdot\sum_{k=0}^n(\sigma_{k+}-\sigma_{k-})\cdot x_k

### 2023-09-11 3.Design/memcell.tex
Worked on xbar + new package for refs + cleaning repo

@@
 
-The circuit is shown in figure \ref{fig:memcellCircuit}. The value/voltage is trapped in the capacitor using \ac{CMOS} switches.
+The circuit is shown in \cref{fig:memcellCircuit}. The value/voltage is trapped in the capacitor using \ac{CMOS} switches.
 
@@
 This is due to the high voltage difference between the two sides of the \ac{CMOS} swicth.
-Using two \ac{CMOS} switches allows for this difference to be mitigated (figure \ref{fig:memcellLoss}).
+Using two \ac{CMOS} switches allows for this difference to be mitigated (\cref{fig:memcellLoss}).
 
@@
 This circuit is pretty straight forward, when the input enable pin is high (up to $V_{dd}$), then the first switch is open and the capacitor is being charged, if it is low, then the switch is closed and the capacitor keeps the voltage and thus the analog data.
-When the output enable is high, the buffer (appendix \ref{apsec:buffer}) forwards the voltage stored in the capactor to the output pin without emptying the capacitor.
+When the output enable is high, the buffer (\cref{apsec:buffer}) forwards the voltage stored in the capactor to the output pin without emptying the capacitor.
 

### 2023-09-11 3.Design/activation.tex
Changed circuit's ref

@@
 
-The circuit used is the same as the one in \cite{thesisRef}, the circuit is the one shown in \cref{fig:afCircuit}.
+The circuit used is the same as the one in \cite{thesisRef}, the circuit is the one shown in \cref{circt:af}.
 The technology used being different, all the parameters had to be determined empirically to best fit a sigmoid shape.
@@
   \caption{Activation functions circuit}
-  \label{fig:afCircuit}
+  \label{circt:af}
 \end{figure}
@@
 
-Note that the actual output of the circuit in \cref{fig:afCircuit} is inverted arround $V_{cm}$ a simple circuit such as the one shown in  (TODO show circuit in annex) does inverts the
+Note that the actual output of the circuit in \cref{circt:af} is inverted arround $V_{cm}$ a simple circuit such as the one shown in  (TODO show circuit in annex) does inverts the
 

### 2023-09-11 3.Design/inverter.tex
onchip area begin + timediagram for memcell
Worked on xbar + new package for refs + cleaning repo
Changed circuit's ref

@@
   \hspace*{2.5cm}
-  \subfloat[Inverter's circuit]{\label{fig:invCircuit}\includesvg[height=3cm]{inverter/invCircuit}}%
+  \subfloat[Inverter's circuit]{\label{circt:inv}\includesvg[height=3cm]{inverter/invCircuit}}%
   \hfill
@@
 
-Figure \ref{fig:inv} combines the circuit and the symbol of the inverter used in this thesis.
+\Cref{fig:inv} combines the circuit and the symbol of the inverter used in this thesis.
+
+The approximate onChip area for the inverter is :
+\begin{equation}
+  A_{inv}=2\cdot A_{CMOS}
+\end{equation}
+
+With $A_{CMOS}= w\cdot l$ TODO check w et l

### 2023-09-12 2.State/crossbar.tex
crossbar advancement, plus changes to state crossbar for unity

@@
 
-First of all, a voltage is applied on the $i^{th}$ line, because every column is virtually grounded, the voltage applied to the memristor, of a set resistance of $R$, is $V_i$.
-As such and using Ohm's law, we know the the current flowing into the memristor is bound by the following equation :
+First of all, a voltage is applied on the $i^{th}$ line, because every column is virtually grounded, the voltage applied to the memristor, of a set resistance $R_i$, is $V_i$.
+As such and using Ohm's law, we know the the current flowing into the memristor ($I_{mem}$) is bound by the following equation :
+
 \begin{equation}
-  V_i = R\cdot I_{mem} \implies I_{mem} = V_i\cdot (\frac{1}{R})=I_{mem} = V_i\cdot G
+  V_i = R\cdot I_{mem} \Rightarrow I_{mem} = V_i\cdot (\frac{1}{R})=I_{mem} = V_i\cdot\sigma_i
 \end{equation}
-With $G$ being the conductance of memristor.
-This line then joins the column where a current of $I_{i,j-1}$ is flowing, then according to kirchhoff's law the resulting current is :
+With $\sigma_i$ being the conductance of memristor, defined as $\sigma_i=\frac{1}{R_i}$.
+
+This line then joins the column where a current of $I_{i,j-1}$ is flowing, then according to kirchhoff's current law the resulting current is :
 \begin{equation}
-  I_{j,i} = I_{j,i-1}+I_{mem} = I_{j,i-1} + V_i\cdot G
+  I_{j,i} = I_{j,i-1}+I_{mem} = I_{j,i-1} + V_i\cdot\sigma_i
 \end{equation}
@@
 \begin{equation}
-  I_1=G_1\cdot V_1 + G_2\cdot V_2 + G_3\cdot V_3
+  I_1=\sigma_1\cdot V_1 + \sigma_2\cdot V_2 + \sigma_3\cdot V_3
 \end{equation}
-With $G_1$, $G_2$ and $G_3$ being the conductance of the 3 memristors in the first column.
+With $\sigma_1$, $\sigma_2$ and $\sigma_3$ being the conductance of the 3 memristors in the first column.
 
-This analog circuit removes the need to copy data from a main memory as the memristors themselves hold the value and the weights are then stored where the computation happens.
+This analog circuit removes the need to copy data from a main memory as the memristors themselves hold the value and the weights are then stored where the computation happens.%TODO elaborate

### 2023-09-12 3.Design/crossbar.tex
Cafouiallage equation, a voir plus tard
worked on area
crossbar advancement, plus changes to state crossbar for unity

@@
 
-First of all, a design with two memristors per synapse has been used. This design is beneficial in two ways :
+\subsection{Two memristors per synapse}
+
+First of all, a two memristors per synapse architecture has been used. This design is beneficial in two ways :
 
@@
 
+% For the sake of simplicity, the following equations considers the ground to be $V_{cm}$.
+
 \begin{equation}
   \label{eq:doubleMem0}
-  V_{opAmp1}=-R\cdot i_+
+  V_{opAmp0}=-R\cdot i_+ \Leftrightarrow
 \end{equation}
@@
   \label{eq:doubleMem1}
-  i_{R_f}=i_-+\frac{V_{opAmp1}}{R}=i_--i_+
+  i_{R_f}=i_-+\frac{V_{opAmp0}}{R}=i_--i_+
 \end{equation}
@@
 
+\subsection{Serialization/Parallelization}
+
+The idea using the system in a serial mode came from \cite{thesisRef}.
+The principle of this system is to save onChip area by reducing the number of \ac{opAmp} required.
+It also, and mainly reduces the point wise part of the \ac{LSTM} circuit.
+
+This serial system has great advantages but comes with a few down sides. Let's assume the number of hidden state ($n_h$) is the major changing element in the final system(TODO ref final onchip area).
+
+\begin{itemize}
+  \item When the system is serialized, the inference time increases with a factor of $O(n_h)$.
+  \item When the system is in parallel, the circuit onChip area increases with a factor of $O(n_h)$.
+\end{itemize}
+
+However, in both cases, the overall onChip area increases with a factor of $O(n_h^2)$.
+That means that using the parallel system won't increase the onChip area by a lot as $n_h$ rises, because the $O(n_h)$ factor of the parallel version is absorbed as $O(n_h^2+n_h)=O(n_h^2)$.
+
+That leads to the conclusion that the best thing to do is use a both serialized and parallel system, using the limiting inference time for our system to set the maximum serial size.
+
+%TODO understand better \subsection{Sneak path problem}
+\subsection{Usage}
 Since the analog system is only simulated in the inference phase of the \ac{NN}, the resistance doesn't need to change during this phase.

### 2023-09-12 3.Design/crossbar.tex
Changes to crossbar final figure + text change

@@
 \subsection{Usage}
-Since the analog system is only simulated in the inference phase of the \ac{NN}, the resistance doesn't need to change during this phase.
-The weights only change during the training phase, and the resistances are just a rendition of the weights (section TODO : ref weigths to res)
+Since the analog system is only simulated in the inference phase of the \ac{NN}, the resistances don't need to change, the circuit used in the simulations are thus using resistors instead of memristors for the sake of simplicity (notably in generating the netlist of the system, TODO ref how hard it is to import parameters using spice netlists).
+The weights only change during the training phase, and the resistances are just a rendition of the weights (TODO : ref weigths to res).

### 2023-09-12 3.Design/crossbar.tex
Done for now, missing figures and a few TODOs

--- 2023-09-12 3.Design/crossbar.tex
@@
-\section{Crossbar circuit}
+\section{Voltage-driven crossbar circuit}
 \label{sec:xbarCircuit}
@@
 That leads to the conclusion that the best thing to do is use a both serialized and parallel system, using the limiting inference time for our system to set the maximum serial size.
+Depending on the system we're using this circuit for, we might have different time requirements.
 
 %TODO understand better \subsection{Sneak path problem}
+\subsection{Symbol}
+This is the symbol defined for the voltage based memristor crossbar array used in this thesis for a better undertsanding of the future circuits.
+It depends on several parameters, the number of inputs ($n_i$), the number of outputs ($n_o$) and the serial size ($n_s$).
+
+%TODO : include figure
+
+The total onChip area for the crossbar circuit depends on the previously defined parameters.
+
+\begin{equation}
+  A_{xbar}(n_i,n_o,n_s)=2\cdot n_i\cdot n_o \cdot n_s\cdot A_{memristor}+2\cdot A_{CMOS}\cdot n_o\cdot n_s +n_o\cdot(2\cdot[A_{opAmp}+A_R]+A_{R_f})
+\end{equation}
+
+With $A_{CMOS}$ being the same than for the memory cell (\cref{sec:memcell}), $A_{CMOS}=1.2\cdot 10^{-14}m^2$, and the feature size of typical memristor that can be fabricated at INESC is $2\mu m$, which would make the approximate area for a memristor $A_{memristor}=4\mu m^2=4\cdot 10^{-12} m^2$.%TODO check info and inesc citation
+
 \subsection{Usage}
-Since the analog system is only simulated in the inference phase of the \ac{NN}, the resistance doesn't need to change during this phase.
-The weights only change during the training phase, and the resistances are just a rendition of the weights (section TODO : ref weigths to res)
+
+TODO Show off results
+
+The analog system is only simulated in the inference phase of the \ac{NN}.
+And since the weights only change during the training phase, and the resistances are just a rendition of the weights (TODO : ref weigths to res), the resistances don't need to change, the circuit used in the simulations is thus using resistors instead of memristors for the sake of simplicity (notably in generating the netlist of the system, TODO ref how hard it is to import parameters using spice netlists).

### 2023-09-12 3.Design/wei2res.tex
Added new file for weight to resistance

\section{Conversion from weights to resistance}
\label{sec:wei2res}

This chapter will contain all the information on the weight conversion to a resistance value.

### 2023-09-13 3.Design/models.tex
Worked on xbar + new package for refs + cleaning repo

@@
 
-This component is the very famous \ac{opAmp}. I didn't design a specific \ac{opAmp} because of the current range of the system I'm trying to build.
+This component is the very famous \ac{opAmp}.
+This specific component wasn't designed for the thesis, it required a current range that didn't allow to use ones already made by members of the research group.
+The choice was then to use a verilog-A model, and then design one if time permits it.
 
@@
 
-This component while a bit less popular than the latter, is just as useful for our specific use.
+This component while far less popular than the latter, is just as useful for our specific use.
 It allows us to multiply, as its name implies, two voltages.
-It is used to compute the pointwise multiplications of the \ac{LSTM} (figure \ref{fig:lstmCell}).
+It is used to compute the pointwise multiplications of the \ac{LSTM} (\cref{fig:lstmCell}).
 
@@
 Indeed, we need to have $1\cdot 1=1$, but if we set $V_{in_1}=V_{in_2}=1V$, we would find $V_{out}=0.91V$.
-To fix this issue, a small circuit added following the output of the voltage multiplier, this circuit needs to do the following :
+To fix this issue, a small circuit has to be added following the output of the voltage multiplier, this circuit needs to do the following :
 
@@
 Where $V_{amp}$ is the output of the amplificator and $V_{out}$ is the output of the voltage multiplier and the input of the amplifier circuit.
-This circuit is shown in appendix \ref{apsec:amp}.
+This circuit is shown in \cref{apsec:amp}.

### 2023-09-13 3.Design/crossbar.tex
Mainly final xbar circuit, super joliegit add 3.Design/crossbar.tex figures/crossbar/crossbarUse.svg git pull
Small changes to crossbar

@@
 The final circuit is the one in \cref{circt:xbar}.
+The circuit depends on three parameters :
+\begin{itemize}
+  \item $n_i$ : The number of input for our crossbar array (not including bias for a more general circuit).
+  \item $n_o$ : The number of parallel output for our crossbar array.
+  \item $n_s$ : The serial size of our crossbar system.
+\end{itemize}
 
@@
   \includesvg[width=\textwidth]{crossbar/crossbarUse}
-  \caption{Crossbar circuit used in the final system}
+  \caption{Simplified circuit of a crossbar array ($n_i$, $n_o$, $n_s$)}
   \label{circt:xbar}
 \end{figure}
+
+The parallel and serial are explained later (\cref{subsec:serpar}).
+
+The enable flags ($e_j,\forall j\in[\![ 0,n_s]\!]$ and similarly $\overline{e_j}$) are there to show when the \ac{CMOS} switches are open, the states of those flags is shown in \cref{tim:serpar}.
 
@@
 
-In the previous equations, $x_k$ is the voltage for the input line $k$.
+In \cref{eq:doubleMem0,eq:doubleMem1,eq:doubleMem2}, $x_k$ is the voltage for the input line $k$.
 
@@
 \subsection{Serialization/Parallelization}
+\label{subsec:serpar}
 
@@
 
+\begin{figure}[H]
+  \centering
+  \begin{tikztimingtable}%TODO change or find other way
+    Input data & x6D{$x_1$}6D{$x_2$}6D{$x_3$}x \\
+    Input enable & xL2H2L3H7L2HLx\\
+    Output enable & x3L2H4L5H1L3Hx\\
+    Memory & xU4D{$x_1$}10D{$x_2$}3D{$x_3$}x \\
+    Output data & x3U2D{$x_1$}4U5D{$x_2$}U3D{$x_3$}x \\
+    \extracode
+    \tablerules
+    %\draw (0,0) circle (2pt); % Origin
+    \begin{pgfonlayer}{background}
+      \vertlines[help lines]{0.6,18.6}
+      %\vertlines[red]{1.6,5.6,15.6}
+      %\vertlines[blue]{3.6,9.6,15.6}
+    \end{pgfonlayer}
+  \end{tikztimingtable}
+  \caption{Time diagram that shows the states of the enable flags for the crossbar circuit}
+  \label{tim:serpar}
+\end{figure}
+
 %TODO understand better \subsection{Sneak path problem}
@@
 
-With $A_{CMOS}$ being the same than for the memory cell (\cref{sec:memcell}), $A_{CMOS}=1.2\cdot 10^{-14}m^2$, and the feature size of typical memristor that can be fabricated at INESC is $2\mu m$, which would make the approximate area for a memristor $A_{memristor}=4\mu m^2=4\cdot 10^{-12} m^2$.%TODO check info and inesc citation
+With $A_{CMOS}$ being the same than for the memory cell (\cref{sec:memcell}), $A_{CMOS}=1.2\cdot 10^{-14}m^2$, and the feature size of typical memristor that can be fabricated at INESC is $3\mu m$, which would make the approximate area for a memristor $A_{memristor}=9\mu m^2=9\cdot 10^{-12} m^2$.%TODO check info and inesc citation
 

### 2023-09-14 3.Design/crossbar.tex
new figures and few lines

@@
   \includesvg[width=\textwidth]{crossbar/crossbarUse}
-  \caption{Simplified circuit of a crossbar array ($n_i$, $n_o$, $n_s$)}
+  \caption{Circuit of the crossbar array used in the final system ($n_i$, $n_o$, $n_s$)}
   \label{circt:xbar}
@@
 \Cref{eq:doubleMem0,eq:doubleMem1,eq:doubleMem2} are describing how this architecture works.
+A simplified version of the double memristors per synapse circuit is also available in \cref{circt:doubleMem}.
 
-% For the sake of simplicity, the following equations considers the ground to be $V_{cm}$.
+\begin{figure}[H]
+  \centering
+  \includesvg[height=8cm]{crossbar/doubleMem}
+  \caption{Simplified circuit of a double memristor per synapse architecture}
+  \label{circt:doubleMem}
+\end{figure}
+
+With $x_k$ being the voltage for the input line $k$. The highest \ac{opAmp} is identified as opamp0 and the lowest opamp1.
+
+For the sake of simplicity, the following equations considers the ground to be $V_{cm}$.
 
@@
   \label{eq:doubleMem2}
-  V_{opAmp2}=V_{out}=-R_f\cdot(i_--i_+)=R_f\cdot(i_+-i_-)=R_f\cdot\sum_{k=0}^n(\sigma_{k+}-\sigma_{k-})\cdot x_k
+  V_{opAmp2}=y_0=-R_f\cdot(i_--i_+)=R_f\cdot(i_+-i_-)=R_f\cdot\sum_{k=0}^n(\sigma_{k+}-\sigma_{k-})\cdot x_k
 \end{equation}
 With $i_+=\sum_{k=0}^n\sigma_{k+}\cdot x_k$ and $i_-=\sum_{k=0}^n\sigma_{k-}\cdot x_k$.
-
-In \cref{eq:doubleMem0,eq:doubleMem1,eq:doubleMem2}, $x_k$ is the voltage for the input line $k$.
 
@@
 
+The \ac{CMOS} switches are here to open the necessary input gates when the output is required, the gates are controlled as in \cref{tim:serpar}.
 \begin{figure}[H]
@@
   \begin{tikztimingtable}%TODO change or find other way
-    Input data & x6D{$x_1$}6D{$x_2$}6D{$x_3$}x \\
-    Input enable & xL2H2L3H7L2HLx\\
-    Output enable & x3L2H4L5H1L3Hx\\
-    Memory & xU4D{$x_1$}10D{$x_2$}3D{$x_3$}x \\
-    Output data & x3U2D{$x_1$}4U5D{$x_2$}U3D{$x_3$}x \\
+    $e_0$ & xH3Lx\\
+    $e_1$ & xLH2Lx\\
+    $e_2$ & x2LHLx\\
+    $e_3$ & x3LHx\\
     \extracode
@@
     \begin{pgfonlayer}{background}
-      \vertlines[help lines]{0.6,18.6}
+      \vertlines[help lines]{0.6,4.6}
       %\vertlines[red]{1.6,5.6,15.6}
@@
   \end{tikztimingtable}
-  \caption{Time diagram that shows the states of the enable flags for the crossbar circuit}
+  \caption{Example of enable flags timing with $n_s=4$ in a single time step}
   \label{tim:serpar}

### 2023-09-14 3.Design/crossbar.tex
Symbol is finished

@@
 \subsection{Symbol}
-This is the symbol defined for the voltage based memristor crossbar array used in this thesis for a better undertsanding of the future circuits.
+The symbol (\cref{sym:xbar}) defined for the voltage based memristor crossbar array used in this thesis is more compact and helps the readability of the final circuit.
 It depends on several parameters, the number of inputs ($n_i$), the number of outputs ($n_o$) and the serial size ($n_s$).
 
-%TODO : include figure
+\begin{figure}[H]
+  \centering
+  \includesvg[height=2.5cm]{crossbar/xbarSymbol}
+  \caption{Symbol used for the crossbar array}
+  \label{symb:xbar}
+\end{figure}
 

### 2023-09-14 3.Design/wei2res.tex
Almost finished wei2res, and changed figures

@@
 This chapter will contain all the information on the weight conversion to a resistance value.
+
+The weight to resistances equations has been found using simple math.
+
+\begin{equation}
+  \label{eq:wei2res0}
+  w=R_f\cdot(\sigma_+-\sigma_-)=\frac{R_f}{R_+}-\frac{R_f}{R_-}
+\end{equation}
+First, we know from \cref{eq:doubleMem2} that the weight is represented by \cref{eq:wei2res0}. This means that the weights are limited in values to :
+
+\begin{itemize}
+  \item $w_{max}=R_f\cdot(\sigma_{max}-\sigma_{min})$
+  \item $w_{min}=-w_{max}=R_f\cdot(\sigma_{min}-\sigma_{max})$
+\end{itemize}
+
+For the rest of this chapter we consider that $R_f$ is set to the middle point of $R_{max}$ and $R_{min}$ meaning that $R_f=\frac{R_{min}+R_{max}}{2}$.
+
+Since we only have one equation (\cref{eq:wei2res0}) for 2 unknowns ($R_+$ and $R_-$), we need to set a second equation.
+This is done by centering the resistances around $R_f$, this means that \cref{eq:wei2res1} is the second equation, that makes our problem now solvable.
+%TODO explain how in future work the weight doesn't have to be centered around R_f, to gain time only one of the two resistance could change
+\begin{equation}
+  \label{eq:wei2res1}
+  R_f=\frac{R_-+R_+}{2}
+\end{equation}
+\begin{equation}
+  \label{eq:wei2res01}
+  \begin{cases}
+    w=\frac{R_f}{R_+}-\frac{R_f}{R_-}\\
+    R_f=\frac{R_-+R_+}{2}
+  \end{cases}
+\end{equation}
+
+By solving \cref{eq:wei2res01}, we find the \cref{eq:wei2res2} that gives the values for $R_-$ and $R_+$. All the steps for solving the equations can be found in .%TODO solve equations in appendix
+
+The real value to voltage conversion (\cref{tab:valConv}) doesn't affect this part of the system as the crossbar array already works in voltages.
+Placing $w$ from \cref{eq:wei2res0} in \cref{eq:doubleMem2} would not change the result.
+
+\begin{equation}
+  \label{eq:wei2res2}
+  \begin{cases}
+    R_+= (w+1-\sqrt{w^2+1})\cdot\frac{R_f}{w}\\
+    R_-=2\cdot R_f -R_+
+  \end{cases}
+\end{equation}
+
+For the thesis, this step is done in python (TODO show function). Since the resistance resolution is of about%TODO check memristor precision
+, the python code then reproduces this by limiting the amount of significant digit of the resistance.

### 2023-09-14 diapo/15-09-23_meeting/main.tex
Few changes
Finished pres demain

--- 2023-06-05 diapo/groupTemplate/main.tex
@@
     \node[left=0cm] at (current page.34){
-      \includegraphics[height=1cm]{../figures/ist.eps}
+      \includegraphics[height=1cm]{logos/ist.eps}
       \hspace{1pt}
-      \includegraphics[height=1cm]{figures/inesc-mn.png}
-      \includegraphics[height=1cm]{../figures/inesc-id.eps}
+      \includegraphics[height=1cm]{logos/inesc-mn.png}
+      \includegraphics[height=1cm]{logos/inesc-id.eps}
     };
@@
     \frametitle{Introduction}
-  Quick introduction to your topic.
 
@@
         \color{text}
-    \item Briefly summarized your project. Some of the attendants may need some refresh.
-    \item Explain in which project stage you are, thus we all know how close you are to achieve your goals.
+      \item I am working on an analog implementation of a LSTM Neural Networks.
+      \item The final objective of my thesis is to have a successful simulation of such a circuit.
     \end{itemize}
 
-  ex: I have to design a PCB to control a magnetic sensor. I have already done the PCB and now I am testing it.
+  \end{frame}
+
+  \begin{frame}
+    \frametitle{Introduction}
+    \centering\includegraphics[width=\textwidth]{lstm/lstm-np}
   \end{frame}
@@
 
-  Last meeting, I promised to do…
-
-  \centering
-  \rowcolors{2}{evenRow}{oddRow}
-  \begin{tabular}{ c m{3cm} c c c }
-    \rowcolor{firstRow}
-    \color{white}\textbf{\#} & \centering\color{white}Task & \color{white}Done? & \color{white}Why not? & \color{white}Help? \\
-    1 & test2 & test23 & test3 & test4\\
-    2 & test2 & test23 & test3 & test4\\
-    3 & test2 & test23 & test3 & test4\\
-    4 & test2 & test23 & test3 & test4\\
-    5 & test2 & test23 & test3 & test4\\
-  \end{tabular}
-
+    There was no meeting last week, and a lot happened but here is what I'm doing :
+    \begin{itemize}
+      \item Mainly focused on writing the thesis
+      \item I also need to run a new simulation with a new problem
+    \end{itemize}
   \end{frame}
@@
             \color{text}
-        \item
+          \item None yet
         \end{itemize}
@@
             \color{text}
-        \item test
+          \item
         \end{itemize}
@@
 
-\begin{frame}{Problems/doubts}
+  \begin{frame}
+    \frametitle{Problems/doubts}
     \begin{itemize}
-    \item Things that must me sorted out in order to continue with your work.\\ Explain the problem.
-    \item Things that you do not understand.
+      \item What is the approximate onChip area of an opAmp
+      \item What is the precision of a memristor when setting the resistance
+      \item I feel like my state of the art is missing content...
+      \item When would my thesis defense happen ?
+      \item I'm afriad i might not have time to finish on time wiht enough content.
     \end{itemize}
@@
     \rowcolors{2}{evenRow}{oddRow}
-  \begin{tabular}{ c m{4cm} }
+    \begin{tabular}{ c m{6cm} }
       \rowcolor{firstRow}
       \color{white}\textbf{\#} & \centering\color{white}Task \cr
-    1 & test23 \\
-    2 & test23 \\
-    3 & test23 \\
-    4 & test23 \\
-    5 & test23 \\
+      0 & Work on thesis \\
+      1 & Figure out the new problem \\
+      2 & Digital inference time \\
     \end{tabular}
@@
     \begin{tikzpicture}
-    \leftRect{Help/Collaboration :}{- If anybody has any idea of an interesting LSTM/NN problem to work on.\\-Also if any phenomen happening are familiar to you, I'll take the help}
-    \rightRect{Recommendations :}{I worked on a latex/beamer version of the template}
+      \leftRect{Help/Collaboration :}{- I'm looking for a software to create nice looking electronic circuit.}
+
+      \rightRect{Recommendations :}{None}
     \end{tikzpicture}

### 2023-09-18 2.State/lstm.tex
Couldn't really work, very bad day...

@@
 
-Figure \ref{fig:lstmCell} shows the complexity of the \ac{LSTM} architecture.
+\Cref{fig:lstmCell} shows the complexity of the \ac{LSTM} architecture.
 In an \ac{LSTM}, each gate is a different \ac{NN} and then activated with either a tanh or a sigmoid activation function.
@@
 Let's start with the more classic gate equations. They are the ones that behave like the more classic \ac{NN}.
-The input (equation \ref{eq:inputG}), forget (equation \ref{eq:forgetG}) and ouput (equation \ref{eq:ouputG}) gates are described below.
+The input (\cref{eq:inputG}), forget (\cref{eq:forgetG}) and ouput (\cref{eq:ouputG}) gates are described below.
 
@@
 
-For the next equation describes the candidate cell state (equation \ref{eq:candCell}), that will next become the cell state (equation \ref{eq:cellS}).
+For the next equation describes the candidate cell state (\cref{eq:candCell}), that will next become the cell state (\cref{eq:cellS}).
 
@@
 
-The final step of the \ac{LSTM} is to compute the hidden state (equation \ref{eq:hiddenS}).
+The final step of the \ac{LSTM} is to compute the hidden state (\cref{eq:hiddenS}).
 \begin{equation}\label{eq:hiddenS}

### 2023-09-18 3.Design/netlist.tex
Couldn't really work, very bad day...

\section{Netlist generation}
\label{sec:netlist}

In order to be able to run the simulation with different kinds of \ac{NN} architecture. The point of the part is thus to explain how the netlist generator tool works.

This is done by generating a SPICE netlist using a python script. This script takes in a few parameters :

\begin{itemize}
\item The first parameter is the number of input we use our system. This is the size of the input vector ($x_t$) from \cref{fig:lstmCell}.
\item The next parameter is the number of time steps. Everything about the \ac{LSTM} time steps is all explained in \cref{sec:lstm}.
\item The number of hidden state of the \ac{LSTM} network.
\item The serial size of the crossbar arrays, as described in \cref{sec:xbarCircuit}, used in the \ac{LSTM} network.
\item The files containing the weights for each part of the circuit. The weights in the files have to be organized using the following model. %TODO
\item Finally the name of the file in which the output of the script (the netlist) will be written to.
\end{itemize}

All the code can be found on my github page \cite{lstmGen}.

### 2023-09-19 3.Design/crossbar.tex
Added a little detail
Almost finished wei2res, and changed figures
wrote in wrong file
Few changes

@@
 \subsection{Two memristors per synapse}
+\label{subsec:doubleMem}
 
@@
 The symbol (\cref{sym:xbar}) defined for the voltage based memristor crossbar array used in this thesis is more compact and helps the readability of the final circuit.
+It contains the size of each parameters on its symbol.
 It depends on several parameters, the number of inputs ($n_i$), the number of outputs ($n_o$) and the serial size ($n_s$).
@@
   \caption{Symbol used for the crossbar array}
-  \label{symb:xbar}
+  \label{sym:xbar}
 \end{figure}
@@
 
-TODO Show off results
+%TODO Show off results ??
+
+The crossbar system, and in a more general manner, the entire system will require an external controller to control the enable flags that are required in the system.
+Adding such a system will increase the power consumption of the overall system.
+However, due to the slow (precision of about $1\mu s$) nature of those signals, the controller will be very low power, and won't change the overall power consumption of the final system.%TODO maybe move to final circuit
 
 The analog system is only simulated in the inference phase of the \ac{NN}.
-And since the weights only change during the training phase, and the resistances are just a rendition of the weights (TODO : ref weigths to res), the resistances don't need to change, the circuit used in the simulations is thus using resistors instead of memristors for the sake of simplicity (notably in generating the netlist of the system, TODO ref how hard it is to import parameters using spice netlists).
+And since the weights only change during the training phase, and the resistances are just a rendition of the weights (\cref{sec:wei2res}), the resistances don't need to change, the circuit used in the simulations is thus using resistors instead of memristors for the sake of simplicity, as importing resistors are way easier to import in Cadence's virutoso.

### 2023-09-19 3.Design/genwei.tex
New files for weight generation

\section{Weights generation}
\label{sec:weigen}

The choice for the weight generation was to use Keras using the tensorflow implementation. However, both tensorflow and pyTorch were tried and were giving similar results, the final choice of using tensorflow was made because it's the most popular among the research group.

Generating the weights simply requires to train the \ac{NN} with the required achitecture. Once the trainning is done, we can run an inference of the \ac{NN} to check whether the training was good. Then the weights are reorganized based the same model required by the netlist generator (\cref{sec:netlist}) and saved as a file. The files can now be taken to generate the netlist.

The code used for all the weights generations is available at \cite{lstmWei}.

### 2023-09-19 3.Design/crossbar.tex
small change

@@
 Adding such a system will increase the power consumption of the overall system.
-However, due to the slow (precision of about $1\mu s$) nature of those signals, the controller will be very low power, and won't change the overall power consumption of the final system.%TODO maybe move to final circuit
+However, due to the slow (precision of about $1\mu s$) nature of those signals, the controller will be very low power, and won't change the overall power consumption of the final system.
+This controller is not needed for the simulation as everything is managed within Cadence's virtuoso.
+In the real circuit, the enable signals need to be controlled somehow.%TODO maybe move to final circuit
 

### 2023-09-19 5.Conclusions/further.tex
Bunch of changes but started the going further

\section{Future work}\label{sec:further}

\subsection{Changing the verilog-A files for real circuit implementation}

Making the system actually one hundred percent implemented for inference is the obvious next step in making a fully analog \ac{LSTM} computer.

This means designing both an \ac{opAmp} for the current flowing through the circuit and a voltage multiplier working in the right voltage range.

\subsection{inSitu trainning}

Of course the first thing that comes to mind to improve the system and making it actually usable for a real world application would be doing inSitu training. Training within the analog system would mean implementing the backpropagation algorithm. Here, two options are available :

\begin{itemize}
\item The easiest option would be to use the external controller (\cref{sec:fullSys}) to make the backprogation computation needed. This is not the best option, as the backpropagation would be happening in digital. This digital computation would have to happen as fast as the analog computation and thus a much faster controller would be required, meaning a bigger power consumption.
\item The ideal way of implementing the backpropagation would be using an analog system similar to the one that is used for the inference. This would still mean having a controller to change the internal resistance of the memristors, this version would require a slower clock and making the controller lower powered. Using an analog implementation might also harm the system, as the backpropagation algorithm will probably be simpler and reduce the efficiency of the system.
\end{itemize}

Figure out how to change the values of the weights for the inSitu training. Different from \cref{sec:wei2res}.

\subsubsection{Only change the value of one of the two memristor}%TODO

\subsection{Permanent training}

Once the training is fully implemented in an analog computer, the next step would be to find a way to permanently training the system. Meaning that any new information will be used to update the weights. This means that the system would need to know if the prediction it has made is correct or not.

\subsection{\ac{LSTM} continuous mode}

This would allow to have a real time results and not having to retake the same inputs several times. Basically it would be like %TODO Make graphs

This might be something that already exist that I didn't find anything about.

### 2023-09-19 3.Design/models.tex
Added symbols in the models

@@
 This specific component wasn't designed for the thesis, it required a current range that didn't allow to use ones already made by members of the research group.
-The choice was then to use a verilog-A model, and then design one if time permits it.
+The choice was then to use a verilog-A model, and then design one if time allows it.
 
-This is the code (figure TODO : in annex) of a perfect \ac{opAmp}. TODO : more description.
-TODO : add symbol.
+\begin{equation}
+  \label{eq:opAmp}
+  V_{out}=\mu \cdot (V_+-V_-)
+\end{equation}
+
+\Cref{eq:opAmp} is the equation used in the verilog-A model. It makes it so it behaves like an ideal \ac{opAmp}.
 
@@
 This circuit is shown in \cref{apsec:amp}.
+
+\subsection{Symbols}
+
+The symbols for those two components are the ones in \cref{sym:models}.
+
+\begin{figure}[H]
+  \centering
+  \hspace*{2cm}
+  \subfloat[\ac{opAmp}'s symbol\label{sym:opAmp}]{\includesvg[height=2.5cm]{models/opAmpSymbol}}%
+  \hfill
+  \subfloat[Voltage multiplier's symbol\label{sym:voltMult}]{\includesvg[height=2.5cm]{models/voltMultSymbol}}%
+  \hspace*{2cm}
+  \caption{Symbols used for the verilog-A components}
+  \label{sym:models}
+\end{figure}
+
+The \ac{opAmp} uses its classic symbol (\cref{sym:opAmp}) while the voltage multiplier uses a custom symbol (\cref{sym:voltMult}).

### 2023-09-19 3.Design/inverter.tex
worked on area
upgraded inv part

@@
   \hfill
-  \subfloat[Inverter's symbol]{\label{fig:invSymbol}\includesvg[height=2.5cm]{inverter/invSymbol}}%
+  \subfloat[Inverter's symbol]{\label{sym:inv}\includesvg[height=2.5cm]{inverter/invSymbol}}%
   \hspace*{1.5cm}
@@
 
-\Cref{fig:inv} combines the circuit and the symbol of the inverter used in this thesis.
+Like mentionned, the circuit for the inverter is very straight forward (\cref{circt:inv}). The symbol is nothing fancy as well as it is the classic symbol (\cref{sym:inv}) for such a circuit.
 
@@
 
-With $A_{CMOS}= w\cdot l$ TODO check w et l
+With $A_{CMOS}= w\cdot l=2\cdot 10^{-7}\cdot6\cdot10^{-8} =1.2\cdot10^{-14}m^2$

### 2023-09-19 2.State/crossbar.tex
Removed the TODOs

@@
 
-This analog circuit removes the need to copy data from a main memory as the memristors themselves hold the value and the weights are then stored where the computation happens.%TODO elaborate
+Using such an analog circuit is a huge gain in time, for two main reasons :
+
+\begin{itemize}
+  \item This circuit being an analog it allows for analog computations which are much faster than their digital counter parts.
+As explained in this section, the circuit only uses physical properties of the different components and can compute large \acp{VMM} in a very short amount of time.
+  \item The other advantage of such a system is the use of the memristors, that make the need to copy data from another memory useless as the memristors have a long term memory themselves.
+This is one of the ways to break down the Von Neumann bottleneck.
+\end{itemize}

### 2023-09-19 3.Design/netlist.tex
Few changes
Working on netlist, to finish

@@
   \item The serial size of the crossbar arrays, as described in \cref{sec:xbarCircuit}, used in the \ac{LSTM} network.
-\item The files containing the weights for each part of the circuit. The weights in the files have to be organized using the following model. %TODO
+  \item The files containing the weights for each part of the circuit. The weights in the files have to be organized using a specific model (\cref{subsec:weiStore}).
   \item Finally the name of the file in which the output of the script (the netlist) will be written to.
@@
 
+\subsection{Weights storage}\label{subsec:weiStore}
+
+The weights are stored in a single file that is the save a numpy array.
+The first dimension contains the weights for the different layers.
+Depending on the layer used, the weights are stored a bit differently :
+
 All the code can be found on my github page \cite{lstmGen}.

### 2023-09-20 3.Design/netlist.tex
Ädded description

@@
 
+\begin{itemize}
+  \item Dense layer : This is basically a \ac{VMM} so the weights are just sorted linearly in a list.
+The list is of size $n^2$, where $n$ is the size of the input vector.
+\Cref{mtrx:wei} shows the position ($i$) of the weight ($w_i$) in the matrix.
+  \item \ac{LSTM} layer : An \ac{LSTM} layer contains four \acp{VMM} which can be assimilated and thus stored like a dense layer. The weights for each \ac{VMM} is then stored in a sub list.
+\end{itemize}
+
+\begin{equation}\label{mtrx:wei}
+  \begin{bmatrix}
+    w_{0} & w_{1} & \dots \\
+    \dots & w_i & \dots \\
+    \dots & w_{n^2-2} & w_{n^2-1}\\
+  \end{bmatrix}
+\end{equation}
+
 All the code can be found on my github page \cite{lstmGen}.

### 2023-09-20 3.Design/crossbar.tex
Finishing at the library

--- 2023-09-19 3.Design/crossbar.tex
@@
 The circuit depends on three parameters :
+
 \begin{itemize}
@@
 
-The crossbar system, and in a more general manner, the entire system will require an external controller to control the enable flags that are required in the system.
-Adding such a system will increase the power consumption of the overall system.
-However, due to the slow (precision of about $1\mu s$) nature of those signals, the controller will be very low power, and won't change the overall power consumption of the final system.%TODO maybe move to final circuit
-
 The analog system is only simulated in the inference phase of the \ac{NN}.

### 2023-09-20 3.Design/fullsys.tex
Finishing at the library

\section{Full analog system}\label{sec:fullsys}

\subsection{Circuit controller}%TODO maybe move in annex further

The final circuit will need an external controller to manage diffenrent things in the circuit. This wasn't created in this thesis but will have to be in order to have a functionning circuit once fabricated %TODO ref annex

The controller will increase the power consumption of the overall system. However, due to the slow (precision of about $1\mu s$) nature of the system requirements, the controller will be very low power, and won't affect the overall power consumption of the final system.

The controller will have several roles in the system :
\begin{itemize}
\item The first role will be to manage the different flags. This means that the controller will have to set those to high or low in order to control the system.
\end{itemize}

### 2023-09-20 3.Design/crossbar.tex
conjugaison

@@
 The analog system is only simulated in the inference phase of the \ac{NN}.
-And since the weights only change during the training phase, and the resistances are just a rendition of the weights (\cref{sec:wei2res}), the resistances don't need to change, the circuit used in the simulations is thus using resistors instead of memristors for the sake of simplicity, as importing resistors are way easier to import in Cadence's virutoso.
+And since the weights only change during the training phase, and the resistances are just a rendition of the weights (\cref{sec:wei2res}), the resistances don't need to change, the circuit used in the simulations is thus using resistors instead of memristors for the sake of simplicity, as importing resistors is way easier to import in Cadence's virutoso.

### 2023-09-20 3.Design/fullsys.tex
Finished external controllers

@@
 \section{Full analog system}\label{sec:fullsys}
+
+\subsection{Inputs}%TODO oen section about datasets ?
 
@@
 \begin{itemize}
-\item The first role will be to manage the different flags. This means that the controller will have to set those to high or low in order to control the system.
+  \item The first role will be to manage the different flags.
+This means that the controller will have to set those to high or low in order to control the system.
+This includes flags like the enable flags of the voltage driven crossbar array (\cref{subsec:serpar}) and those for the memory cell control.
+  \item Anoter use for this controller is to reset the memory cell to $V_{cm}$ ($0$ as real value equivalent, \cref{tab:valConv}).
+This will be done by enabling the input and then applying $V_{cm}$ to the input, this will set the capacitor to $V_{cm}$.
+This has to be done everytime the circuit is ran.
+  \item The controller will also be used to control the memristors internal resistance and control the weights.
+Depending on the type of memristor, the method and time to change the internal resistance differs.
 \end{itemize}

### 2023-09-20 3.Design/crossbar.tex
Potentialnew section + part of the threshold problem dealt with

@@
 
-The analog system is only simulated in the inference phase of the \ac{NN}.
-And since the weights only change during the training phase, and the resistances are just a rendition of the weights (\cref{sec:wei2res}), the resistances don't need to change, the circuit used in the simulations is thus using resistors instead of memristors for the sake of simplicity, as importing resistors is way easier to import in Cadence's virutoso.
+The input voltages must be monitored very carefuly in order not to mess up the entire system.
+Indeed, memristors have a voltage threshold ($V_{threshold}$) above which the internal resistance changes.
+%TODO Find info in the model paper
+This means that the input voltages must follow the inequality in \cref{eq:memThres}.
+
+\begin{equation}\label{eq:memThres}
+  |V_{mem}|= |x_i-V_{cm}|\le V_{threshold}
+\end{equation}
+
+Where $V_{mem}$ is the terminal voltage of a memristor.
+
+The analog system being only simulated for the inference of the \ac{NN}.
+Furthermore, since the weights only change during the training phase, and the resistances are just a rendition of the weights (\cref{sec:wei2res}), the resistances don't need to change, the circuit used in the simulations is thus using resistors instead of memristors for the sake of simplicity, as importing resistors is way easier to import in Cadence's virutoso.

### 2023-09-20 3.Design/dataset.tex
Potentialnew section + part of the threshold problem dealt with

\section{The datasets}\label{sec:dataset}

\subsection{The format}

The values for the data

### 2023-09-20 2.State/memristors.tex
Changed memristor state

@@
 It was first theorized in 1971 by L. Chua from UC Berkeley, in \cite{TheoMemristor}. The name comes from the blend of \textit{memory} and \textit{resistance}.
-The theory behind the component was extracted from a missing component to link the fundamental circuit variables.
-In figure \ref{fig:fundComp}, the four fundamental variables are on each side of the square, with the ones on opposite sides being linked as shown in the following equations :
+The theory behind this component was extracted from a missing component to link the four fundamental circuit variables, voltage ($v$), charge ($q$), current ($i$) and flux ($\phi$).
+\Cref{fig:fundComp} shows the four fundamental variables are on each side of the square, with the ones on opposite sides being linked by the following equations :
 \begin{equation}
@@
 \end{equation}
-Resistors, capacitors and inductors were already discovered and well known, so it was theorized that a fourth device should then exists to physically link flux ($\phi$) and charge ($q$).
-The flux in this case is not a magnetic flux and is defined as such : $ d\phi=V\cdot dt \implies \phi =  \int V(t) \,dt  $.\\
-The component stayed theoretical until 2008 when it's been implemented in a physical device for the first time \cite{memristorFab}. It took 37 years to actually have a working device.\\
+
+Resistors, capacitors and inductors were already very established and well known component, so it was theorized that a fourth device should then exists to physically link flux ($\phi$) and charge ($q$).
+The flux in this case is not a magnetic flux and is defined as such : $ d\phi=V\cdot dt \Rightarrow \phi =  \int V(t) \,dt  $.\\
+The component stayed theoretical until 2008 when it's been implemented in a physical device for the first time \cite{memristorFab}. It took 37 years to have an actual working device.\\
 There is then an extention of this theoretical device to another, the memristive device.
@@
 \subsection{Behavior}
+
 A memristor is defined as a non-linear two-terminal fundamental electrical component.
@@
 This enables us to manipulate the resistance of the component.
-The huge benefit of memristors is the memory component, when you power it, it will have the resistance it had before.
+The huge benefit of memristors its ability to retain its internal resistance, the device can left without power for a long period of time (retention time of minimum 10 years according to \cite{memRetention}).
+When the device is powered backup, it will have the same resistance it had before.
 

### 2023-09-20 diapo/23-09-23_meeting/main.tex
changed date for last weeks missed pres...
Ready for vendredi, ca rime

--- 2023-09-14 diapo/15-09-23_meeting/main.tex
@@
 
-\date{26/05/2023}
+\date{23/09/2023}
 
@@
     \begin{itemize}
-      \item What is the approximate onChip area of an opAmp
-      \item What is the precision of a memristor when setting the resistance
+      \item What is the approximate onChip area of an opAmp ?
+      \item What is the approximate onChip area of an resistor ?
+      \item What is the precision of a memristor when setting the resistance ?
       \item I feel like my state of the art is missing content...
       \item When would my thesis defense happen ?
-      \item I'm afriad i might not have time to finish on time wiht enough content.
+      \item Is a demo of simple system useful ?
     \end{itemize}

### 2023-09-21 3.Design/genwei.tex
Bunch of changes related to the weight and architecture files

@@
 
-The choice for the weight generation was to use Keras using the tensorflow implementation.
+The choice for the weight generation was to use Keras API with the tensorflow framework.
 However, both tensorflow and pyTorch were tried and were giving similar results, the final choice of using tensorflow was made because it's the most popular among the research group.
 
-Generating the weights simply requires to train the \ac{NN} with the required achitecture.
+Generating the weights simply requires to train the \ac{NN} with the wanted achitecture.
 Once the trainning is done, we can run an inference of the \ac{NN} to check whether the training was good.
-Then the weights are reorganized based the same model required by the netlist generator (\cref{sec:netlist}) and saved as a file.
-The files can now be taken to generate the netlist.
+This step is very important and allows to perfect the \ac{NN}.
+Once a correct architecture has been found, the weights are reorganized based the same model required by the netlist generator (\cref{sec:netlist}) and saved in a file along with a quick description of the architecture used for generating those weights for an easier generation of the netlist(\cref{sec:netlist}).
+The files can now be taken and used to generate the netlist.
 

### 2023-09-21 3.Design/lstm.tex
Bunch of changes related to the weight and architecture files

--- 2023-09-20 3.Design/fullsys.tex
@@
-\section{Full analog system}\label{sec:fullsys}
+\section{\ac{LSTM} analog implementation}\label{sec:fullsys}
 
-\subsection{Circuit controller}%TODO maybe move in annex further
-
-The final circuit will need an external controller to manage diffenrent things in the circuit.
-This wasn't created in this thesis but will have to be in order to have a functionning circuit once fabricated %TODO ref annex
-
-The controller will increase the power consumption of the overall system.
-However, due to the slow (precision of about $1\mu s$) nature of the system requirements, the controller will be very low power, and won't affect the overall power consumption of the final system.
-
-The controller will have several roles in the system :
-\begin{itemize}
-\item The first role will be to manage the different flags. This means that the controller will have to set those to high or low in order to control the system.
-\end{itemize}
+\subsection{Inputs}%TODO oen section about datasets ?

### 2023-09-21 3.Design/netlist.tex
Small change
Bunch of changes related to the weight and architecture files

@@
 
-In order to be able to run the simulation with different kinds of \ac{NN} architecture. The point of the part is thus to explain how the netlist generator tool works.
+In order to be able to run the simulation with different kinds of \ac{NN} architecture.
+The point of the part is thus to explain how the netlist generator tool works.
+The tool also work to generate any architecture with the supported layers (listed in the README.md of \cite{lstmGen}).
 
@@
 \begin{itemize}
-  \item The first parameter is the number of input we use our system. This is the size of the input vector ($x_t$) from \cref{fig:lstmCell}.
+  \item The first parameter is the number of input we use our system. This is the size of the first input vector ($x_t$).
   \item The next parameter is the number of time steps. Everything about the \ac{LSTM} time steps is all explained in \cref{sec:lstm}.
-  \item The number of hidden state of the \ac{LSTM} network.
-  \item The serial size of the crossbar arrays, as described in \cref{sec:xbarCircuit}, used in the \ac{LSTM} network.
-  \item The files containing the weights for each part of the circuit. The weights in the files have to be organized using a specific model (\cref{subsec:weiStore}).
+  \item The serial size of the crossbar arrays, as described in \cref{sec:xbarCircuit}, used in the \ac{LSTM} network. This parameter must divide the number of hidden states in the \acp{LSTM} layers.
+  \item The files containing different info about the model.
+This file contains the type of \ac{NN} architecture and the weights associated with each layer.
+The weights in the files have to be organized using a specific model (\cref{subsec:weiStore}).
   \item Finally the name of the file in which the output of the script (the netlist) will be written to.
@@
     w_{0} & w_{1} & \dots \\
-    \dots & w_i & \dots \\
+    \vdots & w_i & \vdots \\
     \dots & w_{n^2-2} & w_{n^2-1}\\

### 2023-09-21 5.Conclusions/further.tex
Working on netlist, to finish
Bunch of changes related to the weight and architecture files

@@
 
-\subsection{inSitu trainning}
+\subsection{Circuit controller}\label{subsec:control}
+
+The final circuit will need an external controller to manage diffenrent things in the circuit.
+This wasn't created in this thesis but will have to be in order to have a functionning circuit once fabricated %TODO ref annex
+
+The controller will increase the power consumption of the overall system.
+However, due to the slow (precision of about $1\mu s$) nature of the system requirements, the controller will be very low power, and won't affect the overall power consumption of the final system.
+
+The controller will have several roles in the system :
+\begin{itemize}
+  \item The first role will be to manage the different flags.
+This means that the controller will have to set those to high or low in order to control the system.
+This includes flags like the enable flags of the voltage driven crossbar array (\cref{subsec:serpar}) and those for the memory cell control.
+  \item Anoter use for this controller is to reset the memory cell to $V_{cm}$ ($0$ as real value equivalent, \cref{tab:valConv}).
+This will be done by enabling the input and then applying $V_{cm}$ to the input, this will set the capacitor to $V_{cm}$.
+This has to be done everytime the circuit is ran.
+  \item The controller will also be used to control the memristors internal resistance and control the weights.
+Depending on the type of memristor, the method and time to change the internal resistance differs.
+That will only be use for inSitu trainning(\cref{subsec:inSitu}).
+\end{itemize}
+
+\subsection{inSitu trainning}\label{subsec:inSitu}
 
@@
 \begin{itemize}
-\item The easiest option would be to use the external controller (\cref{sec:fullSys}) to make the backprogation computation needed.
+  \item The easiest option would be to use the external controller (\cref{subsec:control}) to make the backprogation computation needed.
 This is not the best option, as the backpropagation would be happening in digital.
@@
 This might be something that already exist that I didn't find anything about.
+
+\subsection{Netlist generator script}
+
+A big upgrade to make would be to have a completly customizable script that works similarly to tensorflow.

### 2023-09-21 5.Conclusions/further.tex
More changes, arriving in paris...

@@
 
+Training directly on the chip has a great advange over external training.
+Indeed training within the chip will allow to completely ignore the wires own resistance in the crossbar array.
+This is because they will an integral part of the weights, if the weight needs to be lowered then the resistance will be raised (because the conductances need to be lowered).
+%TODO finish with better explainnation
+
 \subsubsection{Only change the value of one of the two memristor}%TODO

### 2023-09-22 3.Design/run.tex
Added a small part, not sure if usefull

\section{Running the simulation}\label{sec:run}

The simulation is ran using Cadence's virtuoso. In order to run the simulation for all the entire dataset all at once, the corners function of the si;ulqtion is used. It is intended to be used for small variations in the circuit. Howeverm the way it is used here is different. It changes the value of the input variables (being a sperfect square wave generator set to activate once) to the values of the dataset. The next step is to start the corner simulation. The result are recorded by looking at the output net when at the time of output activation. The name of the output net is outputed by the netlist generator script (\cref{sec:netlist}).

### 2023-09-22 3.Design/models.tex
Updated voltMult

@@
 
-It must follow the following equation :
+However this part is tricky.
+Indeed, because of the voltage conversion (\cref{tab:valConv}), the equation needs to take that into consideration.
+The final equation needs to be \cref{eq:finalVoltMult} in order to have $1\cdot 1=1$ ($V_{in_1}\cdot V_{in_2}=1V$, with $V_{in_1}=V_{in_2}=1V$)
 
-\begin{equation}
-  V_{out}=(V_{in_1}-V_{cm})\cdot (V_{in_2}-V_{cm}) + V_{cm}
+\begin{equation}\label{eq:finalVoltMult}
+  V_{out}=10\cdot(V_{in_1}-V_{cm})\cdot (V_{in_2}-V_{cm}) + V_{cm}
 \end{equation}
 
-Where $V_{out}$ is the output voltage of the circuit and $V_{in_1}$ and $V_{in_2}$ are the input voltages.
+This is taken care of using in reality two parts, the actual voltage multiplier (\cref{eq:voltMult}) and a non inverting amplifier (\cref{eq:invAmp}), the circuit of which is available in \cref{apsec:invAmp}.
 
-However, the output of the voltage multiplier needs to be ajusted after because of the voltage value of a unit.
-Indeed, we need to have $1\cdot 1=1$, but if we set $V_{in_1}=V_{in_2}=1V$, we would find $V_{out}=0.91V$.
-To fix this issue, a small circuit has to be added following the output of the voltage multiplier, this circuit needs to do the following :
-
-\begin{equation}
-  V_{amp}=(V_{out}-V_{cm})\cdot10+V_{cm}
+\begin{equation}\label{eq:voltMult}
+  V_{voltMult}=-(V_{in_1}-V_{cm})\cdot (V_{in_2}-V_{cm}) + V_{cm}
+\end{equation}
+\begin{equation}\label{eq:invAmp}
+  V_{out}=-(V_{voltMult}-V_{cm})\cdot10+V_{cm}
 \end{equation}
 
-Where $V_{amp}$ is the output of the amplificator and $V_{out}$ is the output of the voltage multiplier and the input of the amplifier circuit.
-This circuit is shown in \cref{apsec:amp}.
+Where $V_{out}$ is the output voltage the inverting amplifier (\cref{apsec:invAmp}), $V_{voltMult}$ is the out voltage of the voltage multiplier itself and $V_{in_1}$ and $V_{in_2}$ are the input voltages.
+
+The \cref{eq:voltMult} is assumed possible because of the actual voltage multiplier's datasheet available at \cite{actualVoltMult}.
 

### 2023-09-22 Appendices/appendixA.tex
onchip area begin + timediagram for memcell
Updated voltMult

@@
 
-\section{Amplificator}\label{apsec:amp}
+\section{Buffer}\label{apsec:buffer}
 
-This is the circuit that comes after the voltage multiplier (section \ref{sec:models}), it simply consists of two inverting amplifiers with one amplifying the results by 10.
-Figure TODO shows this simple circuit.
+This is a classic circuit that only takes one \ac{opAmp} wired as like in figure TODO. It is used to forward a voltage without impacting the original voltage.
+
+\section{Inverting amplifier}\label{apsec:invAmp}
+
+This is the circuit that comes after the voltage multiplier (section \ref{sec:models}), it is here to amplify the results of the voltage multiplier by 10.
+Sometimes, its use doubles as an adder making it an inverting summing amplifier (TODO circuit).
 

### 2023-09-26 3.Design/lstm.tex
More changes, arriving in paris...
Trying to add the lstm circuit figure
A bunch of changes, can't be detailed

@@
-\section{\ac{LSTM} analog implementation}\label{sec:fullsys}
+\section{\ac{LSTM} analog implementation}\label{sec:lstmCircuit}
+
+\subsection{Circuit}
+
+This section describes the circuit of an \ac{LSTM} with an input vector of size $n_i$, a $n_h$ hidden states, a serial size of $n_s$ and $n_{ts}$ time steps.
+$n_o=n_h/n_s$ is going to be used for future references in this section.
+In order for the crossbar array to be used $n_o$ must be an integer, in other words, $n_s$ must divide $n_h$.
+The circuit (shown in \cref{circt:lstm}) is pretty complex and contains numerous parts that require explaination.
+
+\begin{figure}[H]
+  \centering
+  \includesvg[width=\textwidth,pretex=\tiny]{lstm/lstmCircuit}
+  \caption{\ac{LSTM} circuit}
+  \label{circt:lstm}
+\end{figure}
+
+The system is built using crossbar array (\cref{sec:xbarCircuit}) with ($n_i+n_h+1$,$n_o$, $n_s$) as parameters.
+
+First of all, the different vectors and variables present in the schematic have to be described :
+
+\begin{itemize}
+  \item $\overrightarrow{x_t}$ : This is the input vector for the \ac{LSTM} circuit at time $t$. It has a size of $n_i$.
+  \item $\overrightarrow{h_t}$ : This is the hidden layer vector for the feedback connections, it is defined as $\overrightarrow{h_t}=(\overrightarrow{h_{t,i}}) \forall i\in [\![0,n_o-1]\!]$, with $\overrightarrow{h_{t,i}}=(\overrightarrow{h_{t,i,j}}) \forall j\in [\![0,n_s-1]\!]$.
+  \item $\overrightarrow{z_t}$ is the input of the crossbar but not the input of the \ac{LSTM}.
+This vector is there to lighten the informations on the schematic (\cref{circt:lstm}).
+$\overrightarrow{z_t}$ is defined by $\overrightarrow{z_t}=(\overrightarrow{x_t},\overrightarrow{h_{t-1}},b)$.
+  \item $\overrightarrow{o_{t,i}}$,$\overrightarrow{i_{t,i}}$,$\overrightarrow{\tilde{c}_{t,i}}$,$\overrightarrow{f_{t,i}}$ all work the same way.
+They are each the output of the activated \ac{VMM} and they are in subparts to show that from the output of the
+  \item $\overrightarrow{x_t}$
+\end{itemize}
+
+The wires coming into the crossbar are a bus of size $n_i+n_h+1$ and the output of the crossbar is a bus of size $n_o$.
+This is why everything in the system apart from the crossbar arrays is only shown once in \cref{circt:lstm} but in reality, those components are present $n_o$ times.
+This is to manage the system in parallel.
 

### 2023-09-26 3.Design/lstm.tex
Working on the lstm part, but need to change the figure again ...

@@
 $\overrightarrow{z_t}$ is defined by $\overrightarrow{z_t}=(\overrightarrow{x_t},\overrightarrow{h_{t-1}},b)$.
-  \item $\overrightarrow{o_{t,i}}$,$\overrightarrow{i_{t,i}}$,$\overrightarrow{\tilde{c}_{t,i}}$,$\overrightarrow{f_{t,i}}$ all work the same way.
-They are each the output of the activated \ac{VMM} and they are in subparts to show that from the output of the
-  \item $\overrightarrow{x_t}$
+  \item $e_{j,0}$ and $e_{j,1}$ are two enable flags that respectively represent the first and second half of $e_j$.
+  \item $e_{in}$ and $e_{out}$ are the flags used to enable the hidden state values to go to the input (feedback connection) or to the output of the circuit.
 \end{itemize}
@@
 This is why everything in the system apart from the crossbar arrays is only shown once in \cref{circt:lstm} but in reality, those components are present $n_o$ times.
-This is to manage the system in parallel.
+Those extra components are needed in order for the parallel channels to work.
+
+\subsubsection{Memcells}
+
+The memcells use and their number is going to be
 

### 2023-10-02 1.Introduction/motivation.tex
Much else to add, need inspi

--- 2023-07-04 1.Introduction/1.motivation.tex
@@
 That is useful especially for VR headsets that need to know their location in the room.
-Having an embedded chip to do all this computation.
+The embedded chip would allow this computation to be much faster than other methods, and thus give more realistic results.
+
+The focus of this thesis is to create a working simulation of an analog \ac{LSTM} circuit using Cadence's virtuoso as simulation software.
+Having those preliminary results would allow for fabricating and testing such a chip.
+Once a working prototype is manufactured, such a circuit could highly improve embedded \ac{AI} use.

### 2023-10-02 3.Design/genwei.tex
Bunch of changes moving to bu

@@
 The choice for the weight generation was to use Keras API with the tensorflow framework.
-However, both tensorflow and pyTorch were tried and were giving similar results, the final choice of using tensorflow was made because it's the most popular among the research group.
+However, both tensorflow and pyTorch were tried and were giving similar results, the final choice of using tensorflow was made because it's the most popular among the research group.%INESC ID ???
 
-Generating the weights simply requires to train the \ac{NN} with the wanted achitecture.
-Once the trainning is done, we can run an inference of the \ac{NN} to check whether the training was good.
-This step is very important and allows to perfect the \ac{NN}.
+\subsection{Training the weights}
+
+Generating the weights requires to train a \ac{NN}.
+This is done in Keras by creating the model, in other words the architecture required to solve the problem.
+This model contains the description of the \ac{NN}, such as the type of \ac{RNN} and the Dense layers that would come before of after the \ac{RNN}.
+In Keras, there are only two of the \ac{LSTM} variants (\cref{sec:lstm}) available :
+
+\begin{itemize}
+\item The \ac{NP} \ac{LSTM} : This is most commonly used version and by such one of the two available in Keras.
+\item The \ac{GRU} : This one is present here because it's considered to be different from an \ac{LSTM} \ac{NN} despite their similarities.
+\end{itemize}
+
+The other kinds of \ac{LSTM} are not present in the default Keras librairies.
+
+Training can only be done outside of the circuit as this state of the project.
+However, in order to train the weights as closely as they would have been on a real circuit, the activation functions used for the training are the custom activation functions generated by the dedicated circuit (\cref{sec:af}).
+In other words, the activation functions used are the ones shown in \cref{fig:afGraph}.
+This allows to train the weights as they almost (considering the activation functions are recreated using 51 points) as they would have been in the real circuit.
+
+All the training are done using \ac{MSE} loss function and Adam optimizer \cite{adamOpti}.
+
+\subsection{Exporting the weights}
+
+Once all the parameters (number of hidden states, number of dense layers, etc) have been set. The weights can be exported using the required weight
+
 Once a correct architecture has been found, the weights are reorganized based the same model required by the netlist generator (\cref{sec:netlist}) and saved in a file along with a quick description of the architecture used for generating those weights for an easier generation of the netlist(\cref{sec:netlist}).
@@
 
-The code used for all the weights generations is available at \cite{lstmWei}.
+The code used for all the weights generations is available at \cite{lstmWei}. %TODO README

### 2023-10-02 3.Design/netlist.tex
Bunch of changes moving to bu

@@
 The weights are stored in a single file that is the save a numpy array.
-The first dimension contains the weights for the different layers.
+This file also contains a brief description of the architecture being used.
+The first index stores this desciption.
+the following indexes are for the actual weights, in the order given in the description.
 Depending on the layer used, the weights are stored a bit differently :
@@
 \Cref{mtrx:wei} shows the position ($i$) of the weight ($w_i$) in the matrix.
+This is represented in the description as "Dense($n_o$)" where $n_o$ is the number of output
   \item \ac{LSTM} layer : An \ac{LSTM} layer contains four \acp{VMM} which can be assimilated and thus stored like a dense layer. The weights for each \ac{VMM} is then stored in a sub list.

### 2023-10-02 3.Design/netlist.tex
Finished netlist part

@@
 \Cref{mtrx:wei} shows the position ($i$) of the weight ($w_i$) in the matrix.
-This is represented in the description as "Dense($n_o$)" where $n_o$ is the number of output
-  \item \ac{LSTM} layer : An \ac{LSTM} layer contains four \acp{VMM} which can be assimilated and thus stored like a dense layer. The weights for each \ac{VMM} is then stored in a sub list.
+This is represented in the description as "Dense($n_o$)" where $n_o$ is the number of output of the layer.
+  \item \ac{LSTM} layer : An \ac{LSTM} layer contains four \acp{VMM} which can be assimilated and thus stored like a dense layer.
+The weights for each \ac{VMM} is then stored in a sub list.
+This is represented in the description as "LSTM($n_h$)" where $n_h$ is the number of hidden states of the layer.
 \end{itemize}

### 2023-10-02 3.Design/genwei.tex
Finished updating genwei

@@
 In other words, the activation functions used are the ones shown in \cref{fig:afGraph}.
-This allows to train the weights as they almost (considering the activation functions are recreated using 51 points) as they would have been in the real circuit.
+This allows to train the weights as they almost (considering the activation functions are recreated using 51 simulated points) as they would have been in the real circuit.
 
-All the training are done using \ac{MSE} loss function and Adam optimizer \cite{adamOpti}.
+All the weight trainings are done using \ac{MSE} loss function and Adam optimizer \cite{adamOpti}.
 
@@
 
-Once all the parameters (number of hidden states, number of dense layers, etc) have been set. The weights can be exported using the required weight
-
-Once a correct architecture has been found, the weights are reorganized based the same model required by the netlist generator (\cref{sec:netlist}) and saved in a file along with a quick description of the architecture used for generating those weights for an easier generation of the netlist(\cref{sec:netlist}).
-The files can now be taken and used to generate the netlist.
+Once all the parameters (number of hidden states, number of dense layers, etc) have been set.
+The weights can be exported using the required weight repartition layed out in \cref{sec:netlist}.
+A description of the architecture is also saved along the weights.
+This file can now be used as the input for the netlist generator (\cref{sec:netlist}).
 

### 2023-10-02 5.Conclusions/further.tex
Pause midi

@@
 
-\subsection{Netlist generator script}
+\subsection{Other \acp{LSTM} variants}
 
-A big upgrade to make would be to have a completly customizable script that works similarly to tensorflow.
+As discused before(\cref{sec:genwei}), they are only two of the \ac{LSTM} variants available in Keras.
+Nonetheless, it is possible to write custom layer function for Keras.
+Creating a layer from scratch is very time consuming and not so interresting for the scope the results needed.
+It can be done to extend the reach of the work.

### 2023-10-02 3.Design/memcell.tex
Changed circuit's ref
worked on area
Changed pins

@@
 
-The circuit is shown in \cref{fig:memcellCircuit}. The value/voltage is trapped in the capacitor using \ac{CMOS} switches.
+The circuit is shown in \cref{circt:memcell}. The value/voltage is trapped in the capacitor using \ac{CMOS} switches.
 
@@
   \caption{Memory cell circuit}
-  \label{fig:memcellCircuit}
+  \label{circt:memcell}
 \end{figure}
@@
 
-With $A_{CMOS} = w\cdot l = 9\cdot 10^{-7} \cdot 6 \cdot 10^{-8} = 5.4 \cdot 10^{-14} m^2$ TODO check value w et l
+With $A_{CMOS} = w\cdot l = 2\cdot 10^{-7} \cdot 6 \cdot 10^{-8} = 1.2 \cdot 10^{-14} m^2$
 
@@
   \includesvg[height=2.5cm]{memcell/memcellSymbol}
-  \caption{Memory cell symbol with the input pin (left), the output pin (right), the input enable pin (top) and the output enable pin (bottom)}
-  \label{fig:memcellCircuit}
+  \caption{Memory cell symbol with the input enable pin (top) and the output enable pin (bottom). The left and right pins are interchangebly the input and ouput for the sake of readability}
+  \label{circt:memcell}
 \end{figure}

### 2023-10-02 3.Design/lstm.tex
Changes to thedescription, working on simus..
Bunch of chnages

@@
   \item $\overrightarrow{x_t}$ : This is the input vector for the \ac{LSTM} circuit at time $t$. It has a size of $n_i$.
-  \item $\overrightarrow{h_t}$ : This is the hidden layer vector for the feedback connections, it is defined as $\overrightarrow{h_t}=(\overrightarrow{h_{t,i}}) \forall i\in [\![0,n_o-1]\!]$, with $\overrightarrow{h_{t,i}}=(\overrightarrow{h_{t,i,j}}) \forall j\in [\![0,n_s-1]\!]$.
+  \item $\overrightarrow{h_t}$ : This is the hidden layer vector for the feedback connections, it is defined as $\overrightarrow{h_t}=(\overrightarrow{h_{t,i}}) \forall i\in [\![0,n_o-1]\!]$, with $\overrightarrow{h_{t,i}}=(h_{t,i,j}) \forall j\in [\![0,n_s-1]\!]$.
   \item $\overrightarrow{z_t}$ is the input of the crossbar but not the input of the \ac{LSTM}.
@@
   \item $e_{in}$ and $e_{out}$ are the flags used to enable the hidden state values to go to the input (feedback connection) or to the output of the circuit.
+  \item $e_{next}$ is the enable flag on in between two time steps.
+  \item $R_{amp0}$ and $R_{amp1}$ are the two resistances used to amplify the output voltage of the voltage multipliers.
+The ratio of the resistances value must of $\frac{R_{amp1}}{R_{amp0}}=10$.
+The resistances must stay arround the values of the resistances used around the circuit, especially those of the memristors.
 \end{itemize}
@@
 The wires coming into the crossbar are a bus of size $n_i+n_h+1$ and the output of the crossbar is a bus of size $n_o$.
-This is why everything in the system apart from the crossbar arrays is only shown once in \cref{circt:lstm} but in reality, those components are present $n_o$ times.
+This is why everything in the system apart from the crossbar arrays is only shown once in \cref{circt:lstm} but in reality those components are present $n_o$ times.
 Those extra components are needed in order for the parallel channels to work.
 
-\subsubsection{Memcells}
+\subsection{Doubling memory cells}
 
-The memcells use and their number is going to be
+\subsubsection{Feedback hidden states}
 
-\subsection{Inputs}%TODO oen section about datasets ?
+In order for the system to work in serial mode, the memory cells of the output need to be doubled.
+This is done in \cref{circt:lstm}, and allows for the hidden states to be saved for the next stage.
+Indeed, if using the system in serial mode with only one line of memory cells, the old hidden state ($\overrightarrow{h_{t-1}}$) is slowly overwritten by the current hidden state ($\overrightarrow{h_t}$).
+As soon as the first $n_o$ serial values are computed, they are going to override the old ones, still required for the following serial values.
+\Cref{tim:lstmMemcell} shows that the value in the memory cell changes too early in the cycle and gives a bad input for the next serial values.
+The value given to the input of the \ac{LSTM} should always be the old hidden state ($\overrightarrow{h_{t-1}}$).
+
+\begin{figure}[H]
+  \centering
+  \begin{tikztimingtable}
+    $e_{n_s-1}$ & 3H8Lx \\
+    $e_0$ & 3L4H4Lx \\
+    $e_1$ & 7L4Hx \\
+    $m_{i,0}$ & 3D{$h_{t-1,i,0}$}8D{$h_{t,i,0}$}x \\
+    $m_{i,1}$ & 7D{$h_{t-1,i,1}$}4D{$h_{t,i,1}$}x \\
+    \extracode
+    \tablerules
+    %\draw (0,0) circle (2pt); % Origin
+    \begin{pgfonlayer}{background}
+      \vertlines[help lines]{3.05}
+      %\vertlines[red]{1.6,5.6,15.6}
+      %\vertlines[blue]{3.6,9.6,15.6}
+    \end{pgfonlayer}
+  \end{tikztimingtable}
+  \caption{Time diagram of the values in the different memory cells ($m_{i,j}$ being the value stored in the $i^{th}$ parallel memory cells for the $j^{th}$ serial value).
+It is assumed that $n_s>1$ (the system is in serial mode).
+This graph does not take into account the pauses $e_{next}$ between the time steps.}
+  \label{tim:lstmMemcell}
+\end{figure}
+
+Using two memory cells when in serial mode allows for this issue to be elevated.
+The values are transfered from one level to an other at the end of the time step.
+That way, the values are updated safely during each time step, without changing the final output.
+
+When the system is running is a fully parallel mode, doubling the memory cells serves no purpose, and the system can then use only one line of memory cell to save onChip area.
+
+\subsubsection{Cell states}
+
+Cell states have the same solution to a similar problem.
+Here the problem is that the old cell state value needs to be used to compute the current values for the hidden states ($\overrightarrow{h_t}$) and cell states ($\overrightarrow{c_t}$).
+The issue is that at the value in the memory cell needs to be conserved until the next serial step.
+This means the time restraints for this issue is much smaller than for the latter.
+It could possibly allow for another solution to save onChip area (\cref{subsec:noDoubleMemcell}).
+
+Once again, when the system is running in fully parallel mode, only one memory cell line is required for a normal behavior.
+
+\subsection{}

### 2023-10-02 5.Conclusions/further.tex
Bunch of chnages

@@
 It can be done to extend the reach of the work.
+
+\subsection{Avoiding doubling memory cells}\label{subsec:noDoubleMemcell}
+
+This can be could be solved by adding a single capacitor before the input pin of the memory cell and then enabling the input of the memory cell at the very end of the serial time step.
+
+This is not a working solution, simply a theory on how to optimize onChip area.

### 2023-10-02 3.Design/crossbar.tex
correct mistake in crossbar and better figure, not finished tho
A bunch of changes, can't be detailed
Bunch of chnages
Great rewrite of lstm and xbar, not finished

@@
   \centering
-  \includesvg[height=8cm]{crossbar/doubleMem}
+  \includesvg[height=8cm,pretex=\scriptsize]{crossbar/doubleMem}
   \caption{Simplified circuit of a double memristor per synapse architecture}
@@
   \label{eq:doubleMem0}
-  V_{opAmp0}=-R\cdot i_+ \Leftrightarrow
+  V_{opAmp0}=-R_r\cdot i_+ \Leftrightarrow
 \end{equation}
@@
   \label{eq:doubleMem1}
-  i_{R_f}=i_-+\frac{V_{opAmp0}}{R}=i_--i_+
+  i_{R_f}=i_-+\frac{V_{opAmp0}}{R_r}=i_--i_+
 \end{equation}
@@
 
-The idea using the system in a serial mode came from \cite{thesisRef}.
-The principle of this system is to save onChip area by reducing the number of \ac{opAmp} required.
-It also, and mainly reduces the point wise part of the \ac{LSTM} circuit.
+The idea using the system in a serial mode came from \cite{thesisRef}. The principle of this system is to save onChip area by reducing the number of \ac{opAmp} required.
 
-This serial system has great advantages but comes with a few down sides. Let's assume the number of hidden state ($n_h$) is the major changing element in the final system(TODO ref final onchip area).
+The \ac{CMOS} switches are here to open the necessary input gates when the output is required, the \ac{CMOS} switches are controlled as in \cref{tim:serpar}.
 
-\begin{itemize}
-  \item When the system is serialized, the inference time increases with a factor of $O(n_h)$.
-  \item When the system is in parallel, the circuit onChip area increases with a factor of $O(n_h)$.
-\end{itemize}
-
-However, in both cases, the overall onChip area increases with a factor of $O(n_h^2)$.
-That means that using the parallel system won't increase the onChip area by a lot as $n_h$ rises, because the $O(n_h)$ factor of the parallel version is absorbed as $O(n_h^2+n_h)=O(n_h^2)$.
-
-That leads to the conclusion that the best thing to do is use a both serialized and parallel system, using the limiting inference time for our system to set the maximum serial size.
-Depending on the system we're using this circuit for, we might have different time requirements.
-
-The \ac{CMOS} switches are here to open the necessary input gates when the output is required, the gates are controlled as in \cref{tim:serpar}.
 \begin{figure}[H]
@@
   \begin{tikztimingtable}%TODO change or find other way
-    $e_0$ & xH3Lx\\
-    $e_1$ & xLH2Lx\\
-    $e_2$ & x2LHLx\\
-    $e_3$ & x3LHx\\
+    $e_0$ & x2H2L N(A1) 2L N(A2) 2L N(A3) 2L N(A4) 2Lx\\
+    $e_1$ & x2L2H N(B1) 2L N(B2) 2L N(B3) 2L N(B4) 2Lx\\
+    $e_i$ & x4L N(C1) 2L N(C2) 2H N(C3) 2L N(C4) 2Lx\\
+    $e_{n_s-1}$ & x4L N(D1) 2L N(D2) 2L N(D3) 2L N(D4) 2Hx\\
+    %foo & 2L N(A1)  4H N(A2) L\\
     \extracode
+    \node[gap, at={($(A1|-A2)!0.5!(A2)$)}];
+    \node[gap, at={($(A3|-A4)!0.5!(A4)$)}];
+    \node[gap, at={($(B1|-B2)!0.5!(B2)$)}];
+    \node[gap, at={($(B3|-B4)!0.5!(B4)$)}];
+    \node[gap, at={($(C1|-C2)!0.5!(C2)$)}];
+    \node[gap, at={($(C3|-C4)!0.5!(C4)$)}];
+    \node[gap, at={($(D1|-D2)!0.5!(D2)$)}];
+    \node[gap, at={($(D3|-D4)!0.5!(D4)$)}];
     \tablerules
@@
     \begin{pgfonlayer}{background}
-      \vertlines[help lines]{0.6,4.6}
+      \vertlines[help lines]{0.55,12.55}
       %\vertlines[red]{1.6,5.6,15.6}
@@
   \end{tikztimingtable}
-  \caption{Example of enable flags timing with $n_s=4$ in a single time step}
+  \caption{Enable flags timing for any value of $n_s$ in a single time step}
   \label{tim:serpar}
@@
 
+When the system is used fully in parallel, the \ac{CMOS} switches are not required can then be removed to lower the final onChip area.
+
 %TODO understand better \subsection{Sneak path problem}
 \subsection{Symbol}
-The symbol (\cref{sym:xbar}) defined for the voltage based memristor crossbar array used in this thesis is more compact and helps the readability of the final circuit.
-It contains the size of each parameters on its symbol.
+The symbol (\cref{sym:xbar}) defined for the voltage based memristor crossbar array used in this thesis is more compact and helps the readability of the circuits that require a crossbar array.
 It depends on several parameters, the number of inputs ($n_i$), the number of outputs ($n_o$) and the serial size ($n_s$).
@@
   \includesvg[height=2.5cm]{crossbar/xbarSymbol}
-  \caption{Symbol used for the crossbar array}
+  \caption{Symbol used for the crossbar array, the input pin is a bus of size $n_i$ and the output pin is a bus of size $n_o$}
   \label{sym:xbar}
@@
 
-The total onChip area for the crossbar circuit depends on the previously defined parameters.
+The total onChip area for the crossbar circuit depends on the previously defined parameters. It also slightly changes depending on the value of $n_s$.
 
 \begin{equation}
-  A_{xbar}(n_i,n_o,n_s)=2\cdot n_i\cdot n_o \cdot n_s\cdot A_{memristor}+2\cdot A_{CMOS}\cdot n_o\cdot n_s +n_o\cdot(2\cdot[A_{opAmp}+A_R]+A_{R_f})
+  A_{xbar}(n_i,n_o,n_s)=
+  \begin{cases}
+    2\cdot n_i\cdot n_o \cdot A_{memristor}+n_o\cdot(2\cdot[A_{opAmp}+A_{R_r}]+A_{R_f}),\hfill n_s=1\\
+    \begin{array}{2}
+      2\cdot n_i\cdot n_o \cdot n_s\cdot A_{memristor}+2\cdot A_{CMOS}\cdot n_o\cdot n_s\\
+      +n_o\cdot(2\cdot[A_{opAmp}+A_{R_r}]+A_{R_f})
+    \end{array}
+    ,\hfill \forall n_s>1\\
+  \end{cases}
 \end{equation}
@@
 
-%TODO Show off results ??
+\subsubsection{Input voltages}
 
@@
 
+\subsubsection{Dense layer}
+
+A dense layer is simply a \ac{VMM}.
+This means that a dense layer in a \ac{NN} can be assimilated with a crossbar array.
+Dense layers are mostly used as by fully parallel voltage-driven crossbar array in this thesis to keep the timing simple, otherwise the timing of the enable flags, and later on the \ac{LSTM}'s memory cells (\cref{sec:lstm}), would get too complicated.
+
+\subsubsection{Resistors vs Memristors}
+
 The analog system being only simulated for the inference of the \ac{NN}.

### 2023-10-02 3.Design/lstm.tex
Great rewrite of lstm and xbar, not finished

@@
 
-\subsection{}
+\subsection{Serialization/Parallelization}
+\label{subsec:serpar}
+
+Using an \ac{LSTM} in serial mode is very beneficial as it divides the number of point wise circuit by $n_s$.
+
+This serial system has great advantages but comes with a few down sides.
+If the system we are working with has $n_h$ hidden states.
+This number of hidden states ($n_h$) is the major changing element in the final system(TODO ref final onchip area).
+
+TODO : Rework
+\begin{itemize}
+  \item When the system is serialized, the inference time increases with a factor of $O(n_h)$.
+  \item When the system is in parallel, the pointwise circuit's onChip area increases with a factor of $O(n_h^2)$.
+    This is due to the fact that the onChip area of the pointwise part of the circuit increases with a factor of $O(n_h)$ and the onChip area of a crossbar array increase with a factor of $O(n_h^2)$.
+So the final increase of onChip area is $O(n_h^2+n_h)=O(n_h^2)$.
+\end{itemize}
+
+However, in both cases, the overall onChip area of the crossbar arrays increases with a factor of $O(n_h^2)$.
+That means that using the parallel system won't increase the onChip area by a lot as $n_h$ rises, because the $O(n_h)$ factor of the parallel version is absorbed as $O(n_h^2+n_h)=O(n_h^2)$.
+
+That leads to the conclusion that the best thing to do is use an hybrid system (both serialized and parallel), using the limiting inference time for our system to set the maximum serial size.
+Depending on the system we're using this circuit for, we might have different time requirements.
+
+\subsection{Symbol}
+The symbol of the \ac{LSTM} circuit can be found in \cref{sym:lstm}.
+It contains the size of each parameters on its symbol. It depends on several parameters, the number of inputs ($n_i$), the number of outputs ($n_o$) and the serial size ($n_s$).
+
+\begin{figure}[H]
+  \centering
+  \includesvg[height=2.5cm]{crossbar/xbarSymbol}
+  \caption{Symbol used for the \ac{LSTM} circuit}
+  \label{sym:lstm}
+\end{figure}
+
+The total onChip area for the crossbar circuit depends on the previously defined parameters.
+
+\begin{equation}
+  A_{xbar}(n_i,n_o,n_s)=
+  \begin{cases}
+    2\cdot n_i\cdot n_o \cdot A_{memristor}+n_o\cdot(2\cdot[A_{opAmp}+A_{R_r}]+A_{R_f}),\hfill n_s=1\\
+    \begin{array}{2}
+      2\cdot n_i\cdot n_o \cdot n_s\cdot A_{memristor}+2\cdot A_{CMOS}\cdot n_o\cdot n_s\\
+      +n_o\cdot(2\cdot[A_{opAmp}+A_{R_r}]+A_{R_f})
+    \end{array}
+    ,\hfill \forall n_s>1\\
+  \end{cases}
+\end{equation}
+
+With $A_{CMOS}$ being the same than for the memory cell (\cref{sec:memcell}), $A_{CMOS}=1.2\cdot 10^{-14}m^2$, and the feature size of typical memristor that can be fabricated at INESC is $3\mu m$, which would make the approximate area for a memristor $A_{memristor}=9\mu m^2=9\cdot 10^{-12} m^2$.%TODO check info and inesc citation

### 2023-10-03 2.State/lstm.tex
New figures and stuff

@@
   \centering
-  \includesvg[width=\textwidth]{lstm/lstmCell.svg}
-  \caption{LSTM Cell}
-  \label{fig:lstmCell}
+  \begin{minipage}{\columnwidth}
+    \subfloat[\acs{LSTM} cell\label{fig:lstmCell}]{\includesvg[width=\textwidth,pretex=\large]{lstm/lstmCell.svg}}
+  \end{minipage}
+  \begin{minipage}{\columnwidth}
+    \subfloat[Legend\label{leg:lstm}]{\includesvg[width=\textwidth,pretex=\large]{lstm/lstmLegend.svg}}
+  \end{minipage}
+  \caption{}
 \end{figure}
@@
 \Cref{fig:lstmCell} shows the complexity of the \ac{LSTM} architecture.
-In an \ac{LSTM}, each gate is a different \ac{NN} and then activated with either a tanh or a sigmoid activation function.
+In an \ac{LSTM}, each gate is a different \ac{NN} and then activated with either a \ac{tanh} or a sigmoid activation function.
 Each input to the cell is a vector.
@@
   \includesvg[width=\textwidth]{lstm/lstmUse}
-  \caption{LSTM use case}
+  \caption{LSTM use case, legend in \cref{leg:lstm}}
   \label{fig:lstmUse}

### 2023-10-04 3.Design/lstm.tex
Changes to the lstm circuit

@@
 
-The wires coming into the crossbar are a bus of size $n_i+n_h+1$ and the output of the crossbar is a bus of size $n_o$.
+The wires coming into the crossbar are a bus of size $n_i+n_h+1$ and the output of the crossbar is a bus of size $n_o$ (\cref{sym:xbar}).
 This is why everything in the system apart from the crossbar arrays is only shown once in \cref{circt:lstm} but in reality those components are present $n_o$ times.
@@
 
-When the system is running is a fully parallel mode, doubling the memory cells serves no purpose, and the system can then use only one line of memory cell to save onChip area.
+When the system is running is a fully parallel mode, doubling the memory cells still causes a problem.
+The enable in and enable out flags can't be high at the same time when the memory cell has a feedback connection, the signal would interfere with it self and change its own value.
+Potential solutions are discussed in \cref{subsec:noDoubleMemcell}.
 
@@
 Cell states have the same solution to a similar problem.
-Here the problem is that the old cell state value needs to be used to compute the current values for the hidden states ($\overrightarrow{h_t}$) and cell states ($\overrightarrow{c_t}$).
+Here the problem is that the memory cell has a feedback connection and needs to be activated within one serial step.
+The value of the old cell state needs to be used to compute the current values for the hidden states ($\overrightarrow{h_t}$) and cell states ($\overrightarrow{c_t}$).
 The issue is that at the value in the memory cell needs to be conserved until the next serial step.
-This means the time restraints for this issue is much smaller than for the latter.
-It could possibly allow for another solution to save onChip area (\cref{subsec:noDoubleMemcell}).
+(\cref{subsec:noDoubleMemcell}).
 
@@
 \subsection{Symbol}
+
 The symbol of the \ac{LSTM} circuit can be found in \cref{sym:lstm}.
-It contains the size of each parameters on its symbol. It depends on several parameters, the number of inputs ($n_i$), the number of outputs ($n_o$) and the serial size ($n_s$).
+It depends on several parameters, the number of inputs ($n_i$), the number of hidden states ($n_h$), the serial size of the system ($n_s$) and the number of time steps ($n_{ts}$).
 

### 2023-10-04 5.Conclusions/further.tex
Changes to the lstm circuit

--- 2023-10-02 5.Conclusions/further.tex
@@
 It can be done to extend the reach of the work.
+
+\subsection{Avoiding doubling memory cells}\label{subsec:noDoubleMemcell}
+
+Here there are two different problems that have two similar solutions, using a capacitor to keep the voltage in the node for a longer time :
+\begin{itemize}
+  \item For the hidden state feedback memory cells, a potential solution would be adding a single capacitor before the input pin of the memory cell and then enabling the input during the pause between time steps ($e_{next}$).
+This would allow the enable in and enable out flags to be high at different times.
+This solution only works in fully parallel mode.
+  \item For the cell state memory cells, adding a capacitor after the output pin of each to store the value after the output enable flag goes to low.
+\end{itemize}
+
+This is not a working solution, simply a theory on how to optimize onChip area.

### 2023-10-04 3.Design/lstm.tex
changed serial vs par

@@
 
-This serial system has great advantages but comes with a few down sides.
-If the system we are working with has $n_h$ hidden states.
-This number of hidden states ($n_h$) is the major changing element in the final system(TODO ref final onchip area).
+There is a sweet spot for the serial size ($n_s$) :
 
-TODO : Rework
 \begin{itemize}
-  \item When the system is serialized, the inference time increases with a factor of $O(n_h)$.
-  \item When the system is in parallel, the pointwise circuit's onChip area increases with a factor of $O(n_h^2)$.
-    This is due to the fact that the onChip area of the pointwise part of the circuit increases with a factor of $O(n_h)$ and the onChip area of a crossbar array increase with a factor of $O(n_h^2)$.
-So the final increase of onChip area is $O(n_h^2+n_h)=O(n_h^2)$.
+  \item When the serial size ($n_s)$ increases, the inference time increases with a factor of $O(n_s)$ while the onChip area decrease with the same factor.
+  \item When the serial size ($n_s)$ decreases, similarly, the onChip area increases with a factor of $O(n_s)$ while the tinference time decreases by the same factor.
 \end{itemize}
 
-However, in both cases, the overall onChip area of the crossbar arrays increases with a factor of $O(n_h^2)$.
-That means that using the parallel system won't increase the onChip area by a lot as $n_h$ rises, because the $O(n_h)$ factor of the parallel version is absorbed as $O(n_h^2+n_h)=O(n_h^2)$.
-
-That leads to the conclusion that the best thing to do is use an hybrid system (both serialized and parallel), using the limiting inference time for our system to set the maximum serial size.
-Depending on the system we're using this circuit for, we might have different time requirements.
+The overall onChip area is linked to the number of hidden states ($n_h$) by $O(n_h^2)$.
+This means that the serial size ($n_s$) will be set depending on the limiting factor of our system (onChip area or inference time).
 

### 2023-10-04 5.Conclusions/further.tex
further

@@
 This solution only works in fully parallel mode.
-  \item For the cell state memory cells, adding a capacitor after the output pin of each to store the value after the output enable flag goes to low.
+  \item For the cell state memory cells, adding a capacitor and a buffer (\cref{ap:buffer}) after the output of the memory cell group.
+This capacitor is here to store the voltage after the output enable flag goes to low.
+The buffer is here to output the voltage with a high impedance.
 \end{itemize}
 
-This is not a working solution, simply a theory on how to optimize onChip area.
+Those solutions are purely theorical, they haven't been tested due to lack of time.

### 2023-10-04 3.Design/dataset.tex
Few more words ...
Updates on dataset

@@
 \section{The datasets}\label{sec:dataset}
+
+\subsection{Airline}
+
+The first dataset is very simple and was mostly used to check whether the analog circuit was working.
+It is extracted from the tutorial available at \cite{airline}.
+The dataset is available at \cite{datasets}, the name of the file containing the dataset is \textit{airline-passengers.csv}.
+
+This is a regression problem, meaning that the point is to predict the number of passenger of the next month
+
+\begin{figure}[H]
+\centering
+\includesvg[width=\textwidth]{datasets/airline}
+\caption{Graphical representation of the airline passenger dataset. The vertical lines represent a full year.}
+\label{fig:afGraph}
+\end{figure}
 
@@
 
-The values for the data
+The values for the data needs to be set not to exceed 1. This is to make sure that the terminal voltage doesn't reach the threshold as explained in \cref{sec:xbarCircuit}.

### 2023-10-04 3.Design/dataset.tex
Bunch of things added to dataset

@@
 \section{The datasets}\label{sec:dataset}
+
+\subsection{The format}
+
+The values for the data needs to be set not to exceed 1. This is to make sure that the terminal voltage doesn't reach the threshold as explained in \cref{sec:xbarCircuit}.
 
@@
 
-The first dataset is very simple and was mostly used to check whether the analog circuit was working.
+The first dataset contains a time series of international airline passengers from January 1949 to December 1960, the data is recorded monthly and is in thousands.
+The dataset then contains twelve years of monthly data so $144$ sample points.
+The results of such a problem aren't useful, it was used in this thesis because of its simplicity to check whether the analog circuit is working.
 It is extracted from the tutorial available at \cite{airline}.
@@
 
-This is a regression problem, meaning that the point is to predict the number of passenger of the next month
+This is a regression problem, the role of the \ac{LSTM} is to predict the number of passenger flying the next month being given previous months' passenger count.
 
@@
 
-\subsection{The format}
+\subsubsection{Data format}
 
-The values for the data needs to be set not to exceed 1. This is to make sure that the terminal voltage doesn't reach the threshold as explained in \cref{sec:xbarCircuit}.
+The dataset contains $144$ sample points.
+It has been transformed into $142$ data points for training.
+This has been done by taking each values three by three.
+The first two values are two timesteps for the input vector and the third value being used the target value.
+Two third of the dataset is being used for training and the other third is used for validation.
+
+\subsubsection{Network configuration}
+
+The layers used to solve this problem are listed below :
+
+\begin{itemize}
+  \item \ac{LSTM} with four hidden states ($n_h=4$) and an input with feature size of one and two time steps.
+\end{itemize}
+
+The model used to solve this problem is an \ac{LSTM} with four hidden states, a batch size of one and uses two time steps.
+
+\subsubsection{Results} %TODO Move to results obviously
+
+The weights are trained using both the custom activation function (\cref{sec:genwei}) and without ones and compare their output.
+
+After training on a digital computer using the settings described in \cref{sec:genwei}.

### 2023-10-04 3.Design/run.tex
Working on rewrite

@@
 
-The simulation is ran using Cadence's virtuoso.
-In order to run the simulation for all the entire dataset all at once, the corners function of the si;ulqtion is used.
-It is intended to be used for small variations in the circuit.
-Howeverm the way it is used here is different.
+The simulation is ran using Cadence's virtuoso simulator, using a corner simulation to simulate all the inputs.
+
+In order to run the simulation for the entire dataset in one go, the corner function of the simulation is used.
+Corners are intended to be used for small variations in the circuit.
+However, it is used here as a way to set the values for dataset.
+the way it is used here is different.
 It changes the value of the input variables (being a sperfect square wave generator set to activate once) to the values of the dataset.

### 2023-10-05 3.Design/dataset.tex
New figures and more

@@
   \includesvg[width=\textwidth]{datasets/airline}
-  \caption{Graphical representation of the airline passenger dataset. The vertical lines represent a full year.}
-  \label{fig:afGraph}
+  \caption{the airline dataset in a grphic. This is the curve that the regression is trying to reproduce. The vertical lines represent a full year.}
+  \label{graph:airline}
 \end{figure}
@@
 \begin{itemize}
-  \item \ac{LSTM} with four hidden states ($n_h=4$) and an input with feature size of one and two time steps.
+  \item An \ac{LSTM} with four hidden states ($n_h=4$) and an input with feature size of one and two time steps.
+  \item A Dense layer with an output size of one.
 \end{itemize}
 
-The model used to solve this problem is an \ac{LSTM} with four hidden states, a batch size of one and uses two time steps.
+\Cref{fig:airlineModel} is a graphical representation of the model just described.
+
+\begin{figure}[H]
+  \centering
+  \includesvg[width=\textwidth]{datasets/airlineModel}
+  \caption{Model used to solve the airline passengers problem}
+  \label{fig:airlineModel}
+\end{figure}
 

### 2023-10-05 3.Design/activation.tex
more cleanning, even more coming
Changed pins
Second dataset started

@@
   \hspace*{0.8cm}
-  \caption{Activation functions symbols with the input pin on the left and the output pin on the right}
+  \caption{Activation functions symbols with the input and output pins on either side depending on the flow of the current for better readability}
   \label{fig:afSymbol}
@@
   \centering
-  \includesvg[width=\textwidth]{activation/activGraph}
-  \caption{Input/Output graph of the activation function circuit for both sigmoid and \ac{tanh} functions}
+  \includesvg[width=\textwidth]{activation/afGraph}
+  \caption{Input/Output graph of the activation function circuit for both sigmoid and \acs{tanh} functions}
   \label{fig:afGraph}

### 2023-10-05 3.Design/dataset.tex
Second dataset started

@@
 After training on a digital computer using the settings described in \cref{sec:genwei}.
+
+\subsection{\ac{C. elegans}}
+
+This dataset is far more interresting than the latter.
+This data set aims to use \acp{LSTM} to mimic the behavior of real neurons.
+As explained in \cite{celegans}, \acf{C.
+elegans} are simple organisms that are getting very popular for whole brain organization studies.
+The point of this problem is to reproduce the nervous system of the \ac{C.
+elegans}.
+This is done using recorded data of the input of 4 neurons and the output of 4 other neurons.
+
+The dataset is great for our study because :
+\begin{itemize}
+  \item It comes from a very recent paper, and means the research is happening right now.
+  \item It aims at reproducing the behavior of the brain of a simple organism (\ac{C. elegans}), domain in which having a small and low powered system to mimic is an advantage.
+  \item The paper was written by Dr. Barbulescu who is one of my supervisor.
+\end{itemize}

### 2023-10-05 1.Introduction/motivation.tex
Improved motivation

@@
 \acp{AI} are very power hungry algorithms which limits its use in embedded systems and power efficient devices.
-Furthermore, the time an \ac{AI} takes to compute its output can become quite long.
+Furthermore, the time an \ac{AI} takes to compute its output quickly rises, especially on lower end chip such as the ones used in embedded systems.
 So much so that the most complex ones only run on online and very powerful servers, like the aforementioned \textit{ChatGPT}.
@@
 
+Somre research \cite{celegans}, focuses on reproducing parts of the nervous system of a small organism.
+This could be applied on larger organisms, or, in the long term, reproduce part of the human brain.
+Using analog computers, small chips could be used to replace a part of one's brain non functionnal part.
+
 The focus of this thesis is to create a working simulation of an analog \ac{LSTM} circuit using Cadence's virtuoso as simulation software.

### 2023-10-05 3.Design/crossbar.tex
changed serial vs par
Updated time diagram

@@
 
-The idea using the system in a serial mode came from \cite{thesisRef}. The principle of this system is to save onChip area by reducing the number of \ac{opAmp} required.
+The idea using the system in a serial mode came from \cite{thesisRef}.
+The principle of this system is to save onChip area by reducing the number of \ac{opAmp} required.
+Using the system in serial mode, increases the time it takes to compute the output of the the \ac{VMM}.
+Let's assume a serial size of $n_s\in\mathbb{N}$, the number of \ac{opAmp} and resistor is then divided by $n_s$ at the cost of multipying the computation time by $n_s$.
 
@@
   \begin{tikztimingtable}%TODO change or find other way
-    $e_0$ & x2H2L N(A1) 2L N(A2) 2L N(A3) 2L N(A4) 2Lx\\
-    $e_1$ & x2L2H N(B1) 2L N(B2) 2L N(B3) 2L N(B4) 2Lx\\
-    $e_i$ & x4L N(C1) 2L N(C2) 2H N(C3) 2L N(C4) 2Lx\\
-    $e_{n_s-1}$ & x4L N(D1) 2L N(D2) 2L N(D3) 2L N(D4) 2Hx\\
+    $e_0$ & x2H2L N(A1) L N(A2) 2L N(A3) L N(A4) 2Lx\\
+    $e_1$ & x2L2H N(B1) L N(B2) 2L N(B3) L N(B4) 2Lx\\
+    $e_i$ & x4L N(C1) L N(C2) 2H N(C3) L N(C4) 2Lx\\
+    $e_{n_s-1}$ & x4L N(D1) L N(D2) 2L N(D3) L N(D4) 2Hx\\
     %foo & 2L N(A1)  4H N(A2) L\\
@@
     \begin{pgfonlayer}{background}
-      \vertlines[help lines]{0.55,12.55}
+      \vertlines[help lines]{0.55,10.55}
       %\vertlines[red]{1.6,5.6,15.6}

### 2023-10-06 3.Design/dataset.tex
Adding stuff to celegans model/problem
Changed one detail

@@
 \end{itemize}
+
+\subsubsection{Data format}
+
+The dataset contains $144$ sample points.
+It has been transformed into $142$ data points for training.
+This has been done by taking each values three by three.
+The first two values are two timesteps for the input vector and the third value being used the target value.
+Two third of the dataset is being used for training and the other third is used for validation.
+
+\subsubsection{Network configuration}
+
+The layers used to solve this problem are listed below :
+
+\begin{itemize}
+  \item An \ac{LSTM} with eight hidden states ($n_h=8$) and an input with feature size of four (because of the four input neurons that are considered), one thousand time steps and every time steps outputs a value for the next layer.
+  \item A time distributed dense layer with an output size of four. It is time distributed because it needs to compute every time steps and not only the last.
+\end{itemize}
+
+\Cref{fig:celegansModel} is a graphical representation of the model just described.
+
+\begin{figure}[H]
+  \centering
+  \includesvg[width=\textwidth]{datasets/celegansModel}
+  \caption{Model used to solve the \ac{C. elegans} problem}
+  \label{fig:celegansModel}
+\end{figure}

### 2023-10-06 3.Design/dataset.tex
Missing graphs

@@
 \begin{itemize}
-  \item It comes from a very recent paper, and means the research is happening right now.
-  \item It aims at reproducing the behavior of the brain of a simple organism (\ac{C. elegans}), domain in which having a small and low powered system to mimic is an advantage.
-  \item The paper was written by Dr. Barbulescu who is one of my supervisor.
+  \item It comes from a very recent paper, and means the research is current.
+  \item It aims at reproducing the behavior of the brain of a simple organism (\ac{C. elegans}). In the (very) long term, a full parts of the human brain could be replaced by a very low powered chip.
+  \item The paper was written by Dr. Barbulescu who happens to be one of my supervisor.
 \end{itemize}
 
-\subsubsection{Data format}
+The dataset consists of several sets of input/output, twenty are used for training, ten for testing and ten for validation.
+Those dataset are the exitations of four inputs neurons (PLML2, PLMR, AVBL, AVBR) and the expected outputs from four neurons (DB1, LUAL, PVR, VB1) extracted from a known model of \ac{C.
+elegans} \cite{celegans}.
+
+\subsubsection{Input dataset format}
 

### 2023-10-10 3.Design/dataset.tex
preparing for new graphs

@@
 \begin{itemize}
-  \item An \ac{LSTM} with eight hidden states ($n_h=8$) and an input with feature size of four (because of the four input neurons that are considered), one thousand time steps and every time steps outputs a value for the next layer.
+  \item An \ac{LSTM} with eight hidden states and an input with feature size of four (because of the four input neurons that are considered), one thousand time steps and every time steps outputs a value for the next layer.
   \item A time distributed dense layer with an output size of four. It is time distributed because it needs to compute every time steps and not only the last.
 \end{itemize}
+
+The \ac{LSTM} was chosen with a number of hidden states of eight because it gets the best results \cite{celegans}.
 

### 2023-10-10 3.Design/dataset.tex
Dataset chap update

@@
 
-The dataset consists of several sets of input/output, twenty are used for training, ten for testing and ten for validation.
-Those dataset are the exitations of four inputs neurons (PLML2, PLMR, AVBL, AVBR) and the expected outputs from four neurons (DB1, LUAL, PVR, VB1) extracted from a known model of \ac{C.
-elegans} \cite{celegans}.
+\begin{figure}[H]
+  \centering
+  \begin{minipage}{\columnwidth}
+    \subfloat[Training dataset sample]{\includesvg[width=0.5\textwidth, pretex=\tiny]{datasets/celegans/ioTrain0}}%
+    \hfill
+    \subfloat[Test dataset sample]{\includesvg[width=0.5\textwidth,pretex=\tiny]{datasets/celegans/ioTest0}}%
+  \end{minipage}
+  \begin{minipage}{\columnwidth}
+    \subfloat[Training dataset sample]{\includesvg[width=0.5\textwidth, pretex=\tiny]{datasets/celegans/ioTrain1.svg}}%
+    \hfill
+    \subfloat[Validation dataset sample]{\includesvg[width=0.5\textwidth,pretex=\tiny]{datasets/celegans/ioValid0}}%
+  \end{minipage}
+  \caption{\ac{C. elegans} dataset samples}
+  \label{graph:celegans}
+\end{figure}
 
-\subsubsection{Input dataset format}
+\subsubsection{Dataset format}
 
-The dataset contains $144$ sample points.
-It has been transformed into $142$ data points for training.
-This has been done by taking each values three by three.
-The first two values are two timesteps for the input vector and the third value being used the target value.
-Two third of the dataset is being used for training and the other third is used for validation.
+The dataset contains a set of fourty set of input/output.
+Each input/output set contains $500ms$ of data with a time step of $0.5ms$, making a thousand points for each neurons.
+This data is recorded after applying current to the input neurons (PLML2, PLMR, AVBL, AVBR), and monitoring the output of four neurons (DB1, LUAL, PVR, VB1) that are known to have strong activity during  a specific behavior of the nematode known as \ac{FCM}.
+The simulation is done using a known model of \ac{C.
+elegans} connectome (complete overview of the brains connections) \cite{celegans}.
+
+\Cref{graph:celegans} shows four sets of input/output our of the fourty that were recorded for the dataset.
+
+The inputs for the systems are then going to be the current values for the four input neurons at every of the thousand time steps and plans to get an output as close as possible to the modeled output for all time steps.
 

### 2023-10-10 3.Design/wei2res.tex
Pause midi
New wei2res

@@
 
-By solving \cref{eq:wei2res01}, we find the \cref{eq:wei2res2} that gives the values for $R_-$ and $R_+$. All the steps for solving the equations can be found in .%TODO solve equations in appendix
+By solving \cref{eq:wei2res01}, we find the \cref{eq:wei2res2} that gives the values for $R_-$ and $R_+$. All the steps for solving the equations can be found in \cref{apsec:wei2res}.
 
 The real value to voltage conversion (\cref{tab:valConv}) doesn't affect this part of the system as the crossbar array already works in voltages.
-Placing $w$ from \cref{eq:wei2res0} in \cref{eq:doubleMem2} would not change the result.
+Replacing the resistor variable in \cref{eq:doubleMem2} by $w$ using \cref{eq:wei2res0} would not change the results.
 
@@
 
-For the thesis, this step is done in python (TODO show function). Since the resistance resolution is of about%TODO check memristor precision
+While solving the system (\cref{apsec:wei2res}), we get two potential equations for $R_+$.
+Graphing them shows that one of them is outside of the memristor's resistance range ($[R_{min},R_{max}]$).
+Leaving us with only one equation because the other one is physically unreachable.
+
+For the thesis, this step is done in python and integrated in the netlist generator script (\cref{sec:netlist}). Since the resistance resolution is of about%TODO check memristor precision
 , the python code then reproduces this by limiting the amount of significant digit of the resistance.

### 2023-10-11 4.Results/airline.tex
Started results

\section{Airline dataset}
\label{sec:resAirline}

\subsection{Digital results}
\label{subsec:digitalAirline}

The digital results of the \ac{NN} once trained are very important. They are the results that the analog system will have to reproduce. Furthermore, this part is used to measure the impact of the custom activation function used for training (\cref{sec:af}).

\begin{figure}[H]
  \centering
  \includesvg[width=\linewidth]{datasets/airlineDigital}
  \caption{Graph of the digital predictions for airline dataset. The dotted vertical line shows the limit of the data used for training and the one used for validation.}
  \label{graph:airlineDigital}
\end{figure}

\Cref{graph:airlineDigital} shows the results of the \ac{NN} next to the orinal data. The results are visually very close. The error is quantified using \ac{RMSE} error function. All those curves error to each other are measured and displayed in \cref{tab:airlineDigital}.

\begin{table}[H]
  \centering
  \begin{tabular}{|c|c|c|c|}
    \hline
    %\rowcolor{gray}
    \cellcolor[HTML]{808080}\acs{RMSE} & Original data & Digital output & Custom activation functions\\
    \hline
    Original data &\cellcolor[HTML]{202020} & $45.24$ & $45.94$\\
    \hline
    Digital output & $45.24$ & \cellcolor[HTML]{202020} & $4.04$\\
    \hline
    Custom activation function & $45.94$ & $4.04$ & \cellcolor[HTML]{202020}\\
    \hline
  \end{tabular}
  \caption{\ac{RMSE} errors of each curve to the others}
  \label{tab:airlineDigital}
\end{table}

The error between the digital curves with and without the custom activation functions are visually inexistant. The \ac{RMSE} error between those curves confirms this low error as it is of only $4.04$ thousands of passengers away. Considering the values at play, this error is negligible.

Those two

\subsection{Subsection B}
\label{subsec:subbsectiona}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.5\linewidth]{cover}
  \caption[Dummy Figure Caption for List of Figures.]{Dummy Figure Caption.}
  \label{fig:dummyfigure1}
\end{figure}

Remember you can change the reference style. Another dummy citation \cite{site}.

### 2023-10-11 4.Results/airline.tex
Going to lunch
Few more lines

@@
     %\rowcolor{gray}
-    \cellcolor[HTML]{808080}\acs{RMSE} & Original data & Digital output & Custom activation functions\\
+    \cellcolor[HTML]{808080}\acs{RMSE} & Original data & Classic prediction & Custom activation functions\\
     \hline
@@
 
-Those two
+Those two curves are both very close to the target curve.
+They have a very similar error rate to this target curve, indeed they're \ac{RMSE} error is $45.24$ for the training using the original activation functions, and $45.94$ for the data created with the analog activation function.
+The difference of error between the two prediction is of about seven hundred passengers, which is very low considering the scale.
+
+Computing the \ac{NN} with or without the analog activation functions doesn't affect the digital prediction of the \ac{NN}.
+An error difference so low can be due to the initial weights of the \ac{NN}, before it was trained.
+Furthermore, training the \ac{NN} for a few more epochs would change the errors by enough to make this difference negligible.
 

### 2023-10-11 3.Design/genwei.tex
Pause midi
added weight constraint

@@
 \section{Weights generation}
-\label{sec:weigen}
+\label{sec:genwei}
 
@@
 
+\subsection{Weights constaint}
+
+The weights need to be contrained during the training to make sure that the voltage of the analog circuit stays with in its operating voltage range ($[0,V_{dd}]$).
+To avoid any unwanted behavior when the system supposedly goes out of range, the weights in the \ac{LSTM} are contrained.
+This constrain isn't fixed and depends on the architecture it is generated with.
+
+The worst case scenario for an \ac{LSTM} is when every value reaches the voltage threshold ($V_{threshold}$) (\cref{sec:xbarCircuit}) and every weight is maximized ($w_{max}$).
+When that is the case the output of the \acp{VMM} is found in \cref{eq:weiCons}.
+
+\begin{equation}\label{eq:weiCons}
+  V_{threshold}\cdot w_{max} \cdot(n_i+n_h)+w_{max}+V_{cm}= V_{dd}
+\end{equation}
+
+We can then determine the maximum acceptable value for the weights ($w_{max}$), this value has thus been determined to be the one in \cref{eq:weiConsRes}.
+
+\begin{equation}\label{eq:weiConsRes}
+  w_{max}=\frac{V_{dd}-V_{cm}}{V_{threshold}\cdot(n_i+n_h)+1}
+\end{equation}
+
+The analog system being completly centered and symetrical around $V_{cm}$ the value for the minimum weight is the opposite to the maximum one (\cref{eq:weiConsRes1})
+
+\begin{equation}\label{eq:weiConsRes1}
+  w_{min}=-w_{max}=-\frac{V_{dd}-V_{cm}}{V_{threshold}\cdot(n_i+n_h)+1}
+\end{equation}
+
+A way to remove those constraints is discussed in \cref{subsec:noCons}. Having a constraint on the weights limits the performance of the final \ac{NN}.
+
 \subsection{Exporting the weights}

### 2023-10-11 5.Conclusions/further.tex
No more weights constraints

@@
 Those solutions are purely theorical, they haven't been tested due to lack of time.
+
+\subsection{Removing weight constraints}\label{subsec:noCons}
+
+Removing the weights constraints is tricky since the the \ac{opAmp} is based on a verilog-A model (\cref{subsec:opamp}).
+Indeed, the model needs to be changed in order not to output voltages outside of the system's voltage range ($[0,v_{dd}]$).
+Once that is done, the \ac{opAmp} should behave more like a real one.
+The \acp{opAmp} should now be able to output a saturated signal.
+
+However those saturated levels need to be added to the training of the weights.
+This can be done in Keras by creating a custom layer in the circuit that keeps the values within a the defined range ($[-9,9]$ for this thesis).
+The training will thus happen with the values being saturated as it would happen in the circuit.
+
+This solution is not implemented in the circuit as the saturating signals might affect the real circuit.
+This is would allow the circuit to access a larger range of weights and make the training more performant.

### 2023-10-11 3.Design/dataset.tex
changed dataset
Reformulated

@@
 
-\subsection{The format}
-
-The values for the data needs to be set not to exceed 1. This is to make sure that the terminal voltage doesn't reach the threshold as explained in \cref{sec:xbarCircuit}.
+To make sure that the terminal voltage doesn't reach the threshold as explained in \cref{sec:xbarCircuit}.
+The voltage threshold ($V_{threshold}$) was chosen to be $V_{threshold}=0.1V$ because the memristors used in this thesis are purely theorical.
+The dataset needs to be formated not to exceed this value when converted to voltage (\cref{tab:valConv}), the dataset needs to be trained while not exceeding $1$.
 
@@
   \includesvg[width=\textwidth]{datasets/airline}
-  \caption{the airline dataset in a grphic. This is the curve that the regression is trying to reproduce. The vertical lines represent a full year.}
+  \caption{The airline dataset. The vertical lines represent a full year.}
   \label{graph:airline}

### 2023-10-12 4.Results/airline.tex
Removed duplicate error

@@
 The results are visually very close.
-The error is quantified using \ac{RMSE} error function.
+The error is quantified using \ac{RMSE} function.
 All those curves error to each other are measured and displayed in \cref{tab:airlineDigital}.
@@
   \end{tabular}
-  \caption{\ac{RMSE} errors of each curve to the others}
+  \caption{\acp{RMSE} of each curve to the others}
   \label{tab:airlineDigital}
@@
 The error between the digital curves with and without the custom activation functions are visually inexistant.
-The \ac{RMSE} error between those curves confirms this low error as it is of only $4.04$ thousands of passengers away.
+The \ac{RMSE} between those curves confirms this low error as it is of only $4.04$ thousands of passengers away.
 Considering the values at play, this error is negligible.
@@
 Those two curves are both very close to the target curve.
-They have a very similar error rate to this target curve, indeed they're \ac{RMSE} error is $45.24$ for the training using the original activation functions, and $45.94$ for the data created with the analog activation function.
+They have a very similar error rate to this target curve, indeed they're \ac{RMSE} is $45.24$ for the training using the original activation functions, and $45.94$ for the data created with the analog activation function.
 The difference of error between the two prediction is of about seven hundred passengers, which is very low considering the scale.

### 2023-10-12 4.Results/airline.tex
New figures and text in res

--- 2023-10-11 4.Results/airline.tex
@@
     %\rowcolor{gray}
-    \cellcolor[HTML]{808080}\acs{RMSE} & Original data & Digital output & Custom activation functions\\
+    \cellcolor[HTML]{808080}\acs{RMSE} & Target data & Digital predictions & \specialcell{Digital predictions with\\analog activation function}\\
     \hline
-    Original data &\cellcolor[HTML]{202020} & $45.24$ & $45.94$\\
+    Target data &\cellcolor[HTML]{202020} & $54.60$ & $52.27$\\
     \hline
-    Digital output & $45.24$ & \cellcolor[HTML]{202020} & $4.04$\\
+    Digital predictions  & $54.60$ & \cellcolor[HTML]{202020} & $78.64$\\
     \hline
-    Custom activation function & $45.94$ & $4.04$ & \cellcolor[HTML]{202020}\\
+    \specialcell{Digital predictions with\\analog activation function} & $52.27$ & $78.64$ & \cellcolor[HTML]{202020}\\
     \hline
   \end{tabular}
-  \caption{\ac{RMSE} errors of each curve to the others}
+  \caption{\acp{RMSE} of each curve to the others}
   \label{tab:airlineDigital}
@@
 
-The error between the digital curves with and without the custom activation functions are visually inexistant.
-The \ac{RMSE} error between those curves confirms this low error as it is of only $4.04$ thousands of passengers away.
-Considering the values at play, this error is negligible.
+Both predictions have similar \ac{RMSE} to the target curve.
+They have a \ac{RMSE} of $54.60$ for the \ac{LSTM} using the orignal activation functions (sigmoid and \ac{tanh} functions) and $52.27$ for the one using the analog activation functions (\cref{sec:af}).
+The error difference is so low (about two thousands passengers) that the difference may just come from the training and the initial values of the weights.
+They both approach the target curve pretty well.
+These error could be lowered by training for more epochs.
+This is not done because it diverges too much from the focus of the thesis.
 
-Those two
+The two predictions are very different from each other, the error rate between the two is of $78.64$, which is significantly higher than the \ac{RMSE} of the curves to the target curve.
 
-\subsection{Subsection B}
-\label{subsec:subbsectiona}
+Computing the \ac{NN} with or without the analog activation functions doesn't seem to affect the error of the predictions, experience has shown that the \ac{LSTM} using the original activation functions seem to predict higher curves than the target and the \ac{LSTM} using the analog activation functions tends to predict curves that are under the target, the reasons for such a difference are unknown.
+This is simply an observation after a few different runs.
+
+\subsection{Analog results trained without analog activation function}
+
+This subsection shows the importance of using the analog activation functions in the training phase of the \ac{NN}.
 
@@
   \centering
-  \includegraphics[width=0.5\linewidth]{cover}
-  \caption[Dummy Figure Caption for List of Figures.]{Dummy Figure Caption.}
-  \label{fig:dummyfigure1}
+  \includesvg[width=\linewidth]{results/airline/noCustomAF}
+  \caption{Analog predictions trained with the original activation functions.}
+  \label{graph:airlineAnalogNoC}
 \end{figure}
 
-Remember you can change the reference style. Another dummy citation \cite{site}.
+As expected, the analog predictions are very far off the target curve. Which, in this case, isn't the target dataset but the digital prediction ran with the default activation functions.
+
+\begin{figure}[H]
+  \centering
+  \includesvg[width=\linewidth]{results/airline/noCustomAFZoomed}
+  \caption{Analog predictions trained with the original activation functions.}
+  \label{graph:airlineAnalogNoCZoomed}
+\end{figure}

### 2023-10-13 4.Results/airline.tex
Finished first draft of w/o caf

@@
     \hline
-    %\rowcolor{gray}
-    \cellcolor[HTML]{808080}\acs{RMSE} & Target data & Digital predictions & \specialcell{Digital predictions with\\analog activation function}\\
+    \cellcolor[HTML]{808080}\acs{RMSE} & Target data & \specialcell{Digital prediction with\\default activation function} & \specialcell{Digital prediction with\\analog activation function}\\
     \hline
@@
     \hline
-    Digital predictions  & $54.60$ & \cellcolor[HTML]{202020} & $78.64$\\
+    \specialcell{Digital prediction with\\default activation function}  & $54.60$ & \cellcolor[HTML]{202020} & $78.64$\\
     \hline
-    \specialcell{Digital predictions with\\analog activation function} & $52.27$ & $78.64$ & \cellcolor[HTML]{202020}\\
+    \specialcell{Digital prediction with\\analog activation function} & $52.27$ & $78.64$ & \cellcolor[HTML]{202020}\\
     \hline
@@
 Both predictions have similar \ac{RMSE} to the target curve.
-They have a \ac{RMSE} of $54.60$ for the \ac{LSTM} using the orignal activation functions (sigmoid and \ac{tanh} functions) and $52.27$ for the one using the analog activation functions (\cref{sec:af}).
+They have a \ac{RMSE} of $54.60$ for the \ac{LSTM} using the default activation functions (sigmoid and \ac{tanh} functions) and $52.27$ for the one using the analog activation functions (\cref{sec:af}).
 The error difference is so low (about two thousands passengers) that the difference may just come from the training and the initial values of the weights.
@@
 
-Computing the \ac{NN} with or without the analog activation functions doesn't seem to affect the error of the predictions, experience has shown that the \ac{LSTM} using the original activation functions seem to predict higher curves than the target and the \ac{LSTM} using the analog activation functions tends to predict curves that are under the target, the reasons for such a difference are unknown.
+Computing the \ac{NN} with or without the analog activation functions doesn't seem to affect the error of the predictions, experience has shown that the \ac{LSTM} using the analog activation functions seem to predict higher curves than the target and the \ac{LSTM} using the default activation functions tends to predict curves that are under the target, the reasons for such a difference are unknown.
 This is simply an observation after a few different runs.
 
-\subsection{Analog results trained without analog activation function}
+\subsection{Analog results trained with default activation function}
 
@@
   \includesvg[width=\linewidth]{results/airline/noCustomAF}
-  \caption{Analog predictions trained with the original activation functions.}
+  \caption{Analog predictions trained with the default activation functions.}
   \label{graph:airlineAnalogNoC}
@@
 
-As expected, the analog predictions are very far off the target curve. Which, in this case, isn't the target dataset but the digital prediction ran with the default activation functions.
+As expected, the analog predictions are very far off the target curve.
+Which, in this case, isn't the target dataset but the digital prediction ran with the default activation functions.
+Indeed, the goal of the analog system is to reproduce as closely as possible the digital prediction.
+
+\begin{table}[H]
+  \centering
+  \begin{tabular}{|c|c|c|c|}
+    \hline
+    \cellcolor[HTML]{808080}\acs{RMSE} & Analog prediction ($n_s=1$) & Analog prediction ($n_s=2$) &Analog prediction ($n_s=4$) \\
+    \hline
+    \specialcell{Digital prediction with\\default activation function} & $140.68$ & $143.66$ & $139.71$\\
+    \hline
+  \end{tabular}
+  \caption{\acp{RMSE} of each analog prediction to their associated digital prediction}
+  \label{tab:airlineAnalogNoC}
+\end{table}
+
+As can be observed on \cref{graph:airlineAnalogNoC}, the analog predictions are not on par with what is expected at all.
+This is confirmed by the error rates found in \cref{tab:airlineAnalogNoC}, the error rates are not in an acceptable range at all.
+This error would be even higher to the target dataset as the digital prediction using the default activation functions is already under than the target curve.
+
+However, the prediction don't seem to be completely off. The curves seem to follow the shape of it's target, and by extension of the target data. This deserves a closer look.
 
@@
   \includesvg[width=\linewidth]{results/airline/noCustomAFZoomed}
-  \caption{Analog predictions trained with the original activation functions.}
+  \caption{Analog predictions trained with the default activation functions.}
   \label{graph:airlineAnalogNoCZoomed}
 \end{figure}
+
+As expected, in \cref{graph:airlineAnalogNoCZoomed} the trend of the curves is very similar to the original dataset (\cref{graph:airline}) and the predictions are very close to each others no matter the serial size used ($n_s$).
+
+\begin{table}[H]
+  \centering
+  \begin{tabular}{|c|c|c|c|}
+    \hline
+    \cellcolor[HTML]{808080}Analog prediction & $n_s=1$ & $n_s=2$ & $n_s=4$ \\
+    \hline
+    $n_s=1$ &\cellcolor[HTML]{202020} & $3.97$ & $5.05$\\
+    \hline
+    $n_s=2$  & $3.97$ & \cellcolor[HTML]{202020} & $4.78$\\
+    \hline
+    $n_s=4$ & $5.05$ & $4.78$ & \cellcolor[HTML]{202020}\\
+    \hline
+  \end{tabular}
+  \caption{\acp{RMSE} of each analog prediction to the others depending on the serial size ($n_s$)}
+  \label{tab:airlineAnalogNoCZoomed}
+\end{table}
+
+This similarity is further proven by the errors shown in \cref{tab:airlineAnalogNoCZoomed}.
+Those error are all around four thousands passengers, which is negligible in our case.
+That shows a certain stability of the system when using different values of serial size ($n_s$).
+The error rate is also quite low because of the lower results obtained.
+
+The predicted results generated by the analog system seem to be on a different scale while still following the overall aspect of the dataset.
+The down scale of those analog prediction would also mean that the errors of those predictions to each other is also scaled down.
+
+\subsection{Analog results trained with analog activation function}

### 2023-10-13 4.Results/airline.tex
Added text to res and updated figures

@@
 They are the results that the analog system will have to reproduce.
-Furthermore, this part is used to measure the impact of the custom activation function used for training (\cref{sec:af}).
+Furthermore, this part is used to measure the impact of the custom activation functions used for training (\cref{sec:af}).
 
@@
     \hline
-    \cellcolor[HTML]{808080}\acs{RMSE} & Target data & \specialcell{Digital prediction with\\default activation function} & \specialcell{Digital prediction with\\analog activation function}\\
+    \cellcolor[HTML]{808080}\acs{RMSE} & Target data & \specialcell{Digital prediction with\\default activation functions} & \specialcell{Digital prediction with\\analog activation functions}\\
     \hline
@@
     \hline
-    \specialcell{Digital prediction with\\default activation function}  & $54.60$ & \cellcolor[HTML]{202020} & $78.64$\\
+    \specialcell{Digital prediction with\\default activation functions}  & $54.60$ & \cellcolor[HTML]{202020} & $78.64$\\
     \hline
-    \specialcell{Digital prediction with\\analog activation function} & $52.27$ & $78.64$ & \cellcolor[HTML]{202020}\\
+    \specialcell{Digital prediction with\\analog activation functions} & $52.27$ & $78.64$ & \cellcolor[HTML]{202020}\\
     \hline
@@
 
-\subsection{Analog results trained with default activation function}
+\subsection{Analog results trained with default activation functions}\label{subsec:airlineAnalogNoC}
 
@@
     \hline
-    \cellcolor[HTML]{808080}\acs{RMSE} & Analog prediction ($n_s=1$) & Analog prediction ($n_s=2$) &Analog prediction ($n_s=4$) \\
+    \multirow{2}{*}{\acs{RMSE}} & \multicolumn{3}{|c|}{Analog prediction}\\
+    \cline{2-4}
+    & $n_s=1$ & $n_s=2$ & $n_s=4$ \\
     \hline
-    \specialcell{Digital prediction with\\default activation function} & $140.68$ & $143.66$ & $139.71$\\
+    \specialcell{Digital prediction with\\default activation functions} & $140.68$ & $143.66$ & $139.71$\\
     \hline
@@
   \centering
-  \begin{tabular}{|c|c|c|c|}
+  \begin{tabular}{|c|c|c|c|c|}
     \hline
-    \cellcolor[HTML]{808080}Analog prediction & $n_s=1$ & $n_s=2$ & $n_s=4$ \\
+    \multicolumn{2}{|c}{\multirow{2}{*}{\ac{RMSE}}} & \multicolumn{3}{|c|}{Analog prediction}\\
+    \cline{3-5}
+    \multicolumn{2}{|c}{} & \multicolumn{1}{|c|}{$n_s=1$} & $n_s=2$ & $n_s=4$ \\
     \hline
-    $n_s=1$ &\cellcolor[HTML]{202020} & $3.97$ & $5.05$\\
-    \hline
-    $n_s=2$  & $3.97$ & \cellcolor[HTML]{202020} & $4.78$\\
-    \hline
-    $n_s=4$ & $5.05$ & $4.78$ & \cellcolor[HTML]{202020}\\
+    \multirow{3}{*}{Analog prediction} & $n_s=1$ &\cellcolor[HTML]{202020} & $3.97$ & $5.05$\\
+    \cline{2-5}
+    & $n_s=2$  & $3.97$ & \cellcolor[HTML]{202020} & $4.78$\\
+    \cline{2-5}
+    & $n_s=4$ & $5.05$ & $4.78$ & \cellcolor[HTML]{202020}\\
     \hline
@@
 
-\subsection{Analog results trained with analog activation function}
+\subsection{Analog results trained with analog activation functions}
+
+This part will show the result of the inference when the weights were trained using the analog activation functions.
+
+\begin{figure}[H]
+  \centering
+  \includesvg[width=\linewidth]{results/airline/analog}
+  \caption{Analog predictions trained with the analog activation functions.}
+  \label{graph:airlineAnalog}
+\end{figure}
+
+As can be observed in \cref{graph:airlineAnalog}, the analog results are, in this case, way closer to the target values.
+The curves are very close to eahc other and are a definite improvement from the the ones trained using default activation functions.
+
+\begin{table}[H]
+  \centering
+  \begin{tabular}{|c|c|c|c|}
+    \hline
+    \multirow{2}{*}{\acs{RMSE}} & \multicolumn{3}{|c|}{Analog prediction}\\
+    \cline{2-4}
+    & $n_s=1$ & $n_s=2$ & $n_s=4$ \\
+    \hline
+    \specialcell{Digital prediction with\\analog activation functions} & $28.41$ & $92.57$ & $68.51$\\
+    \hline
+  \end{tabular}
+  \caption{\acp{RMSE} of each analog prediction to their associated digital prediction}
+  \label{tab:airlineAnalog}
+\end{table}
+
+\Cref{tab:airlineAnalog} contains the errors of the analog predictions.
+It shows that, as can be observed, the results the closest to the target are the ones generated using a serial size of one ($n_s=1$).
+Indeed, its error ($28.41$) is lower than the error from the digital results ($52.27$) to the original target values, almost half of it.
+The other ones are still very close to the targeted curve, but are still lower.
+This is probably due to the serialization circuit causing small differences in the signals at each steps.
+
+\begin{table}[H]
+  \centering
+  \begin{tabular}{|c|c|c|c|c|}
+    \hline
+    \multicolumn{2}{|c}{\multirow{2}{*}{\ac{RMSE}}} & \multicolumn{3}{|c|}{Analog prediction}\\
+    \cline{3-5}
+    \multicolumn{2}{|c}{} & \multicolumn{1}{|c|}{$n_s=1$} & $n_s=2$ & $n_s=4$ \\
+    \hline
+    \multirow{3}{*}{Analog prediction} & $n_s=1$ &\cellcolor[HTML]{202020} & $64.61$ & $40.63$\\
+    \cline{2-5}
+    & $n_s=2$  & $64.61$ & \cellcolor[HTML]{202020} & $25.75$\\
+    \cline{2-5}
+    & $n_s=4$ & $40.63$ & $25.75$ & \cellcolor[HTML]{202020}\\
+    \hline
+  \end{tabular}
+  \caption{\acp{RMSE} of each analog prediction to the others depending on the serial size ($n_s$)}
+  \label{tab:airlineAnalogError}
+\end{table}
+
+Once again, the results are quite close to each others, but as expected the errors (\cref{tab:airlineAnalogError}) are higher then the ones found in \cref{tab:airlineAnalogNoCZoomed}.
+The farthest appart ($n_s=1$ to $n_s=2$ with an error of $64.61$ thousands passengers) have an error barely higher than the error of digital results to the target dataset (\cref{tab:airlineDigital})
+
+The overall observation of those results compared to the previous ones (\cref{subsec:airlineAnalogNoC}) are how much closer to their target they are.
+It still isn't quite perfect but simply changing the activation functions used for training improved the results by a lot.
+It can thus be theorized that the error will become null if the system is trained with the same parameters as the ones it is using for inference.
+In other words, the error will probably be nullified when the training will happen inSitu (\cref{subsec:inSitu}).

### 2023-10-13 5.Conclusions/further.tex
Updated further

@@
 
-Of course the first thing that comes to mind to improve the system and making it actually usable for a real world application would be doing inSitu training.
-Training within the analog system would mean implementing the backpropagation algorithm.
+Of course the first thing that comes to mind to improve the system and making it actually usable for a real world application would be doing inSitu training, meaning that the weight training would also be happening on the chip.
+Training on the chip also means that the weights, hence the memristors in the circuit, need to be changed between each epoch.
+That requires to have a backpropagation algorithm.
 Here, two options are available :
@@
 
-Figure out how to change the values of the weights for the inSitu training. Different from \cref{sec:wei2res}.
+In order for the inSitu training to work, the weights also need to change.
+Future work also also include looking for ways to change the weights in the fastest possible way.
+Idealy the algorithm won't need to compute the weight to resistance conversion (\cref{sec:wei2res}) and directly know how the change the resistances' values.
 
@@
 
-\subsubsection{Only change the value of one of the two memristor}%TODO
+\subsubsection{Only change the value of one of the two memristor}
+
+This is a potential idea on how to improve the time it takes to compute a single epoch.
+Indeed changing the resistance value in a memristor, takes time depending on the type of memristor that is being used.
+If the weight to resistance is computed the same way as in the thesis (\cref{sec:wei2res}), so that each pair of resistance is centered around the middle point of the memristor's resistance range.
+
+Instead of having to do that, the idea is to only change one of the memristor's internal resistance so that the final weight is still the same.
+\Cref{eq:wei2res0} needs to be respected in any situation.
+The memristor's resistances won't be centered (\cref{eq:wei2res1}) anymore.
 
@@
 
-\subsection{\ac{LSTM} continuous mode}
-
-This would allow to have a real time results and not having to retake the same inputs several times. Basically it would be like %TODO Make graphs
-
-This might be something that already exist that I didn't find anything about.
+A potential way to implement that, would be in a two phase plan.
+First the prediction, and so on until a verification on the predictions can be done.
+After that the weights are updated based on those previous reviewed results.
 

### 2023-10-16 2.State/rnn.tex
added new chapters

\section{\acs{RNN}}\label{sec:rnn}

\acp{RNN} are, as the name suggests, a type of \ac{NN} using recurrent connections. They are a \ac{NN} with at least one cycle within the structure, where outputs of the previous run is used the next one.

### 2023-10-16 2.State/ai.tex
Added stuff to state art

\section{\acl{AI}}\label{sec:ai}

\acf{AI} is very popular, and is getting more attention, being used as selling point by lots of services, the word is thrown around so much that it lost its meaning. So what are \acp{AI} exactly.

\ac{AI} is defined as theories and stragies used to simulate human intelligence, in other words, a computer or software that can take a decision by itself based on previously set rules. The definition is very wide and have lots of ways to be implemented.

The most commonly used type of \ac{AI} is machine learning, which consists of teaching the software to make decision. It is defined as the use and development of computer systems that are able to learn and adapt without following explicit instructions, by using algorithms and statistical models to analyse and draw inferences from patterns in data.

\subsection{Machine learning}

In the last 10 years, the research concerning machine learning has considerably increased.

%TODO add superbe diagram

### 2023-10-16 2.State/nn.tex
Changed state
Added stuff to state art

@@
-\section{Neural Networks}\label{sec:nn}
+\section{\aclp{NN}}\label{sec:nn}
 
-\ac{AI} is wide topic, it refers to any computer system that can make a descision.
-It can take many forms but usually when we talk about \acp{AI}, we refer to the most common type, Machine Learning.
-Especially, a sub part of Machine Learning called Deep Learning.
-Deep Learning has seen a surge in research the last decade, which lead to great advancement in this field and the numerous \ac{AI} tools that popped up in the last years.
-Deep Learning works using what is called \acp{NN}.
-
-\acfp{NN} are a network of several layers of artificial neurons.
-Figure \ref{fig:snn}, shows a simple representation of a \ac{NN}, the artificial neurons are the represented by the colored circles.
-On figure \ref{fig:snn} each arrow represent a synapse.
+\acfp{NN} are a set of units known as neurons.
+Those neurons are linked to each other with arcs known as synapses, thoses synapses each have a weight associated to them.
+The set of neurons interconnected with their synapses is what is called a \acl{NN}.
+\Cref{fig:snn}, shows a simple representation of a \ac{NN}, the artificial neurons are the represented by the colored circles. On \cref{fig:snn} each arrow represent a synapse.
 
@@
 
-They are several layers to a \ac{NN} :
+\acp{NN} contains several layers :
+
 \begin{itemize}
   \item Input layer : This layer is simply the different inputs.
-  \item Hidden layer : This layer can be (and usually is) wider than the one in figure \ref{fig:snn}, it is there to add more layers and thus more precision to the result
+  \item Hidden layer : This layer can be (and usually is) wider than the one in figure \ref{fig:snn}.
+This is the layer that can be modified the most, by adding layers or increasing the amount of neurons in a layer.
   \item Output layer : This layer is where you can find the result from the \ac{NN}.
@@
 
-In a computational \ac{NN}, synapses are represented by weights.
-Those weights are to be multiplied by the last neuron and then added to each other to produce the next stage.
-Using the names defined in figure \ref{fig:snn}, the equation linking each layer is the following matrix multiplication :
-\begin{equation}
+The model presented in \cref{fig:snn} is modular and can be scaled up as much as required.
+It is generally accepted that the more neurons a \ac{NN} has, the best it can learn and execute the given task.
+
+The weights of the synapses have to be multiplied with the previous neuron and then added to each other to produce the next stage.
+Using the names defined in figure \ref{fig:snn}, the output is linked to the input by \cref{eq:nnHid,eq:nnOut}.
+First, the hidden layers' neurons need to be computed (\cref{eq:nnHid}).
+
+\begin{equation}\label{eq:nnHid}
   \begin{bmatrix}
@@
   \begin{bmatrix}
-    W_{1,1} & W_{1,2} & W_{1,3}\\
-    W_{2,1} & W_{2,2} & W_{2,3}\\
-    W_{3,1} & W_{3,2} & W_{3,3}\\
-    W_{4,1} & W_{4,2} & W_{4,3}\\
+    w_h_{1,1} & w_h_{1,2} & w_h_{1,3}\\
+    w_h_{2,1} & w_h_{2,2} & w_h_{2,3}\\
+    w_h_{3,1} & w_h_{3,2} & w_h_{3,3}\\
+    w_h_{4,1} & w_h_{4,2} & w_h_{4,3}\\
   \end{bmatrix}
@@
 
-And it works the same way to go to the next layer.
+Similarly the output is computed like in \cref{eq:nnOut}
 
-Those matrix multiplication are very power-hungry and their computation time scales up with the size of the \ac{NN} using classical computing.
+\begin{equation}\label{eq:nnOut}
+  \begin{bmatrix}
+    O_1\\ O_2
+  \end{bmatrix}
+  =
+  \begin{bmatrix}
+    w_o_{1,1} & w_o_{1,2} & w_o_{1,3} & w_o_{1,4}\\
+    w_o_{2,1} & w_o_{2,2} & w_o_{2,3} & w_o_{2,4}\\
+  \end{bmatrix}
+  \cdot
+  \begin{bmatrix}
+    H_1\\ H_2\\ H_3\\ H_4\\
+  \end{bmatrix}
+\end{equation}
+
+Those matrix multiplication are called \ac{VMM} because it is the result of the multiplication of a vector and a matrix, thus giving us another vector.
 

### 2023-10-16 2.State/nn.tex
Bunch of changes, going back to nn, rebooting

@@
 
-Analog computation enables those same calculations almost instantly and being far more power-efficient. This can be done using the physical properties of electrical devices.
-We want to reproduce one of those neural network using electrical circuits to perform the same computation that a computer would in much faster manner.
-Using \hyperref[subsec:memristors]{memristors} as weights. This removes the need to copy the weights from the main memory
+\subsection{Training weights}
+
+The weights values are obtained through training. This is done in a few steps :
+
+\begin{enumerate}
+  \item Generate random values for each weight.
+  \item Run the \ac{NN} to get an output vector.\label{step:restart}
+  \item Measure error rate using the chosen error algorithm.
+  \item Use the
+  \item go back to \cref{step:restart}
+\end{enumerate}

### 2023-10-16 2.State/rnn.tex
Bunch of changes, going back to nn, rebooting

@@
 They are a \ac{NN} with at least one cycle within the structure, where outputs of the previous run is used the next one.
+Those feedback connections are what differenciates it from feedforward \ac{NN}.
+
+This type of \ac{NN} is used when dealing with an unknown amount of inputs.
+Especially useful when treating timeseries \cite{rnn}.
+Example of \acp{RNN} uses are speech recognition, automatic language translation \cite{gru} and shape recognition, especially for handwriting recognition.
+
+They are trained the same way \ac{NN} are, using backpropagation to measure the error and try adjusting the weights.
+
+\subsection{Vanishing gradient problem}
+
+The Vanishing gradient problem is a problem that comes when dealing with time depedent data.
+
+\subsection{Simple \ac{RNN}}

### 2023-10-16 2.State/rnn.tex
Worked on rnns

@@
 This type of \ac{NN} is used when dealing with an unknown amount of inputs.
-Especially useful when treating timeseries \cite{rnn}.
+Especially useful when treating time series \cite{rnn}.
 Example of \acp{RNN} uses are speech recognition, automatic language translation \cite{gru} and shape recognition, especially for handwriting recognition.
 
-They are trained the same way \ac{NN} are, using backpropagation to measure the error and try adjusting the weights.
+They are trained the same way \ac{NN} are, measuring the error, backpropagte and adjust the weights accordingly.
 
@@
 
-The Vanishing gradient problem is a problem that comes when dealing with time depedent data.
+The Vanishing gradient problem is a problem that comes when dealing with time depedent data \cite{vanishGrad}.
+When long amount of time depedent data is fed to the \ac{RNN}, the older the data, the lower the impact on the weight change.
+This phenomenen makes the \ac{RNN} less effective when it comes to very long time series.
 

### 2023-10-16 figures/TimeDiagram/wavedromtikz.py
Removed use less files

@@ -1,877 +1,0 @@
-#!/usr/bin/env python
-
-import yaml
-import argparse
-import sys
-import os.path
-from string import Template
-from collections import namedtuple
-
-STANDALONE_TEX_HEADER = r"""
-\documentclass[crop,tikz]{standalone}
-\usepackage{tikz}
-\usepackage{ifthen}
-\usepackage{listings}
-
-\usetikzlibrary{ calc, positioning, decorations.markings, patterns}
-% all other packages and stuff you need for the picture
-%
-\begin{document}
-\begin{tikzpicture}[thick]
-"""
-
-STANDALONE_TEX_FOOTER = r"""
-\end{tikzpicture}
-\end{document}
-"""
-
-# A TikZ (LaTeX) header to be included before a waveform diagram. Defines
-# various primitives for drawing parts of a waveform.
-TIKZ_HEADER = Template(r"""
-% Seperation between lines
-\pgfmathsetlengthmacro{\wavesep}{2.0em}
-
-% Height of a waveform
-\pgfmathsetlengthmacro{\waveheight}{1.2em}
-
-% Width of a brick (half a cycle)
-\pgfmathsetlengthmacro{\wavewidth}{${size}em}
-
-% Width of the slant on slanted signal changes
-\pgfmathsetlengthmacro{\transitionwidth}{0.3em}
-
-% Width of the curve on slow signal changes (e.g. to z)
-\pgfmathsetlengthmacro{\curvedtransitionwidth}{1.0em}
-
-% Special signal styles
-\tikzset{wave x/.style={pattern=north east lines}}
-\tikzset{wave bus/.style={fill=white}}
-\tikzset{wave busyellow/.style={fill=yellow!25!white}}
-\tikzset{wave busorange/.style={fill=orange!25!white}}
-\tikzset{wave busblue/.style={fill=blue!25!white}}
-\tikzset{wave pulled/.style={dotted}}
-
-% Initialise pointer
-\coordinate (last waveform);
-
-% Label for a signal. Arguments:
-%  #1: The human-readable label string
-\newcommand{\signallabel}[1]{
-   \node [ anchor=east
-         , left=0 of last waveform
-         , minimum height=\waveheight
-         ]
-         {#1};
-}
-
-% Advance the "last brick" coordinate to the next position
-%  #1: Width of a brick
[…]

### 2023-10-16 2.State/rnn.tex
Done with vanishing gradient for now

@@
 
+\subsection{Simple \ac{RNN}}
+
 \subsection{Vanishing gradient problem}
@@
 The Vanishing gradient problem is a problem that comes when dealing with time depedent data \cite{vanishGrad}.
-When long amount of time depedent data is fed to the \ac{RNN}, the older the data, the lower the impact on the weight change.
-This phenomenen makes the \ac{RNN} less effective when it comes to very long time series.
+When big amount of time depedent data is fed to the \ac{RNN}, the weights can't be updated properly.
+The older the data, the lower it will impact how much the weight must change.
+Rendering the old input data almost useless.
+Classic \acp{RNN} must be used with relatively short time series.
 
-\subsection{Simple \ac{RNN}}
+Some \acp{RNN} were designed around this issue. This is the case of the \ac{LSTM} and \ac{GRU} which were created with internal mechanisms to regulate the flow of information and gradients.

### 2023-10-16 2.State/memristors.tex
worked on adding equations

@@
 Resistors, capacitors and inductors were already very established and well known component, so it was theorized that a fourth device should then exists to physically link flux ($\phi$) and charge ($q$).
-The flux in this case is not a magnetic flux and is defined as such : $ d\phi=V\cdot dt \Rightarrow \phi =  \int V(t) \,dt  $.\\
+The flux in this case is not a magnetic flux and is defined as such : $ d\phi=v\cdot dt \Rightarrow \phi =  \int v \,dt  $.\\
 The component stayed theoretical until 2008 when it's been implemented in a physical device for the first time \cite{memristorFab}. It took 37 years to have an actual working device.\\

### 2023-10-16 2.State/rnn.tex
worked on adding equations

@@
 
+The simple \ac{RNN} works just like a \ac{tanh} activated feedforward \ac{NN} with a feedback connection.
+
+\Cref{eq:srnn} shows the equation that the simple \ac{RNN} runs at every time step.
+
+\begin{equation}\label{eq:srnn}
+h_t=tanh([x_t,h_{t-1}]\cdot w + b)
+  \end{equation}
+
+  Where
+
   \subsection{Vanishing gradient problem}
@@
 Rendering the old input data almost useless.
-Classic \acp{RNN} must be used with relatively short time series.
+Simple \acp{RNN} must be used with relatively short time series.
 

### 2023-10-17 2.State/gru.tex
Tons of change, added GRU section

\section{\acf{GRU}}
The \ac{GRU} is another type of \ac{RNN}. It is also known to reduce the effect of the vanishing gradient problem. It was first introduced to improve translation techniques \cite{gru}.

The \ac{GRU} is very often compared to the \ac{LSTM}, it is sometimes assimilated as a type of \ac{LSTM} \cite{nbLSTM}. Their performance was found to be very similar in most situations, making those two types of \acp{RNN} coexistant in the modern machine learning world.

There are two versions of the \ac{GRU}, both are found on the internet, they are known as the encoder and decoder version \cite{gru}. They were originally designed to encode the message to translate and then decode in the translation. PyTorch only supports the decoder version \cite{gruPyTorch}, while the keras library supports both \cite{gruKeras} chosen by changing an argument.

\subsection{Encoder \ac{GRU}}

The encoder \ac{GRU} is the version of the \ac{GRU} most widely described on the internet. It contains an update gate (\cref{eq:updateG}), a reset gate (\cref{eq:resetG}), a candidate activation gate (\cref{eq:candActivG}). The hidden state is then computed (\cref{eq:gruHidG}) using the previous results.

\begin{equation}\label{eq:updateG}
\overrightarrow{z_t}=\sigma ([\overrightarrow{x_t},\overrightarrow{h_{t-1}}] \cdot w_z + \overrightarrow{b_z})
  \end{equation}
  \begin{equation}\label{eq:resetG}
\overrightarrow{r_t}=\sigma ([\overrightarrow{x_t},\overrightarrow{h_{t-1}}] \cdot w_r + \overrightarrow{b_r})
  \end{equation}
  \begin{equation}\label{eq:candActivG}
\overrightarrow{\hat{h_t}}=tanh(\overrightarrow{x_t}\cdot w_{hx}+(\overrightarrow{r_t}\odot\overrightarrow{h_{t-1}}) \cdot w_{hh} + \overrightarrow{b_h})
  \end{equation}
  \begin{equation}\label{eq:gruHidG}
  \overrightarrow{h_t}=(\overrightarrow{1}-\overrightarrow{z_t})\odot \overrightarrow{h_{t-1}} + \overrightarrow{z_t}\odot \overrightarrow{\hat{h_t}}
  \end{equation}

  Where ($w_z$,$\overrightarrow{b_z}$), ($w_r$,$\overrightarrow{b_r}$),($w_{hx}$,$w_{hh}$,$\overrightarrow{b_h}$) are the weights matrixes and bias vectors for the update, reset and candidate activation gates respectively.

  Some times the \cref{eq:gruHidG} can be found in another form (\cref{eq:gruHidG1}). This, however, has no impact on the final results because it is mathematically equivalent.

  \begin{equation}\label{eq:gruHidG1}
  \overrightarrow{h_t}=\overrightarrow{z_t}\odot \overrightarrow{h_{t-1}} + (\overrightarrow{1}-\overrightarrow{z_t})\odot \overrightarrow{\hat{h_t}}
  \end{equation}

  \subsection{Decoder \ac{GRU}}

  The decoder \ac{GRU}, while being less described, is the version used in pyTorch, which is getting very popular.

  The candidate activation gate (\cref{eq:candActivG1}) is the only difference from the encoder \ac{GRU}.

  \begin{equation}\label{eq:candActivG1}
\overrightarrow{\hat{h_t}}=tanh(\overrightarrow{x_t}\cdot w_{hx}+ \overrightarrow{b_{hx}}+\overrightarrow{r_t}\odot[\overrightarrow{h_{t-1}} \cdot w_{hh} + \overrightarrow{b_{hh}}])
  \end{equation}

  \subsection{Similarities with \ac{LSTM}}

  The \ac{CIFG} \ac{LSTM} or more commonly known as \ac{GRU}. \cite{gru,gruKeras,gruPyTorch} As it's name implies, \acl{CIFG}, links the input and forget gates as such, $f_t=1-i_t$. Other differences are :
  \begin{itemize}
  \item No peepholes and output activation function.
  \item Combined hidden state and cell state.
  \item The candidate cell state is point wise multiplied with the output gate before the activation function.
  \end{itemize}

  This is clearer with the math equations :
  \begin{equation}\label{eq:cellGGRU}
\tilde{c}_t = tanh(w_{cx}\cdot x_t + w_{ch}\cdot(o_t \odot h_{t-1}) + b_c)
  \end{equation}
  \begin{equation}\label{eq:hiddenGRU}
  h_t=(1-i_t)\cdot h_{t-1} + i_t \cdot \tilde{c}_t
  \end{equation}

### 2023-10-17 2.State/lstm.tex
added new chapters
Bunch of changes, going back to nn, rebooting
Tons of change, added GRU section

@@
-\section{LSTM}\label{sec:lstm}
-\acp{LSTM} are a type of \ac{NN} used to analyze sequence of data.
+\section{\acs{LSTM}}\label{sec:lstm}
+\acp{LSTM} are a type of \ac{RNN} used to analyze sequence of data.
 They are capable of predict data based of the previous results.
@@
 The first \ac{LSTM} was invented in 1997 by Hochreiter and Schmidhuber \cite{firstLSTM}. \acp{LSTM} changed a lot through the years to become what they are now.
-\acp{LSTM} fixed the vanishing/exploding gradient problem.
-Traditionnal \ac{RNN} have the ability to model sequential events by propagating through time, for example forward and backward propagation.
-This is achieved by connecting these sequential events with the hidden state:
-$$h_t=f(h_{t-1},x_t)$$
-The hidden state $h_t$ carries all the past informations for the next time step.
-Even though \acp{RNN} work very well, they are subject to the vanishing/exploding gradient problem. This issue occurs when too many time steps are chained.
 
+\acp{LSTM} was created to fix the vanishing gradient problem.
 \acp{LSTM} alleviate this issue by adding a cell state, this state gives it the ability to choose what input is important and which one isn't, thus giving it a long term memory.
@@
 Let's start with the more classic gate equations. They are the ones that behave like the more classic \ac{NN}.
-The input (\cref{eq:inputG}), forget (\cref{eq:forgetG}) and ouput (\cref{eq:ouputG}) gates are described below.
+The input (\cref{eq:inputG}), forget (\cref{eq:forgetG}) and ouput (\cref{eq:outputG}) gates are described below.
 
 \begin{equation}\label{eq:inputG}
-  i_t=\sigma (w_i\cdot[x_{t_1},h_{t-1}] + b_i)
+  \overrightarrow{i_t}=\sigma ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_i + \overrightarrow{b_i})
 \end{equation}
 \begin{equation}\label{eq:forgetG}
-  f_t=\sigma (w_f\cdot[x_{t_1},h_{t-1}] + b_f)
+  \overrightarrow{f_t}=\sigma ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_f + \overrightarrow{b_f})
 \end{equation}
-\begin{equation}\label{eq:ouputG}
-  o_t=\sigma (w_o\cdot[x_{t_1},h_{t-1}] + b_o)
+\begin{equation}\label{eq:outputG}
+  \overrightarrow{o_t}=\sigma ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_o + \overrightarrow{b_o})
 \end{equation}
 
-Where ($w_i$,$b_i$), ($w_f$,$b_f$) and ($w_o$,$b_o$) are the pair of weights and biases for the input, forget and ouput gates respectively.
-$x_t$ is the input vector and $h_t$ is the hidden state vector.
+Where ($w_i$,$\overrightarrow{b_i}$), ($w_f$,$\overrightarrow{b_f}$) and ($w_o$,$\overrightarrow{b_o}$) are the pair of weights matrixes and biases vectors for the input, forget and ouput gates respectively.
+$\overrightarrow{x_t}$ is the input vector and $\overrightarrow{h_t}$ is the hidden state vector.
 
@@
 \begin{equation}\label{eq:candCell}
-  \tilde{c}_t=tanh(w_c\cdot[x_{t-1},h_{t-1}] + b_c)
+  \overrightarrow{\tilde{c}_t}=tanh([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_c+ \overrightarrow{b_c})
 \end{equation}
 \begin{equation}\label{eq:cellS}
-  c_t=f_t\cdot c_{t-1} + i_t \cdot \tilde{c}_t
+  \overrightarrow{c_t}=\overrightarrow{f_t}\odot \overrightarrow{c_{t-1}} + \overrightarrow{i_t} \odot \overrightarrow{\tilde{c}_t}
 \end{equation}
 
-Where $w_c$ and $b_c$ are the weights and bias for the candidate cell state.
+Where $w_c$ and $\overrightarrow{b_c}$ are the weights matrix and bias vector for the candidate cell state.
 
@@
 \begin{equation}\label{eq:hiddenS}
-  h_t=o_t\cdot tanh(c_t)
+  \overrightarrow{h_t}=\overrightarrow{o_t}\odot tanh(\overrightarrow{c_t})
 \end{equation}
 
-We set $x_0$ as the first input and define $h_0$ as a zero only vector.
+We set $x_1$ as the first input and define $\overrightarrow{h_0}$ as a zero only vector.
 
@@
   \includesvg[width=\textwidth]{lstm/lstmUse}
-  \caption{LSTM use case, legend in \cref{leg:lstm}}
+  \caption{Unfolded \acs{LSTM}, legend in \cref{leg:lstm}}
   \label{fig:lstmUse}
@@
 \begin{equation}\label{eq:inputGVanilla}
-  i_t=\sigma (w_i\cdot[x_{t_1},h_{t-1}] + b_i+c_{t-1}\odot p_f)
+  \overrightarrow{i_t}=\sigma ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_i + \overrightarrow{b_i} +\overrightarrow{c_{t-1}}\odot \overrightarrow{p_f})
 \end{equation}
 \begin{equation}\label{eq:forgetGVanilla}
-  f_t=\sigma (w_f\cdot[x_{t_1},h_{t-1}] + b_f+c_{t-1}\odot p_i)
+  \overrightarrow{f_t}=\sigma ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_f + \overrightarrow{b_f}+\overrightarrow{c_{t-1}}\odot \overrightarrow{p_i})
 \end{equation}
 \begin{equation}\label{eq:ouputGVanilla}
-  o_t=\sigma (w_o\cdot[x_{t_1},h_{t-1}] + b_o+c_{t}\odot p_o)
+  \overrightarrow{o_t}=\sigma ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_o + \overrightarrow{b_o}+\overrightarrow{c_{t}}\odot \overrightarrow{p_o})
 \end{equation}
-With $p_f$, $p_i$ and $p_o$ being the peephole weights vectors. Their size is the same as the size of the hidden state vector ($h_t$). $\odot$ is the pointwise multiplication operator.
+
+With $\overrightarrow{p_f}$, $\overrightarrow{p_i}$ and $\overrightarrow{p_o}$ being the peephole weights vectors.
+Their size is the same as the size of the hidden state vector ($\overrightarrow{h_t}$).
+$\odot$ is the pointwise multiplication operator.
 
 The equations that haven't been rewritten simply stay the same.
-
-\subsubsection{\acf{CIFG} \ac{LSTM} or \acf{GRU}}
-The \ac{CIFG} \ac{LSTM} or more commonly known as \ac{GRU}.
-\cite{gru,gruKeras,gruPyTorch} As it's name implies, \acl{CIFG}, links the input and forget gates as such, $f_t=1-i_t$.
-Other differences are :
-\begin{itemize}
-  \item No peepholes and output activation function.
-  \item Combined hidden state and cell state.
-  \item The candidate cell state is point wise multiplied with the output gate before the activation function.
-\end{itemize}
-
-This is clearer with the math equations :
-\begin{equation}\label{eq:cellGGRU}
-  \tilde{c}_t = tanh(w_{cx}\cdot x_t + w_{ch}\cdot(o_t \odot h_{t-1}) + b_c)
-\end{equation}
-\begin{equation}\label{eq:hiddenGRU}
-  h_t=(1-i_t)\cdot h_{t-1} + i_t \cdot \tilde{c}_t
-\end{equation}
 
@@
 The modified equations are :
+
 \begin{equation}\label{eq:inputGFGR}
-  i_t=\sigma (w_i\cdot[x_{t_1},h_{t-1}]+ R_{ii}\cdot i_{t-1} + R_{if}\cdot f_{t-1} + R{io}\cdot o_{t-1} + b_i+c_{t-1}\odot p_f)
+  \overrightarrow{i_t}=\sigma ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_i  + \overrightarrow{i_{t-1}}\cdot R_{ii} + \overrightarrow{f_{t-1}}\cdot R_{if} + \overrightarrow{o_{t-1}}\cdot R_{io} + \overrightarrow{b_i} +\overrightarrow{c_{t-1}}\odot \overrightarrow{p_i})
 \end{equation}
 \begin{equation}\label{eq:forgetGFGR}
-  f_t=\sigma (w_f\cdot[x_{t_1},h_{t-1}]+ R_{fi}\cdot i_{t-1} + R_{ff}\cdot f_{t-1} + R{fo}\cdot o_{t-1} + b_f+c_{t-1}\odot p_i)
+  \overrightarrow{f_t}=\sigma ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_f  + \overrightarrow{i_{t-1}}\cdot R_{fi} + \overrightarrow{f_{t-1}}\cdot R_{ff} + \overrightarrow{o_{t-1}}\cdot R_{fo} + \overrightarrow{b_f} +\overrightarrow{c_{t-1}}\odot \overrightarrow{p_f})
 \end{equation}
 \begin{equation}\label{eq:ouputGFGR}
-  o_t=\sigma (w_o\cdot[x_{t_1},h_{t-1}]+ R_{oi}\cdot i_{t-1} + R_{of}\cdot f_{t-1} + R{oo}\cdot o_{t-1} + b_o+c_{t-1}\odot p_o)
+  \overrightarrow{o_t}=\sigma ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_o + \overrightarrow{i_{t-1}}\cdot R_{oi} + \overrightarrow{f_{t-1}}\cdot R_{of} + \overrightarrow{o_{t-1}}\cdot R_{oo} + \overrightarrow{b_o} +\overrightarrow{c_{t-1}}\odot \overrightarrow{p_o})
 \end{equation}
+
 With $R_{fi}$ being the weight matrix of the input feedback connection to compute the forget gate, and the same goes for the other weight matrices.

### 2023-10-17 2.State/rnn.tex
Tons of change, added GRU section

@@
 
+Traditionnal \ac{RNN} have the ability to model sequential events by propagating through time, for example forward and backward propagation.
+This is achieved by connecting these sequential events with the hidden state like in \cref{eq:rnn}.
+
+\begin{equation}\label{eq:rnn}
+  \overrightarrow{h_t}=f(\overrightarrow{x_t},\overrightarrow{h_{t-1}})
+\end{equation}
+
+The hidden state ($\overrightarrow{h_t}$) carries all the past informations for the next time step. It also serves as the output of the \ac{RNN}.
+
 They are trained the same way \ac{NN} are, measuring the error, backpropagte and adjust the weights accordingly.
@@
 \begin{equation}\label{eq:srnn}
-h_t=tanh([x_t,h_{t-1}]\cdot w + b)
+  \overrightarrow{h_t}=tanh([\overrightarrow{x_t},\overrightarrow{h_{t-1}}]\cdot w + b)
 \end{equation}
 
-  Where
+Where $t\in\mathbb{N}^*$ is the time index, $w$ the weight matrix, $b$ the bias matrix, $\overrightarrow{x_t}$ the input vector and $\overrightarrow{h_{t}}$ the hidden state of the \ac{RNN}.
 

### 2023-10-17 2.State/gru.tex
changed gru full and no full forms
Change to GRU

@@
-\section{\acf{GRU}}
-The \ac{GRU} is another type of \ac{RNN}. It is also known to reduce the effect of the vanishing gradient problem. It was first introduced to improve translation techniques \cite{gru}.
+\section{\acs{GRU}}
+The \acf{GRU} is another type of \ac{RNN}. It is also known to reduce the effect of the vanishing gradient problem. It was first introduced to improve translation techniques \cite{gru}.
 
@@
 
-  Some times the \cref{eq:gruHidG} can be found in another form (\cref{eq:gruHidG1}). This, however, has no impact on the final results because it is mathematically equivalent.
+Sometimes the \cref{eq:gruHidG} can be found in another form (\cref{eq:gruHidG1})\cite{gruPyTorch}.
+This, however, has no impact on the final results, it means the weights are going to be trained differently for the update gate.
 

### 2023-10-17 1.Introduction/motivation.tex
added more motivattion

@@
 \section{Motivation}
-\label{sec:int_motivation}
+\label{sec:motivation}
 
 Nowadays, it's impossible not to have heard about \acp{AI}, they are everywhere and everybody talks about them.
-The most recent example being \textit{ChatGPT} from \textbf{OpenAI} and is getting very popular even among casual computer users.
-Everybody is using those tools.
-It's clear that \ac{AI} is becoming more and more important in the current world.
+It is common practice to advertise products using \ac{AI} powered tools.
+\textit{ChatGPT} from \textbf{OpenAI} is a recent example and is getting very popular even among casual computer users.
+Everyone is using this tool.
+An older example include home assistants that have been around for about ten years.
+Those products (such as \textbf{Amazon}'s \textit{Alexa}, \textbf{Google}'s \textit{nest} system and others) are not computing their answers locally, hence creating avoidable internet trafic.
+It's clear that \ac{AI} is becoming the predominant technology in the modern world and future innovations will depend on it.
 
@@
 Furthermore, the time an \ac{AI} takes to compute its output quickly rises, especially on lower end chip such as the ones used in embedded systems.
-So much so that the most complex ones only run on online and very powerful servers, like the aforementioned \textit{ChatGPT}.
+So much so that the most complex ones only run on online and very powerful servers, like the aforementioned \textit{ChatGPT} and home assistants.
 
@@
 Using an \ac{FPGA} or \ac{ASIC} are other ways to improve power consumption and execution time.
-The latter is the most restricting (as it's name implies) but has the best results.
+The latter is the most restricting, as it's name implies, but has the best results.
 
-Using an \ac{ASIC} allows the use of analog computation. Analog computation offers great advantages compared to digital computers as it's much faster while being very power-efficient.
+Using an \ac{ASIC} allows the use of analog computation. Analog computation offers great advantages compared to digital computers as it's much faster while being known to be very power-efficient.
 
-Making an analog chip with a very low power consumption allows to use chip like that in embedded systems such as video surveillance cameras.
+Home assistants could greatly benefit from being able to quickly compute answers without having to connect to an online server.
+Locally computing answers would drastically improve the usefulness of the device.
+It could even allow to be used on the go, in a car for example.
+
+Another great use for local \ac{AI} computation would be to use \acp{LLM}, the technology behind \textit{ChatGPT}, for any local use.
+\acp{LLM} are not the focus of the thesis, but are a long term potential objective of using analog computation for \ac{AI}.
+
+Making an analog chip with a very low power consumption also allows to use chip like that in embedded systems such as video surveillance cameras.
 A small chip could be installed inside the system that could be able to do gait detection \cite{gaitDS,gaitDig,gait}.
@@
 
-Somre research \cite{celegans}, focuses on reproducing parts of the nervous system of a small organism.
+Some research \cite{celegans}, focuses on reproducing parts of the nervous system of a small organism.
 This could be applied on larger organisms, or, in the long term, reproduce part of the human brain.
@@
 The focus of this thesis is to create a working simulation of an analog \ac{LSTM} circuit using Cadence's virtuoso as simulation software.
-Having those preliminary results would allow for fabricating and testing such a chip.
+Having those preliminary results would allow to fabricate and test such a chip.
 Once a working prototype is manufactured, such a circuit could highly improve embedded \ac{AI} use.

### 2023-10-17 3.Design/activation.tex
added more things in appendix

@@
 It is the reason why making good Activation function circuits is fundamental.
-Of course the easy way would be to simply design a hard sigmoid (TODO : Ref annex), the issue being that the hard sigmoid is much worse than the regular sigmoid, especially for regression problems \cite{hardSigm}.
+Of course the easy way would be to simply design a hard sigmoid (\cref{apsec:hardFunc}), the issue being that the hard sigmoid is much worse than the regular sigmoid, especially for regression problems \cite{hardSigm}.
 The same goes for the \ac{tanh} and hard \ac{tanh} functions.

### 2023-10-17 Appendices/appendixA.tex
New wei2res
Appendix wei2res
added more things in appendix

@@
 
-A non-inverting amplifier was producing too much difference from the expected result (about a 20\% error) %TODO maybe show something
+\section{Solving weight to resistance}\label{apsec:wei2res}
+
+\begin{equation}\label{eq:wei2res3}
+  \begin{cases}
+    w=\frac{R_f}{R_+}-\frac{R_f}{2\cdot R_f -R_+}\\
+    R_-=2\cdot R_f -R_+
+  \end{cases}
+\end{equation}
+
+\begin{equation}\label{eq:wei2res4}
+  \begin{cases}
+    w=\frac{2\cdot R_f^2-2\cdot R_+\cdot R_f}{2\cdot R_f\cdot R_+ -R_+^2}\\
+    R_-=2\cdot R_f -R_+
+  \end{cases}
+\end{equation}
+
+\begin{equation}\label{eq:wei2res5}
+  \begin{cases}
+    2\cdot R_f\cdot R_+\cdot w -R_+^2\cdot w=2\cdot R_f^2-2\cdot R_+\cdot R_f\\
+    R_-=2\cdot R_f -R_+
+  \end{cases}
+\end{equation}
+
+\begin{equation}\label{eq:wei2res6}
+  \begin{cases}
+    R_+^2\cdot w - R_+\cdot 2\cdot R_f \cdot(w+1) + 2R_f^2 = 0\\
+    R_-=2\cdot R_f -R_+
+  \end{cases}
+\end{equation}
+
+\begin{equation}\label{eq:wei2res61}
+  R_+^2\cdot w - R_+\cdot 2\cdot R_f \cdot(w+1) + 2R_f^2 = 0
+\end{equation}
+
+We must now solve \cref{eq:wei2res61}.
+
+\begin{equation}\label{eq:wei2res61}
+  \Delta = 4\cdot R_f^2\cdot(w+1)^2-8\cdot w\cdot R_f^2 = 4\cdot R_f^2\cdot(w^2+1) \ge 0
+\end{equation}
+
+Giving the two solutions :
+
+\begin{equation}\label{eq:wei2res62}
+  R_{+,0}=\frac{R_f\cdot(w+1)-R_f\cdot\sqrt{w^2+1}}{w}
+\end{equation}
+\begin{equation}\label{eq:wei2res63}
+  R_{+,1}=\frac{R_f\cdot(w+1)+R_f\cdot\sqrt{w^2+1}}{w}
+\end{equation}
+
+\section{Hard sigmoid and \ac{tanh} functions}\label{apsec:hardFunc}
+
+The hard sigmoid and \ac{tanh} functions are simplifications of the the classic sigmoid and \ac{tanh} used for faster computation.
+They come with a downside, they are less efficient then their original functions.
+
+They are defined as \cref{eq:hSigmoid} for the hard sigmoid and \cref{eq:hTanh} for the hard \ac{tanh} \cite{hSigmoid, hTanh}.
+
+\begin{equation}\label{eq:hSigmoid}
+  \sigma (x) =
+  \begin{cases}
+    0, \hfill x< -2.5\\
+    \frac{x}{5}+\frac{1}{2}, \hspace{0.2cm} 2.5>x>-2.5\\
+    1, \hfill x> 2.5\\
+  \end{cases}
+\end{equation}
+\begin{equation}\label{eq:hTanh}
+  tanh(x) =
+  \begin{cases}
+    -1, \hfill x<-1\\
+    x, \hspace{0.2cm} 1>x>-1\\
+    1, \hfill x>1\\
+  \end{cases}
+\end{equation}

### 2023-10-17 2.State/main.tex
Changed state orga
added new chapters
Tons of change, added GRU section
New figures + debut main

@@
 
-\input{2.State/1.memristors.tex}
-\input{2.State/2.crossbar.tex}
-\input{2.State/3.model.tex}
-\input{2.State/4.lstm.tex}
+This chapter will go over the several subject that need to be known in order to understand the thesis.
+
+Math is the best way to universally describe all scientific models. For this reason, this chapter, and the thesis as whole, will contains equations.
+
+It seems only logical to define all the tools that are going to be used in the thesis.
+
+\begin{itemize}
+\item Vectors are displayed as $\overrightarrow{v}$ and are considered to be vertical matrixes when used in an equation.
+\item $\overrightarrow{1}$ represent a vector that only contains ones of variable size.
+\item $\odot$ is used for the Hadamart product, that is the point wise multiplication of each element of the matrixes
+\item $\cdot$ is used in two cases :
+\begin{itemize}
+\item When used with a number, it is the product of those numbers.
+\item When used with matrixes, it is the matrix product
+\end{itemize}
+\end{itemize}
+
+\input{2.State/ai.tex}
+\input{2.State/nn.tex}
+\input{2.State/rnn.tex}
+\input{2.State/lstm.tex}
+\input{2.State/gru.tex}
+\input{2.State/memristors.tex}
+\input{2.State/crossbar.tex}
+\input{2.State/model.tex}
 

### 2023-10-17 2.State/main.tex
finshed preview

@@
 \begin{itemize}
-\item Vectors are displayed as $\overrightarrow{v}$ and are considered to be vertical matrixes when used in an equation.
+  \item Vectors are displayed as $\overrightarrow{v}$ and are considered to be vertical matrixes, matrixes of size ($1\times n$) for vector of size $n$.
   \item $\overrightarrow{1}$ represent a vector that only contains ones of variable size.
@@
     \begin{itemize}
-\item When used with a number, it is the product of those numbers.
+      \item When used with scalars, it is the product of those scalars.
       \item When used with matrixes, it is the matrix product
+    \end{itemize}
+  \item $+$ is also used in two cases :
+    \begin{itemize}
+      \item When used with scalars, it is the sum of those scalars.
+      \item When used with vectors (column matrixes), it is the point wise sum, basically adding each element to each other. The result is a vector of the same size.
     \end{itemize}

### 2023-10-17 Appendices/appendixA.tex
Return to line
Added figure to pdf

@@
 \end{equation}
+
 \begin{equation}\label{eq:hTanh}
@@
 \end{equation}
+
+Some versions of the hard sigmoid are a bit different from \cref{eq:hSigmoid}, the slope depends on the library used \cite{hSigmoid1}.
+
+\begin{figure}[H]
+  \centering
+  \includesvg[width=\textwidth]{activation/hardFunc}
+  \caption{Hard sigmoid and hard \acs{tanh} with their original function}
+  \label{graph:hardFunc}
+\end{figure}
+
+A visual representation helps seeing how close those functions are, \cref{graph:hardFunc} displays the two together with the function they are modeled after.

### 2023-10-17 2.State/ai.tex
Saved figure
Changed desciption
Added one sentence
Bigger figure
Correction

@@
 
-\acf{AI} is very popular, and is getting more attention, being used as selling point by lots of services, the word is thrown around so much that it lost its meaning. So what are \acp{AI} exactly.
+\acf{AI} is very popular, and is getting more attention, being used as selling point by lots of services, the word is thrown around so much that it has lost its meaning. So what are \acp{AI} exactly.
 
@@
 
-The most commonly used type of \ac{AI} is machine learning, which consists of teaching the software to make decision.
+The most promising type of \ac{AI} is machine learning, which consists of teaching the software to make decision.
 It is defined as the use and development of computer systems that are able to learn and adapt without following explicit instructions, by using algorithms and statistical models to analyse and draw inferences from patterns in data.
@@
 
-%TODO add superbe diagram
+\begin{figure}[H]
+  \centering
+  \includesvg[width=\textwidth]{ml.svg}
+  \caption{Machine learning and its subsets}
+  \label{fig:ml}
+\end{figure}
+
+\Cref{fig:ml} is a visual representation of machine learning and all the disciplines of machine learning. The thesis will focus, as the title implies, on the \aclp{RNN}.

### 2023-10-17 2.State/lstm.tex
Full lstm
Bew figgures + updates

@@
 \section{\acs{LSTM}}\label{sec:lstm}
-\acp{LSTM} are a type of \ac{RNN} used to analyze sequence of data.
+\acfp{LSTM} are a type of \ac{RNN} used to analyze sequence of data.
 They are capable of predict data based of the previous results.
@@
   \centering
-  \begin{minipage}{\columnwidth}
-    \subfloat[\acs{LSTM} cell\label{fig:lstmCell}]{\includesvg[width=\textwidth,pretex=\large]{lstm/lstmCell.svg}}
-  \end{minipage}
-  \begin{minipage}{\columnwidth}
-    \subfloat[Legend\label{leg:lstm}]{\includesvg[width=\textwidth,pretex=\large]{lstm/lstmLegend.svg}}
-  \end{minipage}
-  \caption{}
+  \includesvg[width=\textwidth,pretex=\large]{lstm/lstmCell.svg}
+  \label{fig:lstmCell}
+  \caption{\acs{LSTM} cell}
 \end{figure}

### 2023-10-18 2.State/model.tex
Updated figures + started working on memristors mordel

@@
 \section{Memristor's model}\label{sec:model}
-They are several ways to model a memristor (TODO : cite).
+
+They are several ways to model a memristor \cite{memristorFab,memTEAMmodel, memVTEAMmodel, memristorSpiceModels}.
+The way the memristor works depends on the materials used to fabricate the memristor.
+Each type of memristor has slightly different behavior.
+Models were thus created to simulate, as accurately as possible, the memristor in electrical circuit simulation software such as \textbf{Cadence}'s \textit{virtuoso} or SPICE.
+
+Some examples of memristor's model are :
+
+\begin{itemize}
+\item \acf{LIDM}, the model of the first memristor \cite{memristorFab}.
+\item \acf{STBM}, is another model of the $TiO_x$ memristor created in 2008 by \textbf{HP} \cite{memristorFab, memristorSpiceModels}.
+\item \acf{TEAM}, is a model that can easily adapt to several types of memristive devices while focusing on fast computation \cite{memTEAMmodel}.
+\item \acf{VTEAM}, is a later improvement of the \ac{TEAM}. The internal resistance depends on voltage, unlike \ac{TEAM} which depends on current \cite{memVTEAMmodel}.
+\end{itemize}
+
+The \ac{VTEAM} being the most versatile model, its implementation would fit most memristor types.
+
+It's been established that the internal resistance also depends on an internal state $x$ (\cref{eq:memristiveDev}).
+
+\begin{equation}
+\frac{dx}{dt}=f(x,v)
+  \end{equation}
+
+  The model is based on previous work \cite{memCadenceModel}. It defines sevral equations for the model such as \cref{}.
+
+  \begin{equation}
+  i(R,v)=
+  \begin{cases}
+  \frac{a_p}{R}\cdot sinh(b_p\cdot v), \hspace{10px} v\ge 0\\
+    \frac{a_n}{R}\cdot sinh(b_n\cdot v), \hspace{10px} v<0\\
+    \end{cases}
+    \end{equation}
+
+    \subsection{\ac{VTEAM} in \textbf{Cadence}'s \textit{virtuoso}}
 
     One of them, works by applying voltage pulse to the device. The device then change it's internal resistance based on how far from the stable point (usually the initial resistance) it is.
-TODO
+
+    %\subsection{\acf{LIDM}}
+
+    %The \ac{LIDM} is the model of the first memristor eveer created, it was created to fit the behavior of the $TiO_x$ memristor they had just fabricated \cite{memristorFab}.
+The device, of width $D$, has two parts.
+The doped part and the undoped

### 2023-10-18 2.State/model.tex
model almost done, missing behavior

@@
 
+\subsection{Equations}
+
 It's been established that the internal resistance also depends on an internal state $x$ (\cref{eq:memristiveDev}).
@@
 
-  The model is based on previous work \cite{memCadenceModel}. It defines sevral equations for the model such as \cref{}.
+Where $v$ is the voltage, $t$ is time and $f$ is a function.
 
-  \begin{equation}
+The function $f$ is described in its simplest form as in \cref{eq:funcDesc}.
+
+\begin{equation}\label{eq:funcDesc}
+  f(x,v)=g(x)\cdot s(v)
+\end{equation}
+
+Where $g$ is the window function and $s$ the sensitivity function.
+
+\begin{itemize}
+  \item The window function delimits the operational boundaries of the memristor.
+  \item The sensitivity function, also known as voltage sensitivity describes the effect of voltage on the internal state's variations.
+\end{itemize}
+
+The model is based on previous work \cite{memCadenceModel} that aims to reproduce as many memristive device types.
+They have chosen to use the resistive state ($R$) as the internal state \cite{memModelOrigin}.
+Based on this assumption the \cref{eq:memModel0,eq:memModel1} are extracted from the memristive device.
+This model is of course very basic and ignores some physical depedencies such as temperature.
+
+\begin{equation}\label{eq:memModel0}
   i(R,v)=
   \begin{cases}
-  \frac{a_p}{R}\cdot sinh(b_p\cdot v), \hspace{10px} v\ge 0\\
-    \frac{a_n}{R}\cdot sinh(b_n\cdot v), \hspace{10px} v<0\\
+    \frac{a_p}{R}\cdot sinh(b_p\cdot v) & v\ge 0\\
+    \frac{a_n}{R}\cdot sinh(b_n\cdot v) & v<0\\
   \end{cases}
@@
 
-    \subsection{\ac{VTEAM} in \textbf{Cadence}'s \textit{virtuoso}}
+\begin{equation}\label{eq:memModel1}
+  \frac{dR}{dt}=f(R,v)=s(v)\cdot g(R,v)
+\end{equation}
 
-    One of them, works by applying voltage pulse to the device. The device then change it's internal resistance based on how far from the stable point (usually the initial resistance) it is.
+The sensitivity function was chosen to be a voltage-dependent exponential function, like described in \cref{eq:memModel2}.
+
+\begin{equation}\label{eq:memModel2}
+  s(v)=
+  \begin{cases}
+    A_p\cdot (e^{t_p\cdot |v|}-1)& v>0\\
+    A_n\cdot (e^{t_n\cdot |v|}-1)& v<0\\
+    0 &  v=0\\
+  \end{cases}
+\end{equation}
+
+The window function used is a state-dependent quadratic function as shown in \cref{eq:memModel3}.
+
+\begin{equation}\label{eq:memModel3}
+  g(R,v)=
+  \begin{cases}
+    (R-r_p(v))^2& v>0\\
+    (R-r_n(v))^2& v<0\\
+    0 &  v=0\\
+  \end{cases}
+\end{equation}
+
+Where $r_p$ and $r_n$ are voltage depedent in a polynomial manner.
+They represent the boundaries of the state variable.
+All the other parameters in \cref{eq:memModel0,eq:memModel1,eq:memModel2,eq:memModel3} that haven't been declared are free fitting variables.
+They have to be set depending on the type of memristor used.
+
+\subsection{Verilog-A integration}
+
+The model now needs to be adapted to work as Verilog-A code.
+Verilog-A is a hardware desciption language, very popular in the industry for its simplicity and flexibility when using it with circuit simulator such as \textbf{Cadence}'s \textit{virtuoso}.
+
+The focus of this specific implementation will be the resistance range $[20k\Omega, 120k\Omega]$. The boundaries functions are found in \cref{eq:memModel4}.
+
+\begin{equation}\label{eq:memModel4}
+  \begin{cases}
+    r_p(v)= r_{p,0} + r_{p,1}\cdot v& v>0\\
+    r_n(v)= r_{r,0} + r_{n,1}\cdot v& v\le 0\\
+  \end{cases}
+\end{equation}
+
+Where $r_{p,0}$, $r_{p,1}$, $r_{n,0}$ and $r_{n,1}$ are free fitting variables that need to be extracted from the physical device.
+
+The changes of resistive states from the initial resistance ($R_0$) are computed by \cref{eq:memModel5} extracted from \cite{memCadenceModel}.
+It uses the bias voltage ($V_b$) of the pulse applied to the memristor.
+The pulse has a duration of $t$.
+
+\begin{equation}\label{eq:memModel5}
+  R(t)_{|V_b} =
+  \begin{cases}
+    \frac{R_0+s(V_b)\cdot r_p(V_b)\cdot(r_p(V_b)-R_0)\cdot t}{1+s(V_b)\cdot (r_p(V_b)-R_0)\cdot t} & V_b>0, R<r_p(V_b)\\
+    \frac{R_0+s(V_b)\cdot r_n(V_b)\cdot(r_n(V_b)-R_0)\cdot t}{1+s(V_b)\cdot (r_n(V_b)-R_0)\cdot t} & V_b\le 0, R>r_n(V_b)\\
+    R_0 & else\\
+  \end{cases}
+\end{equation}
+
+The final Verilog-A code is available at \cref{apsec:memModel}.
+
+\subsection{Behavior}
+
+The model
 

### 2023-10-18 3.Design/crossbar.tex
Corrected equations
Removed one TODO

@@
   \centering
-  \begin{tikztimingtable}%TODO change or find other way
+  \begin{tikztimingtable}
     $e_0$ & x2H2L N(A1) L N(A2) 2L N(A3) L N(A4) 2Lx\\
@@
   \begin{cases}
-    2\cdot n_i\cdot n_o \cdot A_{memristor}+n_o\cdot(2\cdot[A_{opAmp}+A_{R_r}]+A_{R_f}),\hfill n_s=1\\
+    2\cdot n_i\cdot n_o \cdot A_{memristor}+n_o\cdot(2\cdot[A_{opAmp}+A_{R_r}]+A_{R_f})& n_s=1\\
     \begin{array}{2}
@@
     \end{array}
-    ,\hfill \forall n_s>1\\
+    & \forall n_s>1\\
   \end{cases}
@@
 The input voltages must be monitored very carefuly in order not to mess up the entire system.
-Indeed, memristors have a voltage threshold ($V_{threshold}$) above which the internal resistance changes.
-%TODO Find info in the model paper
+Indeed, memristors have a voltage threshold ($V_{threshold}$) above which the internal resistance changes \cite{memristorSpiceModels,memCadenceModel,memTEAMmodel,memVTEAMmodel}.
 This means that the input voltages must follow the inequality in \cref{eq:memThres}.

### 2023-10-19 2.State/model.tex
Almost finished model

@@
 Based on this assumption the \cref{eq:memModel0,eq:memModel1} are extracted from the memristive device.
-This model is of course very basic and ignores some physical depedencies such as temperature.
+This model is of course very basic and ignores some physical dependencies such as temperature.
 
@@
 
-Where $r_p$ and $r_n$ are voltage depedent in a polynomial manner.
+Where $r_p$ and $r_n$ are voltage dependent in a polynomial manner.
 They represent the boundaries of the state variable.
@@
 
-The model
+The model was tested then tested.
+As with physical memristors, it has a threshold voltage value from which its resistive states starts changing.
+This threshold value is $V_{read}$.
+For this specific model the threshold voltage is set to $V_{read}=500mV$.
+
+Every pulse is composed of 2 parts and last for a total of $1 ms$ :
+\begin{itemize}
+  \item The reading of the resistive state, by applying a triangular wave that in bound by the threshold value $V_{read}$.
+  \item The writing to the memristor. During this part the resistive state will change. This is done by apply a square wave of variable width ($t_{\Delta}$) with a set bias voltage ($V_{bias}$).
+\end{itemize}
+
+\begin{figure}[H]
+  \centering
+  \includesvg[width=\textwidth]{memristor/input.svg}
+  \label{graph:memInput}
+  \caption{Pulse shape graph}
+\end{figure}
+
+The \cref{graph:memInput} is the visual representation of what each pulse, the value is read during the first half of the pulse.
+
+The resistive state of the memristor changes depending on the pulse width ($t_{\Delta}$) and the bias voltage ($V_{bias}$) used for each pulse.
+
+\begin{figure}[H]
+  \centering
+  \includesvg[width=\textwidth]{memristor/memPulseChange.svg}
+  \label{graph:pulseChange}
+  \caption{Memristor's resistive state under different pulses width (Fixed bias voltage $V_{bias}=1.8V$)}
+\end{figure}
+
+The effect of the pulse width variation is shown in \cref{graph:pulseChange}.
+Indeed, the larger the pulse widths, the faster the internal resistance reaches the bounding resistance (\cref{eq:memModel4}).
+The bounding resistance being the same as all curve use the same bias voltage ($V_{bias}=1.8V$), the curves will all eventually reach the max resistance.
+
+\begin{figure}[H]
+  \centering
+  \includesvg[width=\textwidth]{memristor/memVoltChange.svg}
+  \label{graph:voltChange}
+  \caption{Memristor's resistive state under different voltages (Fixed pulse width $t_{\Delta}=100\mu s$)}
+\end{figure}
+
+The voltage dependence of the memristor's model is described in \cref{graph:voltChange}.
+The influence of the bias voltage is, as \cref{eq:memModel3, eq:memModel4} show, the way to change the resistance range of the memristor, until the physical limit is reached of course.
+Applying higher voltage also allows to reach a set resistance faster as the slopes are bigger the higher the bias voltage gets.
 

### 2023-10-19 conclusions/further.tex
Renamed files and folder

--- 2023-10-13 5.Conclusions/further.tex
@@
 The final circuit will need an external controller to manage diffenrent things in the circuit.
-This wasn't created in this thesis but will have to be in order to have a functionning circuit once fabricated %TODO ref annex
+This wasn't created in this thesis but will have to be in order to have a functionning circuit once fabricated.
 
@@
 Idealy the algorithm won't need to compute the weight to resistance conversion (\cref{sec:wei2res}) and directly know how the change the resistances' values.
-
-Training directly on the chip has a great advange over external training.
-Indeed training within the chip will allow to completely ignore the wires own resistance in the crossbar array.
-This is because they will an integral part of the weights, if the weight needs to be lowered then the resistance will be raised (because the conductances need to be lowered).
-%TODO finish with better explainnation
 

### 2023-10-19 design/dataset.tex
Renamed files and folder

--- 2023-10-11 3.Design/dataset.tex
@@
 
-\subsubsection{Results} %TODO Move to results obviously
+\subsubsection{Timing}
 
-The weights are trained using both the custom activation function (\cref{sec:genwei}) and without ones and compare their output.
+The timing of the flags are quite hard to keep track of, for this reason all of them are going to be shown for the full duration of the system execution to get one output.
 
-After training on a digital computer using the settings described in \cref{sec:genwei}.
+\Cref{tim:airline} shows the time diagram of the entire execution using the parameters discussed earlier.
+
+\begin{figure}[H]
+  \centering
+  \begin{tikztimingtable}%TODO change or find other way
+    $e_0$       & x 2H 6L     L 2H 6L     L L x\\
+    $e_1$       & x 2L 2H 4L  L 2L 2H 4L  L L x\\
+    $e_2$       & x 4L 2H 2L  L 4L 2H 2L  L L x\\
+    $e_3$       & x 6L 2H     L 6L 2H     L L x\\
+    $e_{next}$  & x 8L        H 8L        H L x\\
+    $e_{in}$    & x 8H        L 8H        L L x\\
+    $e_{out}$   & x 8L        L 8L        L H x\\
+    \extracode
+    \tablerules
+    \begin{pgfonlayer}{background}
+      %\vertlines[help lines]{0.55,10.55}
+      %\vertlines[red]{1.6,5.6,15.6}
+      %\vertlines[blue]{3.6,9.6,15.6}
+    \end{pgfonlayer}
+  \end{tikztimingtable}
+  \caption{Flags time diagram for the airline problem with $n_s=4$}
+  \label{tim:airline}
+\end{figure}
 

### 2023-10-19 design/netlist.tex
Renamed files and folder

--- 2023-10-02 3.Design/netlist.tex
@@
 
-This is done by generating a SPICE netlist using a python script. This script takes in a few parameters :
+This is done by generating a SPICE netlist using a python script.
+A SPICE netlist was chosen because of Cadence's virtuoso limitations.
+Indeed, Verilog netlists can be imported with a downside, component's parameters can't be imported.
+This is very limiting for this thesis' use case (required to set the resistors' resistance and thus the weights).
+SPICE netlists can import parameters for the components and is open source and by such well documented.
+The generator script takes in a few parameters :
 
@@
   \item The serial size of the crossbar arrays, as described in \cref{sec:xbarCircuit}, used in the \ac{LSTM} network. This parameter must divide the number of hidden states in the \acp{LSTM} layers.
-  \item The files containing different info about the model.
+  \item The files containing different informations about the model.
 This file contains the type of \ac{NN} architecture and the weights associated with each layer.
@@
 \end{itemize}
+
+\subsection{Available layers}
+
+As of the writing of the thesis, the netlist generator can generate the netlist for :
+
+\begin{itemize}
+  \item \ac{NP} \ac{LSTM}
+  \item Vanilla \ac{LSTM}
+  \item \ac{FGR} \ac{LSTM}
+  \item \ac{GRU}
+\end{itemize}
+
+Nevertheless, the weights can only be generated for the \ac{NP} \ac{LSTM} and the \ac{GRU} (\cref{sec:genwei}).
+While the script can tecnically generate the Vanilla and \ac{FGR} \acp{LSTM}, since there is no way to generate weights for those version of the \ac{LSTM}, they are rendered useless.
 

### 2023-10-19 design/run.tex
Renamed files and folder

\section{Running the simulation}\label{sec:run}

The simulation is ran using Cadence's virtuoso simulator, using a corner simulation to simulate all the inputs.

In case the \ac{LSTM} we want to run requires several outputs, the use of the corner functionality. Corners are intended to be used for small variations of the variables in the circuit. However, it used differently in this thesis. It runs with the value for each input variable changing for the current \ac{LSTM} run and then moves on to the next with the variables changed.

In order to run the simulation some variables need to be set. The time step ($T$) value was chosen to be $T=8\mu s$, the pause between two \ac{LSTM} steps is set to be $\frac{T}{8}=1\mu s$. The value for the maximum and minimum resistances value of the memristor used ($R_{max}$ and $R_{min}$, respectively) are set to an arbitrary value of $R_{max}=10^6 \Omega=1M\Omega$ and $R_{min}=10^4 \Omega = 10k\Omega$.

Once everything is set, the simulation can be started. The result are recorded by looking at the output net when at the time of output activation. The name of the output nets is outputed by the netlist generator script (\cref{sec:netlist}).

### 2023-10-19 design/wei2res.tex
Renamed files and folder

--- 2023-10-10 3.Design/wei2res.tex
@@
 
-For the thesis, this step is done in python and integrated in the netlist generator script (\cref{sec:netlist}). Since the resistance resolution is of about%TODO check memristor precision
-, the python code then reproduces this by limiting the amount of significant digit of the resistance.
+For the thesis, this step is done in python and integrated in the netlist generator script (\cref{sec:netlist}).
+The resolution of the memristor (the precision of the resistances at which the memristor can be set) is simulated by limiting the generated resistances values to having only two significant figures.

### 2023-10-19 introduction/contributions.tex
Renamed files and folder

\section{Main Contributions}
\label{sec:contributions}

In the thesis the following points will worked on in order to create, improve or clarify any of those topics :

\begin{itemize}
\item A script that can generate spice netlists for \acps{LSTM} or feed forward \aclp{NN} of any size and importing weights as resistance values from a specified file containing the architecture and the weights.
\item An \ac{LSTM} capable analog circuit.
\item Analog activation function circuit that supports \ac{tanh} and sigmoid ranges.
\item A \ac{VMM} capable analog circuit using a matrix of memristors.
\item Memory cells to store analog data for a short amount of time.
\item A python (tensorflow library) code to use piece wise functions, to reproduce in python the analog activation functions.
\item A way to run software simulation of different \acp{NN}.
\end{itemize}

### 2023-10-19 results/airline.tex
Renamed files and folder

--- 2023-10-13 4.Results/airline.tex
@@
 
+coincidentally, the curve is closer to the target curve than the digital prediction itself, this is mostly due to the lack of precision of the analog prediction, but it is nevertheless interresting to mention.
+\Cref{graph:airlineCoin} shows the graph and the \ac{RMSE} is the lowest, even lower than the ones from the digital predictions.
+
+\begin{figure}[H]
+  \centering
+  \includesvg[width=\linewidth]{results/airline/closeRes}
+  \caption{Analog prediction ($n_s=1$) with the target dataset.}
+  \label{graph:airlineCoin}
+\end{figure}
+
 The overall observation of those results compared to the previous ones (\cref{subsec:airlineAnalogNoC}) are how much closer to their target they are.

### 2023-10-19 state/gru.tex
Renamed files and folder

--- 2023-10-17 2.State/gru.tex
@@
 
+\begin{figure}[H]
+  \centering
+  \includesvg[width=\textwidth,pretex=\large]{gru/encoderCell.svg}
+  \label{fig:lstmCell}
+  \caption{Encoder \acs{GRU} cell, legend in \cref{leg:cells}}
+\end{figure}
+
 Sometimes the \cref{eq:gruHidG} can be found in another form (\cref{eq:gruHidG1})\cite{gruPyTorch}.
@@
 
+\begin{figure}[H]
+  \centering
+  \includesvg[width=\textwidth,pretex=\large]{gru/decoderCell.svg}
+  \label{fig:lstmCell}
+  \caption{Decoder \acs{GRU} cell, legend in \cref{leg:cells}}
+\end{figure}
+
 \subsection{Similarities with \ac{LSTM}}
 
-The \ac{CIFG} \ac{LSTM} or more commonly known as \ac{GRU}.
-\cite{gru,gruKeras,gruPyTorch} As it's name implies, \acl{CIFG}, links the input and forget gates as such, $f_t=1-i_t$.
-Other differences are :
+While the \ac{GRU} is its own type of \ac{RNN}, it has been compared and assimilated to an \ac{LSTM} \cite{nbLSTM}.
+Indeed, the subtype of \ac{LSTM} it has been associated with is called the \ac{CIFG} \ac{LSTM}.
+If the input and forget gates are combined into one ($f_t=\overrightarrow{1}-i_t$), they can be assimilated as the update gate of the \ac{GRU}.
+The reset gate would then be the \ac{LSTM}'s output gate.
+
+\Cref{eq:lstmGru0,eq:lstmGru1,eq:lstmGru2,eq:lstmGru3,eq:lstmGru4} are the \ac{GRU}'s equations with the \ac{LSTM}'s names.
+
+\begin{equation}\label{eq:lstmGru0}
+  \overrightarrow{f_t}=\sigma ([\overrightarrow{x_t},\overrightarrow{h_{t-1}}] \cdot w_f + \overrightarrow{b_f})
+\end{equation}
+\begin{equation}\label{eq:lstmGru1}
+  \overrightarrow{o_t}=\sigma ([\overrightarrow{x_t},\overrightarrow{h_{t-1}}] \cdot w_o + \overrightarrow{b_o})
+\end{equation}
+\begin{equation}\label{eq:lstmGru2}
+  \overrightarrow{i_t}=\overrightarrow{1}-\overrightarrow{f_t}
+\end{equation}
+\begin{equation}\label{eq:lstmGru3}
+  \overrightarrow{\tilde{c_t}}=tanh(\overrightarrow{x_t}\cdot w_{cx}+(\overrightarrow{o_t}\odot\overrightarrow{h_{t-1}}) \cdot w_{ch} + \overrightarrow{b_c})
+\end{equation}
+\begin{equation}\label{eq:lstmGru4}
+  \overrightarrow{h_t}=(\overrightarrow{i_t})\odot \overrightarrow{h_{t-1}} + \overrightarrow{f_t}\odot \overrightarrow{\tilde{h_t}}
+\end{equation}
+
+Differences between the two \acp{RNN} are :
 \begin{itemize}
-  \item No peepholes and output activation function.
-  \item Combined hidden state and cell state.
-  \item The candidate cell state is point wise multiplied with the output gate before the activation function.
+  \item The \ac{GRU} has no output activation function.
+  \item The \ac{GRU} has combined hidden state and cell state.
+  \item The \ac{GRU}'s candidate hidden state is point wise multiplied with the reset gate before passing into the dense layer.
 \end{itemize}
-
-This is clearer with the math equations :
-\begin{equation}\label{eq:cellGGRU}
-  \tilde{c}_t = tanh(w_{cx}\cdot x_t + w_{ch}\cdot(o_t \odot h_{t-1}) + b_c)
-\end{equation}
-\begin{equation}\label{eq:hiddenGRU}
-  h_t=(1-i_t)\cdot h_{t-1} + i_t \cdot \tilde{c}_t
-\end{equation}

### 2023-10-19 state/model.tex
Renamed files and folder

--- 2023-10-19 2.State/model.tex
@@
 The voltage dependence of the memristor's model is described in \cref{graph:voltChange}.
-The influence of the bias voltage is, as \cref{eq:memModel3, eq:memModel4} show, the way to change the resistance range of the memristor, until the physical limit is reached of course.
+The influence of the bias voltage is, as \cref{eq:memModel3,eq:memModel4} show, the way to change the resistance range of the memristor, until the physical limit is reached of course.
 Applying higher voltage also allows to reach a set resistance faster as the slopes are bigger the higher the bias voltage gets.
+
+The resistive state can also be lowered by applying negative voltages.TODO
 

### 2023-10-19 state/nn.tex
Renamed files and folder

--- 2023-10-16 2.State/nn.tex
@@
 
-The weights values are obtained through training. This is done in a few steps :
+The weights values are obtained through training.
+To train the weights it is required to have a dataset of inputs/outputs values that are correct.
+The outputs are the target values, the values that the \ac{NN} need to compute when being executed.
+Once the weights have been initialized (usually randomly picked values).
+The training happens by running the following steps (known as an epoch) :
 
 \begin{enumerate}
-  \item Generate random values for each weight.
   \item Run the \ac{NN} to get an output vector.\label{step:restart}
   \item Measure error rate using the chosen error algorithm.
-  \item Use the
-  \item go back to \cref{step:restart}
+  \item Run the backpropagation algorithm to determine how to change each individual weight.
 \end{enumerate}
+
+A \ac{NN} is trained for at least hundreds of epochs to get decent results.
+
+Once trained, the \ac{NN} can be tested by feeding the \ac{NN} input data it hasn't been trained with.
+
+One layer of this feedforward \ac{NN} is also known as a dense layer.

### 2023-10-19 introduction/outline.tex
Forgot to rename the folders in the mains...

\section{Thesis Outline}
\label{sec:outline}

This thesis is going to tackle the work that went into creating a software simulation of an analog computer capable of running \aclp{RNN}.

All the technologies used and, thus are required to know in order to understand the thesis are detailed in the state of the art.

Next the work that was done for the thesis is goign to be thoroughly explained. Starting with the different circuit that were designed for the thesis. Then the different scripts and python code that were created, either to run the digital \acp{NN} or to generate the analog system into a netlist.

### 2023-10-19 0.Intro/abstract.tex
first draft of abstract

\begin{abstract}

  \acp{AI} are one of the main research domain in computer sciences and is used in other scientific domains. One great application of \ac{AI}, is in embedded systems. Lots of small electronic devices around us could benefit us. \textbf{Google} and \textbf{Apple} are pushing \ac{AI} in phones with their in house designed ARM processors having, very advertised, tensor cores.
  Being able to run low powered \acp{AI} is thus one of the biggest computer objective of the deceny.
  Analog computers are known to be a low powered technology that most importantly give almost instant results. An analog \ac{ASIC} could then be fabricated for specific embedded applications. This would definitely improve digital technologies in terms of time, and maybe in power consumption as well. Which are two prerequisite for embedded system use.
  For this reason this work aims at creating a functioning software simulation of analog circuit capable of running \ac{AI}. The thesis will focus on the software simulation, the first step of the work in order to have a fully working chip able to run \ac{AI} algorithms.

\end{abstract}

### 2023-10-19 0.Intro/acknowledgments.tex
thanking people and all

\pdfbookmark{Acknowledgments}{Acknowledgments}
\begin{acknowledgments}

  I would like to thank the \ac{IST} for allowing to join their double degree program, that made it possible for me to widen my computer knowledge from computer science to electronics.
  I would also like to thank \ac{INESC} to have let me join their rank, to both \ac{INESC}-Id and \ac{INESC}-MN groups to have given the necessary tools to succeed with this thesis.
  I would also like to thank all the other student working in the \ac{ASIC} group with whom I could share problems and were happy to help.
  I would lastly like to thank Dr Ruxandra Barlulescu and Dr Diogo Caetano, to have trusted with this very interresting topic despite my recent arrival in the univerty.

\end{acknowledgments}
\clearpage
\thispagestyle{empty}
\cleardoublepage

### 2023-10-20 results/celegans.tex
Celegans analog results beginning

--- 2023-10-19 results/airline.tex
@@
-\section{Airline dataset}
-\label{sec:resAirline}
+\section{\acs{C. elegans} dataset}
+\label{sec:resCelegans}
+
+Due do the complexity of this dataset and the results associated are only trained using the analog activation function.
+Furthermore, the usefulness of using those activation functions was already demonstrated in \cref{sec:resAirline}.
 
 \subsection{Digital results}
-\label{subsec:digitalAirline}
+\label{subsec:digitalCelegans}
 
-The digital results of the \ac{NN} once trained are very important.
-They are the results that the analog system will have to reproduce.
-Furthermore, this part is used to measure the impact of the custom activation functions used for training (\cref{sec:af}).
-
-\begin{figure}[H]
-  \centering
-  \includesvg[width=\linewidth]{datasets/airlineDigital}
-  \caption{Graph of the digital predictions for airline dataset. The dotted vertical line shows the limit of the data used for training and the one used for validation.}
-  \label{graph:airlineDigital}
-\end{figure}
-
-\Cref{graph:airlineDigital} shows the results of the \ac{NN} next to the orinal data.
-The results are visually very close.
-The error is quantified using \ac{RMSE} function.
-All those curves error to each other are measured and displayed in \cref{tab:airlineDigital}.
-
-\begin{table}[H]
-  \centering
-  \begin{tabular}{|c|c|c|c|}
-    \hline
-    \cellcolor[HTML]{808080}\acs{RMSE} & Target data & \specialcell{Digital prediction with\\default activation functions} & \specialcell{Digital prediction with\\analog activation functions}\\
-    \hline
-    Target data &\cellcolor[HTML]{202020} & $54.60$ & $52.27$\\
-    \hline
-    \specialcell{Digital prediction with\\default activation functions}  & $54.60$ & \cellcolor[HTML]{202020} & $78.64$\\
-    \hline
-    \specialcell{Digital prediction with\\analog activation functions} & $52.27$ & $78.64$ & \cellcolor[HTML]{202020}\\
-    \hline
-  \end{tabular}
-  \caption{\acp{RMSE} of each curve to the others}
-  \label{tab:airlineDigital}
-\end{table}
-
-Both predictions have similar \ac{RMSE} to the target curve.
-They have a \ac{RMSE} of $54.60$ for the \ac{LSTM} using the default activation functions (sigmoid and \ac{tanh} functions) and $52.27$ for the one using the analog activation functions (\cref{sec:af}).
-The error difference is so low (about two thousands passengers) that the difference may just come from the training and the initial values of the weights.
-They both approach the target curve pretty well.
-These error could be lowered by training for more epochs.
-This is not done because it diverges too much from the focus of the thesis.
-
-The two predictions are very different from each other, the error rate between the two is of $78.64$, which is significantly higher than the \ac{RMSE} of the curves to the target curve.
-
-Computing the \ac{NN} with or without the analog activation functions doesn't seem to affect the error of the predictions, experience has shown that the \ac{LSTM} using the analog activation functions seem to predict higher curves than the target and the \ac{LSTM} using the default activation functions tends to predict curves that are under the target, the reasons for such a difference are unknown.
-This is simply an observation after a few different runs.
-
-\subsection{Analog results trained with default activation functions}\label{subsec:airlineAnalogNoC}
-
-This subsection shows the importance of using the analog activation functions in the training phase of the \ac{NN}.
-
-\begin{figure}[H]
-  \centering
-  \includesvg[width=\linewidth]{results/airline/noCustomAF}
-  \caption{Analog predictions trained with the default activation functions.}
-  \label{graph:airlineAnalogNoC}
-\end{figure}
-
-As expected, the analog predictions are very far off the target curve.
-Which, in this case, isn't the target dataset but the digital prediction ran with the default activation functions.
-Indeed, the goal of the analog system is to reproduce as closely as possible the digital prediction.
-
-\begin{table}[H]
-  \centering
-  \begin{tabular}{|c|c|c|c|}
-    \hline
-    \multirow{2}{*}{\acs{RMSE}} & \multicolumn{3}{|c|}{Analog prediction}\\
-    \cline{2-4}
-    & $n_s=1$ & $n_s=2$ & $n_s=4$ \\
-    \hline
-    \specialcell{Digital prediction with\\default activation functions} & $140.68$ & $143.66$ & $139.71$\\
-    \hline
-  \end{tabular}
-  \caption{\acp{RMSE} of each analog prediction to their associated digital prediction}
-  \label{tab:airlineAnalogNoC}
-\end{table}
-
-As can be observed on \cref{graph:airlineAnalogNoC}, the analog predictions are not on par with what is expected at all.
-This is confirmed by the error rates found in \cref{tab:airlineAnalogNoC}, the error rates are not in an acceptable range at all.
-This error would be even higher to the target dataset as the digital prediction using the default activation functions is already under than the target curve.
-
-However, the prediction don't seem to be completely off. The curves seem to follow the shape of it's target, and by extension of the target data. This deserves a closer look.
-
-\begin{figure}[H]
-  \centering
-  \includesvg[width=\linewidth]{results/airline/noCustomAFZoomed}
-  \caption{Analog predictions trained with the default activation functions.}
-  \label{graph:airlineAnalogNoCZoomed}
-\end{figure}
-
-As expected, in \cref{graph:airlineAnalogNoCZoomed} the trend of the curves is very similar to the original dataset (\cref{graph:airline}) and the predictions are very close to each others no matter the serial size used ($n_s$).
-
-\begin{table}[H]
-  \centering
-  \begin{tabular}{|c|c|c|c|c|}
-    \hline
-    \multicolumn{2}{|c}{\multirow{2}{*}{\ac{RMSE}}} & \multicolumn{3}{|c|}{Analog prediction}\\
-    \cline{3-5}
-    \multicolumn{2}{|c}{} & \multicolumn{1}{|c|}{$n_s=1$} & $n_s=2$ & $n_s=4$ \\
-    \hline
-    \multirow{3}{*}{Analog prediction} & $n_s=1$ &\cellcolor[HTML]{202020} & $3.97$ & $5.05$\\
-    \cline{2-5}
-    & $n_s=2$  & $3.97$ & \cellcolor[HTML]{202020} & $4.78$\\
-    \cline{2-5}
-    & $n_s=4$ & $5.05$ & $4.78$ & \cellcolor[HTML]{202020}\\
-    \hline
-  \end{tabular}
-  \caption{\acp{RMSE} of each analog prediction to the others depending on the serial size ($n_s$)}
-  \label{tab:airlineAnalogNoCZoomed}
-\end{table}
-
-This similarity is further proven by the errors shown in \cref{tab:airlineAnalogNoCZoomed}.
-Those error are all around four thousands passengers, which is negligible in our case.
-That shows a certain stability of the system when using different values of serial size ($n_s$).
-The error rate is also quite low because of the lower results obtained.
-
-The predicted results generated by the analog system seem to be on a different scale while still following the overall aspect of the dataset.
-The down scale of those analog prediction would also mean that the errors of those predictions to each other is also scaled down.
-
-\subsection{Analog results trained with analog activation functions}
+\subsection{Analog results}
 
@@
   \centering
-  \includesvg[width=\linewidth]{results/airline/analog}
-  \caption{Analog predictions trained with the analog activation functions.}
-  \label{graph:airlineAnalog}
+  \includesvg[width=\linewidth]{results/celegans/out5.svg}
+  \caption{Analog predictions of the 5th data sequence.}
+  \label{graph:celegansAnalog0}
 \end{figure}

### 2023-10-20 results/celegans.tex
Going to lunch

@@
 
+Let's focus on a single data sequence, for simplicity. The training data sequence found in \cref{graph:io5Celegans}.
+
 This part will show the result of the inference when the weights were trained using the analog activation functions.
@@
   \centering
-  \includesvg[width=\linewidth]{results/celegans/out5.svg}
+  \includesvg[width=\linewidth]{results/celegans/out5}
   \caption{Analog predictions of the 5th data sequence.}
   \label{graph:celegansAnalog0}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \begin{minipage}{\columnwidth}
+    \subfloat[Training dataset sample\label{graph:io5Celegans}]{\includesvg[width=0.5\textwidth, pretex=\tiny]{datasets/celegans/ioTrain0}}%
+    \hfill
+    \subfloat[Test dataset sample]{\includesvg[width=0.5\textwidth,pretex=\tiny]{datasets/celegans/ioTest0}}%
+  \end{minipage}
+  \begin{minipage}{\columnwidth}
+    \subfloat[Training dataset sample]{\includesvg[width=0.5\textwidth, pretex=\tiny]{datasets/celegans/ioTrain1.svg}}%
+    \hfill
+    \subfloat[Validation dataset sample]{\includesvg[width=0.5\textwidth,pretex=\tiny]{datasets/celegans/ioValid0}}%
+  \end{minipage}
+  \caption{\ac{C. elegans} dataset samples}
+  \label{graph:celegans}
 \end{figure}

### 2023-10-20 results/celegans.tex
Worked on graphs and results of the celegans ds

--- 2023-10-20 results/celegans.tex
@@
   \centering
-  \includesvg[width=\linewidth]{results/celegans/out5.svg}
-  \caption{Analog predictions of the 5th data sequence.}
+  \includesvg[width=\linewidth]{results/celegans/out5}
+  \caption{\acs{C. elegans} analog responses with their digital counterparts}
   \label{graph:celegansAnalog0}
@@
 
-As can be observed in \cref{graph:airlineAnalog}, the analog results are, in this case, way closer to the target values.
-The curves are very close to eahc other and are a definite improvement from the the ones trained using default activation functions.
+\Cref{graph:celegansAnalog0} contains the predictions of both the digital and analog \ac{NN} for all four output neurons that are being monitored.
+The having all predicted responses to the stimuli on the same graph makes things very unintelligeable.
 
-\begin{table}[H]
-  \centering
-  \begin{tabular}{|c|c|c|c|}
-    \hline
-    \multirow{2}{*}{\acs{RMSE}} & \multicolumn{3}{|c|}{Analog prediction}\\
-    \cline{2-4}
-    & $n_s=1$ & $n_s=2$ & $n_s=4$ \\
-    \hline
-    \specialcell{Digital prediction with\\analog activation functions} & $28.41$ & $92.57$ & $68.51$\\
-    \hline
-  \end{tabular}
-  \caption{\acp{RMSE} of each analog prediction to their associated digital prediction}
-  \label{tab:airlineAnalog}
-\end{table}
+The response have thus been sperated into four graphs.
+One for every neuron for better readability.
+The target response was also added to all four graphs to highlight the objective of the \ac{NN}.
+Those graphs can be found in \cref{graph:spread5Celegans}.
 
-\Cref{tab:airlineAnalog} contains the errors of the analog predictions.
-It shows that, as can be observed, the results the closest to the target are the ones generated using a serial size of one ($n_s=1$).
-Indeed, its error ($28.41$) is lower than the error from the digital results ($52.27$) to the original target values, almost half of it.
-The other ones are still very close to the targeted curve, but are still lower.
-This is probably due to the serialization circuit causing small differences in the signals at each steps.
-
-\begin{table}[H]
-  \centering
-  \begin{tabular}{|c|c|c|c|c|}
-    \hline
-    \multicolumn{2}{|c}{\multirow{2}{*}{\ac{RMSE}}} & \multicolumn{3}{|c|}{Analog prediction}\\
-    \cline{3-5}
-    \multicolumn{2}{|c}{} & \multicolumn{1}{|c|}{$n_s=1$} & $n_s=2$ & $n_s=4$ \\
-    \hline
-    \multirow{3}{*}{Analog prediction} & $n_s=1$ &\cellcolor[HTML]{202020} & $64.61$ & $40.63$\\
-    \cline{2-5}
-    & $n_s=2$  & $64.61$ & \cellcolor[HTML]{202020} & $25.75$\\
-    \cline{2-5}
-    & $n_s=4$ & $40.63$ & $25.75$ & \cellcolor[HTML]{202020}\\
-    \hline
-  \end{tabular}
-  \caption{\acp{RMSE} of each analog prediction to the others depending on the serial size ($n_s$)}
-  \label{tab:airlineAnalogError}
-\end{table}
-
-Once again, the results are quite close to each others, but as expected the errors (\cref{tab:airlineAnalogError}) are higher then the ones found in \cref{tab:airlineAnalogNoCZoomed}.
-The farthest appart ($n_s=1$ to $n_s=2$ with an error of $64.61$ thousands passengers) have an error barely higher than the error of digital results to the target dataset (\cref{tab:airlineDigital})
-
-coincidentally, the curve is closer to the target curve than the digital prediction itself, this is mostly due to the lack of precision of the analog prediction, but it is nevertheless interresting to mention.
-\Cref{graph:airlineCoin} shows the graph and the \ac{RMSE} is the lowest, even lower than the ones from the digital predictions.
+A big problem that is immediatly visible is that the LUAL output neuron's analogly predicted response is negative when it should not be.
+For an unknown reason the analog results of the LUAL neuron seem to be negative instead of positive.
+This is the case for the analog response to all sequences.
+This issue most likely comes from dense layer that comes after the \ac{LSTM}.
+In order to keep using the data the output of this neuron will be flipped over the x-axis.
 
@@
   \centering
-  \includesvg[width=\linewidth]{results/airline/closeRes}
-  \caption{Analog prediction ($n_s=1$) with the target dataset.}
-  \label{graph:airlineCoin}
+  \begin{minipage}{\columnwidth}
+    \subfloat[Analog prediction for the DB1 neuron\label{graph:out5BD1Celegans}]{\includesvg[width=0.5\textwidth, pretex=\tiny]{results/celegans/out5DB1}}%
+    \hfill
+    \subfloat[Analog prediction for the LUAL neuron\label{graph:out5LUALCelegans}]{\includesvg[width=0.5\textwidth,pretex=\tiny]{results/celegans/out5LUAL}}%
+  \end{minipage}
+  \begin{minipage}{\columnwidth}
+    \subfloat[Analog prediction for the PVR neuron\label{graph:out5PVRCelegans}]{\includesvg[width=0.5\textwidth, pretex=\tiny]{results/celegans/out5PVR}}%
+    \hfill
+    \subfloat[Analog prediction for the VB1 neuron\label{graph:out5VB1Celegans}]{\includesvg[width=0.5\textwidth,pretex=\tiny]{results/celegans/out5VB1}}%
+  \end{minipage}
+  \caption{\ac{C. elegans} analog responses with their digital counterparts spread out for each output neuron}
+  \label{graph:spread5Celegans}
 \end{figure}
 
-The overall observation of those results compared to the previous ones (\cref{subsec:airlineAnalogNoC}) are how much closer to their target they are.
-It still isn't quite perfect but simply changing the activation functions used for training improved the results by a lot.
-It can thus be theorized that the error will become null if the system is trained with the same parameters as the ones it is using for inference.
-In other words, the error will probably be nullified when the training will happen inSitu (\cref{subsec:inSitu}).
+All the analog predicted responses seem to have the same problem.
+When the response is flat, the signal looks very good, while with a small vertical displacement.
+That displacement could be attributed to the difference of computations from the digital training to the analog inference (\cref{sec:resAirline}).
+However, when the analog response sends out a pulse, the said pulse is very unstable.
+It continuously flickers up and down.
+This can be observed by zooming on the graph like in \cref{graph:zoom5Celegans}.
+
+This phenomenen is very strange and hasn't been identified so far.
+The main theory is that the data isn't stored properly in the memory cells, thus at every \ac{LSTM} time step the supposed output is slightly altered.
+This would create this flickering of the outputed data.
+
+The data isn't useless for that matter.
+It was decided to smooth the data by averaging every output point with its neighboring points.
+This simple average will smooth out the curves to get a decently readable result, that can then be compared with the digitally predicted response.
+The python code for this function is available at \cref{}.
+%TODO
+
+\begin{figure}[H]
+  \centering
+  \includesvg[width=\linewidth]{results/celegans/zoom5}
+  \caption{\acs{C. elegans} DB1 analog response with zoom on the pulses}
+  \label{graph:zoom5Celegans}
+\end{figure}
+
+The smoothed out and exploitable analog predictions are available in \cref{graph:smooth5Celegans}. Theses graphs allow for a much easier visualisation of the results.
+
+The best looking output of the four predicted neurons responses is the one from PVR.
+That is confirmed by its \ac{RMSE} with the digital prediction.
+It is the lowest of the four, however, it isn't close enough to the digital prediction.
+
+All four analog predictions are very close to each others
+
+\begin{figure}[H]
+  \centering
+  \begin{minipage}{\columnwidth}
+    \subfloat[Averaged analog prediction for the DB1 neuron\label{graph:smooth5BD1Celegans}]{\includesvg[width=0.5\textwidth, pretex=\tiny]{results/celegans/smooth5DB1}}%
+    \hfill
+    \subfloat[Averaged analog prediction for the LUAL neuron\label{graph:smooth5LUALCelegans}]{\includesvg[width=0.5\textwidth,pretex=\tiny]{results/celegans/smooth5LUAL}}%
+  \end{minipage}
+  \begin{minipage}{\columnwidth}
+    \subfloat[Averaged analog prediction for the PVR neuron\label{graph:smooth5PVRCelegans}]{\includesvg[width=0.5\textwidth, pretex=\tiny]{results/celegans/smooth5PVR}}%
+    \hfill
+    \subfloat[Averaged analog prediction for the VB1 neuron\label{graph:smooth5VB1Celegans}]{\includesvg[width=0.5\textwidth,pretex=\tiny]{results/celegans/smooth5VB1}}%
+  \end{minipage}
+  \caption{\ac{C. elegans} smoothed out analog responses averaged with $21$ neighbors, spread out for each output neuron}
+  \label{graph:smooth5Celegans}
+\end{figure}

### 2023-10-23 results/celegans.tex
No work was done

@@
 
+The results will be analyzed by focusing on a single data sequence, the training data sequence available in \cref{graph:io5Celegans}.
+
 \subsection{Digital results}
@@
 \subsection{Analog results}
+\label{subsec:analogCelegans}
 
-Let's focus on a single data sequence, for simplicity. The training data sequence found in \cref{graph:io5Celegans}.
+It's been established in \cref{sec:resAirline}, that the analog circuit with a serial size of one ($n_s=1$) show a lower \ac{RMSE} to the digital predictions than the digital predictions.
+
+Due to the complexity of the dataset, the simulation were only ran using a serial size of one ($n_s=1$) to get the best looking results.
 
@@
 The best looking output of the four predicted neurons responses is the one from PVR.
-That is confirmed by its \ac{RMSE} with the digital prediction.
+That is confirmed by its error rate to the digital prediction.
 It is the lowest of the four, however, it isn't close enough to the digital prediction.

### 2023-10-23 design/gru.tex
Bunch of changes, added design gru

--- 2023-10-04 3.Design/lstm.tex
@@
-\section{\ac{LSTM} analog implementation}\label{sec:lstmCircuit}
+\section{\acs{GRU} analog implementation}\label{sec:gruCircuit}
 
@@
 
-This section describes the circuit of an \ac{LSTM} with an input vector of size $n_i$, a $n_h$ hidden states, a serial size of $n_s$ and $n_{ts}$ time steps.
-$n_o=n_h/n_s$ is going to be used for future references in this section.
-In order for the crossbar array to be used $n_o$ must be an integer, in other words, $n_s$ must divide $n_h$.
-The circuit (shown in \cref{circt:lstm}) is pretty complex and contains numerous parts that require explaination.
+This section describes the circuit of an encoder \ac{GRU} with an input vector of size $n_i$, a $n_h$ hidden states and $n_{ts}$ time steps.
+The decoder \ac{GRU} could also be implemented in an analog circuit, but the choice was made to focus on the encoder \ac{GRU}.
+The \ac{GRU} is by its nature very similar to the \ac{LSTM}.
 
@@
 
-The system is built using crossbar array (\cref{sec:xbarCircuit}) with ($n_i+n_h+1$,$n_o$, $n_s$) as parameters.
+The system is built, once again, using crossbar array (\cref{sec:xbarCircuit}) with ($n_i+n_h+1$,$n_o$, $n_s$) as parameters.
 
@@
 
-The wires coming into the crossbar are a bus of size $n_i+n_h+1$ and the output of the crossbar is a bus of size $n_o$ (\cref{sym:xbar}).
+The wires coming into the crossbar are a bus of size $n_i+n_h+1$ and the output of the crossbar is a bus of size $n_h$ (\cref{sym:xbar}).
 This is why everything in the system apart from the crossbar arrays is only shown once in \cref{circt:lstm} but in reality those components are present $n_o$ times.
@@
 
-\subsection{Doubling memory cells}
+\subsection{Serialization/Parallelization}
+\label{subsec:gruSerPar}
 
-\subsubsection{Feedback hidden states}
+As opposed to the to the \ac{LSTM}, the \ac{GRU} can not be serialized. This is due to a mathematical issue.
 
-In order for the system to work in serial mode, the memory cells of the output need to be doubled.
-This is done in \cref{circt:lstm}, and allows for the hidden states to be saved for the next stage.
-Indeed, if using the system in serial mode with only one line of memory cells, the old hidden state ($\overrightarrow{h_{t-1}}$) is slowly overwritten by the current hidden state ($\overrightarrow{h_t}$).
-As soon as the first $n_o$ serial values are computed, they are going to override the old ones, still required for the following serial values.
-\Cref{tim:lstmMemcell} shows that the value in the memory cell changes too early in the cycle and gives a bad input for the next serial values.
-The value given to the input of the \ac{LSTM} should always be the old hidden state ($\overrightarrow{h_{t-1}}$).
+Indeed in the equation of the candidate hidden state is a problem for the serialization of the circuit.
+The \cref{eq:candActivG} from \cref{sec:gru} is repeated in \cref{eq:serProb} for better lisibility of the thesis.
+
+\begin{equation}\label{eq:serProb}
+  \overrightarrow{\hat{h_t}}=tanh(\overrightarrow{x_t}\cdot w_{hx}+(\overrightarrow{r_t}\odot\overrightarrow{h_{t-1}}) \cdot w_{hh} + \overrightarrow{b_h})
+\end{equation}
+
+As \cref{eq:serProb} shows, the reset gate's results vector needs to be multiplied with the the hidden state's previous vector.
+However, while the hidden state's previous vector is fully available in the memory cells, the reset gate's vector needs to be computed using the current inputs.
+
+If the circuit were serialized, a full cycle would be required to compute the reset gate, that is itself required to compute the candidate hidden state.
+
+That is why, the \ac{GRU} would take twice as much time to compute a time step if it were serialized. It is not worth it to double the computation time to save a bit of onChip area.
+
+\subsection{Symbol}
+
+The symbol of the \ac{GRU} circuit is available in \cref{sym:lstm}.
+It depends on a few parameters namely, the number of inputs ($n_i$), the number of hidden states ($n_h$) and the number of time steps ($n_{ts}$).
 
@@
   \centering
-  \begin{tikztimingtable}
-    $e_{n_s-1}$ & 3H8Lx \\
-    $e_0$ & 3L4H4Lx \\
-    $e_1$ & 7L4Hx \\
-    $m_{i,0}$ & 3D{$h_{t-1,i,0}$}8D{$h_{t,i,0}$}x \\
-    $m_{i,1}$ & 7D{$h_{t-1,i,1}$}4D{$h_{t,i,1}$}x \\
-    \extracode
-    \tablerules
-    %\draw (0,0) circle (2pt); % Origin
-    \begin{pgfonlayer}{background}
-      \vertlines[help lines]{3.05}
-      %\vertlines[red]{1.6,5.6,15.6}
-      %\vertlines[blue]{3.6,9.6,15.6}
-    \end{pgfonlayer}
-  \end{tikztimingtable}
-  \caption{Time diagram of the values in the different memory cells ($m_{i,j}$ being the value stored in the $i^{th}$ parallel memory cells for the $j^{th}$ serial value).
-It is assumed that $n_s>1$ (the system is in serial mode).
-This graph does not take into account the pauses $e_{next}$ between the time steps.}
-  \label{tim:lstmMemcell}
-\end{figure}
-
-Using two memory cells when in serial mode allows for this issue to be elevated.
-The values are transfered from one level to an other at the end of the time step.
-That way, the values are updated safely during each time step, without changing the final output.
-
-When the system is running is a fully parallel mode, doubling the memory cells still causes a problem.
-The enable in and enable out flags can't be high at the same time when the memory cell has a feedback connection, the signal would interfere with it self and change its own value.
-Potential solutions are discussed in \cref{subsec:noDoubleMemcell}.
-
-\subsubsection{Cell states}
-
-Cell states have the same solution to a similar problem.
-Here the problem is that the memory cell has a feedback connection and needs to be activated within one serial step.
-The value of the old cell state needs to be used to compute the current values for the hidden states ($\overrightarrow{h_t}$) and cell states ($\overrightarrow{c_t}$).
-The issue is that at the value in the memory cell needs to be conserved until the next serial step.
-(\cref{subsec:noDoubleMemcell}).
-
-Once again, when the system is running in fully parallel mode, only one memory cell line is required for a normal behavior.
-
-\subsection{Serialization/Parallelization}
-\label{subsec:serpar}
-
-Using an \ac{LSTM} in serial mode is very beneficial as it divides the number of point wise circuit by $n_s$.
-
-There is a sweet spot for the serial size ($n_s$) :
-
-\begin{itemize}
-  \item When the serial size ($n_s)$ increases, the inference time increases with a factor of $O(n_s)$ while the onChip area decrease with the same factor.
-  \item When the serial size ($n_s)$ decreases, similarly, the onChip area increases with a factor of $O(n_s)$ while the tinference time decreases by the same factor.
-\end{itemize}
-
-The overall onChip area is linked to the number of hidden states ($n_h$) by $O(n_h^2)$.
-This means that the serial size ($n_s$) will be set depending on the limiting factor of our system (onChip area or inference time).
-
-\subsection{Symbol}
-
-The symbol of the \ac{LSTM} circuit can be found in \cref{sym:lstm}.
-It depends on several parameters, the number of inputs ($n_i$), the number of hidden states ($n_h$), the serial size of the system ($n_s$) and the number of time steps ($n_{ts}$).
-
-\begin{figure}[H]
-  \centering
-  \includesvg[height=2.5cm]{crossbar/xbarSymbol}
-  \caption{Symbol used for the \ac{LSTM} circuit}
+  \includesvg[height=2.5cm]{gru/gruSymbol}
+  \caption{Symbol used for the \ac{GRU} circuit}
   \label{sym:lstm}
@@
 \begin{equation}
-  A_{xbar}(n_i,n_o,n_s)=
-  \begin{cases}
-    2\cdot n_i\cdot n_o \cdot A_{memristor}+n_o\cdot(2\cdot[A_{opAmp}+A_{R_r}]+A_{R_f}),\hfill n_s=1\\
-    \begin{array}{2}
-      2\cdot n_i\cdot n_o \cdot n_s\cdot A_{memristor}+2\cdot A_{CMOS}\cdot n_o\cdot n_s\\
-      +n_o\cdot(2\cdot[A_{opAmp}+A_{R_r}]+A_{R_f})
+  \begin{array}{3}
+    A_{gru}(n_i,n_h)=4\cdot A_{xbar}(n_i,\frac{n_h}{n_s},n_s)\\+ \frac{n_h}{n_s}\cdot (5\cdot A_{af}+3\cdot A_{voltMult}+3\cdot A_{R_{amp0}}+2\cdot A_{R_{amp1}}+2\cdot A_{opAmp})\\+5\cdot n_h\cdot A_{memcell}
   \end{array}
-    ,\hfill \forall n_s>1\\
-  \end{cases}
 \end{equation}
-
-With $A_{CMOS}$ being the same than for the memory cell (\cref{sec:memcell}), $A_{CMOS}=1.2\cdot 10^{-14}m^2$, and the feature size of typical memristor that can be fabricated at INESC is $3\mu m$, which would make the approximate area for a memristor $A_{memristor}=9\mu m^2=9\cdot 10^{-12} m^2$.%TODO check info and inesc citation

### 2023-10-23 design/dataset.tex
Going to lunch
Moved part to results

--- 2023-10-11 3.Design/dataset.tex
@@
 
-\subsubsection{Network configuration}
-
-The layers used to solve this problem are listed below :
-
-\begin{itemize}
-  \item An \ac{LSTM} with four hidden states ($n_h=4$) and an input with feature size of one and two time steps.
-  \item A Dense layer with an output size of one.
-\end{itemize}
-
-\Cref{fig:airlineModel} is a graphical representation of the model just described.
-
-\begin{figure}[H]
-  \centering
-  \includesvg[width=\textwidth]{datasets/airlineModel}
-  \caption{Model used to solve the airline passengers problem}
-  \label{fig:airlineModel}
-\end{figure}
-
-\subsubsection{Results} %TODO Move to results obviously
-
-The weights are trained using both the custom activation function (\cref{sec:genwei}) and without ones and compare their output.
-
-After training on a digital computer using the settings described in \cref{sec:genwei}.
-
 \subsection{\ac{C. elegans}}
@@
   \begin{minipage}{\columnwidth}
-    \subfloat[Training dataset sample]{\includesvg[width=0.5\textwidth, pretex=\tiny]{datasets/celegans/ioTrain0}}%
+    \subfloat[Training dataset sample\label{graph:io5Celegans}]{\includesvg[width=0.5\textwidth, pretex=\tiny]{datasets/celegans/ioTrain0}}%
     \hfill
@@
 The inputs for the systems are then going to be the current values for the four input neurons at every of the thousand time steps and plans to get an output as close as possible to the modeled output for all time steps.
-
-\subsubsection{Network configuration}
-
-The layers used to solve this problem are listed below :
-
-\begin{itemize}
-  \item An \ac{LSTM} with eight hidden states and an input with feature size of four (because of the four input neurons that are considered), one thousand time steps and every time steps outputs a value for the next layer.
-  \item A time distributed dense layer with an output size of four. It is time distributed because it needs to compute every time steps and not only the last.
-\end{itemize}
-
-The \ac{LSTM} was chosen with a number of hidden states of eight because it gets the best results \cite{celegans}.
-
-\Cref{fig:celegansModel} is a graphical representation of the model just described.
-
-\begin{figure}[H]
-  \centering
-  \includesvg[width=\textwidth]{datasets/celegansModel}
-  \caption{Model used to solve the \ac{C. elegans} problem}
-  \label{fig:celegansModel}
-\end{figure}

### 2023-10-23 results/airline.tex
Moved part to results

@@
 \label{sec:resAirline}
+
+\subsubsection{Network configuration}
+
+The layers used to solve this problem are listed below :
+
+\begin{itemize}
+  \item An \ac{LSTM} with four hidden states ($n_h=4$) and an input with feature size of one and two time steps.
+  \item A Dense layer with an output size of one.
+\end{itemize}
+
+\Cref{fig:airlineModel} is a graphical representation of the model just described.
+
+\begin{figure}[H]
+  \centering
+  \includesvg[width=\textwidth]{datasets/airlineModel}
+  \caption{Model used to solve the airline passengers problem}
+  \label{fig:airlineModel}
+\end{figure}
+
+\subsubsection{Timing}
+
+The timing of the flags are quite hard to keep track of, for this reason all of them are going to be shown for the full duration of the system execution to get one output.
+
+\Cref{tim:airline} shows the time diagram of the entire execution using the parameters discussed earlier.
+
+\begin{figure}[H]
+  \centering
+  \begin{tikztimingtable}%TODO change or find other way
+    $e_0$       & x 2H 6L     L 2H 6L     L L x\\
+    $e_1$       & x 2L 2H 4L  L 2L 2H 4L  L L x\\
+    $e_2$       & x 4L 2H 2L  L 4L 2H 2L  L L x\\
+    $e_3$       & x 6L 2H     L 6L 2H     L L x\\
+    $e_{next}$  & x 8L        H 8L        H L x\\
+    $e_{in}$    & x 8H        L 8H        L L x\\
+    $e_{out}$   & x 8L        L 8L        L H x\\
+    \extracode
+    \tablerules
+    \begin{pgfonlayer}{background}
+      %\vertlines[help lines]{0.55,10.55}
+      %\vertlines[red]{1.6,5.6,15.6}
+      %\vertlines[blue]{3.6,9.6,15.6}
+    \end{pgfonlayer}
+  \end{tikztimingtable}
+  \caption{Flags time diagram for the airline problem with $n_s=4$}
+  \label{tim:airline}
+\end{figure}
 

### 2023-10-23 results/celegans.tex
Moved part to results

@@
 \label{sec:resCelegans}
+
+\subsubsection{Network configuration}
+
+The layers used to solve this problem are listed below :
+
+\begin{itemize}
+  \item An \ac{LSTM} with eight hidden states and an input with feature size of four (because of the four input neurons that are considered), one thousand time steps and every time steps outputs a value for the next layer.
+  \item A time distributed dense layer with an output size of four. It is time distributed because it needs to compute every time steps and not only the last.
+\end{itemize}
+
+The \ac{LSTM} was chosen with a number of hidden states of eight because it gets the best results \cite{celegans}.
+
+\Cref{fig:celegansModel} is a graphical representation of the model just described.
+
+\begin{figure}[H]
+  \centering
+  \includesvg[width=\textwidth]{datasets/celegansModel}
+  \caption{Model used to solve the \ac{C. elegans} problem}
+  \label{fig:celegansModel}
+\end{figure}
 

### 2023-10-23 results/celegans.tex
Added time diagram for celegans

@@
 The results will be analyzed by focusing on a single data sequence, the training data sequence available in \cref{graph:io5Celegans}.
+
+\subsubsection{Timing}
+
+The timing of the flags are quite hard to keep track of, for this reason all of them are going to be shown for the full duration of the system execution to get one output.
+
+\Cref{tim:airline} shows the time diagram of the entire execution using the parameters discussed earlier.
+
+\begin{figure}[H]
+  \centering
+  \begin{tikztimingtable}%TODO change or find other way
+    $e_0$       & x 4H    L 4H    N(A1) 3L N(A2) 4H    x\\
+    $e_{next}$  & x 4L    H 4L    N(B1) 3H N(B2) 4L    x\\
+    $e_{in}$    & x 4H    L 4H    N(C1) 3L N(C2) 4H    x\\
+    $e_{out}$   & x 2H 2L L 2H 2L N(D1) 3L N(D2) 2H 2L x\\
+    \extracode
+    \node[gap, at={($(A1|-A2)!0.5!(A2)$)}];
+    \node[gap, at={($(B1|-B2)!0.5!(B2)$)}];
+    \node[gap, at={($(C1|-C2)!0.5!(C2)$)}];
+    \node[gap, at={($(D1|-D2)!0.5!(D2)$)}];
+    \tablerules
+    \begin{pgfonlayer}{background}
+      \vertlines[help lines]{0.55,4.55,5.55,9.55,12.55,16.55}
+      %\vertlines[red]{1.6,5.6,15.6}
+      %\vertlines[blue]{3.6,9.6,15.6}
+    \end{pgfonlayer}
+  \end{tikztimingtable}
+  \caption{Flags time diagram for the celegans problem with $n_s=1$}
+  \label{tim:airline}
+\end{figure}
 

### 2023-10-23 results/celegans.tex
Going home, headache ...

@@
 \end{figure}
-
-Due do the complexity of this dataset and the results associated are only trained using the analog activation function.
-Furthermore, the usefulness of using those activation functions was already demonstrated in \cref{sec:resAirline}.
-
-The results will be analyzed by focusing on a single data sequence, the training data sequence available in \cref{graph:io5Celegans}.
 
@@
 \label{subsec:digitalCelegans}
+
+The digital results of the \ac{C.
+elegans} dataset are quite numerous, not all of them can be displayed and analyzed.
+For this reason, only a selected four sequences will be displayed as the results of the dataset are not the focus of this thesis.
+The full work around the \ac{C.
+elegans} dataset and problem is done in \cite{celegans}.
+
+\begin{figure}[H]
+  \centering
+  \begin{minipage}{\columnwidth}
+    \subfloat[Analog prediction for the DB1 neuron\label{graph:digital5BD1Celegans}]{\includesvg[width=0.5\textwidth, pretex=\tiny]{results/celegans/digital5DB1}}%
+    \hfill
+    \subfloat[Analog prediction for the LUAL neuron\label{graph:digital5LUALCelegans}]{\includesvg[width=0.5\textwidth,pretex=\tiny]{results/celegans/digital5LUAL}}%
+  \end{minipage}
+  \begin{minipage}{\columnwidth}
+    \subfloat[Analog prediction for the PVR neuron\label{graph:digital5PVRCelegans}]{\includesvg[width=0.5\textwidth, pretex=\tiny]{results/celegans/digital5PVR}}%
+    \hfill
+    \subfloat[Analog prediction for the VB1 neuron\label{graph:digital5VB1Celegans}]{\includesvg[width=0.5\textwidth,pretex=\tiny]{results/celegans/digital5VB1}}%
+  \end{minipage}
+  \caption{\ac{C. elegans} digital responses with their digital counterparts spread out for each output neuron}
+  \label{graph:spread5Celegans}
+\end{figure}
+
+Due do the complexity of this dataset and the results associated are only trained using the analog activation function.
+Furthermore, the usefulness of using those activation functions was already demonstrated in \cref{sec:resAirline}.
+
+The results will be analyzed by focusing on a single data sequence, the training data sequence available in \cref{graph:io5Celegans}.
 

### 2023-10-24 state/gru.tex
Bunch of changes, added design gru
Gru changes

@@
-\section{\acs{GRU}}
+\section{\acs{GRU}}\label{sec:gru}
 The \acf{GRU} is another type of \ac{RNN}. It is also known to reduce the effect of the vanishing gradient problem. It was first introduced to improve translation techniques \cite{gru}.
@@
 Indeed, the subtype of \ac{LSTM} it has been associated with is called the \ac{CIFG} \ac{LSTM}.
-If the input and forget gates are combined into one ($f_t=\overrightarrow{1}-i_t$), they can be assimilated as the update gate of the \ac{GRU}.
+If the input and forget gates are combined into one ($\overrightarrow{f_t}=\overrightarrow{1}-\overrightarrow{i_t}$), they can be assimilated as the update gate of the \ac{GRU}.
 The reset gate would then be the \ac{LSTM}'s output gate.

### 2023-10-24 0.Intro/abstract.tex
correction ruxandra

@@
 An analog \ac{ASIC} could then be fabricated for specific embedded applications.
-This would definitely improve digital technologies in terms of time, and maybe in power consumption as well.
-Which are two prerequisite for embedded system use.
+This would definitely improve digital technologies in terms of time, and maybe in power consumption as well, these being two prerequisite for embedded system use.
   For this reason this work aims at creating a functioning software simulation of analog circuit capable of running \ac{AI}.

### 2023-10-24 0.Intro/acknowledgments.tex
correction ruxandra

@@
   I would like to thank the \ac{IST} for allowing to join their double degree program, that made it possible for me to widen my computer knowledge from computer science to electronics.
-  I would also like to thank \ac{INESC} to have let me join their rank, to both \ac{INESC}-Id and \ac{INESC}-MN groups to have given the necessary tools to succeed with this thesis.
+  I would also like to thank \ac{INESC} to have let me join their rank, to both \ac{INESC}-ID and \ac{INESC}-MN groups to have given the necessary tools to succeed with this thesis.
   I would also like to thank all the other student working in the \ac{ASIC} group with whom I could share problems and were happy to help.
-  I would lastly like to thank Dr Ruxandra Barlulescu and Dr Diogo Caetano, to have trusted with this very interresting topic despite my recent arrival in the univerty.
+  I would lastly like to thank Dr Ruxandra Barbulescu and Dr Diogo Caetano, to have trusted me with this very interesting topic despite my recent arrival in the university.
 

### 2023-10-24 introduction/motivation.tex
correction ruxandra

--- 2023-10-17 1.Introduction/motivation.tex
@@
 
-Nowadays, it's impossible not to have heard about \acp{AI}, they are everywhere and everybody talks about them.
+Nowadays, it is impossible not to have heard about \acp{AI}, they are everywhere and everybody talks about them.
 It is common practice to advertise products using \ac{AI} powered tools.
@@
 Those products (such as \textbf{Amazon}'s \textit{Alexa}, \textbf{Google}'s \textit{nest} system and others) are not computing their answers locally, hence creating avoidable internet trafic.
-It's clear that \ac{AI} is becoming the predominant technology in the modern world and future innovations will depend on it.
+It is clear that \ac{AI} is becoming the predominant technology in the modern world and future innovations will depend on it.
 
-\acp{AI} are very power hungry algorithms which limits its use in embedded systems and power efficient devices.
-Furthermore, the time an \ac{AI} takes to compute its output quickly rises, especially on lower end chip such as the ones used in embedded systems.
+\acp{AI} are very power hungry algorithms which limits their use in embedded systems and power efficient devices.
+Furthermore, the time an \ac{AI} algorithm takes to compute its output quickly rises, especially on lower end chip such as the ones used in embedded systems.
 So much so that the most complex ones only run on online and very powerful servers, like the aforementioned \textit{ChatGPT} and home assistants.
@@
 There are several options to reduce execution speed and energy consumption such as running the algorithm on a \ac{GPU} rather than on a \ac{CPU}.
-Using an \ac{FPGA} or \ac{ASIC} are other ways to improve power consumption and execution time.
-The latter is the most restricting, as it's name implies, but has the best results.
+Using a \ac{FPGA} or \ac{ASIC} are other ways to improve power consumption and execution time.
+The latter is the most restricting, as it is name implies, but has the best results.
 
-Using an \ac{ASIC} allows the use of analog computation. Analog computation offers great advantages compared to digital computers as it's much faster while being known to be very power-efficient.
+Using an \ac{ASIC} allows the use of analog computation. Analog computation offers great advantages compared to digital computers as it is much faster while being known to be very power-efficient.
 
@@
 Another great use for local \ac{AI} computation would be to use \acp{LLM}, the technology behind \textit{ChatGPT}, for any local use.
-\acp{LLM} are not the focus of the thesis, but are a long term potential objective of using analog computation for \ac{AI}.
+\acp{LLM} are not the focus of this thesis, but are a long term potential objective of using analog computation for \ac{AI}.
 
-Making an analog chip with a very low power consumption also allows to use chip like that in embedded systems such as video surveillance cameras.
+Making analog chips with a very low power consumption also allows to use them in embedded systems such as video surveillance cameras.
 A small chip could be installed inside the system that could be able to do gait detection \cite{gaitDS,gaitDig,gait}.
-The video surveillance camera would then be able to know who is showing up on the camera and if it's not a known person send an alert.
+The video surveillance camera would then be able to know who is showing up on the camera and if it is not a known person send an alert.
 
@@
 It would allow to directly store stabilized video, and not have to use online servers like described in \cite{videoStab}.
-Any action camera would benefit greatly from such an improvement since they're mostly used with a lot of movements.
+Any action camera would benefit greatly from such an improvement since they are mostly used to film scenes with lots of movements.
 
 Another use for \ac{AI} that works great in embedded use is for camera relocalization \cite{videoReloc}.
-That is useful especially for VR headsets that need to know their location in the room.
+That is useful especially for \ac{VR} headsets that need to know their location in the room.
 The embedded chip would allow this computation to be much faster than other methods, and thus give more realistic results.
@@
 This could be applied on larger organisms, or, in the long term, reproduce part of the human brain.
-Using analog computers, small chips could be used to replace a part of one's brain non functionnal part.
+Using analog computers, small chips could be used to replace a part of one's brain non functional part.
 
-The focus of this thesis is to create a working simulation of an analog \ac{LSTM} circuit using Cadence's virtuoso as simulation software.
+The focus of this thesis is to create a working simulation of an analog \ac{LSTM} circuit using Cadence's Virtuoso as simulation software.
 Having those preliminary results would allow to fabricate and test such a chip.

### 2023-10-24 introduction/contributions.tex
Started contibs

@@
 
-In the thesis the following points will worked on in order to create, improve or clarify any of those topics :
+The thesis bring contributions to the following points:
 
 \begin{itemize}
-\item A script that can generate spice netlists for \acps{LSTM} or feed forward \aclp{NN} of any size and importing weights as resistance values from a specified file containing the architecture and the weights.
-\item An \ac{LSTM} capable analog circuit.
-\item Analog activation function circuit that supports \ac{tanh} and sigmoid ranges.
+  \item I created a script that can generate spice netlists for \acps{LSTM} or feed forward \aclp{NN} of any size and importing weights as resistance values from a specified file containing the architecture and the weights.
+  \item An \ac{LSTM} capable analog circuit was created and ran in virtuoso giving encouraging results.
+  \item I developed an analog activation function circuit that supports \ac{tanh} and sigmoid ranges working at $1.8V$.
   \item A \ac{VMM} capable analog circuit using a matrix of memristors.

### 2023-10-25 introduction/contributions.tex
Contributions corrected

@@
   \item I created a script that can generate spice netlists for \acps{LSTM} or feed forward \aclp{NN} of any size and importing weights as resistance values from a specified file containing the architecture and the weights.
-  \item An \ac{LSTM} capable analog circuit was created and ran in virtuoso giving encouraging results.
+  \item The script can generate an \ac{LSTM} capable analog circuit was created and ran in virtuoso giving encouraging results.
+  \item The script can also generate an \ac{GRU} capable analog circuit was created and ran in virtuoso.
   \item I developed an analog activation function circuit that supports \ac{tanh} and sigmoid ranges working at $1.8V$.
-  \item A \ac{VMM} capable analog circuit using a matrix of memristors.
-  \item Memory cells to store analog data for a short amount of time.
-  \item A python (tensorflow library) code to use piece wise functions, to reproduce in python the analog activation functions.
-  \item A way to run software simulation of different \acp{NN}.
+  \item I recreated a \acp{VMM} capable analog circuit using a matrix of memristors.
+  \item I designed a memory cells used to store analog data for a short amount of time, they are used the \ac{LSTM} and \ac{GRU} circuits.
+  \item I coded in python \cite{python} (tensorflow library\cite{tensorflow}) to create a piece wise functions to reproduce in the tensorflow the analog activation functions.
+  \item I used Cadence's Virtuoso to run software simulation of different \acp{NN} capable analog circuits.
 \end{itemize}

### 2023-10-25 introduction/contributions.tex
Forgot s s

@@
   \item I created a script that can generate spice netlists for \acps{LSTM} or feed forward \aclp{NN} of any size and importing weights as resistance values from a specified file containing the architecture and the weights.
-  \item The script can generate an \ac{LSTM} capable analog circuit was created and ran in virtuoso giving encouraging results.
-  \item The script can also generate an \ac{GRU} capable analog circuit was created and ran in virtuoso.
+  \item The script can generate an \ac{LSTM} capable analog circuits was created and ran in virtuoso giving encouraging results.
+  \item The script can also generate an \ac{GRU} capable analog circuits was created and ran in virtuoso.
   \item I developed an analog activation function circuit that supports \ac{tanh} and sigmoid ranges working at $1.8V$.

### 2023-10-25 introduction/outline.tex
corrected outline

@@
 
-All the technologies used and, thus are required to know in order to understand the thesis are detailed in the state of the art.
+All the technologies used and thus required to know in order to understand the thesis are detailed in the state of the art.
 
-Next the work that was done for the thesis is goign to be thoroughly explained.
+Next the work that was done for the thesis is going to be thoroughly explained.
 Starting with the different circuit that were designed for the thesis.
 Then the different scripts and python code that were created, either to run the digital \acp{NN} or to generate the analog system into a netlist.
+
+The following chapter contains the results, obtained from the digital inference and the analog simulation, from both the number of airline passenger and the \ac{C. elegans} datasets.
+
+Finally the conclusions and improvements that can be made to the different parts of the thesis.

### 2023-10-25 state/ai.tex
Renamed files and folder
Corrected state of the art

--- 2023-10-17 2.State/ai.tex
@@
 
-\acf{AI} is very popular, and is getting more attention, being used as selling point by lots of services, the word is thrown around so much that it has lost its meaning. So what are \acp{AI} exactly.
+\acf{AI} is very popular, and is getting more attention, being used as selling point by lots of services, the word is thrown around so much that it has lost its meaning.
+So let us see what are \acp{AI} exactly.
 
-\ac{AI} is defined as theories and stragies used to simulate human intelligence, in other words, a computer or software that can take a decision by itself based on previously set rules.
-The definition is very wide and have lots of ways to be implemented.
+\ac{AI} is defined as theories and strategies used to simulate human intelligence, in other words, a computer or software that can take a decision by itself based on previously set rules.
+The definition is very wide and has lots of ways to be implemented.
 
-The most promising type of \ac{AI} is machine learning, which consists of teaching the software to make decision.
+The most promising type of \ac{AI} is machine learning, which in a broad sense means teaching the software to make decisions.
 It is defined as the use and development of computer systems that are able to learn and adapt without following explicit instructions, by using algorithms and statistical models to analyse and draw inferences from patterns in data.
@@
   \centering
-  \includesvg[width=\textwidth]{ml.svg}
+  \includesvg[width=\textwidth,pretex=\small]{ml.svg}
   \caption{Machine learning and its subsets}

### 2023-10-25 state/crossbar.tex
Renamed files and folder
Corrected state of the art

--- 2023-09-19 2.State/crossbar.tex
@@
 
-Setting memristors in a crossbar array allows to perform analog \ac{VMM} also called Multiply and Accumulate.
+Setting memristors in a crossbar array allows to perform analog \ac{VMM}, also called Multiply and Accumulate.
 This circuit works great for this use as all the computation is done almost instantly and using physics laws.
@@
 
-It uses physical properties of electrical systems to perform analog computing. Lets use the circuit node in figure \ref{fig:crossNode} to explain the theory behind the memristor crossbar array.
+It uses physical properties of electrical systems to perform analog computing.
+In what follows I will use the circuit node in figure \ref{fig:crossNode} to explain the theory behind the memristor crossbar array.
 \begin{figure}[H]
@@
 First of all, a voltage is applied on the $i^{th}$ line, because every column is virtually grounded, the voltage applied to the memristor, of a set resistance $R_i$, is $V_i$.
-As such and using Ohm's law, we know the the current flowing into the memristor ($I_{mem}$) is bound by the following equation :
+As such and using Ohm's law, we know that the current flowing into the memristor ($I_{mem}$) is bound by the following equation :
 
@@
 
-This line then joins the column where a current of $I_{i,j-1}$ is flowing, then according to kirchhoff's current law the resulting current is :
+This line then joins the column where a current of $I_{i,j-1}$ is flowing, then according to Kirchhoff's current law the resulting current is :
 \begin{equation}
@@
 \end{equation}
-By applying this process to the whole system we find that the current as the bottom of one column is :
+By applying this process to the whole system we find that the current at the bottom of one column is :
 \begin{equation}
@@
 \begin{itemize}
-  \item This circuit being an analog it allows for analog computations which are much faster than their digital counter parts.
+  \item By being analog, this circuit allows for analog computations which are much faster than their digital counterparts.
 As explained in this section, the circuit only uses physical properties of the different components and can compute large \acp{VMM} in a very short amount of time.

### 2023-10-25 state/gru.tex
Corrected state of the art

@@
 
-The \ac{GRU} is very often compared to the \ac{LSTM}, it is sometimes assimilated as a type of \ac{LSTM} \cite{nbLSTM}.
+The \ac{GRU} is very often compared to the \ac{LSTM}, being sometimes assimilated as a type of \ac{LSTM} \cite{nbLSTM}.
 Their performance was found to be very similar in most situations, making those two types of \acp{RNN} coexistant in the modern machine learning world.
@@
 They were originally designed to encode the message to translate and then decode in the translation.
-PyTorch only supports the decoder version \cite{gruPyTorch}, while the keras library supports both \cite{gruKeras} chosen by changing an argument.
+PyTorch only supports the decoder version \cite{gruPyTorch}, while the Keras library supports both \cite{gruKeras} chosen by changing an argument.
 
@@
 
+A visual representation of the encoder \ac{GRU} cell is available in \cref{fig:encoderGruCell}.
+
 \begin{figure}[H]
@@
   \includesvg[width=\textwidth,pretex=\large]{gru/encoderCell.svg}
-  \label{fig:lstmCell}
+  \label{fig:encoderGruCell}
   \caption{Encoder \acs{GRU} cell, legend in \cref{leg:cells}}
@@
 
+A visual representation of the decoder \ac{GRU} cell can be found in \cref{fig:decoderGruCell}.
+
 \begin{figure}[H]
@@
   \includesvg[width=\textwidth,pretex=\large]{gru/decoderCell.svg}
-  \label{fig:lstmCell}
+  \label{fig:decoderGruCell}
   \caption{Decoder \acs{GRU} cell, legend in \cref{leg:cells}}

### 2023-10-25 state/lstm.tex
Corrected state of the art

--- 2023-10-17 2.State/lstm.tex
@@
 \section{\acs{LSTM}}\label{sec:lstm}
-\acfp{LSTM} are a type of \ac{RNN} used to analyze sequence of data.
-They are capable of predict data based of the previous results.
-\acp{LSTM} are part of \acp{RNN}.
-\acp{RNN} are different than feedfoward \acl{NN} because of their feedback connections.
-Meaning that the results from the last time step have an impact on the next time step.
+\acfp{LSTM} are a type of \ac{RNN} used to analyze sequences of data. They are capable of predicting data based on previous data points.
 
@@
 \acp{LSTM} was created to fix the vanishing gradient problem.
-\acp{LSTM} alleviate this issue by adding a cell state, this state gives it the ability to choose what input is important and which one isn't, thus giving it a long term memory.
+\acp{LSTM} alleviate this issue by adding a cell state, this state gives it the ability to choose what input is important and which one is not, thus giving it a long term memory.
 This ability gave the uncommon name of \acl{LSTM} as it has both long and short term memory.
-This what gives \acp{LSTM} their use for sequence data.
+This is what makes \acp{LSTM} adequate for sequence data.
 They can analyze the data and keep the information from the last time step to make a better decision afterwards.
 The most comprehensible example is considering a sentence.
-(TODO : find example)
+%TODO : find example)
 
-An \ac{LSTM} is more complicated than just a simple feedforward \acl{NN}, they have several gates, which are all technically a \ac{NN} themselves.
-There is also a cell state which job is to hold a value for the next step.
+An \ac{LSTM} is more complicated than just a simple feedforward \acl{NN}, it has several gates, which all compute a \ac{VMM}. There is also a cell state whose job is to hold a value for the next step.
 
@@
 Each input to the cell is a vector.
-Those vector are of varying sizes depending on several factors.
-For example, both $h_t$ and $c_t$ are of the same size as the number of hidden state (sometimes refered to as cell state) for any $t\geq 0$.
+Those vectors are of varying sizes depending on several factors.
+For example, both $h_t$ and $c_t$ are of the same size as the number of hidden states (sometimes referred to as cell state) for any $t\geq 0$.
 The input vector ($x_t$) is of size of the sample we want to feed for each time step.
@@
 Let's start with the more classic gate equations. They are the ones that behave like the more classic \ac{NN}.
-The input (\cref{eq:inputG}), forget (\cref{eq:forgetG}) and ouput (\cref{eq:outputG}) gates are described below.
+The input (\cref{eq:inputG}), forget (\cref{eq:forgetG}) and output (\cref{eq:outputG}) gates are described below.
 
@@
 
-Where ($w_i$,$\overrightarrow{b_i}$), ($w_f$,$\overrightarrow{b_f}$) and ($w_o$,$\overrightarrow{b_o}$) are the pair of weights matrixes and biases vectors for the input, forget and ouput gates respectively.
+Where ($w_i$,$\overrightarrow{b_i}$), ($w_f$,$\overrightarrow{b_f}$) and ($w_o$,$\overrightarrow{b_o}$) are the pair of weights matrixes and biases vectors for the input, forget and output gates respectively.
 $\overrightarrow{x_t}$ is the input vector and $\overrightarrow{h_t}$ is the hidden state vector.
 
-For the next equation describes the candidate cell state (\cref{eq:candCell}), that will next become the cell state (\cref{eq:cellS}).
+The next equation describes the candidate cell state (\cref{eq:candCell}), that will next be used to compute the cell state (\cref{eq:cellS}).
 
@@
 Using \ac{LSTM} can be a bit tricky.
-Due to it's sequential nature, it takes several time steps.
+Due to its sequential nature, it takes several time steps.
 Every hidden state ($h_t$) is passed to the next time step as an input.
 This hidden state can be used to compute an output at any time step $t$.
+\Cref{fig:lstmUse} shows a visual representation of an \ac{LSTM} going from the current time step to the following time step.
 
@@
   \includesvg[width=\textwidth]{lstm/lstmUse}
-  \caption{Unfolded \acs{LSTM}, legend in \cref{leg:lstm}}
+  \caption{Unfolded \acs{LSTM}, legend in \cref{leg:cells}}
   \label{fig:lstmUse}
@@
 
-\acp{LSTM} come in reality in a few different flavor.
+\acp{LSTM} come in a few different flavors of implementations.
 Usually, when \acp{LSTM} are mentionned, the version used is the \ac{NP} version.
-This is the most common version as those are the equations described on the wikipedia page \cite{wikiLSTM} and used in librairies like Keras \cite{Keras} or pyTorch \cite{PyTorch}.
-In fact, there are total of 8 other variations of \acp{LSTM} \cite{nbLSTM}.
-The difference between each of them varies from one to the other, but they can all be described with new equations.
+This is the most common version as those are the equations described on the wikipedia page \cite{wikiLSTM} and used in libraries like Keras \cite{Keras} or PyTorch \cite{PyTorch}.
+In fact, there are at least 8 other variations of \acp{LSTM} \cite{nbLSTM}.
+The difference between them varies from one to the other, some of them are detailled next.
 
 \subsubsection{Vanilla \ac{LSTM}}
-This variant of the \ac{LSTM} was origanally the only \ac{LSTM} network, hence it's name.
-It differs from the classic \ac{LSTM} with it's use of peephole weights, hence the name of the classic \ac{LSTM} being \acl{NP}.
-The Vanilla \ac{LSTM} is thus defined by those equations \cite{vanillaLSTM, nbLSTM} :
+This variant of the \ac{LSTM} was originally the only \ac{LSTM} network, hence its name.
+It differs from the classic \ac{LSTM} with its use of peephole weights, hence the name of the classic \ac{LSTM} being \acl{NP}.
+The Vanilla \ac{LSTM} is thus defined by the following equations \cite{vanillaLSTM, nbLSTM} :
 
@@
 
-With $R_{fi}$ being the weight matrix of the input feedback connection to compute the forget gate, and the same goes for the other weight matrices.
+With $R_{fi}$ being the weight matrix of the input feedback connection to compute the forget gate, and the same goes for the other weight matrixes.
 

### 2023-10-25 state/main.tex
Forgot to rename the folders in the mains...
Corrected state of the art

--- 2023-10-17 2.State/main.tex
@@
 
-This chapter will go over the several subject that need to be known in order to understand the thesis.
+This chapter will go over the several subjects that need to be known in order to understand the thesis.
 
@@
       \item When used with scalars, it is the sum of those scalars.
-      \item When used with vectors (column matrixes), it is the point wise sum, basically adding each element to each other. The result is a vector of the same size.
+      \item When used with vectors (vertical matrixes), it is the point wise sum, basically adding each element to each other. The result is a vector of the same size.
     \end{itemize}
@@
 
-\input{2.State/ai.tex}
-\input{2.State/nn.tex}
-\input{2.State/rnn.tex}
-\input{2.State/lstm.tex}
-\input{2.State/gru.tex}
-\input{2.State/memristors.tex}
-\input{2.State/crossbar.tex}
-\input{2.State/model.tex}
+\input{state/ai.tex}
+\input{state/nn.tex}
+\input{state/rnn.tex}
+\input{state/lstm.tex}
+\input{state/gru.tex}
+\input{state/memristors.tex}
+\input{state/crossbar.tex}
+\input{state/model.tex}
 

### 2023-10-25 state/memristors.tex
Renamed files and folder
Corrected state of the art

--- 2023-10-16 2.State/memristors.tex
@@
 \Cref{fig:fundComp} shows the four fundamental variables are on each side of the square, with the ones on opposite sides being linked by the following equations :
+
 \begin{equation}
@@
 
-Resistors, capacitors and inductors were already very established and well known component, so it was theorized that a fourth device should then exists to physically link flux ($\phi$) and charge ($q$).
-The flux in this case is not a magnetic flux and is defined as such : $ d\phi=v\cdot dt \Rightarrow \phi =  \int v \,dt  $.\\
-The component stayed theoretical until 2008 when it's been implemented in a physical device for the first time \cite{memristorFab}. It took 37 years to have an actual working device.\\
-There is then an extention of this theoretical device to another, the memristive device.
+Resistors, capacitors and inductors were already very established and well known components, so it was theorized that a fourth device should then exist to physically link flux ($\phi$) and charge ($q$).
+The flux in this case is not a magnetic flux and is defined as such : $ d\phi=v\cdot dt \Rightarrow \phi =  \int v \,dt  $.
+
+The component stayed theoretical until 2008 when it was implemented in a physical device for the first time \cite{memristorFab}. It took 37 years to have an actual working device.
+
+There is then an extension of this theoretical device to another, the memristive device.
 It was theorized in 1976 by L.
@@
 Kang \cite{memrestiveDev}.
-The difference between the memristor and the memristive device is its internal behavior.
-Memristive device are commonly referred to as memristor as well.
+The difference between the memristor and the memristive devices is its internal behavior.
+Memristive device are commonly referred to as memristors as well.
 
@@
 \subsection{Equations}
-A memristor links the flux ($\phi$) and charge ($q$) and creates the memristance.
+A memristor links the flux ($\phi$) and charge ($q$) by the memristance.
 This memristance is defined with the following equation :
@@
 A memristive device is slightly differently defined, it still uses a memristance, but here the memristance also depends on an internal state called $x$. This gives us this equation :
-\begin{equation}
+\begin{equation}\label{eq:memristiveDev}
   v(t)=M(x,i)\cdot i(t)
@@
 \end{equation}
-In all of the previous equations, $v$ is the voltage in Volt ($V$), $i$ is the current in Ampere ($A$), $\phi$ is the flux in Weber ($Wb$), $q$ is the charge in Coulomb ($C$), $M$ is the memristance in Ohm ($\Omega$) and $W$ is the memductance in Siesmens ($S$ or $\Omega^{-1}$).
+In all of the previous equations, $v$ is the voltage in Volt ($V$), $i$ is the current in Ampere ($A$), $\phi$ is the flux in Weber ($Wb$), $q$ is the charge in Coulomb ($C$), $M$ is the memristance in Ohm ($\Omega$) and $W$ is the memductance in Siemens ($S$ or $\Omega^{-1}$).
 
@@
 This enables us to manipulate the resistance of the component.
-The huge benefit of memristors its ability to retain its internal resistance, the device can left without power for a long period of time (retention time of minimum 10 years according to \cite{memRetention}).
+The huge benefit of memristors is the ability to retain its internal resistance, the device can be left without power for a long period of time (retention time of minimum 10 years according to \cite{memRetention}).
 When the device is powered backup, it will have the same resistance it had before.
@@
 It is better than traditionnal DRAM because it keeps the information even when turned off.
-This makes ReRAM a good replacement for both RAM and HDDs/SSDs, thus eliminating Von Neumann bottleneck due to the Von Neumann architecture that is used in all modern computers.
+This makes ReRAM a good replacement for both RAM and HDDs/SSDs, thus eliminating von Neumann bottleneck due to the von Neumann architecture that is used in all modern computers.
 

### 2023-10-25 state/model.tex
Corrected state of the art

@@
 
-It's been established that the internal resistance also depends on an internal state $x$ (\cref{eq:memristiveDev}).
+It has been established that the internal resistance also depends on an internal state $x$ (\cref{eq:memristiveDev}).
 
@@
 
-\subsection{Behavior}
+\subsection{Behavior}\label{subsec:memModelBehave}
 
-The model was tested then tested.
-As with physical memristors, it has a threshold voltage value from which its resistive states starts changing.
+The model implements, as with physical memristors, a threshold voltage value from which its resistive states starts changing.
 This threshold value is $V_{read}$.
@@
 
-Every pulse is composed of 2 parts and last for a total of $1 ms$ :
+Every pulse is composed of 2 parts and lasts for a total of $1 ms$ :
 \begin{itemize}
-  \item The reading of the resistive state, by applying a triangular wave that in bound by the threshold value $V_{read}$.
-  \item The writing to the memristor. During this part the resistive state will change. This is done by apply a square wave of variable width ($t_{\Delta}$) with a set bias voltage ($V_{bias}$).
+  \item The reading of the resistive state, by applying a triangular wave that is bound by the threshold value $V_{read}$.
+The triangular wave starts from $0V$, goes to $V_{read}$ then to $-V_{read}$ before going back to $0V$.
+  \item The writing to the memristor. During this part the resistive state will change. This is done by applying a square wave of variable width ($t_{\Delta}$) with a set bias voltage ($V_{bias}$).
 \end{itemize}
@@
   \includesvg[width=\textwidth]{memristor/input.svg}
+  \caption{Pulse shape graph}
   \label{graph:memInput}
-  \caption{Pulse shape graph}
 \end{figure}
@@
   \includesvg[width=\textwidth]{memristor/memPulseChange.svg}
+  \caption{Memristor's resistive state under different pulses width (Fixed bias voltage $V_{bias}=1.8V$)}
   \label{graph:pulseChange}
-  \caption{Memristor's resistive state under different pulses width (Fixed bias voltage $V_{bias}=1.8V$)}
 \end{figure}
@@
 Indeed, the larger the pulse widths, the faster the internal resistance reaches the bounding resistance (\cref{eq:memModel4}).
-The bounding resistance being the same as all curve use the same bias voltage ($V_{bias}=1.8V$), the curves will all eventually reach the max resistance.
+The bounding resistance being the same as all curves use the same bias voltage ($V_{bias}=1.8V$), the curves will all eventually reach the max resistance.
 
@@
   \includesvg[width=\textwidth]{memristor/memVoltChange.svg}
+  \caption{Memristor's resistive state under different voltages (Fixed pulse width $t_{\Delta}=100\mu s$)}
   \label{graph:voltChange}
-  \caption{Memristor's resistive state under different voltages (Fixed pulse width $t_{\Delta}=100\mu s$)}
 \end{figure}
@@
 
-The resistive state can also be lowered by applying negative voltages.TODO
+The resistive state can also be lowered by applying negative voltages.%TODO
 

### 2023-10-25 state/nn.tex
Corrected state of the art

--- 2023-10-16 2.State/nn.tex
@@
 The model presented in \cref{fig:snn} is modular and can be scaled up as much as required.
-It is generally accepted that the more neurons a \ac{NN} has, the best it can learn and execute the given task.
+It is generally accepted that the more neurons a \ac{NN} has, the more complex the problems  it can solve can be.
 
@@
 
-The weights values are obtained through training. This is done in a few steps :
+The weights values are obtained through training.
+In supervised learning, in order train the weights it is required to have a dataset of inputs/outputs values that are correct.
+The outputs are the target values, the values that the \ac{NN} needs to compute when being executed.
+The training starts once the weights have been initialized (usually randomly picked values).
+The training is an iteravtive process, each iteration being called an epoch.
+Each of those epoch usally consists of the following steps :
 
 \begin{enumerate}
-  \item Generate random values for each weight.
   \item Run the \ac{NN} to get an output vector.\label{step:restart}
-  \item Measure error rate using the chosen error algorithm.
-  \item Use the
-  \item go back to \cref{step:restart}
+  \item Measure the error (known as loss) by comparing the current prediction with the targeted output using the chosen error algorithm (\ac{RMSE}, \ac{MSE}, etc).
+  \item Run the backpropagation algorithm to change each individual weight.
 \end{enumerate}
+
+The number of epochs required depends on the complexity of the problem the \ac{NN} is solving.
+
+Once trained, the \ac{NN} can be tested by feeding the \ac{NN} input data it hasn't seen before.

### 2023-10-25 state/rnn.tex
Corrected state of the art

--- 2023-10-17 2.State/rnn.tex
@@
 \acp{RNN} are, as the name suggests, a type of \ac{NN} using recurrent connections.
-They are a \ac{NN} with at least one cycle within the structure, where outputs of the previous run is used the next one.
-Those feedback connections are what differenciates it from feedforward \ac{NN}.
+They are a \ac{NN} with at least one cycle within the structure, outputs of previous time step are used as input for the next time step, these outputs are generally known are hidden states.
+Those feedback connections are the main difference from feedforward \ac{NN}.
 
@@
 
-Traditionnal \ac{RNN} have the ability to model sequential events by propagating through time, for example forward and backward propagation.
+Traditionnal \acp{RNN} have the ability to model sequential events by propagating through time, for example forward and backward propagation.
 This is achieved by connecting these sequential events with the hidden state like in \cref{eq:rnn}.
@@
 
-They are trained the same way \ac{NN} are, measuring the error, backpropagte and adjust the weights accordingly.
+They are trained the same way \ac{NN} are, measuring the error, backpropagate and adjust the weights accordingly.
 
@@
 The simple \ac{RNN} works just like a \ac{tanh} activated feedforward \ac{NN} with a feedback connection.
+
+\begin{figure}[H]
+  \centering
+  \begin{minipage}{\columnwidth}
+    \subfloat[\acs{RNN} cell\label{fig:rnnCell}]{\includesvg[width=\textwidth,pretex=\large]{rnn/rnnCell.svg}}
+  \end{minipage}
+  \begin{minipage}{\columnwidth}
+    \subfloat[Legend\label{leg:cells}]{\includesvg[width=\textwidth,pretex=\large]{cellsLegend.svg}}
+  \end{minipage}
+  \caption{}
+\end{figure}
 
@@
 
-The Vanishing gradient problem is a problem that comes when dealing with time depedent data \cite{vanishGrad}.
-When big amount of time depedent data is fed to the \ac{RNN}, the weights can't be updated properly.
+The Vanishing gradient problem is a problem that comes when dealing with time dependent data \cite{vanishGrad}.
+When a big amount of time dependent data is fed to the \ac{RNN}, the weights cannot be updated properly.
 The older the data, the lower it will impact how much the weight must change.
 Rendering the old input data almost useless.
-Simple \acp{RNN} must be used with relatively short time series.
+That is the reasom why, simple \acp{RNN} must be used with relatively short time series.
 
-Some \acp{RNN} were designed around this issue. This is the case of the \ac{LSTM} and \ac{GRU} which were created with internal mechanisms to regulate the flow of information and gradients.
+Some \acp{RNN} were designed to tackle this issue. This is the case of the \ac{LSTM} and \ac{GRU} which were created with internal mechanisms to regulate the flow of information and gradients.

### 2023-10-25 results/airline.tex
Changed to subsec
All afternoon work

@@
 
-\subsubsection{Network configuration}
+\subsection{Network configuration}
 
@@
     \begin{pgfonlayer}{background}
-      %\vertlines[help lines]{0.55,10.55}
+      \vertlines[help lines]{0.55,8.55,9.55,17.55,18.55}
       %\vertlines[red]{1.6,5.6,15.6}
@@
     \hline
-    Target data &\cellcolor[HTML]{202020} & $54.60$ & $52.27$\\
+    Target data &\cellcolor[HTML]{202020} & $54.6$ & $52.3$\\
     \hline
-    \specialcell{Digital prediction with\\default activation functions}  & $54.60$ & \cellcolor[HTML]{202020} & $78.64$\\
+    \specialcell{Digital prediction with\\default activation functions}  & $54.6$ & \cellcolor[HTML]{202020} & $78.6$\\
     \hline
-    \specialcell{Digital prediction with\\analog activation functions} & $52.27$ & $78.64$ & \cellcolor[HTML]{202020}\\
+    \specialcell{Digital prediction with\\analog activation functions} & $52.3$ & $78.6$ & \cellcolor[HTML]{202020}\\
     \hline
@@
 Both predictions have similar \ac{RMSE} to the target curve.
-They have a \ac{RMSE} of $54.60$ for the \ac{LSTM} using the default activation functions (sigmoid and \ac{tanh} functions) and $52.27$ for the one using the analog activation functions (\cref{sec:af}).
+They have a \ac{RMSE} of $54.6$ for the \ac{LSTM} using the default activation functions (sigmoid and \ac{tanh} functions) and $52.3$ for the one using the analog activation functions (\cref{sec:af}).
 The error difference is so low (about two thousands passengers) that the difference may just come from the training and the initial values of the weights.
@@
 
-The two predictions are very different from each other, the error rate between the two is of $78.64$, which is significantly higher than the \ac{RMSE} of the curves to the target curve.
+The two predictions are very different from each other, the error rate between the two is of $78.6$, which is significantly higher than the \ac{RMSE} of the curves to the target curve.
 
@@
     \hline
-    \specialcell{Digital prediction with\\default activation functions} & $140.68$ & $143.66$ & $139.71$\\
+    \specialcell{Digital prediction with\\default activation functions} & $141$ & $144$ & $140$\\
     \hline
@@
     \hline
-    \specialcell{Digital prediction with\\analog activation functions} & $28.41$ & $92.57$ & $68.51$\\
+    \specialcell{Digital prediction with\\analog activation functions} & $28.4$ & $92.6$ & $68.5$\\
     \hline
@@
     \hline
-    \multirow{3}{*}{Analog prediction} & $n_s=1$ &\cellcolor[HTML]{202020} & $64.61$ & $40.63$\\
+    \multirow{3}{*}{Analog prediction} & $n_s=1$ &\cellcolor[HTML]{202020} & $64.6$ & $40.6$\\
     \cline{2-5}
-    & $n_s=2$  & $64.61$ & \cellcolor[HTML]{202020} & $25.75$\\
+    & $n_s=2$  & $64.6$ & \cellcolor[HTML]{202020} & $25.8$\\
     \cline{2-5}
-    & $n_s=4$ & $40.63$ & $25.75$ & \cellcolor[HTML]{202020}\\
+    & $n_s=4$ & $40.6$ & $25.8$ & \cellcolor[HTML]{202020}\\
     \hline

### 2023-10-25 results/celegans.tex
All afternoon work

@@
 
-\subsubsection{Network configuration}
+Due do the complexity of this dataset and the results associated are only trained using the analog activation function.
+Furthermore, the usefulness of using those activation functions was already demonstrated in \cref{sec:resAirline}.
+
+\subsection{Network configuration}
 
@@
 
+The results will be analyzed by focusing on two data sequence, the first being simple inputs expecting simple outputs and the second being more complex, those sequences are the ones in \cref{graph:io5Celegans,graph:io15Celegans}.
+We'll identify those as sequence 5 and sequence 15 respectively.
+
 The digital results of the \ac{C.
 elegans} dataset are quite numerous, not all of them can be displayed and analyzed.
-For this reason, only a selected four sequences will be displayed as the results of the dataset are not the focus of this thesis.
-The full work around the \ac{C.
+For this reason, only the two previously selected sequences will be displayed as the results of the dataset are not the focus of this thesis.
+The full work concerning the \ac{C.
 elegans} dataset and problem is done in \cite{celegans}.
@@
   \end{minipage}
-  \caption{\ac{C. elegans} digital responses with their digital counterparts spread out for each output neuron}
-  \label{graph:spread5Celegans}
+  \caption{\ac{C. elegans} digital responses spread out for each output neuron}
+  \label{graph:digital5Celegans}
 \end{figure}
 
-Due do the complexity of this dataset and the results associated are only trained using the analog activation function.
-Furthermore, the usefulness of using those activation functions was already demonstrated in \cref{sec:resAirline}.
+\Cref{graph:digital5Celegans,graph:digital15Celegans} are the digital predictions of the selected dat sequences.
+They all show a decent results, their \ac{RMSE} can be found at \cref{tab:celegansDigital}.
 
-The results will be analyzed by focusing on a single data sequence, the training data sequence available in \cref{graph:io5Celegans}.
+\begin{table}[H]
+  \centering
+  \begin{tabular}{|c|c|c|c|c|}
+    \hline
+    \cellcolor[HTML]{808080}\acs{RMSE} & BD1 & LUAL & PVR & VB1\\
+    \hline
+    Sequence 5 & $5.02\cdot 10^{-2}$ & $1.68\cdot 10^{-1}$ & $1.32\cdot 10^{-1}$ & $5.07\cdot 10^{-2}$\\
+    \hline
+    Sequence 15 & $6.98\cdot 10^{-2}$ & $1.39\cdot 10^{-1}$ & $1.86\cdot 10^{-1}$ & $6.63\cdot 10^{-2}$\\
+    \hline
+  \end{tabular}
+  \caption{\acp{RMSE} of each neuron's prediction for the two selected sequences}
+  \label{tab:celegansDigital}
+\end{table}
 
-\subsection{Analog results}
-\label{subsec:analogCelegans}
+A pattern can be observed here, the error rates are lower for the DB1 and VB1 neurons.
+
+\begin{figure}[H]
+  \centering
+  \begin{minipage}{\columnwidth}
+    \subfloat[Analog prediction for the DB1 neuron\label{graph:digital15BD1Celegans}]{\includesvg[width=0.5\textwidth, pretex=\tiny]{results/celegans/digital15DB1}}%
+    \hfill
+    \subfloat[Analog prediction for the LUAL neuron\label{graph:digital15LUALCelegans}]{\includesvg[width=0.5\textwidth,pretex=\tiny]{results/celegans/digital15LUAL}}%
+  \end{minipage}
+  \begin{minipage}{\columnwidth}
+    \subfloat[Analog prediction for the PVR neuron\label{graph:digital15PVRCelegans}]{\includesvg[width=0.5\textwidth, pretex=\tiny]{results/celegans/digital15PVR}}%
+    \hfill
+    \subfloat[Analog prediction for the VB1 neuron\label{graph:digital15VB1Celegans}]{\includesvg[width=0.5\textwidth,pretex=\tiny]{results/celegans/digital15VB1}}%
+  \end{minipage}
+  \caption{\ac{C. elegans} digital responses spread out for each output neuron}
+  \label{graph:digital15Celegans}
+\end{figure}
+
+\subsection{Analog results of sequence 5}
+\label{subsec:analog5Celegans}
 
@@
 
-A big problem that is immediatly visible is that the LUAL output neuron's analogly predicted response is negative when it should not be.
-For an unknown reason the analog results of the LUAL neuron seem to be negative instead of positive.
-This is the case for the analog response to all sequences.
-This issue most likely comes from dense layer that comes after the \ac{LSTM}.
-In order to keep using the data the output of this neuron will be flipped over the x-axis.
+All curves seem to all have a similar issue, the signal is late to go up.
+However, the reason for this behavior is still unknown.
+This can be adjusted manually in the graph until the reason for this issue is found.
 
@@
 
-The smoothed out and exploitable analog predictions are available in \cref{graph:smooth5Celegans}. Theses graphs allow for a much easier visualisation of the results.
+The adjusted (smoothed out and slided to the left to match the digital prediction best) and exploitable analog predictions are available in \cref{graph:smooth5Celegans}.
+Theses graphs allow for a much easier and better visualisation of the results.
 
-The best looking output of the four predicted neurons responses is the one from PVR.
-That is confirmed by its error rate to the digital prediction.
-It is the lowest of the four, however, it isn't close enough to the digital prediction.
+\begin{figure}[H]
+  \centering
+  \includesvg[width=\linewidth]{results/celegans/rmse5}
+  \caption{\acp{RMSE} of each neuron's analog prediction and adjusted analog prediction for sequence 5}
+  \label{bar:rmse5}
+\end{figure}
 
-All four analog predictions are very close to each others
+The \cref{bar:rmse5} shows that adjusted or not, the errors are in the expected range.
+Adjusting the results reduces the error a bit, mainly due to the sliding of the results.
+Smoothing out the curve doesn't affect the error as it simply does a local average of each value.
+The erros are, once again higher when obtained with the analog simulation than with the digital inference.
 
@@
   \end{minipage}
-  \caption{\ac{C. elegans} smoothed out analog responses averaged with $21$ neighbors, spread out for each output neuron}
+  \caption{\ac{C. elegans} smoothed out and slided analog responses averaged with $21$ neighbors, spread out for each output neuron}
   \label{graph:smooth5Celegans}
 \end{figure}
+
+\subsection{Analog results of sequence 15}
+\label{subsec:analog15Celegans}
+
+\begin{figure}[H]
+  \centering
+  \includesvg[width=\linewidth]{results/celegans/out15}
+  \caption{\acs{C. elegans} analog responses with their digital counterparts}
+  \label{graph:celegansAnalog0}
+\end{figure}
+
+\Cref{graph:celegansAnalog0} contains the predictions of both the digital and analog \ac{NN} for all four output neurons that are being monitored.
+The having all predicted responses to the stimuli on the same graph makes things very unintelligeable.
+
+The response have thus been sperated into four graphs.
+One for every neuron for better readability.
+The target response was also added to all four graphs to highlight the objective of the \ac{NN}.
+Those graphs can be found in \cref{graph:spread15Celegans}.
+
+All curves seem to all have a similar issue, the signal is late to go up.
+However, the reason for this behavior is still unknown.
+This can be adjusted manually in the graph until the reason for this issue is found.
+
+\begin{figure}[H]
+  \centering
+  \begin{minipage}{\columnwidth}
+    \subfloat[Analog prediction for the DB1 neuron\label{graph:out15BD1Celegans}]{\includesvg[width=0.5\textwidth, pretex=\tiny]{results/celegans/out15DB1}}%
+    \hfill
+    \subfloat[Analog prediction for the LUAL neuron\label{graph:out15LUALCelegans}]{\includesvg[width=0.5\textwidth,pretex=\tiny]{results/celegans/out15LUAL}}%
+  \end{minipage}
+  \begin{minipage}{\columnwidth}
+    \subfloat[Analog prediction for the PVR neuron\label{graph:out15PVRCelegans}]{\includesvg[width=0.5\textwidth, pretex=\tiny]{results/celegans/out15PVR}}%
+    \hfill
+    \subfloat[Analog prediction for the VB1 neuron\label{graph:out15VB1Celegans}]{\includesvg[width=0.5\textwidth,pretex=\tiny]{results/celegans/out15VB1}}%
+  \end{minipage}
+  \caption{\ac{C. elegans} analog responses with their digital counterparts spread out for each output neuron}
+  \label{graph:spread15Celegans}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includesvg[width=\linewidth]{results/celegans/rmse15}
+  \caption{\acp{RMSE} of each neuron's analog prediction and adjusted analog prediction for sequence 15}
+  \label{bar:rmse15}
+\end{figure}
+
+The \cref{bar:rmse15} shows that adjusted or not, the errors are in the expected range.
+The PVR output neuron seems to be hardest to reproduce during this sequence as its error is much greater than for the others.
+
+\begin{figure}[H]
+  \centering
+  \begin{minipage}{\columnwidth}
+    \subfloat[Averaged analog prediction for the DB1 neuron\label{graph:smooth15BD1Celegans}]{\includesvg[width=0.5\textwidth, pretex=\tiny]{results/celegans/smooth15DB1}}%
+    \hfill
+    \subfloat[Averaged analog prediction for the LUAL neuron\label{graph:smooth15LUALCelegans}]{\includesvg[width=0.5\textwidth,pretex=\tiny]{results/celegans/smooth15LUAL}}%
+  \end{minipage}
+  \begin{minipage}{\columnwidth}
+    \subfloat[Averaged analog prediction for the PVR neuron\label{graph:smooth15PVRCelegans}]{\includesvg[width=0.5\textwidth, pretex=\tiny]{results/celegans/smooth15PVR}}%
+    \hfill
+    \subfloat[Averaged analog prediction for the VB1 neuron\label{graph:smooth15VB1Celegans}]{\includesvg[width=0.5\textwidth,pretex=\tiny]{results/celegans/smooth15VB1}}%
+  \end{minipage}
+  \caption{\ac{C. elegans} smoothed out and slided analog responses averaged with $21$ neighbors, spread out for each output neuron}
+  \label{graph:smooth15Celegans}
+\end{figure}
+
+The results are a bit worse for this second sequence. That was expected as the inputs and the expected outputs are far more complex.

### 2023-10-26 results/airline.tex
Morning writing

@@
 \end{figure}
+
+The same network configuration using a \ac{GRU} layer instead of an \ac{LSTM} layer. The number of hidden states is the same.
 
@@
 Furthermore, this part is used to measure the impact of the custom activation functions used for training (\cref{sec:af}).
+
+\subsubsection{\acs{LSTM} predictions}
 
@@
 
+\subsubsection{\acs{GRU} predictions}
+
+Unlike for the predictions using an \ac{LSTM} layer, the weights were only trained using the analog activation functions.
+The effect of using the default activation functions is already shown with the \ac{LSTM} layer.
+
+The graph of the resulting prediction is the one in \cref{graph:airlineDigitalGRU}, the error rate is also in the figure.
+
+Interestingly, the results look much better when using the \ac{GRU} layer.
+The error rate confirms that as it is almost half of what it was with the \ac{LSTM} layer (\cref{tab:airlineDigital}).
+That simply means that the \ac{GRU} layer is better at handling this problem, as was previously stated in \cref{sec:gru}.
+
+\begin{figure}[H]
+  \centering
+  \includesvg[width=\linewidth]{datasets/airlineDigitalGRU}
+  \caption{Graph of the digital predictions along with the \ac{RMSE} for the airline dataset using a \ac{GRU} layer.
+The dotted vertical line shows the limit of the data used for training and the one used for validation.}
+  \label{graph:airlineDigitalGRU}
+\end{figure}
+
 \subsection{Analog results trained with default activation functions}\label{subsec:airlineAnalogNoC}
@@
 This subsection shows the importance of using the analog activation functions in the training phase of the \ac{NN}.
+
+The \ac{LSTM} layer being the only to have been trained using the default activation functions the only results presented here are using an \ac{LSTM} layer.
 
@@
     \hline
-    \multirow{2}{*}{\acs{RMSE}} & \multicolumn{3}{|c|}{Analog prediction}\\
+    \multirow{2}{*}{\acs{RMSE}} & \multicolumn{3}{c|}{Analog prediction}\\
     \cline{2-4}
@@
 
+\subsubsection{\ac{LSTM} predictions}
+
 This part will show the result of the inference when the weights were trained using the analog activation functions.
@@
 Once again, the results are quite close to each others, but as expected the errors (\cref{tab:airlineAnalogError}) are higher then the ones found in \cref{tab:airlineAnalogNoCZoomed}.
-The farthest appart ($n_s=1$ to $n_s=2$ with an error of $64.61$ thousands passengers) have an error barely higher than the error of digital results to the target dataset (\cref{tab:airlineDigital})
+The farthest appart ($n_s=1$ to $n_s=2$ with an error of $64.6$ thousands passengers) have an error barely higher than the error of digital results to the target dataset (\cref{tab:airlineDigital})
 
@@
 In other words, the error will probably be nullified when the training will happen inSitu (\cref{subsec:inSitu}).
+
+\subsubsection{\ac{GRU} predictions}
+
+Like explained in \cref{sec:gruCircuit}, the \ac{GRU} layer cannot be serialized in an analog circuit. The simulation were only ran once, for the only version available.
+
+\begin{figure}[H]
+  \centering
+  \includesvg[width=\linewidth]{results/airline/analogGRU}
+  \caption{Analog prediction of the \ac{GRU} analog circuit}
+  \label{graph:airlineAnalogGRU}
+\end{figure}
+
+\Cref{graph:airlineAnalogGRU} contains the analog predcitions of the \ac{GRU} circuit.
+The results are even worse than the results of simulation using weights generated with the default activation functions.
+Indeed, the error rate found in \cref{graph:airlineAnalogGRU} is much greater than the ones in \cref{tab:airlineAnalogNoC}.
+This means that while the circuit produces the right output shape, there is something wrong with the current version of the circuit.

### 2023-10-26 state/gru.tex
Morning writing

@@
 The \ac{GRU} is very often compared to the \ac{LSTM}, being sometimes assimilated as a type of \ac{LSTM} \cite{nbLSTM}.
-Their performance was found to be very similar in most situations, making those two types of \acp{RNN} coexistant in the modern machine learning world.
+Their performance was found to be very similar in most situations \cite{gruVSlstm}, making those two types of \acp{RNN} coexistant in the modern machine learning world.
 

### 2023-10-26 0.Intro/acknowledgments.tex
Asic error

@@
   I would also like to thank \ac{INESC} to have let me join their rank, to both \ac{INESC}-ID and \ac{INESC}-MN groups to have given the necessary tools to succeed with this thesis.
-  I would also like to thank all the other student working in the \ac{ASIC} group with whom I could share problems and were happy to help.
+  I would also like to thank all the other student working in the ASIC (Advanced Sensor Interfaces and Circuits) group with whom I could share problems and were happy to help.
   I would lastly like to thank Dr Ruxandra Barbulescu and Dr Diogo Caetano, to have trusted me with this very interesting topic despite my recent arrival in the university.

### 2023-10-26 design/crossbar.tex
New figures + fixed mistake
area stuff

--- 2023-10-18 3.Design/crossbar.tex
@@
 \end{equation}
-With $i_{opAmp}=0A$ because we assume an ideal \ac{opamp}.
+With $i_{opAmp}=0A$ because we assume an ideal \ac{opAmp}.
 \begin{equation}
@@
 
-The total onChip area for the crossbar circuit depends on the previously defined parameters. It also slightly changes depending on the value of $n_s$.
+The approximate onChip area for the crossbar circuit depends on the previously defined parameters. It also slightly changes depending on the value of $n_s$.
 
@@
 
-With $A_{CMOS}$ being the same than for the memory cell (\cref{sec:memcell}), $A_{CMOS}=1.2\cdot 10^{-14}m^2$, and the feature size of typical memristor that can be fabricated at INESC is $3\mu m$, which would make the approximate area for a memristor $A_{memristor}=9\mu m^2=9\cdot 10^{-12} m^2$.%TODO check info and inesc citation
+With $A_{CMOS}=w\cdot l$ and the feature size of typical memristor that can be fabricated at \ac{INESC} is $3\mu m$, which would make the approximate area for a memristor $A_{memristor}=9\mu m^2=9\cdot 10^{-12} m^2$.
 

### 2023-10-26 design/gru.tex
Changed circuit gru
area stuff

@@
   \centering
-  \includesvg[width=\textwidth,pretex=\tiny]{lstm/lstmCircuit}
+  \includesvg[width=\textwidth,pretex=\tiny]{gru/gruCircuit}
   \caption{\ac{LSTM} circuit}
@@
 
-The total onChip area for the crossbar circuit depends on the previously defined parameters.
+The approximate onChip area for the crossbar circuit depends on the previously defined parameters.
 
@@
   \begin{array}{3}
-    A_{gru}(n_i,n_h)=4\cdot A_{xbar}(n_i,\frac{n_h}{n_s},n_s)\\+ \frac{n_h}{n_s}\cdot (5\cdot A_{af}+3\cdot A_{voltMult}+3\cdot A_{R_{amp0}}+2\cdot A_{R_{amp1}}+2\cdot A_{opAmp})\\+5\cdot n_h\cdot A_{memcell}
+    A_{gru}(n_i,n_h)=4\cdot A_{xbar}(n_i,n_h,1)\\+ \frac{n_h}{n_s}\cdot (5\cdot A_{af}+3\cdot A_{voltMult}+3\cdot A_{R_{amp0}}+2\cdot A_{R_{amp1}}+2\cdot A_{opAmp})\\+5\cdot n_h\cdot A_{memcell}
   \end{array}

### 2023-10-26 design/lstm.tex
area stuff

--- 2023-10-04 3.Design/lstm.tex
@@
   \centering
-  \includesvg[height=2.5cm]{crossbar/xbarSymbol}
+  \includesvg[height=2.5cm]{lstm/lstmSymbol}
   \caption{Symbol used for the \ac{LSTM} circuit}
@@
 
-The total onChip area for the crossbar circuit depends on the previously defined parameters.
+The approximate onChip area for the crossbar circuit depends on the previously defined parameters.
 
 \begin{equation}
-  A_{xbar}(n_i,n_o,n_s)=
-  \begin{cases}
-    2\cdot n_i\cdot n_o \cdot A_{memristor}+n_o\cdot(2\cdot[A_{opAmp}+A_{R_r}]+A_{R_f}),\hfill n_s=1\\
-    \begin{array}{2}
-      2\cdot n_i\cdot n_o \cdot n_s\cdot A_{memristor}+2\cdot A_{CMOS}\cdot n_o\cdot n_s\\
-      +n_o\cdot(2\cdot[A_{opAmp}+A_{R_r}]+A_{R_f})
+  \begin{array}{3}
+    A_{lstm}(n_i,n_h,n_s)=4\cdot A_{xbar}(n_i,\frac{n_h}{n_s},n_s)\\+ \frac{n_h}{n_s}\cdot (5\cdot A_{af}+3\cdot A_{voltMult}+3\cdot A_{R_{amp0}}+2\cdot A_{R_{amp1}}+2\cdot A_{opAmp})\\+5\cdot n_h\cdot A_{memcell}
   \end{array}
-    ,\hfill \forall n_s>1\\
-  \end{cases}
 \end{equation}
-
-With $A_{CMOS}$ being the same than for the memory cell (\cref{sec:memcell}), $A_{CMOS}=1.2\cdot 10^{-14}m^2$, and the feature size of typical memristor that can be fabricated at INESC is $3\mu m$, which would make the approximate area for a memristor $A_{memristor}=9\mu m^2=9\cdot 10^{-12} m^2$.%TODO check info and inesc citation

### 2023-10-26 0.Intro/abstract.tex
Changed abstract

--- 2023-10-19 0.Intro/abstract.tex
@@
 
-  \acp{AI} are one of the main research domain in computer sciences and is used in other scientific domains.
-One great application of \ac{AI}, is in embedded systems.
-Lots of small electronic devices around us could benefit us.
-\textbf{Google} and \textbf{Apple} are pushing \ac{AI} in phones with their in house designed ARM processors having, very advertised, tensor cores.
-  Being able to run low powered \acp{AI} is thus one of the biggest computer objective of the deceny.
-  Analog computers are known to be a low powered technology that most importantly give almost instant results.
-An analog \ac{ASIC} could then be fabricated for specific embedded applications.
-This would definitely improve digital technologies in terms of time, and maybe in power consumption as well.
-Which are two prerequisite for embedded system use.
-  For this reason this work aims at creating a functioning software simulation of analog circuit capable of running \ac{AI}.
-The thesis will focus on the software simulation, the first step of the work in order to have a fully working chip able to run \ac{AI} algorithms.
+  In this work we propose an analog structure for memristor based recurrent modules targeting neural computing.
+The system is fully analog and implements a working \ac{LSTM} circuit block and a work in progress \ac{GRU} circuit block.
+Both of those blocks contain memristors to be used as weights in a analog \ac{VMM} capable circuit.
+These circuit blocks allow to run very fast computation of \aclp{RNN} of any size, in a relatively small integrated circuit.
+As part of the \ac{LSTM} and \ac{GRU} blocks, an analog activation function circuit was designed.
+This specific circuit is capable of reproducing sigmoid and \ac{tanh} like functions, with similar shapes and the same output ranges.
+The work also include the implementation of a memory cell used to store an analog value for a short period of time.
+The \ac{LSTM} block can be serialized or not with the ability to choose the level of serialization.
+Serializing the system allows to save onChip area at the cost of execution time.
+To the author's knwoledge this is the first analog implementation of the behavior of \acs{C.
+elegans} using the \ac{LSTM} block.
+Such an analog system provides ground for real time implementation of nervous systems.
 

### 2023-10-26 conclusions/conclusion.tex
working on conclusion

\section{Conclusion}\label{sec:conc}

\subsection{Performances}

\subsubsection{\ac{LSTM}}

\subsection{onChip area}

The \ac{LSTM} and \ac{GRU} blocks's onChip area is impossible to estimate, as part of the circuit is made of verilog-A models. However, to get a general idea of the area of the chip, it can be assumed that the area of the circuit mainly depends on the the area of a memrisor. Since the number of memristors is the number of weights in the circuit, the minimum area of any \ac{NN} can be determined.

For example, the \ac{NN} used to solve the airline problem using an \ac{LSTM} uses $101$ weights. This means that the minimum onChip area for this circuit is $A=101\cdot A_{memristors} = 101\cdot 9\cdot 10^{-12} = 9.09 \cdot 10^{-10} = 909 \mu m^2$.

The same parameters using a \ac{GRU} layer instead has $77$ weights so has a minimum onChip area of $A=693\mu m^2$.

A good exercise would be to take bigger systems like \textit{ChatGPT} which has

### 2023-10-27 conclusions/conclusion.tex
Worked on conclusion, missing table

@@
 \subsubsection{\ac{LSTM}}
+
+The circuit's performance has been evaluated in the previous chapter.
+The results highly depend on the parameters at play.
+For example the complexity of the dataset used, out of the two used in the thesis, the airline dataset is the simple dataset while the \ac{C.
+elegans} dataset is the more complicated datset.
+
+The best performance is the with the airline dataset, using the circuit with a serial size of one ($n_s=1$) (\cref{tab:airlineAnalog}).
+While the results are not quite right, it is assumed that this issue will be elevated once inSitu training is implemented (\cref{subsec:inSitu}).
+
+More complex datsets like \ac{C.
+elegans}, the issue is getting more present.
+Indeed, the dataset feeds to the circuit four inputs across a thousand time steps and outputs just as much data.
+Any inaccuracy is scaled up.
+
+With a simple input sequence, such as sequence 5 (\cref{graph:io5Celegans}), the resulting predictions are quite good and on par with what is expected.
+The predictions are a bit late and it is not clear what is causing this specific issue.
+A theory is that the memory cells deteriorate its stored value by a very small amount every time step, thus the stored value is affect a lot after a large amount of time steps.
+
+More complex input sequences produce even more inaccuracies.
+Indeed, when working with sequence 15 (\cref{graph:io15Celegans}), the inaccuracies spiral up and gives out a barely usable prediction.
+The curve generated doesn't fit its digital counterpart anymore.
+
+The results obtained demonstrate that the \ac{LSTM} circuit block can be run with low error when dealing with simple inputs.
+The circuit is assumed to be ready for a full simulation, meaning also training with the analog circuit.
+
+\subsubsection{\ac{GRU}}
+
+The \ac{GRU} circuit, while being present, is still a work in progress.
+The outputed predictions are scaled down for a still unknown reason, but show the right output shape.
+Once those issues are dealt with and have been fixed, the \ac{GRU} circuit will be ready for an analog training as well.
 
@@
 
+\begin{table}[H]
+  \centering
+  \begin{tabular}{|c|c|c|}
+    \hline
+    \rowcolor{gray}
+    & Sigmoid & \ac{tanh} \\
+    \hline
+    $V_1$ & \multicolumn{2}{c|}{$1.1V$}\\
+    \hline
+    $V_2$ & \multicolumn{2}{c|}{$635mV$}\\
+    \hline
+    $V_3$ & $0.8V$ & $550mV$\\
+    \hline
+    $i_{dc}$ & \multicolumn{2}{c|}{$150uA$}\\
+    \hline
+    $w$ & \multicolumn{2}{c|}{$900nm$}\\
+    \hline
+    $l$ & \multicolumn{2}{c|}{$60nm$}\\
+    \hline
+    $R_1$ & \multicolumn{2}{c|}{$5k\Omega$}\\
+    \hline
+    $R_2$ & \multicolumn{2}{c|}{$10k\Omega$}\\
+    \hline
+    $R_3$ & $2k\Omega$ & $4k\Omega$\\
+    \hline
+  \end{tabular}
+  \caption{Estimated onChip areas}
+  \label{tab:areas}
+\end{table}
+
+%do a table
 For example, the \ac{NN} used to solve the airline problem using an \ac{LSTM} uses $101$ weights.

### 2023-10-27 conclusions/conclusion.tex
On chip area almost done

@@
 
+\Cref{tab:areas} contains several \acp{NN} models with their number of parameters, a minimum onChip area and an estimated feature size.
+None of those values have any scientific ground, they are just here to give general idea of the kind of circuit thta could be fabricated.
+Those areas only considers the area of the memristors.
+
+The feature size represents the length of the side of the chip is the chip was manufactured in a square.
+
+The values of \acp{LLM} were included to show how the area of the chip scales up.
+Those models don't use \acp{RNN}, and futhermore, give out ridiculous areas.
+Fabricating such cicuits in analog is not feasible witht the current technologies.
+
 \begin{table}[H]
   \centering
-  \begin{tabular}{|c|c|c|}
+  \begin{tabular}{|c|c|c|c|}
     \hline
     \rowcolor{gray}
-    & Sigmoid & \ac{tanh} \\
+    Model & Parameters & Minimum area & Approximate feature size \\
     \hline
-    $V_1$ & \multicolumn{2}{c|}{$1.1V$}\\
+    Airline \ac{GRU} $n_h=4$ & $77$ & $693\mu m^2$ & $26.3\mu m$\\
     \hline
-    $V_2$ & \multicolumn{2}{c|}{$635mV$}\\
+    Airline \ac{LSTM} $n_h=4$ & $101$ & $909\mu m^2$ & $30.1\mu m$\\
     \hline
-    $V_3$ & $0.8V$ & $550mV$\\
+    \ac{C. elegans} \ac{GRU} $n_h=8$ & $348$ & $3132\mu m^2$ & $56.0\mu m$\\
     \hline
-    $i_{dc}$ & \multicolumn{2}{c|}{$150uA$}\\
+    \ac{C. elegans} \ac{LSTM} $n_h=8$ & $452$ & $4068\mu m^2$ & $63.8\mu m$\\
     \hline
-    $w$ & \multicolumn{2}{c|}{$900nm$}\\
+    \multirow{3}{*}{\acs{LLaMA}-2} & $7\cdot 10^9$ & $630 cm^2$ & $25.1 cm$ \\
+    \cline{2-4}
+    & $13\cdot 10^9$ & $1170 cm^2$ & $34.2 cm$ \\
+    \cline{2-4}
+    & $70\cdot 10^9$ & $6300 cm^2$ & $79.4 cm$ \\
     \hline
-    $l$ & \multicolumn{2}{c|}{$60nm$}\\
+    \acs{GPT}-3 & $175\cdot 10^9$ & $1.57 m^2$ & $1.25m$\\
     \hline
-    $R_1$ & \multicolumn{2}{c|}{$5k\Omega$}\\
-    \hline
-    $R_2$ & \multicolumn{2}{c|}{$10k\Omega$}\\
-    \hline
-    $R_3$ & $2k\Omega$ & $4k\Omega$\\
+    \acs{GPT}-4 & $1\cdot 10^{12}$ & $9.00 m^2$ & $3m$\\
     \hline
@@
 
-%do a table
-For example, the \ac{NN} used to solve the airline problem using an \ac{LSTM} uses $101$ weights.
-This means that the minimum onChip area for this circuit is $A=101\cdot A_{memristors} = 101\cdot 9\cdot 10^{-12} = 9.09 \cdot 10^{-10} = 909 \mu m^2$.
-
-The same parameters using a \ac{GRU} layer instead has $77$ weights so has a minimum onChip area of $A=693\mu m^2$.
-
-A good exercise would be to take bigger systems like \textit{ChatGPT} which has
+Nevertheless, it is interresting to look at those approximation.

### 2023-10-27 conclusions/conclusion.tex
Finished this part

@@
 
-\subsection{onChip area}
+\subsection{onChip area}\label{subsec:area}
 
 The \ac{LSTM} and \ac{GRU} blocks's onChip area is impossible to estimate, as part of the circuit is made of verilog-A models.
-However, to get a general idea of the area of the chip, it can be assumed that the area of the circuit mainly depends on the the area of a memrisor.
+However, to get a general idea of the area of the chip, it can be assumed that the area of the circuit mainly depends on the the area of a memrisor, because the area of a memristor is much greater than the ones of the other components.
 Since the number of memristors is the number of weights in the circuit, the minimum area of any \ac{NN} can be determined.
+The area for a memristor was determined using the feature size of the memristors that can be fabricated at \ac{INESC}.
+The typical memristor that can be fabricated at \ac{INESC} is $3\mu m$, which would make the approximate area for a memristor $A_{memristor}=9\mu m^2=9\cdot 10^{-12} m^2$
 
@@
 
-The feature size represents the length of the side of the chip is the chip was manufactured in a square.
+The feature size represents the length of the side of the chip if the chip was manufactured in a square.
 
@@
   \end{tabular}
-  \caption{Estimated onChip areas}
+  \caption{Estimated onChip area for different models of \acp{NN}}
   \label{tab:areas}
@@
 Nevertheless, it is interresting to look at those approximation.
+It shows that the simple models used in this work would take a very small onChip area, even though this a very raw number.
+It also shows the current limit with state of the art \acp{NN} at the current time.

### 2023-10-27 paper/paper.tex
Working on paper, will do over the we

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{Memristor-based recurrent modules for neural computing}

\author{\IEEEauthorblockN{Valentin BARBAZA}
%\IEEEauthorblockA{\textit{Master student} \\
%\textit{name of organization (of Aff.)}\\
Lisbon, Portugal \\
valentin@barbaza.org}

\maketitle

\begin{abstract}
  In this work we propose an analog structure for memristor based recurrent modules targeting neural computing. The system is fully analog and implements a working \ac{LSTM} circuit block and a work in progress \ac{GRU} circuit block. Both of those blocks contain memristors to be used as weights in a analog \ac{VMM} capable circuit. These circuit blocks allow to run very fast computation of \aclp{RNN} of any size, in a relatively small integrated circuit. As part of the \ac{LSTM} and \ac{GRU} blocks, an analog activation function circuit was designed. This specific circuit is capable of reproducing sigmoid and \ac{tanh} like functions, with similar shapes and the same output ranges. The work also include the implementation of a memory cell used to store an analog value for a short period of time. The \ac{LSTM} block can be serialized or not with the ability to choose the level of serialization. Serializing the system allows to save onChip area at the cost of execution time. To the author's knwoledge this is the first analog implementation of the behavior of \acs{C. elegans} using the \ac{LSTM} block. Such an analog system provides ground for real time implementation of nervous systems.
\end{abstract}

\begin{IEEEkeywords}
  Memristors, Crossbar array, VMM, Machine Learning, RNN, LSTM, GRU, Analog computing, Analog circuits, Analog activation functions, Tensorflow, Keras, Embedded systems, Neural computing
\end{IEEEkeywords}

\section{Introduction}

AIs (Artificial Intelligences) are one of the main research domain in computer sciences and is used in other scientific domains. One great application of AI, is in embedded systems. Lots of small electronic devices around us could benefit us. \textbf{Google} and \textbf{Apple} are pushing AI in phones with their in house designed ARM processors having, very advertised, tensor cores.
Being able to run low powered AIs is thus one of the biggest computer objective of the deceny.
Analog computers are known to be a low powered technology that most importantly give almost instant results. An analog ASIC (Application Specific Integrated Circuit) could then be fabricated for specific embedded applications. This would definitely improve digital technologies in terms of time, and maybe in power consumption as well, these being two prerequisite for embedded system use.
For this reason this work aims at creating a functioning software simulation of analog circuit capable of running AI. The thesis will focus on the software simulation, the first step of the work in order to have a fully working chip able to run AI algorithms.

\section{State of the art}

\subsection{Recurrent Neural Networks (RNN)}

RNNs are a family of neural networks that differentiate themselves by having feedback connections. It alloww it to be used with sequences of data, sometimes, with varying amount of input data.

The IEEEtran class file is used to format your paper and style the text. All margins,
column widths, line spaces, and text fonts are prescribed; please do not
alter them. You may note peculiarities. For example, the head margin
measures proportionately more than is customary. This measurement
and others are deliberate, using specifications that anticipate your paper
as one part of the entire proceedings, and not as an independent document.
Please do not revise any of the current designations.

\section{Prepare Your Paper Before Styling}
Before you begin to format your paper, first write and save the content as a
separate text file. Complete all content and organizational editing before
formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on
proofreading, spelling and grammar.

Keep your text and graphic files separate until after the text has been
formatted and styled. Do not number text heads---{\LaTeX} will do that
for you.

\subsection{Abbreviations and Acronyms}\label{AA}
Define abbreviations and acronyms the first time they are used in the text,
even after they have been defined in the abstract. Abbreviations such as
IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use
abbreviations in the title or heads unless they are unavoidable.

\subsection{Units}
\begin{itemize}
  \item Use either SI (MKS) or CGS as primary units. (SI units are encouraged.) English units may be used as secondary units (in parentheses). An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
  \item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds. This often leads to confusion because equations do not balance dimensionally. If you must use mixed units, clearly state the units for each quantity that you use in an equation.
  \item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''. Spell out units when they appear in text: ``. . . a few henries'', not ``. . . a few H''.
  \item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
\end{itemize}

\subsection{Equations}
Number equations consecutively. To make your
equations more compact, you may use the solidus (~/~), the exp function, or
appropriate exponents. Italicize Roman symbols for quantities and variables,
but not Greek symbols. Use a long dash rather than a hyphen for a minus
sign. Punctuate equations with commas or periods when they are part of a
sentence, as in:
\begin{equation}
  a+b=\gamma\label{eq}
\end{equation}

Be sure that the
symbols in your equation have been defined before or immediately following
the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at
the beginning of a sentence: ``Equation \eqref{eq} is . . .''

\subsection{\LaTeX-Specific Advice}

Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
of ``hard'' references (e.g., \verb|(1)|). That will make it possible
to combine sections, add equations, or change the order of figures or
citations without having to go through the file line by line.

Please don't use the \verb|{eqnarray}| equation environment. Use
\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
environment leaves unsightly spaces around relation symbols.

Please note that the \verb|{subequations}| environment in {\LaTeX}
will increment the main equation counter even when there are no
equation numbers displayed. If you forget that, you might write an
article in which the equation numbers skip from (17) to (20), causing
the copy editors to wonder if you've discovered a new method of
counting.

{\BibTeX} does not work by magic. It doesn't get the bibliographic
data from thin air but from .bib files. If you use {\BibTeX} to produce a
bibliography you must send the .bib files.

{\LaTeX} can't read your mind. If you assign the same label to a
subsubsection and a table, you might find that Table I has been cross
referenced as Table IV-B3.

{\LaTeX} does not have precognitive abilities. If you put a
\verb|\label| command before the command that updates the counter it's
supposed to be using, the label will pick up the last counter to be
cross referenced instead. In particular, a \verb|\label| command
should not go before the caption of a figure or a table.

Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
will not stop equation numbers inside \verb|{array}| (there won't be
any anyway) and it might stop a wanted equation number in the
surrounding equation.

\subsection{Some Common Mistakes}\label{SCM}
\begin{itemize}
  \item The word ``data'' is plural, not singular.
  \item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
  \item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation. When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks. A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this). (A parenthetical sentence is punctuated within the parentheses.)
  \item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
  \item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
  \item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
  \item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
  \item Do not confuse ``imply'' and ``infer''.
  \item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
  \item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
  \item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
\end{itemize}
An excellent style manual for science writers is \cite{b7}.

\subsection{Authors and Affiliations}
\textbf{The class file is designed for, but not limited to, six authors.} A
minimum of one author is required for all conference articles. Author names
should be listed starting from left to right and then moving down to the
next line. This is the author sequence that will be used in future citations
and by indexing services. Names should not be listed in columns nor group by
affiliation. Please keep your affiliations as succinct as possible (for
example, do not differentiate among departments of the same organization).

\subsection{Identify the Headings}
Headings, or heads, are organizational devices that guide the reader through
your paper. There are two types: component heads and text heads.

Component heads identify the different components of your paper and are not
topically subordinate to each other. Examples include Acknowledgments and
References and, for these, the correct style to use is ``Heading 5''. Use
``figure caption'' for your Figure captions, and ``table head'' for your
table title. Run-in heads, such as ``Abstract'', will require you to apply a
style (in this case, italic) in addition to the style provided by the drop
down menu to differentiate the head from the text.

Text heads organize the topics on a relational, hierarchical basis. For
example, the paper title is the primary text head because all subsequent
material relates and elaborates on this one topic. If there are two or more
sub-topics, the next level head (uppercase Roman numerals) should be used
and, conversely, if there are not at least two sub-topics, then no subheads
should be introduced.

\subsection{Figures and Tables}
\paragraph{Positioning Figures and Tables} Place figures and tables at the top and
bottom of columns. Avoid placing them in the middle of columns. Large
figures and tables may span across both columns. Figure captions should be
below the figures; table heads should appear above the tables. Insert
figures and tables after they are cited in the text. Use the abbreviation
``Fig.~\ref{fig}'', even at the beginning of a sentence.

\begin{table}[htbp]
  \caption{Table Type Styles}
  \begin{center}
    \begin{tabular}{|c|c|c|c|}
      \hline
      \textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
      \cline{2-4}
      \textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
      \hline
      copy& More table copy$^{\mathrm{a}}$& &  \\
      \hline
      \multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
    \end{tabular}
    \label{tab1}
  \end{center}
\end{table}

\begin{figure}[htbp]
  \centerline{\includegraphics{fig1.png}}
  \caption{Example of a figure caption.}
  \label{fig}
\end{figure}

Figure Labels: Use 8 point Times New Roman for Figure labels. Use words
rather than symbols or abbreviations when writing Figure axis labels to
avoid confusing the reader. As an example, write the quantity
``Magnetization'', or ``Magnetization, M'', not just ``M''. If including
units in the label, present them within parentheses. Do not label axes only
with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization
\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of
quantities and units. For example, write ``Temperature (K)'', not
``Temperature/K''.

\section*{Acknowledgment}

The preferred spelling of the word ``acknowledgment'' in America is without
an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B.
G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor
acknowledgments in the unnumbered footnote on the first page.

\section*{References}

Please number citations consecutively within brackets \cite{b1}. The
sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference
number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at
the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''

Number footnotes separately in superscripts. Place the actual footnote at
the bottom of the column in which it was cited. Do not put footnotes in the
abstract or reference list. Use letters for table footnotes.

Unless there are six authors or more give all authors' names; do not use
``et al.''. Papers that have not been published, even if they have been
submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers
that have been accepted for publication should be cited as ``in press'' \cite{b5}.
Capitalize only the first word in a paper title, except for proper nouns and
element symbols.

For papers published in translation journals, please give the English
citation first, followed by the original foreign-language citation \cite{b6}.

\begin{thebibliography}{00}
  \bibitem{b1} G. Eason, B. Noble, and I. N. Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil. Trans. Roy. Soc. London, vol. A247, pp. 529--551, April 1955.
  \bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
  \bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
  \bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
  \bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
  \bibitem{b6} Y. Yorozu, M. Hirano, K. Oka, and Y. Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl. J. Magn. Japan, vol. 2, pp. 740--741, August 1987 [Digests 9th Annual Conf. Magnetics Japan, p. 301, 1982].
  \bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
\end{thebibliography}
\vspace{12pt}
\color{red}
IEEE conference templates contain guidance text for composing and formatting conference papers. Please ensure that all template text is removed from your conference paper prior to submission to the conference. Failure to remove the template text from your paper may result in your paper not being published.

\end{document}

### 2023-10-27 introduction/contributions.tex
Diogo correction introduction

--- 2023-10-25 introduction/contributions.tex
@@
   \item I created a script that can generate spice netlists for \acps{LSTM} or feed forward \aclp{NN} of any size and importing weights as resistance values from a specified file containing the architecture and the weights.
-  \item The script can generate an \ac{LSTM} capable analog circuit was created and ran in virtuoso giving encouraging results.
-  \item The script can also generate an \ac{GRU} capable analog circuit was created and ran in virtuoso.
+  \item The script can generate an \ac{LSTM} capable analog circuits and then run a simulation of the circuit in virtuoso giving encouraging results.
+  \item The script can also generate an \ac{GRU} capable analog circuits and then run a simulation of the circuit in virtuoso.
   \item I developed an analog activation function circuit that supports \ac{tanh} and sigmoid ranges working at $1.8V$.
@@
   \item I used Cadence's Virtuoso to run software simulation of different \acp{NN} capable analog circuits.
+  \item A paper is being prepared to report the advancements made in this work.
 \end{itemize}

### 2023-10-27 introduction/motivation.tex
Diogo correction introduction

@@
 
-Another use for \ac{AI} that works great in embedded use is for camera relocalization \cite{videoReloc}.
+Another use for \ac{AI} that works great in embedded use is for camera enhanced 3D camera localization \cite{videoReloc}.
 That is useful especially for \ac{VR} headsets that need to know their location in the room.
@@
 
-The focus of this thesis is to create a working simulation of an analog \ac{LSTM} circuit using Cadence's Virtuoso as simulation software.
+The focus of this thesis is to create a working simulation of an analog \ac{LSTM} and \ac{GRU} circuit using Cadence's Virtuoso as simulation software.
 Having those preliminary results would allow to fabricate and test such a chip.

### 2023-10-27 introduction/outline.tex
Diogo correction introduction

--- 2023-10-19 introduction/outline.tex
@@
 
-All the technologies used and, thus are required to know in order to understand the thesis are detailed in the state of the art.
+In \cref{cap:state}, all the technologies used and thus required to know in order to understand the thesis are detailed in the state of the art.
 
-Next the work that was done for the thesis is goign to be thoroughly explained.
+Next, in \cref{cap:design}, the work that was done for the thesis is going to be thoroughly explained.
 Starting with the different circuit that were designed for the thesis.
 Then the different scripts and python code that were created, either to run the digital \acp{NN} or to generate the analog system into a netlist.
+
+Then \cref{cap:results} contains the results, obtained from the digital inference and the analog simulation, from both the number of airline passenger and the \ac{C. elegans} datasets.
+
+Finally, in \cref{cap:conclusions}, the conclusions and improvements that can be made to the different parts of the thesis.

### 2023-10-27 state/lstm.tex
Correctiion diogo state

@@
   \label{fig:lstmCell}
-  \caption{\acs{LSTM} cell}
+  \caption{\acs{LSTM} cell, adapted from \cite{wikiLSTM}}
 \end{figure}
@@
     \begin{itemize}
-      \item the candidate cell and thus becomes $ \tilde{c}_t=(w_c\cdot[x_{t-1},h_{t-1}] + b_c) $, this is called the \ac{NIAF}
-      \item the output of the previous cell state and thus becomes $ h_t=o_t\cdot c_t $, and is called \ac{NOAF}
+      \item the candidate cell and thus becomes $ \overrightarrow{\tilde{c}_t}=(w_c\cdot[\overrightarrow{x_{t-1}},\overrightarrow{h_{t-1}}] + \overrightarrow{b_c}) $, this is called the \ac{NIAF}
+      \item the output of the previous cell state and thus becomes $ \overrightarrow{h_t}=\overrightarrow{o_t}\odot \overrightarrow{c_t} $, and is called \ac{NOAF}
     \end{itemize}
@@
     \begin{itemize}
-      \item \ac{NIG} : $i_t=1$
-      \item \ac{NOG} : $o_t=1$
-      \item \ac{NFG} : $f_t=1$
+      \item \ac{NIG} : $\overrightarrow{i_t}=1$
+      \item \ac{NOG} : $\overrightarrow{o_t}=1$
+      \item \ac{NFG} : $\overrightarrow{f_t}=1$
     \end{itemize}

### 2023-10-27 state/rnn.tex
Correctiion diogo state

@@
 \begin{equation}\label{eq:srnn}
-  \overrightarrow{h_t}=tanh([\overrightarrow{x_t},\overrightarrow{h_{t-1}}]\cdot w + b)
+  \overrightarrow{h_t}=tanh([\overrightarrow{x_t},\overrightarrow{h_{t-1}}]\cdot w + \overrightarrow{b})
 \end{equation}
 
-Where $t\in\mathbb{N}^*$ is the time index, $w$ the weight matrix, $b$ the bias matrix, $\overrightarrow{x_t}$ the input vector and $\overrightarrow{h_{t}}$ the hidden state of the \ac{RNN}.
+Where $t\in\mathbb{N}^*$ is the time index, $w$ the weight matrix, $\overrightarrow{b}$ the bias vector, $\overrightarrow{x_t}$ the input vector and $\overrightarrow{h_{t}}$ the hidden state of the \ac{RNN}.
 

### 2023-10-27 design/activation.tex
area stuff
Correction digoo

--- 2023-10-17 3.Design/activation.tex
@@
 Designing an analog activation function as close to the original is very important for the final result's quality.
+
+\subsection{Circuit}
+
+The circuit used is the same as the one in \cite{thesisRef}, the circuit is the one shown in \cref{circt:af}.
+The technology used being different, all the parameters had to be determined empirically to best fit a sigmoid shape.
+The parameters can be found in \cref{tab:afPar}.
+
+Due to the nature of the functions we want to generate, we will use the same circuit for both a sigmoid and a \ac{tanh} like functions.
+The two different functions are generated by changing two parameters.
+
+The functions generated are the same shape and only differ by their output range.
+
+\begin{figure}[H]
+  \centering
+  \includesvg[width=\textwidth]{activation/afCircuit}
+  \caption{Activation functions circuit}
+  \label{circt:af}
+\end{figure}
 
@@
 
-\subsection{Circuit}
+\subsection{Symbols}
 
-The circuit used is the same as the one in \cite{thesisRef}, the circuit is the one shown in \cref{circt:af}.
-The technology used being different, all the parameters had to be determined empirically to best fit a sigmoid shape.
-The parameters can be found in \cref{tab:afPar}.
-
-Due to the nature of the functions we want to generate, we will use the same circuit for both a sigmoid and a \ac{tanh} like functions.
-The two different functions are generated by changing two parameters.
-
-The functions generated are the same shape and only differ by their output range.
-
-\begin{figure}[H]
-  \centering
-  \includesvg[width=\textwidth]{activation/afCircuit}
-  \caption{Activation functions circuit}
-  \label{circt:af}
-\end{figure}
-
-The approximate onChip area of this circuit is :
-
-\begin{equation}
-  A_{af}=2\cdot A_{R_1} + 2\cdot A_{CMOS} + 5\cdot A_{R_2} + A_{R_3} +2\cdot A_{opAmp}
-\end{equation}
-With $A_{CMOS} = w\cdot l = 9\cdot 10^{-7} \cdot 6 \cdot 10^{-8} = 5.4 \cdot 10^{-14} m^2$
-TODO : Finish calc
-
-\subsection{Symbols}
 The symbols for the sigmoid and the \ac{tanh} are separated for better understanding.
@@
 \end{figure}
+
+The approximate onChip area of this circuit is :
+
+\begin{equation}
+  A_{af}=2\cdot A_{R_1} + 2\cdot A_{CMOS} + 5\cdot A_{R_2} + A_{R_3} +2\cdot A_{opAmp}
+\end{equation}
+With $A_{CMOS} = w\cdot l$
+
+In order to combine the the onChip area of both the sigmoid and \ac{tanh} activation functions, it is assumed that the variation of area taken by the resistor ($R_3$) is negligible compared to the overall active area.
 
@@
 These functions are still a bit different from the original functions (especially for the \ac{tanh}).
-However that doesn't matter too much as the trainning will be happening within the final circuit, all weights will be set in the circuit.
+However that doesn't matter too much as the trainning will be happening with the extracted analog functions, all weights will be set in the circuit.
 This is the reason why such a difference doesn't matter.

### 2023-10-27 design/crossbar.tex
Finished this part
Correction digoo

@@
 
-First of all, a two memristors per synapse architecture has been used. This design is beneficial in two ways :
-
-\begin{itemize}
-  \item It doubles the range of weights (\cref{tab:synapses})
-  \item It allows to easily use negative weights, as this design simply allows the output values to be centered around $V_{cm}$ to be compliant with the standard set in \cref{tab:valConv}.
-\end{itemize}
+We could choose between one or two memristor per synapse.
+Using two memristor per synapse doubles the area but doubles the weight range (\cref{tab:synapses}) and allows to easily use negavite weights.
+The output voltage will be centered around $V_{cm}$ and be compliant with the standard set in \cref{tab:valConv}.
 
@@
 
-With $A_{CMOS}=w\cdot l$ and the feature size of typical memristor that can be fabricated at \ac{INESC} is $3\mu m$, which would make the approximate area for a memristor $A_{memristor}=9\mu m^2=9\cdot 10^{-12} m^2$.
+With $A_{CMOS}=w\cdot l$.
 
@@
 
-The input voltages must be monitored very carefuly in order not to mess up the entire system.
-Indeed, memristors have a voltage threshold ($V_{threshold}$) above which the internal resistance changes \cite{memristorSpiceModels,memCadenceModel,memTEAMmodel,memVTEAMmodel}.
+The input voltages must be monitored very carefuly in order not to accidentally change a memristor's internal resistance.
+Indeed, memristors have a voltage threshold ($V_{read}$) above which the internal resistance changes \cite{memristorSpiceModels,memCadenceModel,memTEAMmodel,memVTEAMmodel}.
 This means that the input voltages must follow the inequality in \cref{eq:memThres}.
@@
 \begin{equation}\label{eq:memThres}
-  |V_{mem}|= |x_i-V_{cm}|\le V_{threshold}
+  |V_{mem}|= |x_i-V_{cm}|\le V_{read}
 \end{equation}
@@
 
-The analog system being only simulated for the inference of the \ac{NN}.
-Furthermore, since the weights only change during the training phase, and the resistances are just a rendition of the weights (\cref{sec:wei2res}), the resistances don't need to change, the circuit used in the simulations is thus using resistors instead of memristors for the sake of simplicity, as importing resistors is way easier to import in Cadence's virutoso.
+In this work, the analog system will be simulated for the inference of the \ac{NN}, thus the weights will not have to change during the simulation.
+Because the weights are represented in the analog circuit by the internal resistances of the memristors, the memristors can be replaced by resistors with a set resistance for the simulation.

### 2023-10-27 design/genwei.tex
Correction digoo

--- 2023-10-11 3.Design/genwei.tex
@@
 The choice for the weight generation was to use Keras API with the tensorflow framework.
-However, both tensorflow and pyTorch were tried and were giving similar results, the final choice of using tensorflow was made because it's the most popular among the research group.%INESC ID ???
+However, both tensorflow and pyTorch were tried and were giving similar results, the final choice of using tensorflow was made because it's the most popular among the research group.
 
@@
 
-The weights need to be contrained during the training to make sure that the voltage of the analog circuit stays with in its operating voltage range ($[0,V_{dd}]$).
+The weights need to be contrained during the training to make sure that the voltage of the analog circuit stays within its operating voltage range ($[0,V_{dd}]$).
 To avoid any unwanted behavior when the system supposedly goes out of range, the weights in the \ac{LSTM} are contrained.

### 2023-10-27 design/lstm.tex
Correctiion diogo state
Correction digoo

@@
 This vector is there to lighten the informations on the schematic (\cref{circt:lstm}).
-$\overrightarrow{z_t}$ is defined by $\overrightarrow{z_t}=(\overrightarrow{x_t},\overrightarrow{h_{t-1}},b)$.
+$\overrightarrow{z_t}$ is defined by $\overrightarrow{z_t}=(\overrightarrow{x_t},\overrightarrow{h_{t-1}},1)$.
   \item $e_{j,0}$ and $e_{j,1}$ are two enable flags that respectively represent the first and second half of $e_j$.
@@
 
-Using an \ac{LSTM} in serial mode is very beneficial as it divides the number of point wise circuit by $n_s$.
-
-There is a sweet spot for the serial size ($n_s$) :
+Using an \ac{LSTM} in serial mode is very beneficial as it divides the number of point wise circuit by $n_s$. There is a compromise to do in order to choose the serial size ($n_s$) :
 
@@
   \item When the serial size ($n_s)$ increases, the inference time increases with a factor of $O(n_s)$ while the onChip area decrease with the same factor.
-  \item When the serial size ($n_s)$ decreases, similarly, the onChip area increases with a factor of $O(n_s)$ while the tinference time decreases by the same factor.
+  \item When the serial size ($n_s)$ decreases, similarly, the onChip area increases with a factor of $O(n_s)$ while the inference time decreases by the same factor.
 \end{itemize}

### 2023-10-27 design/netlist.tex
Correction digoo

@@
 
-As of the writing of the thesis, the netlist generator can generate the netlist for :
-
-\begin{itemize}
-  \item \ac{NP} \ac{LSTM}
-  \item Vanilla \ac{LSTM}
-  \item \ac{FGR} \ac{LSTM}
-  \item \ac{GRU}
-\end{itemize}
-
-Nevertheless, the weights can only be generated for the \ac{NP} \ac{LSTM} and the \ac{GRU} (\cref{sec:genwei}).
-While the script can tecnically generate the Vanilla and \ac{FGR} \acp{LSTM}, since there is no way to generate weights for those version of the \ac{LSTM}, they are rendered useless.
+As of the writing of the thesis, the netlist generator can generate the netlist for the \ac{NP} \ac{LSTM} and the \ac{GRU}. Other type of \acp{NN} can be added to the code.
 
@@
 
-The weights are stored in a single file that is the save a numpy array.
-This file also contains a brief description of the architecture being used.
+The weights are stored in a single file.
+This file contains a brief description of the architecture being used.
 The first index stores this desciption.
-the following indexes are for the actual weights, in the order given in the description.
+The following indexes are for the weights of each layer, in the order given in the description.
 Depending on the layer used, the weights are stored a bit differently :

### 2023-10-27 state/lstm.tex
plural to singular

@@
 
-Where ($w_i$,$\overrightarrow{b_i}$), ($w_f$,$\overrightarrow{b_f}$) and ($w_o$,$\overrightarrow{b_o}$) are the pair of weights matrixes and biases vectors for the input, forget and output gates respectively.
+Where ($w_i$,$\overrightarrow{b_i}$), ($w_f$,$\overrightarrow{b_f}$) and ($w_o$,$\overrightarrow{b_o}$) are the pair of weights matrixes and bias vectors for the input, forget and output gates respectively.
 $\overrightarrow{x_t}$ is the input vector and $\overrightarrow{h_t}$ is the hidden state vector.

### 2023-10-28 state/crossbar.tex
Reorga + cite figures

@@
   \centering
-  \subfloat[Schematics]{\includesvg[width=.45\linewidth]{crossbar/crossbar}}%\qquad
+  \subfloat[Schematics, inspired from \cite{xbarFigures}]{\includesvg[width=.45\linewidth]{crossbar/crossbar}}%\qquad
   \hfill
-  \subfloat[3 dimensional representation]{\includegraphics[width=.45\linewidth]{crossbar/crossbar3D.jpg}}\\
+  \subfloat[3 dimensional representation, from \cite{xbarFigures}]{\includegraphics[width=.45\linewidth]{crossbar/crossbar3D.jpg}}\\
   \caption{Memristor crossbar array}

### 2023-10-28 conclusions/conclusion.tex
Changed table with double mem
Bunch of changes, to gru notably

@@
 None of those values have any scientific ground, they are just here to give general idea of the kind of circuit thta could be fabricated.
-Those areas only considers the area of the memristors.
+Those areas only considers the area of the memristors, using a two memristor per synapse architecture like the one used in the thesis.
 
@@
 
-The values of \acp{LLM} were included to show how the area of the chip scales up.
+The values of \acp{LLM}, such as \ac{LLaMA}-2 \textbf{Meta}'s \ac{LLM} and the two last versions of \acp{GPT} the \ac{LLM} that powers \textit{ChatGPT}, were included to show how the area of the chip scales up.
 Those models don't use \acp{RNN}, and futhermore, give out ridiculous areas.
@@
     \hline
-    Airline \ac{GRU} $n_h=4$ & $77$ & $693\mu m^2$ & $26.3\mu m$\\
+    Airline \ac{GRU} $n_h=4$ & $77$ & $1386\mu m^2$ & $37.2\mu m$\\
     \hline
-    Airline \ac{LSTM} $n_h=4$ & $101$ & $909\mu m^2$ & $30.1\mu m$\\
+    Airline \ac{LSTM} $n_h=4$ & $101$ & $1818\mu m^2$ & $42.6\mu m$\\
     \hline
-    \ac{C. elegans} \ac{GRU} $n_h=8$ & $348$ & $3132\mu m^2$ & $56.0\mu m$\\
+    \ac{C. elegans} \ac{GRU} $n_h=8$ & $348$ & $6264\mu m^2$ & $79.1\mu m$\\
     \hline
-    \ac{C. elegans} \ac{LSTM} $n_h=8$ & $452$ & $4068\mu m^2$ & $63.8\mu m$\\
+    \ac{C. elegans} \ac{LSTM} $n_h=8$ & $452$ & $8136\mu m^2$ & $90.2\mu m$\\
     \hline
-    \multirow{3}{*}{\acs{LLaMA}-2} & $7\cdot 10^9$ & $630 cm^2$ & $25.1 cm$ \\
+    \multirow{3}{*}{\acs{LLaMA}-2} & $7\cdot 10^9$ & $1260 cm^2$ & $35.5 cm$ \\
     \cline{2-4}
-    & $13\cdot 10^9$ & $1170 cm^2$ & $34.2 cm$ \\
+    & $13\cdot 10^9$ & $2340 cm^2$ & $48.4 cm$ \\
     \cline{2-4}
-    & $70\cdot 10^9$ & $6300 cm^2$ & $79.4 cm$ \\
+    & $70\cdot 10^9$ & $1.26 m^2$ & $1.12 m$ \\
     \hline
-    \acs{GPT}-3 & $175\cdot 10^9$ & $1.57 m^2$ & $1.25m$\\
+    \acs{GPT}-3 & $175\cdot 10^9$ & $3.14 m^2$ & $1.77m$\\
     \hline
-    \acs{GPT}-4 & $1\cdot 10^{12}$ & $9.00 m^2$ & $3m$\\
+    \acs{GPT}-4 & $1\cdot 10^{12}$ & $18.0 m^2$ & $4.24m$\\
     \hline

### 2023-10-28 design/gru.tex
Correction digoo
Bunch of changes, to gru notably

@@
   \includesvg[width=\textwidth,pretex=\tiny]{gru/gruCircuit}
-  \caption{\ac{LSTM} circuit}
+  \caption{\ac{GRU} circuit}
   \label{circt:lstm}
@@
 
-The system is built, once again, using crossbar array (\cref{sec:xbarCircuit}) with ($n_i+n_h+1$,$n_o$, $n_s$) as parameters.
+The system is built, once again, using crossbar array (\cref{sec:xbarCircuit}) with ($n_i+n_h+1$,$n_h$, $1$) as parameters.
 
@@
   \item $\overrightarrow{x_t}$ : This is the input vector for the \ac{LSTM} circuit at time $t$. It has a size of $n_i$.
-  \item $\overrightarrow{h_t}$ : This is the hidden layer vector for the feedback connections, it is defined as $\overrightarrow{h_t}=(\overrightarrow{h_{t,i}}) \forall i\in [\![0,n_o-1]\!]$, with $\overrightarrow{h_{t,i}}=(h_{t,i,j}) \forall j\in [\![0,n_s-1]\!]$.
+  \item $\overrightarrow{h_t}$ : This is the hidden layer vector for the feedback connections, it is defined as $\overrightarrow{h_{t}}=(h_{t,i}) \forall i\in [\![0,n_h-1]\!]$.
   \item $\overrightarrow{z_t}$ is the input of the crossbar but not the input of the \ac{LSTM}.
 This vector is there to lighten the informations on the schematic (\cref{circt:lstm}).
-$\overrightarrow{z_t}$ is defined by $\overrightarrow{z_t}=(\overrightarrow{x_t},\overrightarrow{h_{t-1}},b)$.
-  \item $e_{j,0}$ and $e_{j,1}$ are two enable flags that respectively represent the first and second half of $e_j$.
+$\overrightarrow{z_t}$ is defined by $\overrightarrow{z_t}=(\overrightarrow{x_t},\overrightarrow{h_{t-1}},\overrightarrow{b})$.
   \item $e_{in}$ and $e_{out}$ are the flags used to enable the hidden state values to go to the input (feedback connection) or to the output of the circuit.
+  \item $e_{in,end}$ is the second half of the $e_{in}$ flag.
   \item $e_{next}$ is the enable flag on in between two time steps.
@@
 The resistances must stay arround the values of the resistances used around the circuit, especially those of the memristors.
+  \item $V_{inv}$ is the voltage used to perform the $1-x$ operation.
 \end{itemize}
 
-The wires coming into the crossbar are a bus of size $n_i+n_h+1$ and the output of the crossbar is a bus of size $n_h$ (\cref{sym:xbar}).
-This is why everything in the system apart from the crossbar arrays is only shown once in \cref{circt:lstm} but in reality those components are present $n_o$ times.
-Those extra components are needed in order for the parallel channels to work.
+The wires are all buses of different sizes. All the components expect from the crossbar, are only shown once but are in the real circuit present for every wire of the bus.
+
+\subsubsection{The $1-x$ operation}
+
+A little trick had to be emplyed in order to succesfully compute the function. The function is defined as in \cref{eq:1minus0}.
+\begin{equation}\label{eq:1minus0}
+  f(x)=1-x
+\end{equation}
+
+The solution found to do compute the operation consist of using an \ac{opAmp} as an inverter around $V_{inv}$, the formula is available in \cref{eq:1minus1}.
+\begin{equation}\label{eq:1minus1}
+  f_v(v)=-(v-V_{inv})+V_{inv}=2\cdot V_{inv} -v
+\end{equation}
+
+Writing the same operation in voltage gives us \cref{eq:1minus2}.
+
+\begin{equation}\label{eq:1minus2}
+  f_v(v)=(1-V_{cm})-(v-V_{cm})+V_{cm}= 0.1-v+2\cdot V_{cm}
+\end{equation}
+
+It can thus be found that the value for $V_{inv}$ is $V_{inv}=V_{cm}+0.05$
 
@@
 
-As opposed to the to the \ac{LSTM}, the \ac{GRU} can not be serialized. This is due to a mathematical issue.
-
+As opposed to the to the \ac{LSTM}, the \ac{GRU} can not be serialized.
+This is due to a mathematical issue.
 Indeed in the equation of the candidate hidden state is a problem for the serialization of the circuit.
@@
 
-That is why, the \ac{GRU} would take twice as much time to compute a time step if it were serialized. It is not worth it to double the computation time to save a bit of onChip area.
+That is why, the \ac{GRU} would take twice as much time to compute a time step if it were serialized. It is not worth it to double the computation time to save a small amount of onChip area.
 
@@
   \caption{Symbol used for the \ac{GRU} circuit}
-  \label{sym:lstm}
+  \label{sym:gru}
 \end{figure}

### 2023-10-28 design/main.tex
Forgot to rename the folders in the mains...
Bunch of changes, added design gru
Bunch of changes, to gru notably
Changed intro design

--- 2023-07-12 3.Design/main.tex
@@
 A step of one was chosen to be $0.1V$ in the analog circuit.
-Table \ref{tab:valConv} shows the conversions from a real number to it's voltage equivalent.
+\Cref{tab:valConv} shows the conversions from a real number to it's voltage equivalent.
 
@@
   \centering
-\caption{Real/Voltage Conversion Table.}
   \begin{tabular}{|c|c|}
@@
     \hline
-       0 & 0.9V \\
+    $0$ & $0.9V$ \\
     \hline
-         1 & 1.0V \\
+    $1$ & $1.0V$ \\
     \hline
-           $x$ & $\frac{x}{10}+V_{cm}$\\
+    $x$ & $V(x)=\frac{x}{10}+V_{cm}$\\
     \hline
-             $(V-0.9)\cdot 10$ & $V$\\
+    $real(v)=(v-0.9)\cdot 10$ & $v$\\
     \hline
   \end{tabular}
+  \caption{Real/Voltage Conversion Table.}
   \label{tab:valConv}
@@
 
-               \input{3.Design/activation.tex}
-               \input{3.Design/1.sectiona.tex}
-               \input{3.Design/2.sectionb.tex}
+In order to create a fully functional \ac{LSTM} or \ac{GRU} circuit, there are a few components that are required.
+
+An analog activation function circuit to have an analog equivalent to the activation function necessary to the good behavior of \acp{LSTM} and \acp{GRU}.
+
+A memory cell to allow to store the analog value from one time step to the next.
+
+A circuit for the crossbar array, to have a more general circuit that can easily be adapted.
+
+Other, less significant, components required for the circuits to work.
+
+\input{design/activation.tex}
+\input{design/memcell.tex}
+\input{design/crossbar.tex}
+\input{design/inverter.tex}
+\input{design/models.tex}
+\input{design/lstm.tex}
+\input{design/gru.tex}
+\input{design/genwei.tex}
+\input{design/netlist.tex}
+\input{design/wei2res.tex}
+\input{design/dataset.tex}
+\input{design/run.tex}
 \cleardoublepage

### 2023-10-28 conclusions/further.tex
Correcting inappropriate raccourcis and correcting design

@@
 The final circuit will need an external controller to manage diffenrent things in the circuit.
-This wasn't created in this thesis but will have to be in order to have a functionning circuit once fabricated.
+This was not created in this thesis but will have to be in order to have a functionning circuit once fabricated.
 
 The controller will increase the power consumption of the overall system.
-However, due to the slow (precision of about $1\mu s$) nature of the system requirements, the controller will be very low power, and won't affect the overall power consumption of the final system.
+However, due to the slow (precision of about $1\mu s$) nature of the system requirements, the controller will be very low power, and will not affect the overall power consumption of the final system.
 
@@
 Future work also also include looking for ways to change the weights in the fastest possible way.
-Idealy the algorithm won't need to compute the weight to resistance conversion (\cref{sec:wei2res}) and directly know how the change the resistances' values.
+Idealy the algorithm will not need to compute the weight to resistance conversion (\cref{sec:wei2res}) and directly know how the change the resistances' values.
 
@@
 \Cref{eq:wei2res0} needs to be respected in any situation.
-The memristor's resistances won't be centered (\cref{eq:wei2res1}) anymore.
+The memristor's resistances will not be centered (\cref{eq:wei2res1}) anymore.
 
@@
 
-Those solutions are purely theorical, they haven't been tested due to lack of time.
+Those solutions are purely theorical, they have not been tested due to lack of time.
 

### 2023-10-28 design/activation.tex
Changed intro design
Correcting inappropriate raccourcis and correcting design

@@
 Activation functions play a great role in the results obtained \cite{af}.
-It is the reason why making good Activation function circuits is fundamental.
+It is the reason why making good activation function circuits is fundamental.
 Of course the easy way would be to simply design a hard sigmoid (\cref{apsec:hardFunc}), the issue being that the hard sigmoid is much worse than the regular sigmoid, especially for regression problems \cite{hardSigm}.
@@
 These functions are still a bit different from the original functions (especially for the \ac{tanh}).
-However that doesn't matter too much as the trainning will be happening with the extracted analog functions, all weights will be set in the circuit.
-This is the reason why such a difference doesn't matter.
-As long as the curves have the similar shape, the result won't be drastically affected.
+However that does not matter too much as the trainning will be happening with the extracted analog functions, all weights will be set in the circuit.
+This is the reason why such a difference does not matter.
+As long as the curves have the similar shape, the result will not be drastically affected.

### 2023-10-28 design/crossbar.tex
New figures + trying to make good looking graph
New version of graoh
Corrected crosbar
Correcting inappropriate raccourcis and correcting design

@@
 This design is the one that is used in \cite{doubleMem}.
-Let's assume that a given memristor has a resistance range of $R\in[R_{min},R_{max}]$, that means it's conductance range is $\sigma \in [\sigma_{min},\sigma_{max}]$ (with $\sigma_{min}= \frac{1}{R_{max}}$ and $\sigma_{max}= \frac{1}{R_{min}}$).
+Let's assume that a given memristor has a resistance range of $R\in[R_{min},R_{max}]$, that means its conductance range is $\sigma \in [\sigma_{min},\sigma_{max}]$ (with $\sigma_{min}= \frac{1}{R_{max}}$ and $\sigma_{max}= \frac{1}{R_{min}}$).
 This design works using two \ac{opAmp} connected to $V_{cm}$ with the positive pin and the negative pin to the output of the crossbar array.
@@
   \centering
-  \includesvg[height=8cm,pretex=\scriptsize]{crossbar/doubleMem}
+  \includesvg[width=\textwidth,pretex=\scriptsize]{crossbar/doubleMem}
   \caption{Simplified circuit of a double memristor per synapse architecture}
@@
 
-With $x_k$ being the voltage for the input line $k$. The highest \ac{opAmp} is identified as opamp0 and the lowest opamp1.
+With $x_k$ being the voltage for the input line $k$. The highest \ac{opAmp} is identified as $opamp_0$ and the lowest $opamp_1$.
 

### 2023-10-28 design/dataset.tex
All afternoon work
Correcting inappropriate raccourcis and correcting design

@@
 
-To make sure that the terminal voltage doesn't reach the threshold as explained in \cref{sec:xbarCircuit}.
+To make sure that the terminal voltage does not reach the threshold as explained in \cref{sec:xbarCircuit}.
 The voltage threshold ($V_{threshold}$) was chosen to be $V_{threshold}=0.1V$ because the memristors used in this thesis are purely theorical.
@@
 The dataset then contains twelve years of monthly data so $144$ sample points.
-The results of such a problem aren't useful, it was used in this thesis because of its simplicity to check whether the analog circuit is working.
+Although the esults of this problem are not necessarely tailored to the system, in the sense that it is not time sensitive information, it is nonetheless a problem that can, in a straightforward manner, demonstrate the predictive capacity of the proposed model.
 It is extracted from the tutorial available at \cite{airline}.
@@
 
-This dataset is far more interresting than the latter.
+This dataset is far more interesting and complex than the latter.
 This data set aims to use \acp{LSTM} to mimic the behavior of real neurons.
@@
 \begin{itemize}
-  \item It comes from a very recent paper, and means the research is current.
+  \item It comes from a very recent paper from the research group in which this work is also being developed.
   \item It aims at reproducing the behavior of the brain of a simple organism (\ac{C. elegans}). In the (very) long term, a full parts of the human brain could be replaced by a very low powered chip.
-  \item The paper was written by Dr. Barbulescu who happens to be one of my supervisor.
 \end{itemize}
@@
     \hfill
-    \subfloat[Validation dataset sample]{\includesvg[width=0.5\textwidth,pretex=\tiny]{datasets/celegans/ioValid0}}%
+    \subfloat[Validation dataset sample\label{graph:io15Celegans}]{\includesvg[width=0.5\textwidth,pretex=\tiny]{datasets/celegans/ioValid0}}%
   \end{minipage}
@@
 
-\Cref{graph:celegans} shows four sets of input/output our of the fourty that were recorded for the dataset.
+\Cref{graph:celegans} shows four sets of input/output out of the fourty that were recorded for the dataset.
 

### 2023-10-28 design/genwei.tex
Correcting inappropriate raccourcis and correcting design

@@
 The choice for the weight generation was to use Keras API with the tensorflow framework.
-However, both tensorflow and pyTorch were tried and were giving similar results, the final choice of using tensorflow was made because it's the most popular among the research group.
+However, both tensorflow and pyTorch were tried and were giving similar results, the final choice of using tensorflow was made because it is the most popular among the research group.
 
@@
   \item The \ac{NP} \ac{LSTM} : This is most commonly used version and by such one of the two available in Keras.
-  \item The \ac{GRU} : This one is present here because it's considered to be different from an \ac{LSTM} \ac{NN} despite their similarities.
+  \item The \ac{GRU} : This one is present here because it is considered to be different from an \ac{LSTM} \ac{NN} despite their similarities.
 \end{itemize}
@@
 To avoid any unwanted behavior when the system supposedly goes out of range, the weights in the \ac{LSTM} are contrained.
-This constrain isn't fixed and depends on the architecture it is generated with.
+This constrain is not fixed and depends on the architecture it is generated with.
 

### 2023-10-28 design/lstm.tex
Correcting inappropriate raccourcis and correcting design

@@
 When the system is running is a fully parallel mode, doubling the memory cells still causes a problem.
-The enable in and enable out flags can't be high at the same time when the memory cell has a feedback connection, the signal would interfere with it self and change its own value.
+The enable in and enable out flags cannot be high at the same time when the memory cell has a feedback connection, the signal would interfere with it self and change its own value.
 Potential solutions are discussed in \cref{subsec:noDoubleMemcell}.

### 2023-10-28 design/models.tex
Correcting inappropriate raccourcis and correcting design

--- 2023-09-22 3.Design/models.tex
@@
 Due to a lack of time, some of the more common component were not designed by me but instead were simulated using a Verilog model.
+Verilog is a hardware description laguage very popular in the industry.
 The only components that use a verilog model are the voltage multiplier and the \ac{opAmp}.
 
-\subsection{Operational amplifier}
+\subsection{Operational amplifier}\label{subsec:opamp}
 
 This component is the very famous \ac{opAmp}.
-This specific component wasn't designed for the thesis, it required a current range that didn't allow to use ones already made by members of the research group.
+This specific component was not designed for the thesis, it required a current range that did not allow to use ones already made by members of the research group.
 The choice was then to use a verilog-A model, and then design one if time allows it.
@@
 
-\Cref{eq:opAmp} is the equation used in the verilog-A model. It makes it so it behaves like an ideal \ac{opAmp}.
+\Cref{eq:opAmp} is the equation used in the verilog-A model. It makes it so it behaves like an ideal \ac{opAmp}. For this thesis $\mu$ has been set to $\mu=10^5$.
 
-\subsection{Voltage multiplier}
+\subsection{Voltage multiplier}\label{subsec:voltmult}
 

### 2023-10-28 design/netlist.tex
Correcting inappropriate raccourcis and correcting design

@@
 This is done by generating a SPICE netlist using a python script.
-A SPICE netlist was chosen because of Cadence's virtuoso limitations.
-Indeed, Verilog netlists can be imported with a downside, component's parameters can't be imported.
+A SPICE netlist was chosen because of \textbf{Cadence}'s \textit{Virtuoso} limitations.
+Indeed, Verilog netlists can be imported with a downside, component's parameters cannot be imported.
 This is very limiting for this thesis' use case (required to set the resistors' resistance and thus the weights).

### 2023-10-28 design/run.tex
Correcting inappropriate raccourcis and correcting design

@@
 
-The simulation is ran using Cadence's virtuoso simulator, using a corner simulation to simulate all the inputs.
-
-In case the \ac{LSTM} we want to run requires several outputs, the use of the corner functionality.
-Corners are intended to be used for small variations of the variables in the circuit.
-However, it used differently in this thesis.
-It runs with the value for each input variable changing for the current \ac{LSTM} run and then moves on to the next with the variables changed.
+The simulation is ran using \textbf{Cadence}'s \textit{Virtuoso} simulator, using a parametric simulation to simulate all the inputs.
+The use of the parametric simulation allows to change the values for the data we want to use as input.
 
@@
 The time step ($T$) value was chosen to be $T=8\mu s$, the pause between two \ac{LSTM} steps is set to be $\frac{T}{8}=1\mu s$.
-The value for the maximum and minimum resistances value of the memristor used ($R_{max}$ and $R_{min}$, respectively) are set to an arbitrary value of $R_{max}=10^6 \Omega=1M\Omega$ and $R_{min}=10^4 \Omega = 10k\Omega$.
+The value for the maximum and minimum resistances value of the memristor used ($R_{max}$ and $R_{min}$, respectively) are set to be in acordance with the \textbf{KNOWN}'s memristors.
+The resistance values used for the simulation are thus $R_{max}=10^6 \Omega=1M\Omega$ and $R_{min}=10^4 \Omega = 10k\Omega$.
 

### 2023-10-28 design/wei2res.tex
Correcting inappropriate raccourcis and correcting design

@@
 
-This chapter will contain all the information on the weight conversion to a resistance value.
-
-The weight to resistances equations has been found using simple math.
+The weight to resistances equations has been found using \cref{eq:wei2res0}.
 
@@
 This is done by centering the resistances around $R_f$, this means that \cref{eq:wei2res1} is the second equation, that makes our problem now solvable.
-%TODO explain how in future work the weight doesn't have to be centered around R_f, to gain time only one of the two resistance could change
+
 \begin{equation}
@@
 
-The real value to voltage conversion (\cref{tab:valConv}) doesn't affect this part of the system as the crossbar array already works in voltages.
+The real value to voltage conversion (\cref{tab:valConv}) does not affect this part of the system as the crossbar array already works in voltages.
 Replacing the resistor variable in \cref{eq:doubleMem2} by $w$ using \cref{eq:wei2res0} would not change the results.
@@
 
-For the thesis, this step is done in python and integrated in the netlist generator script (\cref{sec:netlist}).
+In the present work, this step is done in python and integrated in the netlist generator script (\cref{sec:netlist}).
 The resolution of the memristor (the precision of the resistances at which the memristor can be set) is simulated by limiting the generated resistances values to having only two significant figures.

### 2023-10-28 introduction/contributions.tex
Correcting inappropriate raccourcis and correcting design

@@
   \item I created a script that can generate spice netlists for \acps{LSTM} or feed forward \aclp{NN} of any size and importing weights as resistance values from a specified file containing the architecture and the weights.
-  \item The script can generate an \ac{LSTM} capable analog circuits and then run a simulation of the circuit in virtuoso giving encouraging results.
-  \item The script can also generate an \ac{GRU} capable analog circuits and then run a simulation of the circuit in virtuoso.
+  \item The script can generate an \ac{LSTM} capable analog circuits and then run a simulation of the circuit in \textit{Virtuoso} giving encouraging results.
+  \item The script can also generate an \ac{GRU} capable analog circuits and then run a simulation of the circuit in \textit{Virtuoso}.
   \item I developed an analog activation function circuit that supports \ac{tanh} and sigmoid ranges working at $1.8V$.
@@
   \item I coded in python \cite{python} (tensorflow library\cite{tensorflow}) to create a piece wise functions to reproduce in the tensorflow the analog activation functions.
-  \item I used Cadence's Virtuoso to run software simulation of different \acp{NN} capable analog circuits.
+  \item I used \textbf{Cadence}'s \textit{Virtuoso} to run software simulation of different \acp{NN} capable analog circuits.
   \item A paper is being prepared to report the advancements made in this work.

### 2023-10-28 introduction/motivation.tex
Correcting inappropriate raccourcis and correcting design

@@
 
-The focus of this thesis is to create a working simulation of an analog \ac{LSTM} and \ac{GRU} circuit using Cadence's Virtuoso as simulation software.
+The focus of this thesis is to create a working simulation of an analog \ac{LSTM} and \ac{GRU} circuit using \textbf{Cadence}'s \textit{Virtuoso} as simulation software.
 Having those preliminary results would allow to fabricate and test such a chip.

### 2023-10-28 results/airline.tex
Correcting inappropriate raccourcis and correcting design

@@
 
-Computing the \ac{NN} with or without the analog activation functions doesn't seem to affect the error of the predictions, experience has shown that the \ac{LSTM} using the analog activation functions seem to predict higher curves than the target and the \ac{LSTM} using the default activation functions tends to predict curves that are under the target, the reasons for such a difference are unknown.
+Computing the \ac{NN} with or without the analog activation functions does not seem to affect the error of the predictions, experience has shown that the \ac{LSTM} using the analog activation functions seem to predict higher curves than the target and the \ac{LSTM} using the default activation functions tends to predict curves that are under the target, the reasons for such a difference are unknown.
 This is simply an observation after a few different runs.
@@
 As expected, the analog predictions are very far off the target curve.
-Which, in this case, isn't the target dataset but the digital prediction ran with the default activation functions.
+Which, in this case, is not the target dataset but the digital prediction ran with the default activation functions.
 Indeed, the goal of the analog system is to reproduce as closely as possible the digital prediction.
@@
 
-However, the prediction don't seem to be completely off. The curves seem to follow the shape of it's target, and by extension of the target data. This deserves a closer look.
+However, the prediction do not seem to be completely off. The curves seem to follow the shape of its target, and by extension of the target data. This deserves a closer look.
 
@@
 The overall observation of those results compared to the previous ones (\cref{subsec:airlineAnalogNoC}) are how much closer to their target they are.
-It still isn't quite perfect but simply changing the activation functions used for training improved the results by a lot.
+It still is not quite perfect but simply changing the activation functions used for training improved the results by a lot.
 It can thus be theorized that the error will become null if the system is trained with the same parameters as the ones it is using for inference.

### 2023-10-28 results/celegans.tex
Correcting inappropriate raccourcis and correcting design

@@
 The results will be analyzed by focusing on two data sequence, the first being simple inputs expecting simple outputs and the second being more complex, those sequences are the ones in \cref{graph:io5Celegans,graph:io15Celegans}.
-We'll identify those as sequence 5 and sequence 15 respectively.
+We will identify those as sequence 5 and sequence 15 respectively.
 
@@
 
-It's been established in \cref{sec:resAirline}, that the analog circuit with a serial size of one ($n_s=1$) show a lower \ac{RMSE} to the digital predictions than the digital predictions.
+It has been established in \cref{sec:resAirline}, that the analog circuit with a serial size of one ($n_s=1$) show a lower \ac{RMSE} to the digital predictions than the digital predictions.
 
@@
 
-This phenomenen is very strange and hasn't been identified so far.
-The main theory is that the data isn't stored properly in the memory cells, thus at every \ac{LSTM} time step the supposed output is slightly altered.
+This phenomenen is very strange and has not been identified so far.
+The main theory is that the data is not stored properly in the memory cells, thus at every \ac{LSTM} time step the supposed output is slightly altered.
 This would create this flickering of the outputed data.
 
-The data isn't useless for that matter.
+The data is not useless for that matter.
 It was decided to smooth the data by averaging every output point with its neighboring points.
@@
 Adjusting the results reduces the error a bit, mainly due to the sliding of the results.
-Smoothing out the curve doesn't affect the error as it simply does a local average of each value.
+Smoothing out the curve does not affect the error as it simply does a local average of each value.
 The erros are, once again higher when obtained with the analog simulation than with the digital inference.

### 2023-10-28 state/model.tex
Correctiion diogo state
Correcting inappropriate raccourcis and correcting design

@@
 
-They are several ways to model a memristor \cite{memristorFab,memTEAMmodel, memVTEAMmodel, memristorSpiceModels}.
+There are several ways to model a memristor \cite{memristorFab,memTEAMmodel, memVTEAMmodel, memristorSpiceModels}.
 The way the memristor works depends on the materials used to fabricate the memristor.
 Each type of memristor has slightly different behavior.
-Models were thus created to simulate, as accurately as possible, the memristor in electrical circuit simulation software such as \textbf{Cadence}'s \textit{virtuoso} or SPICE.
+Models were thus created to simulate, as accurately as possible, the memristor in electrical circuit simulation software such as \textbf{Cadence}'s \textit{Virtuoso} or SPICE.
 
@@
 They represent the boundaries of the state variable.
-All the other parameters in \cref{eq:memModel0,eq:memModel1,eq:memModel2,eq:memModel3} that haven't been declared are free fitting variables.
+All the other parameters in \cref{eq:memModel0,eq:memModel1,eq:memModel2,eq:memModel3} that have not been declared are free fitting variables.
 They have to be set depending on the type of memristor used.
@@
 The model now needs to be adapted to work as Verilog-A code.
-Verilog-A is a hardware desciption language, very popular in the industry for its simplicity and flexibility when using it with circuit simulator such as \textbf{Cadence}'s \textit{virtuoso}.
+Verilog-A is a hardware desciption language, very popular in the industry for its simplicity and flexibility when using it with circuit simulator such as \textbf{Cadence}'s \textit{Virtuoso}.
 

### 2023-10-28 appendices/appendixA.tex
corrected results

--- 2023-10-17 Appendices/appendixA.tex
@@
 
-This is the circuit that comes after the voltage multiplier (section \ref{sec:models}), it is here to amplify the results of the voltage multiplier by 10.
+This is the circuit that comes after the voltage multiplier (\cref{sec:models}), it is here to amplify the results of the voltage multiplier by 10.
 Sometimes, its use doubles as an adder making it an inverting summing amplifier (TODO circuit).
@@
   \begin{cases}
-    0, \hfill x< -2.5\\
-    \frac{x}{5}+\frac{1}{2}, \hspace{0.2cm} 2.5>x>-2.5\\
-    1, \hfill x> 2.5\\
+    0& x< -2.5\\
+    \frac{x}{5}+\frac{1}{2}&  2.5>x>-2.5\\
+    1& x> 2.5\\
   \end{cases}
@@
   \begin{cases}
-    -1, \hfill x<-1\\
-    x, \hspace{0.2cm} 1>x>-1\\
-    1, \hfill x>1\\
+    -1& x<-1\\
+    x& 1>x>-1\\
+    1& x>1\\
   \end{cases}
@@
 
-Some versions of the hard sigmoid are a bit different from \cref{eq:hSigmoid}, the slope depends on the library used \cite{hSigmoid1}.
+Some versions of the hard sigmoid are defined differently from \cref{eq:hSigmoid}, the slope depends on the library used \cite{hSigmoid1}.
 
@@
 A visual representation helps seeing how close those functions are, \cref{graph:hardFunc} displays the two together with the function they are modeled after.
+
+\section{Smoothing the curve}\label{apsec:smoothFunc}
+
+\begin{lstlisting}[language=Python]
+Put your code here.
+\end{lstlisting}

### 2023-10-28 results/airline.tex
corrected results

@@
 Interestingly, the results look much better when using the \ac{GRU} layer.
-The error rate confirms that as it is almost half of what it was with the \ac{LSTM} layer (\cref{tab:airlineDigital}).
+The error rate of $36,6$ confirms that as it is almost half of what it was with the \ac{LSTM} layer (\cref{tab:airlineDigital}).
 That simply means that the \ac{GRU} layer is better at handling this problem, as was previously stated in \cref{sec:gru}.
@@
 This subsection shows the importance of using the analog activation functions in the training phase of the \ac{NN}.
-
 The \ac{LSTM} layer being the only to have been trained using the default activation functions the only results presented here are using an \ac{LSTM} layer.
@@
   \includesvg[width=\linewidth]{results/airline/noCustomAFZoomed}
-  \caption{Analog predictions trained with the default activation functions.}
+  \caption{Analog predictions trained with the default activation functions zoomed on the the analog predictions.}
   \label{graph:airlineAnalogNoCZoomed}
@@
 This similarity is further proven by the errors shown in \cref{tab:airlineAnalogNoCZoomed}.
+The erros are $3.97$, $4.78$ and $5.05$.
 Those error are all around four thousands passengers, which is negligible in our case.
@@
 It shows that, as can be observed, the results the closest to the target are the ones generated using a serial size of one ($n_s=1$).
-Indeed, its error ($28.41$) is lower than the error from the digital results ($52.27$) to the original target values, almost half of it.
+Indeed, its error ($28.4$) is lower than the error from the digital results ($52.3$) to the original target values, almost half of it.
 The other ones are still very close to the targeted curve, but are still lower.
-This is probably due to the serialization circuit causing small differences in the signals at each steps.
+The errors are higher when running the system in serial mode.
+This higher inaccuracy is probably due to the data staying for longer in the memory cells as the time steps get longer.
 
@@
 Once again, the results are quite close to each others, but as expected the errors (\cref{tab:airlineAnalogError}) are higher then the ones found in \cref{tab:airlineAnalogNoCZoomed}.
+This is due to the fact that the errors depend on the scale of the results, the predcitions here having a much greater range (\cref{tab:airlineAnalog}).
 The farthest appart ($n_s=1$ to $n_s=2$ with an error of $64.6$ thousands passengers) have an error barely higher than the error of digital results to the target dataset (\cref{tab:airlineDigital})
-
-coincidentally, the curve is closer to the target curve than the digital prediction itself, this is mostly due to the lack of precision of the analog prediction, but it is nevertheless interresting to mention.
-\Cref{graph:airlineCoin} shows the graph and the \ac{RMSE} is the lowest, even lower than the ones from the digital predictions.
-
-\begin{figure}[H]
-  \centering
-  \includesvg[width=\linewidth]{results/airline/closeRes}
-  \caption{Analog prediction ($n_s=1$) with the target dataset.}
-  \label{graph:airlineCoin}
-\end{figure}
 
@@
 It still is not quite perfect but simply changing the activation functions used for training improved the results by a lot.
-It can thus be theorized that the error will become null if the system is trained with the same parameters as the ones it is using for inference.
-In other words, the error will probably be nullified when the training will happen inSitu (\cref{subsec:inSitu}).
+It can thus be theorized that the error will greatly decrease if the circuit is trained with the same parameters as the ones it is using for inference.
+In other words, the error will probably be very low when the training will happen inSitu (\cref{subsec:inSitu}).
 

### 2023-10-28 results/celegans.tex
corrected results

@@
 
-It has been established in \cref{sec:resAirline}, that the analog circuit with a serial size of one ($n_s=1$) show a lower \ac{RMSE} to the digital predictions than the digital predictions.
+It has been established in \cref{sec:resAirline}, that the analog circuit with a serial size of one ($n_s=1$) shows a lower \ac{RMSE} to the digital predictions than the digital predictions.
 
@@
 \Cref{graph:celegansAnalog0} contains the predictions of both the digital and analog \ac{NN} for all four output neurons that are being monitored.
-The having all predicted responses to the stimuli on the same graph makes things very unintelligeable.
+\Cref{graph:celegansAnalog0} having all predicted responses to the stimuli on the same graph makes things very unintelligeable.
 
@@
 This simple average will smooth out the curves to get a decently readable result, that can then be compared with the digitally predicted response.
-The python code for this function is available at \cref{}.
-%TODO
+The python code for this function is available in \cref{apsec:smoothFunc}.
 
@@
 The \cref{bar:rmse5} shows that adjusted or not, the errors are in the expected range.
-Adjusting the results reduces the error a bit, mainly due to the sliding of the results.
+Adjusting the results slightly reduces the error, mainly due to the sliding of the results.
 Smoothing out the curve does not affect the error as it simply does a local average of each value.
@@
   \caption{\acs{C. elegans} analog responses with their digital counterparts}
-  \label{graph:celegansAnalog0}
+  \label{graph:celegansAnalog1}
 \end{figure}
 
-\Cref{graph:celegansAnalog0} contains the predictions of both the digital and analog \ac{NN} for all four output neurons that are being monitored.
-The having all predicted responses to the stimuli on the same graph makes things very unintelligeable.
+\Cref{graph:celegansAnalog1} contains the predictions of both the digital and analog \ac{NN} for all four output neurons that are being monitored.
+For the reasons mentioned in the previous section, for the sake of readability, the predictions are spread out on separate graphs.
 

### 2023-10-28 code/smooth.py
Started adding code

@@ -1,0 +1,11 @@
+def smooth_line(df, n):
+    size = len(df)
+    out = df.copy()
+    for i in range(size):
+        if (i < n):
+            out[i] = df[0:i+n].mean()
+        elif (size-i-1 < n):
+            out[i] = df[i-n:size-1].mean()
+        else:
+            out[i] = df[i-n:i+n].mean()
+    return out

### 2023-10-28 conclusions/further.tex
Changed color and removed a further part

@@
 
-\subsection{Permanent training}
-
-Once the training is fully implemented in an analog computer, the next step would be to find a way to permanently training the system.
-Meaning that any new information will be used to update the weights.
-This means that the system would need to know if the prediction it has made is correct or not.
-
-A potential way to implement that, would be in a two phase plan.
-First the prediction, and so on until a verification on the predictions can be done.
-After that the weights are updated based on those previous reviewed results.
-
 \subsection{Other \acp{LSTM} variants}

### 2023-10-28 appendices/circuits.tex
Working on appendices

--- 2023-10-17 Appendices/appendixA.tex
@@
-\fancychapter{TODO determine name}
-\label{ap:a}
+\fancychapter{Less important circuits}
+\label{ap:circuit}
 
@@
 
-This is the circuit that comes after the voltage multiplier (section \ref{sec:models}), it is here to amplify the results of the voltage multiplier by 10.
+This is the circuit that comes after the voltage multiplier (\cref{sec:models}), it is here to amplify the results of the voltage multiplier by 10.
 Sometimes, its use doubles as an adder making it an inverting summing amplifier (TODO circuit).
-
-\section{Solving weight to resistance}\label{apsec:wei2res}
-
-\begin{equation}\label{eq:wei2res3}
-  \begin{cases}
-    w=\frac{R_f}{R_+}-\frac{R_f}{2\cdot R_f -R_+}\\
-    R_-=2\cdot R_f -R_+
-  \end{cases}
-\end{equation}
-
-\begin{equation}\label{eq:wei2res4}
-  \begin{cases}
-    w=\frac{2\cdot R_f^2-2\cdot R_+\cdot R_f}{2\cdot R_f\cdot R_+ -R_+^2}\\
-    R_-=2\cdot R_f -R_+
-  \end{cases}
-\end{equation}
-
-\begin{equation}\label{eq:wei2res5}
-  \begin{cases}
-    2\cdot R_f\cdot R_+\cdot w -R_+^2\cdot w=2\cdot R_f^2-2\cdot R_+\cdot R_f\\
-    R_-=2\cdot R_f -R_+
-  \end{cases}
-\end{equation}
-
-\begin{equation}\label{eq:wei2res6}
-  \begin{cases}
-    R_+^2\cdot w - R_+\cdot 2\cdot R_f \cdot(w+1) + 2R_f^2 = 0\\
-    R_-=2\cdot R_f -R_+
-  \end{cases}
-\end{equation}
-
-\begin{equation}\label{eq:wei2res61}
-  R_+^2\cdot w - R_+\cdot 2\cdot R_f \cdot(w+1) + 2R_f^2 = 0
-\end{equation}
-
-We must now solve \cref{eq:wei2res61}.
-
-\begin{equation}\label{eq:wei2res61}
-  \Delta = 4\cdot R_f^2\cdot(w+1)^2-8\cdot w\cdot R_f^2 = 4\cdot R_f^2\cdot(w^2+1) \ge 0
-\end{equation}
-
-Giving the two solutions :
-
-\begin{equation}\label{eq:wei2res62}
-  R_{+,0}=\frac{R_f\cdot(w+1)-R_f\cdot\sqrt{w^2+1}}{w}
-\end{equation}
-\begin{equation}\label{eq:wei2res63}
-  R_{+,1}=\frac{R_f\cdot(w+1)+R_f\cdot\sqrt{w^2+1}}{w}
-\end{equation}
-
-\section{Hard sigmoid and \ac{tanh} functions}\label{apsec:hardFunc}
-
-The hard sigmoid and \ac{tanh} functions are simplifications of the the classic sigmoid and \ac{tanh} used for faster computation.
-They come with a downside, they are less efficient then their original functions.
-
-They are defined as \cref{eq:hSigmoid} for the hard sigmoid and \cref{eq:hTanh} for the hard \ac{tanh} \cite{hSigmoid, hTanh}.
-
-\begin{equation}\label{eq:hSigmoid}
-  \sigma (x) =
-  \begin{cases}
-    0, \hfill x< -2.5\\
-    \frac{x}{5}+\frac{1}{2}, \hspace{0.2cm} 2.5>x>-2.5\\
-    1, \hfill x> 2.5\\
-  \end{cases}
-\end{equation}
-\begin{equation}\label{eq:hTanh}
-  tanh(x) =
-  \begin{cases}
-    -1, \hfill x<-1\\
-    x, \hspace{0.2cm} 1>x>-1\\
-    1, \hfill x>1\\
-  \end{cases}
-\end{equation}

### 2023-10-28 appendices/codes.tex
Working on appendices

--- 2023-10-17 Appendices/appendixA.tex
@@
-\fancychapter{TODO determine name}
-\label{ap:a}
+\fancychapter{Important codes}
+\label{ap:codes}
 
-\section{Buffer}\label{apsec:buffer}
+\section{Smoothing the curve}\label{apsec:smoothFunc}
 
-This is a classic circuit that only takes one \ac{opAmp} wired as like in figure TODO. It is used to forward a voltage without impacting the original voltage.
-
-\section{Inverting amplifier}\label{apsec:invAmp}
-
-This is the circuit that comes after the voltage multiplier (section \ref{sec:models}), it is here to amplify the results of the voltage multiplier by 10.
-Sometimes, its use doubles as an adder making it an inverting summing amplifier (TODO circuit).
-
-\section{Solving weight to resistance}\label{apsec:wei2res}
-
-\begin{equation}\label{eq:wei2res3}
-  \begin{cases}
-    w=\frac{R_f}{R_+}-\frac{R_f}{2\cdot R_f -R_+}\\
-    R_-=2\cdot R_f -R_+
-  \end{cases}
-\end{equation}
-
-\begin{equation}\label{eq:wei2res4}
-  \begin{cases}
-    w=\frac{2\cdot R_f^2-2\cdot R_+\cdot R_f}{2\cdot R_f\cdot R_+ -R_+^2}\\
-    R_-=2\cdot R_f -R_+
-  \end{cases}
-\end{equation}
-
-\begin{equation}\label{eq:wei2res5}
-  \begin{cases}
-    2\cdot R_f\cdot R_+\cdot w -R_+^2\cdot w=2\cdot R_f^2-2\cdot R_+\cdot R_f\\
-    R_-=2\cdot R_f -R_+
-  \end{cases}
-\end{equation}
-
-\begin{equation}\label{eq:wei2res6}
-  \begin{cases}
-    R_+^2\cdot w - R_+\cdot 2\cdot R_f \cdot(w+1) + 2R_f^2 = 0\\
-    R_-=2\cdot R_f -R_+
-  \end{cases}
-\end{equation}
-
-\begin{equation}\label{eq:wei2res61}
-  R_+^2\cdot w - R_+\cdot 2\cdot R_f \cdot(w+1) + 2R_f^2 = 0
-\end{equation}
-
-We must now solve \cref{eq:wei2res61}.
-
-\begin{equation}\label{eq:wei2res61}
-  \Delta = 4\cdot R_f^2\cdot(w+1)^2-8\cdot w\cdot R_f^2 = 4\cdot R_f^2\cdot(w^2+1) \ge 0
-\end{equation}
-
-Giving the two solutions :
-
-\begin{equation}\label{eq:wei2res62}
-  R_{+,0}=\frac{R_f\cdot(w+1)-R_f\cdot\sqrt{w^2+1}}{w}
-\end{equation}
-\begin{equation}\label{eq:wei2res63}
-  R_{+,1}=\frac{R_f\cdot(w+1)+R_f\cdot\sqrt{w^2+1}}{w}
-\end{equation}
-
-\section{Hard sigmoid and \ac{tanh} functions}\label{apsec:hardFunc}
-
-The hard sigmoid and \ac{tanh} functions are simplifications of the the classic sigmoid and \ac{tanh} used for faster computation.
-They come with a downside, they are less efficient then their original functions.
-
-They are defined as \cref{eq:hSigmoid} for the hard sigmoid and \cref{eq:hTanh} for the hard \ac{tanh} \cite{hSigmoid, hTanh}.
-
-\begin{equation}\label{eq:hSigmoid}
-  \sigma (x) =
-  \begin{cases}
-    0, \hfill x< -2.5\\
-    \frac{x}{5}+\frac{1}{2}, \hspace{0.2cm} 2.5>x>-2.5\\
-    1, \hfill x> 2.5\\
-  \end{cases}
-\end{equation}
-\begin{equation}\label{eq:hTanh}
-  tanh(x) =
-  \begin{cases}
-    -1, \hfill x<-1\\
-    x, \hspace{0.2cm} 1>x>-1\\
-    1, \hfill x>1\\
-  \end{cases}
-\end{equation}
+\pythonstyle
+\lstinputlisting[language=python]{code/smooth.py}

### 2023-10-28 appendices/other.tex
Working on appendices

--- 2023-10-28 appendices/appendixA.tex
@@
-\fancychapter{TODO determine name}
-\label{ap:a}
-
-\section{Buffer}\label{apsec:buffer}
-
-This is a classic circuit that only takes one \ac{opAmp} wired as like in figure TODO. It is used to forward a voltage without impacting the original voltage.
-
-\section{Inverting amplifier}\label{apsec:invAmp}
-
-This is the circuit that comes after the voltage multiplier (\cref{sec:models}), it is here to amplify the results of the voltage multiplier by 10.
-Sometimes, its use doubles as an adder making it an inverting summing amplifier (TODO circuit).
+\fancychapter{Formulas and functions}
+\label{ap:other}
 
@@
 A visual representation helps seeing how close those functions are, \cref{graph:hardFunc} displays the two together with the function they are modeled after.
-
-\section{Smoothing the curve}\label{apsec:smoothFunc}
-
-\begin{lstlisting}[language=Python]
-Put your code here.
-\end{lstlisting}

### 2023-10-28 appendices/circuits.tex
New file plus working on circuits

@@
 
-This is a classic circuit that only takes one \ac{opAmp} wired as like in figure TODO. It is used to forward a voltage without impacting the original voltage.
+This is a classic circuit that only takes one \ac{opAmp} wired as \cref{fig:buffer}. It is used to forward a voltage without interfering the node of the input voltage.
+
+\begin{figure}[H]
+\centering
+\includesvg[width=.6\textwidth]{appendix/buffer}
+\caption{Buffer circuit}
+\label{fig:buffer}
+\end{figure}
+
+The input is $V_{in}$ and the output is $V_{out}$.
 

### 2023-10-28 appendices/circuits.tex
Appendix circuit + fixed gru circuit issue

@@
 
-This is a classic circuit that only takes one \ac{opAmp} wired as \cref{fig:buffer}. It is used to forward a voltage without interfering the node of the input voltage.
+This is a classic circuit that only takes one \ac{opAmp} wired as \cref{circt:buffer}. It is used to forward a voltage without interfering the node of the input voltage.
 
@@
   \caption{Buffer circuit}
-\label{fig:buffer}
+  \label{circt:buffer}
 \end{figure}
@@
 
-\section{Inverting amplifier}\label{apsec:invAmp}
+\section{Summing inverter}\label{apsec:sumInv}
 
-This is the circuit that comes after the voltage multiplier (\cref{sec:models}), it is here to amplify the results of the voltage multiplier by 10.
-Sometimes, its use doubles as an adder making it an inverting summing amplifier (TODO circuit).
+The summing amplifier in a very general circuit that is used in a plethera of different situations. A modulable circuit with a variable number of input ($n$) in available in \cref{circt:sumInv}.
+
+\begin{figure}[H]
+  \centering
+  \includesvg[width=.6\textwidth]{appendix/invAmp}
+  \caption{Summing inverter circuit}
+  \label{circt:sumInv}
+\end{figure}
+
+This circuit links the inputs and its output with \cref{eq:sumInv}.
+
+\begin{equation}\label{eq:sumInv}
+  V_{out}=V_{cm}-R\cdot\sum_{i=0}^{n-1}\frac{V_i-V_{cm}}{R_i}
+\end{equation}
+
+In this work itself it is used in at least three different use cases :
+
+\begin{itemize}
+  \item As a voltage inverter around $V_{cm}$, using a single input ($n=1$) and using $R_0=R$. This is used in \cref{sec:af}.
+  \item As a inverting amplifier after a voltage multiplier (\cref{subsec:voltmult}). Using one input ($n=1$) and $R>R_0$. It can be found in \cref{sec:lstmCircuit,sec:gruCircuit}.
+  \item It is also used as a summing inverter with amplification after a voltage multiplier.
+Using two inputs ($n=2$) and $R_0=R_1<R$ for the resistances.
+This is also found in the circuit in \cref{sec:lstmCircuit,sec:gruCircuit}.
+\end{itemize}

### 2023-10-28 design/activation.tex
Fixed cafouillage

@@
 
-Note that the actual output of the circuit in \cref{circt:af} is inverted arround $V_{cm}$ a simple circuit such as the one shown in  (TODO show circuit in annex) does inverts the
+Note that the actual output of the circuit in \cref{circt:af} is inverted arround $V_{cm}$ a circuit such as the one shown in \cref{apsec:sumInv} inverts the the signal to its intended values.
 

### 2023-10-29 conclusions/conclusion.tex
Correcting inappropriate raccourcis and correcting design
added stuff to conclusion

@@
 Indeed, when working with sequence 15 (\cref{graph:io15Celegans}), the inaccuracies spiral up and gives out a barely usable prediction.
-The curve generated doesn't fit its digital counterpart anymore.
+The curve generated does not fit its digital counterpart anymore.
 
@@
 Once those issues are dealt with and have been fixed, the \ac{GRU} circuit will be ready for an analog training as well.
+
+\subsection{Execution time}
+
+The inference time is quite tricky to deal with.
+While the analog circuit's inference time is set, it gets more complicated when trying to get the digital inference time.
+The digital inference time depends on the computer it is being run on.
+The \acp{CPU} used in embedded sytems, the main application for this work, are usually not very fast as they are intended to be low powered.
+This parts will thus only dicuss of the inference of the analog circuit.
+
+\subsubsection{Airline inference}
+
+The airline problem feeding two inputs to the circuit to get a single output, it will have to be run $144$ times to get all the output data.
+The inference time depends on the serial size being used :
+
+\begin{itemize}
+  \item $n_s=1$ : The output can be read as a voltage on the output net from $t_0=18\mu s$ to $t_1=22\mu s$.
+  \item $n_s=2$ : The output can be read as a voltage on the output net from $t_0=34\mu s$ to $t_1=38\mu s$.
+  \item $n_s=4$ : The output can be read as a voltage on the output net from $t_0=66\mu s$ to $t_1=70\mu s$.
+\end{itemize}
+
+Those are the times it takes to compute one of the outputs. It takes $144$ times longer to get all of the results.
+
+\subsubsection{\ac{C. elegans} inference}
+
+This sequence is a bit more complicated, for every input sequence with a thousand time steps, there are a thousand output vectors being read. Here, the inference time also depends on the serial size.
+
+\begin{table}[H]
+  \centering
+  \begin{tabular}{|c|c|c|c|c|}
+    \hline
+    \rowcolor{gray}
+    Serial size & $n_s=1$ & $n_s=2$ & $n_s=4$ & $n_s=8$\\
+    \hline
+    Start of first output & $9\mu s$ & $17\mu s$ & $33\mu s$ & $65\mu s$\\
+    \hline
+    End of first output & $13\mu s$ & $21\mu s$ & $37\mu s$ & $69\mu s$\\
+    \hline
+    Output period & $9\mu s$ & $17\mu s$ & $33\mu s$ & $65\mu s$\\
+    \hline
+    Start of last output & $9ms$ & $17ms$ & $33ms$ & $65ms$\\
+    \hline
+    End of last output & $9ms +4\mu s$ & $17ms+4\mu s$ & $33ms+4\mu s$ & $65ms+4\mu s$\\
+    \hline
+  \end{tabular}
+  \caption{First read values for the different serial sizes}
+  \label{tab:readTimesCelegans}
+\end{table}
+
+\Cref{tab:readTimesCelegans} shows the output time ranges with thier period.
 
@@
 The values of \acp{LLM}, such as \ac{LLaMA}-2 \textbf{Meta}'s \ac{LLM} and the two last versions of \acp{GPT} the \ac{LLM} that powers \textit{ChatGPT}, were included to show how the area of the chip scales up.
-Those models don't use \acp{RNN}, and futhermore, give out ridiculous areas.
+Those models do not use \acp{RNN}, and futhermore, give out ridiculous areas.
 Fabricating such cicuits in analog is not feasible witht the current technologies.

### 2023-10-29 conclusions/conclusion.tex
Possibly finished conc

@@
   \end{tabular}
-  \caption{First read values for the different serial sizes}
+  \caption{First and last read times for the different serial sizes}
   \label{tab:readTimesCelegans}
@@
 
-\Cref{tab:readTimesCelegans} shows the output time ranges with thier period.
+\Cref{tab:readTimesCelegans} shows the first and last output time range with the period.
+All of the output times can be determined using the first time range and using the period.
+A sequence consists of $500ms$ as explained in \cref{subsec:celegans}, the time steps are physically $500\mu s$ appart, in the case where the circuit would be used to do real time inference, the circuit would have to be slowed down to be on sync with the inputs.
+The circuit could could be used in a real time use for this time sensitive problem.
 

### 2023-10-29 design/crossbar.tex
Tried to fix cross bar

@@
 
-\subsection{Serialization/Parallelization}
+\subsection{Serialization}
 \label{subsec:serpar}
 
-The idea using the system in a serial mode came from \cite{thesisRef}.
-The principle of this system is to save onChip area by reducing the number of \ac{opAmp} required.
-Using the system in serial mode, increases the time it takes to compute the output of the the \ac{VMM}.
-Let's assume a serial size of $n_s\in\mathbb{N}$, the number of \ac{opAmp} and resistor is then divided by $n_s$ at the cost of multipying the computation time by $n_s$.
+This circuit has the option to be serialized with varying degrees.
+The idea of serializing the circuit came from \cite{thesisRef}.
+Serializing the circuit reduces the number of components required and thus reduces the final onChip area.
+Serializing the system increases the time it takes to compute the output.
+
+Serializing the system means not computing values of the output vector all at the same time, but instead computing group by group, the groups' size are $n_o$.
+The first output group is computed during $e_0$ and the $i^{th}$ group is computed during $e_i$.
+The timing of when the outputs are available is found in \cref{tim:serpar}.
+Those flag control the \ac{CMOS} swicthes present in \cref{circt:xbar}, the switches control which output group is outputed.
 

### 2023-10-29 conclusions/conclusion.tex
More stuff in conclusion

@@
 
-The \ac{LSTM} and \ac{GRU} blocks's onChip area is impossible to estimate, as part of the circuit is made of verilog-A models.
+The \ac{LSTM} and \ac{GRU} blocks' onChip area are impossible to estimate, as parts of the circuits are made of verilog-A models.
 However, to get a general idea of the area of the chip, it can be assumed that the area of the circuit mainly depends on the the area of a memrisor, because the area of a memristor is much greater than the ones of the other components.
@@
 It also shows the current limit with state of the art \acp{NN} at the current time.
+
+The onChip area of small \acp{NN} such as the ones used in this work are very reasonable and could easily be fabricated for embedded use.

### 2023-10-29 results/celegans.tex
changed table
removed useless files and todos

@@
   \centering
-  \begin{tikztimingtable}%TODO change or find other way
+  \begin{tikztimingtable}
     $e_0$       & x 4H    L 4H    N(A1) 3L N(A2) 4H    x\\
@@
     \hline
-    \cellcolor[HTML]{808080}\acs{RMSE} & BD1 & LUAL & PVR & VB1\\
+    \rowcolor{gray}
+    Neuron & BD1 & LUAL & PVR & VB1\\
     \hline

### 2023-10-29 appendices/codes.tex
Worked on appendix

@@
 \label{ap:codes}
+\pythonstyle
+
+\section{Memristor verilog-A model}\label{apsec:memModel}
+
+The code for the memristor model described in this thesis is available just below. The code was extracted from \cite{memCadenceModel}.
+
+\lstinputlisting[language=Verilog, caption=Memristor Verilog-A model]{code/memristor.va}
 
@@
 
-\pythonstyle
-\lstinputlisting[language=python]{code/smooth.py}
+The code for smoothing the curve takes in a list and the averaging range. The function returns the averaged list.
+
+\lstinputlisting[language=Python, caption=Python function to smooth out curve]{code/smooth.py}
+
+\section{Saving weights to a file}\label{apsec:saveWei}
+
+There are a few functions that show how the weights are saved into a file.
+First depending on the type of layer, the weights need to be set in a list in the correct order.
+Once those list are correctly set everything is saved in a binary file.
+
+\lstinputlisting[language=Python, caption=Python functions used to save Keras' weights to a file]{code/saveWeight.py}
+
+\section{Weights to resistances}\label{apsec:wei2res}
+
+The function takes in a weight and returns a tuple with the two resistances values. The global variables are set to set values chosen in this work.
+
+\lstinputlisting[language=Python, caption=Python function to convert a weight in two resistances]{code/wei2res.py}

### 2023-10-29 code/saveWeight.py

@@ -1,0 +1,68 @@
+import pickle
+
+import tensorflow as tf
+from tensorflow.keras.layers import Layer
+from tensorflow.keras.layers import Dense, LSTM, GRU
+from keras.layers import TimeDistributed
+
+
+def getLSTMWeights(lstm, nbHidden):
+    W, U, b = lstm.get_weights()
+    nbGates = 4
+    out = [None] * nbGates
+
+    for i in range(nbGates):
+        out[i] = []
+        for j in range(nbHidden):
+            out[i].extend(W[:, i * nbGates + j])
+            out[i].extend(U[:, i * nbGates + j])
+            out[i].append(b[i * nbGates + j])
+    return out
+
+
+def getGRUWeights(layer, nbHidden):
+    W, U, b = layer.get_weights()
+    nbGates = 3
+    out = [None] * nbGates
+
+    for i in range(nbGates):
+        out[i] = []
+        for j in range(nbHidden):
+            out[i].extend(W[:, i * nbGates + j])
+            out[i].extend(U[:, i * nbGates + j])
+            out[i].append(b[i * nbGates + j])
+    return out
+
+
+def getDenseWeights(nn, nbOutput):
+    W = nn.get_weights()[0]
+    b = list(nn.get_weights()[1])
+    out = []
+    for i in range(nbOutput):
+        out.extend(W[:, i])
+        out.append(b[i])
+    return out
+
+
+def saveTofile(layers, filename):
+    out = [[]]
+    for layer in layers:
+        if type(layer) == TimeDistributed:
+            nbOut = layer.output_shape[-1]
+            out[0].append("tDense(" + str(nbOut) + ")")
+            out.append(getDenseWeights(layer, nbOut))
+        elif type(layer) == Dense:
+            nbOut = layer.output_shape[1]
+            out[0].append("Dense(" + str(nbOut) + ")")
+            out.append(getDenseWeights(layer, nbOut))
+        elif type(layer) == LSTM:
+            nbHid = layer.units
+            out[0].append("LSTM(" + str(nbHid) + ")")
+            out.append(getLSTMWeights(layer, nbHid))
+        elif type(layer) == GRU:
+            nbHid = layer.units
+            out[0].append("GRU(" + str(nbHid) + ")")
+            out.append(getGRUWeights(layer, nbHid))
+
+    with open(filename, "wb") as file:  # Pickling
+        pickle.dump(out, file)

### 2023-10-29 code/smooth.py

@@ -2,10 +2,10 @@ def smooth_line(df, n):
     size = len(df)
     out = df.copy()
     for i in range(size):
-        if (i < n):
-            out[i] = df[0:i+n].mean()
-        elif (size-i-1 < n):
-            out[i] = df[i-n:size-1].mean()
+        if i < n:
+            out[i] = mean(df[0 : i + n])
+        elif size - i - 1 < n:
+            out[i] = mean(df[i - n : size - 1])
         else:
-            out[i] = df[i-n:i+n].mean()
+            out[i] = mean(df[i - n : i + n])
     return out

### 2023-10-29 code/wei2res.py

@@ -1,0 +1,10 @@
+Rmin = 10**4
+Rmax = 10**6
+Rf = (Rmax + Rmin) / 2
+
+
+def wei2res(w):
+    Rpos = (w + 1 - np.sqrt(w**2 + 1)) * Rf / w
+    Rpos = float("%.2g" % Rpos)
+    Rneg = 2 * Rf - Rpos
+    return (Rpos, Rneg)

### 2023-10-29 results/airline.tex
removed useless files and todos
Fixing not number of epochs

@@
 
+The \ac{NN} was trained for $300$ epochs.
+
 \subsubsection{Timing}
@@
   \centering
-  \begin{tikztimingtable}%TODO change or find other way
+  \begin{tikztimingtable}
     $e_0$       & x 2H 6L     L 2H 6L     L L x\\

### 2023-10-29 conclusions/conclusion.tex
Reading to make sure nothing is missing
More corrections

@@
 More complex input sequences produce even more inaccuracies.
-Indeed, when working with sequence 15 (\cref{graph:io15Celegans}), the inaccuracies spiral up and gives out a barely usable prediction.
+Indeed, when working with sequence 15 (\cref{graph:io15Celegans}), the inaccuracies explode and gives out a barely usable prediction.
 The curve generated does not fit its digital counterpart anymore.
@@
 The results obtained demonstrate that the \ac{LSTM} circuit block can be run with low error when dealing with simple inputs.
-The circuit is assumed to be ready for a full simulation, meaning also training with the analog circuit.
+The circuit is ready for a full simulation, meaning also simulating training with the analog circuit.
 

### 2023-10-30 conclusions/further.tex
Worked on paper+corrected errors

@@
 This is would allow the circuit to access a larger range of weights and make the training more performant.
+
+\subsection{Serializing the \acs{GRU} circuit}
+
+The \ac{GRU} circuit, once fully working and implemented in its current form, can be serialized.
+
+The \ac{GRU} circuit cannot be serialized because of the candidate hidden state needs a fully computed reset gate to be computed.
+The solution would be to use the reset gate crossbar array with a serial size of one.
+While all other crossbar arrays can be serialized.
+This will allow to reduce the amount of pointwise components.

### 2023-10-30 design/crossbar.tex
More corrections
Worked on paper+corrected errors

@@
 
-With $x_k$ being the voltage for the input line $k$. The highest \ac{opAmp} is identified as $opamp_0$ and the lowest $opamp_1$.
+Using $x_k$ as the voltage for the input line $k$. The highest \ac{opAmp} is identified as $opAmp_0$ and the lowest $opAmp_1$.
 
@@
   \label{eq:doubleMem0}
-  V_{opAmp0}=-R_r\cdot i_+ \Leftrightarrow
+  V_{opAmp_0}=-R_r\cdot i_+
 \end{equation}
@@
   \label{eq:doubleMem1}
-  i_{R_f}=i_-+\frac{V_{opAmp0}}{R_r}=i_--i_+
+  i_{R_f}=i_-+\frac{V_{opAmp_0}}{R_r}=i_--i_+
 \end{equation}
-With $i_{opAmp}=0A$ because we assume an ideal \ac{opAmp}.
 \begin{equation}
   \label{eq:doubleMem2}
-  V_{opAmp2}=y_0=-R_f\cdot(i_--i_+)=R_f\cdot(i_+-i_-)=R_f\cdot\sum_{k=0}^n(\sigma_{k+}-\sigma_{k-})\cdot x_k
+  V_{opAmp_1}=y_0=-R_f\cdot(i_--i_+)=R_f\cdot(i_+-i_-)=R_f\cdot\sum_{k=0}^n(\sigma_{k+}-\sigma_{k-})\cdot x_k
 \end{equation}
@@
 
-Serializing the system means not computing values of the output vector all at the same time, but instead computing group by group, the groups' size are $n_o$.
+Serializing the system means not computing all values of the output vector at the same time, but instead computing group by group, the groups' size are $n_o$.
 The first output group is computed during $e_0$ and the $i^{th}$ group is computed during $e_i$.
@@
 
-The approximate onChip area for the crossbar circuit depends on the previously defined parameters. It also slightly changes depending on the value of $n_s$.
+The approximate onChip area for the crossbar circuit depends on the previously defined parameters.
 

### 2023-10-30 design/genwei.tex
Worked on appendix
Worked on paper+corrected errors

@@
 This model contains the description of the \ac{NN}, such as the type of \ac{RNN} and the Dense layers that would come before of after the \ac{RNN}.
-In Keras, there are only two of the \ac{LSTM} variants (\cref{sec:lstm}) available :
-
-\begin{itemize}
-  \item The \ac{NP} \ac{LSTM} : This is most commonly used version and by such one of the two available in Keras.
-  \item The \ac{GRU} : This one is present here because it is considered to be different from an \ac{LSTM} \ac{NN} despite their similarities.
-\end{itemize}
-
-The other kinds of \ac{LSTM} are not present in the default Keras librairies.
+Two out of the three \acp{RNN} available in Keras are trained in this work, the \ac{LSTM} and the \ac{GRU}.
 
@@
 
-The worst case scenario for an \ac{LSTM} is when every value reaches the voltage threshold ($V_{threshold}$) (\cref{sec:xbarCircuit}) and every weight is maximized ($w_{max}$).
+The worst case scenario for an \ac{LSTM} is when every value reaches the voltage threshold ($V_{read}$) (\cref{sec:xbarCircuit}) and every weight is maximized ($w_{max}$).
 When that is the case the output of the \acp{VMM} is found in \cref{eq:weiCons}.
@@
 \begin{equation}\label{eq:weiCons}
-  V_{threshold}\cdot w_{max} \cdot(n_i+n_h)+w_{max}+V_{cm}= V_{dd}
+  V_{read}\cdot w_{max} \cdot(n_i+n_h)+w_{max}+V_{cm}= V_{dd}
 \end{equation}
@@
 \begin{equation}\label{eq:weiConsRes}
-  w_{max}=\frac{V_{dd}-V_{cm}}{V_{threshold}\cdot(n_i+n_h)+1}
+  w_{max}=\frac{V_{dd}-V_{cm}}{V_{read}\cdot(n_i+n_h)+1}
 \end{equation}
@@
 \begin{equation}\label{eq:weiConsRes1}
-  w_{min}=-w_{max}=-\frac{V_{dd}-V_{cm}}{V_{threshold}\cdot(n_i+n_h)+1}
+  w_{min}=-w_{max}=-\frac{V_{dd}-V_{cm}}{V_{read}\cdot(n_i+n_h)+1}
 \end{equation}
@@
 
-The code used for all the weights generations is available at \cite{lstmWei}. %TODO README
+The code used for all the weights generations is available at \cite{lstmWei}. The code containing the functions to save the weights to a file is available in \cref{apsec:saveWei}.

### 2023-10-30 design/gru.tex
Worked on paper+corrected errors

@@
 
-First of all, the different vectors and variables present in the schematic have to be described :
+The different vectors and variables present in the schematic have to be described :
 
@@
 
-A little trick had to be emplyed in order to succesfully compute the function. The function is defined as in \cref{eq:1minus0}.
+A clever trick had to be employed in order to succesfully compute the function. The function is defined as in \cref{eq:1minus0}.
 \begin{equation}\label{eq:1minus0}

### 2023-10-30 design/models.tex
Worked on paper+corrected errors

@@
-\section{Verilog models}
+\section{Verilog-A models}
 \label{sec:models}
 
-Due to a lack of time, some of the more common component were not designed by me but instead were simulated using a Verilog model.
-Verilog is a hardware description laguage very popular in the industry.
+Due to a lack of time, some of the more common component were not designed by me but instead were simulated using a Verilog-A model.
+Verilog-A is a hardware description laguage very popular in the industry.
 The only components that use a verilog model are the voltage multiplier and the \ac{opAmp}.
@@
 It allows us to multiply, as its name implies, two voltages.
-It is used to compute the pointwise multiplications of the \ac{LSTM} (\cref{fig:lstmCell}).
+It is used to compute the pointwise multiplications of the \ac{LSTM} or \ac{GRU} (\cref{fig:lstmCell,fig:encoderGruCell,fig:decoderGruCell}).
 
 However this part is tricky.
-Indeed, because of the voltage conversion (\cref{tab:valConv}), the equation needs to take that into consideration.
+Indeed, because of the voltage conversion (\cref{tab:valConv}), the equation needs to take the conversion into consideration.
 The final equation needs to be \cref{eq:finalVoltMult} in order to have $1\cdot 1=1$ ($V_{in_1}\cdot V_{in_2}=1V$, with $V_{in_1}=V_{in_2}=1V$)

### 2023-10-30 design/netlist.tex
Worked on paper+corrected errors

@@
 \begin{itemize}
-  \item The first parameter is the number of input we use our system. This is the size of the first input vector ($x_t$).
+  \item The first parameter is the number of input we use our system. This is the size of the first input vector ($\overrightarrow{x_t}$).
   \item The next parameter is the number of time steps. Everything about the \ac{LSTM} time steps is all explained in \cref{sec:lstm}.
@@
 This is represented in the description as "LSTM($n_h$)" where $n_h$ is the number of hidden states of the layer.
+  \item \ac{GRU} layer : An \ac{GRU} layer contains three \acp{VMM} which can be assimilated and thus stored like a dense layer.
+The weights for each \ac{VMM} is then stored in a sub list.
+This is represented in the description as "GRU($n_h$)" where $n_h$ is the number of hidden states of the layer.
 \end{itemize}

### 2023-10-30 paper/circuit.tex
Worked on paper+corrected errors

\section{Design}

The system will be working with a $V_{dd}$ of $1.8V$. Such a value was chosen because this is a low power system.

The way values are encoded in the analog system will be descibed here as it serves for the entire thesis.
In order for the system to support negative numbers we're going to use a $V_{cm}$ set to $\frac{V_{dd}}{2}$. That means that $V_{cm}=\frac{V_{dd}}{2}=0.9V$. This $V_{cm}$ will then describe a zero. A step of one was chosen to be $0.1V$ in the analog circuit.
\Cref{tab:valConv} shows the conversions from a real number to its voltage equivalent.

\begin{table}[h]
  \centering
  \begin{tabular}{|c|c|}
    \hline
    \rowcolor{gray}
    Real value & Voltage \\
    \hline
    $0$ & $0.9V$ \\
    \hline
    $1$ & $1.0V$ \\
    \hline
    $x$ & $V(x)=\frac{x}{10}+V_{cm}$\\
    \hline
    $real(v)=(v-0.9)\cdot 10$ & $v$\\
    \hline
  \end{tabular}
  \caption{Real/Voltage Conversion Table.}
  \label{tab:valConv}
\end{table}

Since the system cannot reach voltage outside of the operating range with the intended behavior, the voltage is then restricted to $V\in [0,1.8]$. This means that the range of real value that the systems can handle is $x\in [-9,9]$.

\subsection{Activation functions}\label{subsec:af}

Producing analog activation functions is quite important as using hard sigmoid or hard hyperbolic tangent (tanh) functions impacts the results \cite{af,hardSigm}.

The analog activation circuit thus plays an important role in the final result's quality.

The circuit used is the same as the one in \cite{thesisRef}, and is shown in \cref{circt:af}. The circuit's technology used being different, all the parameters had to be determined empirically to best fit a sigmoid shape. The parameters can be found in \cref{tab:afPar}.

Due to the nature of the functions we want to generate, we will use the same circuit for both a sigmoid and a \ac{tanh} like functions. The two different functions are generated by changing two parameters. The functions generated are the same shape and only differ by their output range.

\begin{figure}[htbp]
  \centering
  \includesvg[width=\columnwidth,pretex=\small]{activation/afCircuit}
  \caption{Activation functions circuit}
  \label{circt:af}
\end{figure}

\begin{table}[htbp]
  \caption{Circuits parameters}
  \label{tab:afPar}
  \centering
  \begin{tabular}{|c|c|c|}
    \hline
    \rowcolor{gray}
    Parameter & Sigmoid & \ac{tanh} \\
    \hline
    $V_1$ & \multicolumn{2}{c|}{$1.1V$}\\
    \hline
    $V_2$ & \multicolumn{2}{c|}{$635mV$}\\
    \hline
    $V_3$ & $0.8V$ & $550mV$\\
    \hline
    $i_{dc}$ & \multicolumn{2}{c|}{$150uA$}\\
    \hline
    $w$ & \multicolumn{2}{c|}{$900nm$}\\
    \hline
    $l$ & \multicolumn{2}{c|}{$60nm$}\\
    \hline
    $R_1$ & \multicolumn{2}{c|}{$5k\Omega$}\\
    \hline
    $R_2$ & \multicolumn{2}{c|}{$10k\Omega$}\\
    \hline
    $R_3$ & $2k\Omega$ & $4k\Omega$\\
    \hline
  \end{tabular}
\end{table}

Here, $w$ and $l$ are, respectively, the width and length of the two NMOS of the circuit.

The outputed voltage depends on the input passed on. The results obtained by are rather convincing and can be found in \cref{graph:af}. \Cref{graph:af} also shows the Root Mean Square Error (RMSE) of the analog results with the ideal result.

\begin{figure}[h]
  \centering
  \includesvg[width=\columnwidth, pretex=\scriptsize]{activation/afGraph}
  \caption{Input/Output graph of the activation function circuit for both sigmoid and \acs{tanh} functions}
  \label{graph:af}
\end{figure}

The symbols for the sigmoid and the \ac{tanh} are separated for better understanding and are available in \cref{sym:af}.

\begin{figure}[htbp]
  \centering
  \hspace*{0.8cm}
  \subfloat[Sigmoid symbol]{\includesvg[height=1cm,pretex=\small]{activation/sigmoidSymbol}}%
  \hfill
  \subfloat[\ac{tanh} symbol]{\includesvg[height=1cm,pretex=\small]{activation/tanhSymbol}}%
  \hspace*{0.8cm}
  \caption{Activation functions symbols with the input and output pins on either side depending on the flow of the current for better readability}
  \label{sym:af}
\end{figure}

These functions are still a bit different from the original functions (especially for the \ac{tanh}). However that does not matter too much as the trainning will be happening with the extracted analog functions, all weights will be set in the circuit. This is the reason why such a difference does not matter. As long as the curves have the similar shape, the result will not be drastically affected.

\subsection{Memory cell}

The memory cell is a circuit that is able to store an analog value for a limited time. It works using capacitors that have the ability to store a voltage for a short period of time.

The circuit is shown in \cref{circt:memcell}. The voltage is stored in the capacitor and is kept using \ac{CMOS} switches.

The \ac{CMOS} switches found in the circuit have width of $w=200nm$ and a length of $l=60nm$.

\begin{figure}[h]
  \centering
  \includesvg[width=\columnwidth,pretex=\small]{memcell/memCircuit}
  \caption{Memory cell circuit}
  \label{circt:memcell}
\end{figure}

%Remove first if too big
The cirucit has a two \ac{CMOS} switches design to avoid voltage leakage through the swicthes. The system has voltage leakage when only one \ac{CMOS} swicth, and thus leads to a large memory leak. This is due to the high voltage difference between the two sides of the \ac{CMOS} swicth. Using two \ac{CMOS} switches allows for this difference to be mitigated (\cref{fig:memcellLoss}).

\begin{figure}[h]
  \centering
  \includesvg[width=\columnwidth,pretex=\scriptsize]{memcell/data-loss}
  \caption{Memory conservation in a memory cell with 1 \ac{CMOS} switch vs 2 \ac{CMOS} swicthes}
  \label{fig:memcellLoss}
\end{figure}

The symbol, shown in \cref{sym:memcell}, for this circuit is designed to show a capacitor because it is its memory mechanism.

\begin{figure}[h]
  \centering
  \includesvg[height=1cm,pretex=\small]{memcell/memcellSymbol}
  \caption{Memory cell symbol with the input enable pin (top) and the output enable pin (bottom). The left and right pins are interchangebly the input and ouput for the sake of readability}
  \label{sym:memcell}
\end{figure}

\subsection{Voltage-driven crossbar circuit}\label{subsec:xbarCircuit}

The crossbar circuit theory has already been explained in the \cref{sec:state}. The actual implementation in the circuit is described here. The circuit is the one in \cref{circt:xbar}. The circuit depends on three parameters :

\begin{itemize}
  \item $n_i$ : The number of input for our crossbar array (not including bias for a more general circuit).
  \item $n_o$ : The number of parallel output for our crossbar array.
  \item $n_s$ : The serial size of our crossbar system.
\end{itemize}

\begin{figure}[h]
  \centering
  \includesvg[width=\columnwidth,pretex=\scriptsize]{crossbar/crossbarUse}
  \caption{Circuit of the crossbar array used in the final system ($n_i$, $n_o$, $n_s$)}
  \label{circt:xbar}
\end{figure}

In this work, the choice was made to use a two memristor per synapse architecture. Using two memristor per synapse doubles the area but doubles the weight range \cite{doubleMem} and allows to easily use negavite weights. The output voltage will be centered around $V_{cm}$ and be compliant with the standard set in \cref{tab:valConv}.

This design is the one that is used in \cite{doubleMem}. Let's assume that a given memristor has a resistance range of $R\in[R_{min},R_{max}]$, that means its conductance range is $\sigma \in [\sigma_{min},\sigma_{max}]$ (with $\sigma_{min}= \frac{1}{R_{max}}$ and $\sigma_{max}= \frac{1}{R_{min}}$). This design works using two \ac{opAmp} connected to $V_{cm}$ with the positive pin and the negative pin to the output of the crossbar array. \Cref{eq:doubleMem0,eq:doubleMem1,eq:doubleMem2} are describing how this architecture works. A simplified version of the double memristors per synapse circuit is also available in \cref{circt:doubleMem}.

\begin{figure}[h]
  \centering
  \includesvg[width=\columnwidth,pretex=\tiny]{crossbar/doubleMem}
  \caption{Simplified circuit of a double memristor per synapse architecture}
  \label{circt:doubleMem}
\end{figure}

Using $x_k$ as the voltage for the input line $k$. The highest \ac{opAmp} is identified as $opAmp_0$ and the lowest $opAmp_1$.

For the sake of simplicity, the following equations considers the ground to be $V_{cm}$.

\begin{equation}
  \label{eq:doubleMem0}
  V_{opAmp_0}=-R_r\cdot i_+
\end{equation}
\begin{equation}
  \label{eq:doubleMem1}
  i_{R_f}=i_-+\frac{V_{opAmp_0}}{R_r}=i_--i_+
\end{equation}
\begin{equation}
  \label{eq:doubleMem2}
  \begin{array}{2}
    V_{opAmp_1}=y_0=-R_f\cdot(i_--i_+)=R_f\cdot(i_+-i_-)\\
    =R_f\cdot\sum_{k=0}^n(\sigma_{k+}-\sigma_{k-})\cdot x_k
  \end{array}
\end{equation}
With $i_+=\sum_{k=0}^n\sigma_{k+}\cdot x_k$ and $i_-=\sum_{k=0}^n\sigma_{k-}\cdot x_k$.

This circuit has the option to be serialized with varying degrees. The idea of serializing the circuit came from \cite{thesisRef}. Serializing the circuit reduces the number of components required and thus reduces the final onChip area. Serializing the system increases the time it takes to compute the output.

Serializing the system means not computing all values of the output vector at the same time, but instead computing group by group, the groups' size are $n_o$. The first output group is computed during $e_0$ and the $i^{th}$ group is computed during $e_i$. The timing of when the outputs are available is found in \cref{tim:serpar}. Those flag control the \ac{CMOS} swicthes present in \cref{circt:xbar}, the switches control which output group is outputed.

The \ac{CMOS} switches are here to open the necessary input gates when the output is required, the \ac{CMOS} switches are controlled as in \cref{tim:serpar}.

\begin{figure}[h]
  \centering
  \begin{tikztimingtable}
    $e_0$ & x2H2L N(A1) L N(A2) 2L N(A3) L N(A4) 2Lx\\
    $e_1$ & x2L2H N(B1) L N(B2) 2L N(B3) L N(B4) 2Lx\\
    $e_i$ & x4L N(C1) L N(C2) 2H N(C3) L N(C4) 2Lx\\
    $e_{n_s-1}$ & x4L N(D1) L N(D2) 2L N(D3) L N(D4) 2Hx\\
    %foo & 2L N(A1)  4H N(A2) L\\
    \extracode
    \node[gap, at={($(A1|-A2)!0.5!(A2)$)}];
    \node[gap, at={($(A3|-A4)!0.5!(A4)$)}];
    \node[gap, at={($(B1|-B2)!0.5!(B2)$)}];
    \node[gap, at={($(B3|-B4)!0.5!(B4)$)}];
    \node[gap, at={($(C1|-C2)!0.5!(C2)$)}];
    \node[gap, at={($(C3|-C4)!0.5!(C4)$)}];
    \node[gap, at={($(D1|-D2)!0.5!(D2)$)}];
    \node[gap, at={($(D3|-D4)!0.5!(D4)$)}];
    \tablerules
    %\draw (0,0) circle (2pt); % Origin
    \begin{pgfonlayer}{background}
      \vertlines[help lines]{0.55,10.55}
      %\vertlines[red]{1.6,5.6,15.6}
      %\vertlines[blue]{3.6,9.6,15.6}
    \end{pgfonlayer}
  \end{tikztimingtable}
  \caption{Enable flags timing for any value of $n_s$ in a single time step}
  \label{tim:serpar}
\end{figure}

When the system is used fully in parallel, the \ac{CMOS} switches are not required can then be removed to lower the final onChip area.

In this work, the analog system will be simulated for the inference of the \ac{NN}, thus the weights will not have to change during the simulation. Because the weights are represented in the analog circuit by the internal resistances of the memristors, the memristors can be replaced by resistors with a set resistance for the simulation.

The symbol (\cref{sym:xbar}) defined for the voltage based memristor crossbar array used in this thesis is more compact and helps the readability of the circuits that require a crossbar array. It depends on several parameters, the number of inputs ($n_i$), the number of outputs ($n_o$) and the serial size ($n_s$).

\begin{figure}[h]
  \centering
  \includesvg[height=1cm]{crossbar/xbarSymbol}
  \caption{Symbol used for the crossbar array, the input pin is a bus of size $n_i$ and the output pin is a bus of size $n_o$}
  \label{sym:xbar}
\end{figure}

\subsection{Verilog-A models}

This work uses Verilog-A components to replace component that could not be designed due to lack of time. Those components are the operational amplifier (opAmp) and a voltage multiplier.

\subsubsection{Operational amplifier}

This component is the very famous \ac{opAmp}. This specific component was not designed for the thesis, it required a current range that did not allow to use ones already made by members of the research group. The choice was then to use a verilog-A model, and then design one if time allows it.

\begin{equation}
  \label{eq:opAmp}
  V_{out}=\mu \cdot (V_+-V_-)
\end{equation}

\Cref{eq:opAmp} is the equation used in the verilog-A model. It makes it so it behaves like an ideal \ac{opAmp}. For this thesis $\mu$ has been set to $\mu=10^5$.

\subsubsection{Voltage multiplier}

This component while far less popular than the latter, is just as useful for our specific use. It allows us to multiply, as its name implies, two voltages. It is used to compute the pointwise multiplications of the \ac{LSTM} or \ac{GRU} (\cref{fig:lstmCell,fig:encoderGruCell}).

In oder for the circuit to be compliant with the real value to voltage conversion (\cref{tab:valConv}), a multiplication needs to be as in \cref{eq:finalVoltMult}.

\begin{equation}\label{eq:finalVoltMult}
  V_{out}=10\cdot(V_{in_1}-V_{cm})\cdot (V_{in_2}-V_{cm}) + V_{cm}
\end{equation}

This is taken care of using in reality two parts, the actual voltage multiplier (\cref{eq:voltMult}) and a non inverting amplifier (\cref{eq:invAmp}).

\begin{equation}\label{eq:voltMult}
  V_{voltMult}=-(V_{in_1}-V_{cm})\cdot (V_{in_2}-V_{cm}) + V_{cm}
\end{equation}
\begin{equation}\label{eq:invAmp}
  V_{out}=-(V_{voltMult}-V_{cm})\cdot10+V_{cm}
\end{equation}

Where $V_{out}$ is the output voltage the inverting amplifier, $V_{voltMult}$ is the out voltage of the voltage multiplier itself and $V_{in_1}$ and $V_{in_2}$ are the input voltages.

The \cref{eq:voltMult} is assumed possible because of the actual voltage multiplier's datasheet available at \cite{actualVoltMult}.

The symbols for those two components are the ones in \cref{sym:models}.

\begin{figure}[h]
  \centering
  \hspace*{1.5cm}
  \subfloat[\ac{opAmp}'s symbol\label{sym:opAmp}]{\includesvg[height=1cm]{models/opAmpSymbol}}%
  \hfill
  \subfloat[Voltage multiplier's symbol\label{sym:voltMult}]{\includesvg[height=1cm]{models/voltMultSymbol}}%
  \hspace*{1.5cm}
  \caption{Symbols used for the verilog-A components}
  \label{sym:models}
\end{figure}

The \ac{opAmp} uses its IEEE symbol (\cref{sym:opAmp}) while the voltage multiplier uses a custom symbol (\cref{sym:voltMult}).

\subsection{LSTM analog implementation}

This section describes the circuit of an \ac{LSTM} with an input vector of size $n_i$, a $n_h$ hidden states, a serial size of $n_s$ and $n_{ts}$ time steps. $n_o=n_h/n_s$ is going to be used for future references in this section. In order for the crossbar array to be used $n_o$ must be an integer, in other words, $n_s$ must divide $n_h$. The circuit used for the LSTM is shown in \cref{circt:lstm}.

\begin{figure*}[h]
  \centering
  \includesvg[width=\textwidth,pretex=\tiny]{lstm/lstmCircuit}
  \caption{\ac{LSTM} circuit}
  \label{circt:lstm}
\end{figure*}

The system is built using the crossbar array from \cref{subsec:xbarCircuit} with ($n_i+n_h+1$,$n_o$, $n_s$) as parameters.

\begin{itemize}
  \item $\overrightarrow{z_t}$ is the input of the crossbar but not the input of the \ac{LSTM}. It is defined by $\overrightarrow{z_t}=(\overrightarrow{x_t},\overrightarrow{h_{t-1}},\overrightarrow{b})$.
  \item $e_{j,0}$ and $e_{j,1}$ are two enable flags that respectively represent the first and second half of $e_j$.
  \item $e_{in}$ and $e_{out}$ are the flags used to enable the hidden state values to go to the input (feedback connection) or to the output of the circuit.
  \item $e_{next}$ is the enable flag on in between two time steps.
  \item $R_{amp0}$ and $R_{amp1}$ are the two resistances used to amplify the output voltage of the voltage multipliers. Their value is set so that $\frac{R_{amp1}}{R_{amp0}}=10$.
\end{itemize}

The wires coming into the crossbar are a bus of size $n_i+n_h+1$ and the output of the crossbar is a bus of size $n_o$ (\cref{sym:xbar}). This is why everything in the system apart from the crossbar arrays is only shown once in \cref{circt:lstm} but in reality those components are present $n_o$ times. Those extra components are needed in order for the parallel channels to work.

The circuit uses two memory cells for its feedback connections not to overwrite the value of the current stored value by the value of the next time step.

The LSTM circuit can be serialized allowing to reduce the number of pointwise components in the circuit by a factor $n_s$. However serializing the circuit increases the times it takes to compute the outputs by a factor of $n_s$.

\subsection{GRU analog implementation}

This section describes the circuit of an encoder \ac{GRU} with an input vector of size $n_i$, a $n_h$ hidden states and $n_{ts}$ time steps. The decoder \ac{GRU} could also be implemented in an analog circuit, but the choice was made to focus on the encoder \ac{GRU}. The \ac{GRU} is by its nature very similar to the \ac{LSTM}. For that reason it has a similar circuit to LSTM circuit (\cref{circt:lstm}). The system is built, once again, using crossbar array (\cref{sec:xbarCircuit}) with ($n_i+n_h+1$,$n_h$, $1$) as parameters.

The GRU equations requires to compute the function defined in \cref{eq:1minus0}.
\begin{equation}\label{eq:1minus0}
  f(x)=1-x
\end{equation}
The solution found to do compute the operation consist of using an \ac{opAmp} as an inverter around $V_{inv}$, the formula is available in \cref{eq:1minus1}.
\begin{equation}\label{eq:1minus1}
  f_v(v)=-(v-V_{inv})+V_{inv}=2\cdot V_{inv} -v
\end{equation}
With $V_{inv}=V_{cm}+0.05$.

The GRU circuit cannot be serialized. This is because the candidate hidden state requires a fully computed reset vector to be computed.

### 2023-10-30 paper/dataset.tex
Worked on paper+corrected errors

--- 2023-10-28 design/dataset.tex
@@
-\section{The datasets}\label{sec:dataset}
+\section{Datasets}
 
-To make sure that the terminal voltage does not reach the threshold as explained in \cref{sec:xbarCircuit}.
-The voltage threshold ($V_{threshold}$) was chosen to be $V_{threshold}=0.1V$ because the memristors used in this thesis are purely theorical.
-The dataset needs to be formated not to exceed this value when converted to voltage (\cref{tab:valConv}), the dataset needs to be trained while not exceeding $1$.
-
-\subsection{Airline}
+\subsection{Airline passengers}
 
@@
 \end{figure}
-
-\subsubsection{Data format}
-
-The dataset contains $144$ sample points.
-It has been transformed into $142$ data points for training.
-This has been done by taking each values three by three.
-The first two values are two timesteps for the input vector and the third value being used the target value.
-Two third of the dataset is being used for training and the other third is used for validation.
-
-\subsection{\ac{C. elegans}}
-
-This dataset is far more interesting and complex than the latter.
-This data set aims to use \acp{LSTM} to mimic the behavior of real neurons.
-As explained in \cite{celegans}, \acf{C.
-elegans} are simple organisms that are getting very popular for whole brain organization studies.
-The point of this problem is to reproduce the nervous system of the \ac{C.
-elegans}.
-This is done using recorded data of the input of 4 neurons and the output of 4 other neurons.
-
-The dataset is great for our study because :
-\begin{itemize}
-  \item It comes from a very recent paper from the research group in which this work is also being developed.
-  \item It aims at reproducing the behavior of the brain of a simple organism (\ac{C. elegans}). In the (very) long term, a full parts of the human brain could be replaced by a very low powered chip.
-\end{itemize}
-
-\begin{figure}[H]
-  \centering
-  \begin{minipage}{\columnwidth}
-    \subfloat[Training dataset sample\label{graph:io5Celegans}]{\includesvg[width=0.5\textwidth, pretex=\tiny]{datasets/celegans/ioTrain0}}%
-    \hfill
-    \subfloat[Test dataset sample]{\includesvg[width=0.5\textwidth,pretex=\tiny]{datasets/celegans/ioTest0}}%
-  \end{minipage}
-  \begin{minipage}{\columnwidth}
-    \subfloat[Training dataset sample]{\includesvg[width=0.5\textwidth, pretex=\tiny]{datasets/celegans/ioTrain1.svg}}%
-    \hfill
-    \subfloat[Validation dataset sample\label{graph:io15Celegans}]{\includesvg[width=0.5\textwidth,pretex=\tiny]{datasets/celegans/ioValid0}}%
-  \end{minipage}
-  \caption{\ac{C. elegans} dataset samples}
-  \label{graph:celegans}
-\end{figure}
-
-\subsubsection{Dataset format}
-
-The dataset contains a set of fourty set of input/output.
-Each input/output set contains $500ms$ of data with a time step of $0.5ms$, making a thousand points for each neurons.
-This data is recorded after applying current to the input neurons (PLML2, PLMR, AVBL, AVBR), and monitoring the output of four neurons (DB1, LUAL, PVR, VB1) that are known to have strong activity during  a specific behavior of the nematode known as \ac{FCM}.
-The simulation is done using a known model of \ac{C.
-elegans} connectome (complete overview of the brains connections) \cite{celegans}.
-
-\Cref{graph:celegans} shows four sets of input/output out of the fourty that were recorded for the dataset.
-
-The inputs for the systems are then going to be the current values for the four input neurons at every of the thousand time steps and plans to get an output as close as possible to the modeled output for all time steps.

### 2023-10-30 paper/paper.tex
Worked on paper+corrected errors

@@
 \usepackage{textcomp}
-\usepackage{xcolor}
+\usepackage{tikz-timing}
+\usepackage[table]{xcolor}
+\usepackage{cleveref}
+\crefname{figure}{Fig.}{Figs.}
+\Crefname{figure}{Fig.}{Figs.}
+\crefname{table}{TABLE}{TABLES}
+\Crefname{table}{TABLE}{TABLES}
+\crefname{equation}{}{}
+\Crefname{equation}{Equation}{Equations}
+\usepackage{svg}
+\usepackage{subfig}
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
 T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+\graphicspath{{../figures/}}
+\input{../definitions/timingsetup.tex}
 \begin{document}
@@
 Both of those blocks contain memristors to be used as weights in a analog \ac{VMM} capable circuit.
-These circuit blocks allow to run very fast computation of \aclp{RNN} of any size, in a relatively small integrated circuit.
+These circuit blocks allow to run very fast computation of \aclp{RNNs} of any size, in a relatively small integrated circuit.
 As part of the \ac{LSTM} and \ac{GRU} blocks, an analog activation function circuit was designed.
@@
 
-\section{State of the art}
+\input{state.tex}
+\input{circuit.tex}
+\input{tools.tex}
+\input{dataset.tex}
 
-\subsection{Recurrent Neural Networks (RNN)}
-
-RNNs are a family of neural networks that differentiate themselves by having feedback connections. It alloww it to be used with sequences of data, sometimes, with varying amount of input data.
-
-The IEEEtran class file is used to format your paper and style the text. All margins,
-column widths, line spaces, and text fonts are prescribed; please do not
-alter them. You may note peculiarities. For example, the head margin
-measures proportionately more than is customary. This measurement
-and others are deliberate, using specifications that anticipate your paper
-as one part of the entire proceedings, and not as an independent document.
-Please do not revise any of the current designations.
-
-\section{Prepare Your Paper Before Styling}
-Before you begin to format your paper, first write and save the content as a
-separate text file. Complete all content and organizational editing before
-formatting. Please note sections \ref{AA}--\ref{SCM} below for more information on
-proofreading, spelling and grammar.
-
-Keep your text and graphic files separate until after the text has been
-formatted and styled. Do not number text heads---{\LaTeX} will do that
-for you.
-
-\subsection{Abbreviations and Acronyms}\label{AA}
-Define abbreviations and acronyms the first time they are used in the text,
-even after they have been defined in the abstract. Abbreviations such as
-IEEE, SI, MKS, CGS, ac, dc, and rms do not have to be defined. Do not use
-abbreviations in the title or heads unless they are unavoidable.
-
-\subsection{Units}
-\begin{itemize}
-  \item Use either SI (MKS) or CGS as primary units.
-(SI units are encouraged.) English units may be used as secondary units (in parentheses).
-An exception would be the use of English units as identifiers in trade, such as ``3.5-inch disk drive''.
-  \item Avoid combining SI and CGS units, such as current in amperes and magnetic field in oersteds.
-This often leads to confusion because equations do not balance dimensionally.
-If you must use mixed units, clearly state the units for each quantity that you use in an equation.
-  \item Do not mix complete spellings and abbreviations of units: ``Wb/m\textsuperscript{2}'' or ``webers per square meter'', not ``webers/m\textsuperscript{2}''.
-Spell out units when they appear in text: ``.
-.
-.
-a few henries'', not ``.
-.
-.
-a few H''.
-  \item Use a zero before decimal points: ``0.25'', not ``.25''. Use ``cm\textsuperscript{3}'', not ``cc''.)
-\end{itemize}
-
-\subsection{Equations}
-Number equations consecutively. To make your
-equations more compact, you may use the solidus (~/~), the exp function, or
-appropriate exponents. Italicize Roman symbols for quantities and variables,
-but not Greek symbols. Use a long dash rather than a hyphen for a minus
-sign. Punctuate equations with commas or periods when they are part of a
-sentence, as in:
-\begin{equation}
-  a+b=\gamma\label{eq}
-\end{equation}
-
-Be sure that the
-symbols in your equation have been defined before or immediately following
-the equation. Use ``\eqref{eq}'', not ``Eq.~\eqref{eq}'' or ``equation \eqref{eq}'', except at
-the beginning of a sentence: ``Equation \eqref{eq} is . . .''
-
-\subsection{\LaTeX-Specific Advice}
-
-Please use ``soft'' (e.g., \verb|\eqref{Eq}|) cross references instead
-of ``hard'' references (e.g., \verb|(1)|). That will make it possible
-to combine sections, add equations, or change the order of figures or
-citations without having to go through the file line by line.
-
-Please don't use the \verb|{eqnarray}| equation environment. Use
-\verb|{align}| or \verb|{IEEEeqnarray}| instead. The \verb|{eqnarray}|
-environment leaves unsightly spaces around relation symbols.
-
-Please note that the \verb|{subequations}| environment in {\LaTeX}
-will increment the main equation counter even when there are no
-equation numbers displayed. If you forget that, you might write an
-article in which the equation numbers skip from (17) to (20), causing
-the copy editors to wonder if you've discovered a new method of
-counting.
-
-{\BibTeX} does not work by magic. It doesn't get the bibliographic
-data from thin air but from .bib files. If you use {\BibTeX} to produce a
-bibliography you must send the .bib files.
-
-{\LaTeX} can't read your mind. If you assign the same label to a
-subsubsection and a table, you might find that Table I has been cross
-referenced as Table IV-B3.
-
-{\LaTeX} does not have precognitive abilities. If you put a
-\verb|\label| command before the command that updates the counter it's
-supposed to be using, the label will pick up the last counter to be
-cross referenced instead. In particular, a \verb|\label| command
-should not go before the caption of a figure or a table.
-
-Do not use \verb|\nonumber| inside the \verb|{array}| environment. It
-will not stop equation numbers inside \verb|{array}| (there won't be
-any anyway) and it might stop a wanted equation number in the
-surrounding equation.
-
-\subsection{Some Common Mistakes}\label{SCM}
-\begin{itemize}
-  \item The word ``data'' is plural, not singular.
-  \item The subscript for the permeability of vacuum $\mu_{0}$, and other common scientific constants, is zero with subscript formatting, not a lowercase letter ``o''.
-  \item In American English, commas, semicolons, periods, question and exclamation marks are located within quotation marks only when a complete thought or name is cited, such as a title or full quotation.
-When quotation marks are used, instead of a bold or italic typeface, to highlight a word or phrase, punctuation should appear outside of the quotation marks.
-A parenthetical phrase or statement at the end of a sentence is punctuated outside of the closing parenthesis (like this).
-(A parenthetical sentence is punctuated within the parentheses.)
-  \item A graph within a graph is an ``inset'', not an ``insert''. The word alternatively is preferred to the word ``alternately'' (unless you really mean something that alternates).
-  \item Do not use the word ``essentially'' to mean ``approximately'' or ``effectively''.
-  \item In your paper title, if the words ``that uses'' can accurately replace the word ``using'', capitalize the ``u''; if not, keep using lower-cased.
-  \item Be aware of the different meanings of the homophones ``affect'' and ``effect'', ``complement'' and ``compliment'', ``discreet'' and ``discrete'', ``principal'' and ``principle''.
-  \item Do not confuse ``imply'' and ``infer''.
-  \item The prefix ``non'' is not a word; it should be joined to the word it modifies, usually without a hyphen.
-  \item There is no period after the ``et'' in the Latin abbreviation ``et al.''.
-  \item The abbreviation ``i.e.'' means ``that is'', and the abbreviation ``e.g.'' means ``for example''.
-\end{itemize}
-An excellent style manual for science writers is \cite{b7}.
-
-\subsection{Authors and Affiliations}
-\textbf{The class file is designed for, but not limited to, six authors.} A
-minimum of one author is required for all conference articles. Author names
-should be listed starting from left to right and then moving down to the
-next line. This is the author sequence that will be used in future citations
-and by indexing services. Names should not be listed in columns nor group by
-affiliation. Please keep your affiliations as succinct as possible (for
-example, do not differentiate among departments of the same organization).
-
-\subsection{Identify the Headings}
-Headings, or heads, are organizational devices that guide the reader through
-your paper. There are two types: component heads and text heads.
-
-Component heads identify the different components of your paper and are not
-topically subordinate to each other. Examples include Acknowledgments and
-References and, for these, the correct style to use is ``Heading 5''. Use
-``figure caption'' for your Figure captions, and ``table head'' for your
-table title. Run-in heads, such as ``Abstract'', will require you to apply a
-style (in this case, italic) in addition to the style provided by the drop
-down menu to differentiate the head from the text.
-
-Text heads organize the topics on a relational, hierarchical basis. For
-example, the paper title is the primary text head because all subsequent
-material relates and elaborates on this one topic. If there are two or more
-sub-topics, the next level head (uppercase Roman numerals) should be used
-and, conversely, if there are not at least two sub-topics, then no subheads
-should be introduced.
-
-\subsection{Figures and Tables}
-\paragraph{Positioning Figures and Tables} Place figures and tables at the top and
-bottom of columns. Avoid placing them in the middle of columns. Large
-figures and tables may span across both columns. Figure captions should be
-below the figures; table heads should appear above the tables. Insert
-figures and tables after they are cited in the text. Use the abbreviation
-``Fig.~\ref{fig}'', even at the beginning of a sentence.
-
-\begin{table}[htbp]
-  \caption{Table Type Styles}
-  \begin{center}
-    \begin{tabular}{|c|c|c|c|}
-      \hline
-      \textbf{Table}&\multicolumn{3}{|c|}{\textbf{Table Column Head}} \\
-      \cline{2-4}
-      \textbf{Head} & \textbf{\textit{Table column subhead}}& \textbf{\textit{Subhead}}& \textbf{\textit{Subhead}} \\
-      \hline
-      copy& More table copy$^{\mathrm{a}}$& &  \\
-      \hline
-      \multicolumn{4}{l}{$^{\mathrm{a}}$Sample of a Table footnote.}
-    \end{tabular}
-    \label{tab1}
-  \end{center}
-\end{table}
-
-\begin{figure}[htbp]
-  \centerline{\includegraphics{fig1.png}}
-  \caption{Example of a figure caption.}
-  \label{fig}
-\end{figure}
-
-Figure Labels: Use 8 point Times New Roman for Figure labels. Use words
-rather than symbols or abbreviations when writing Figure axis labels to
-avoid confusing the reader. As an example, write the quantity
-``Magnetization'', or ``Magnetization, M'', not just ``M''. If including
-units in the label, present them within parentheses. Do not label axes only
-with units. In the example, write ``Magnetization (A/m)'' or ``Magnetization
-\{A[m(1)]\}'', not just ``A/m''. Do not label axes with a ratio of
-quantities and units. For example, write ``Temperature (K)'', not
-``Temperature/K''.
-
-\section*{Acknowledgment}
-
-The preferred spelling of the word ``acknowledgment'' in America is without
-an ``e'' after the ``g''. Avoid the stilted expression ``one of us (R. B.
-G.) thanks $\ldots$''. Instead, try ``R. B. G. thanks$\ldots$''. Put sponsor
-acknowledgments in the unnumbered footnote on the first page.
-
-\section*{References}
-
-Please number citations consecutively within brackets \cite{b1}. The
-sentence punctuation follows the bracket \cite{b2}. Refer simply to the reference
-number, as in \cite{b3}---do not use ``Ref. \cite{b3}'' or ``reference \cite{b3}'' except at
-the beginning of a sentence: ``Reference \cite{b3} was the first $\ldots$''
-
-Number footnotes separately in superscripts. Place the actual footnote at
-the bottom of the column in which it was cited. Do not put footnotes in the
-abstract or reference list. Use letters for table footnotes.
-
-Unless there are six authors or more give all authors' names; do not use
-``et al.''. Papers that have not been published, even if they have been
-submitted for publication, should be cited as ``unpublished'' \cite{b4}. Papers
-that have been accepted for publication should be cited as ``in press'' \cite{b5}.
-Capitalize only the first word in a paper title, except for proper nouns and
-element symbols.
-
-For papers published in translation journals, please give the English
-citation first, followed by the original foreign-language citation \cite{b6}.
-
-\begin{thebibliography}{00}
-  \bibitem{b1} G.
-Eason, B.
-Noble, and I.
-N.
-Sneddon, ``On certain integrals of Lipschitz-Hankel type involving products of Bessel functions,'' Phil.
-Trans.
-Roy.
-Soc.
-London, vol.
-A247, pp.
-529--551, April 1955.
-  \bibitem{b2} J. Clerk Maxwell, A Treatise on Electricity and Magnetism, 3rd ed., vol. 2. Oxford: Clarendon, 1892, pp.68--73.
-  \bibitem{b3} I. S. Jacobs and C. P. Bean, ``Fine particles, thin films and exchange anisotropy,'' in Magnetism, vol. III, G. T. Rado and H. Suhl, Eds. New York: Academic, 1963, pp. 271--350.
-  \bibitem{b4} K. Elissa, ``Title of paper if known,'' unpublished.
-  \bibitem{b5} R. Nicole, ``Title of paper with only first word capitalized,'' J. Name Stand. Abbrev., in press.
-  \bibitem{b6} Y.
-Yorozu, M.
-Hirano, K.
-Oka, and Y.
-Tagawa, ``Electron spectroscopy studies on magneto-optical media and plastic substrate interface,'' IEEE Transl.
-J.
-Magn.
-Japan, vol.
-2, pp.
-740--741, August 1987 [Digests 9th Annual Conf.
-Magnetics Japan, p.
-301, 1982].
-  \bibitem{b7} M. Young, The Technical Writer's Handbook. Mill Valley, CA: University Science, 1989.
-\end{thebibliography}
-\vspace{12pt}
-\color{red}
-IEEE conference templates contain guidance text for composing and formatting conference papers.
-Please ensure that all template text is removed from your conference paper prior to submission to the conference.
-Failure to remove the template text from your paper may result in your paper not being published.
+\bibliographystyle{IEEEtran}
+\bibliography{../definitions/biblio}
 

### 2023-10-30 paper/state.tex
Worked on paper+corrected errors

\section{State of the art}\label{sec:state}

\subsection{Recurrent Neural Networks (RNN)}

RNNs are a family of neural networks that differentiate themselves by having feedback connections. It is often used with sequences of data \cite{rnn}, sometimes, with varying amount of input data. RNNs are used for handwriting recognition and language translation \cite{gru}.

The feedback connection is caracterized by the hidden state vector ($\overrightarrow{h_t}$) which serves as the output and as part of the input for the next time step, a RNN is defined as \cref{eq:rnn}.

\begin{equation}\label{eq:rnn}
  \overrightarrow{h_t}=f(\overrightarrow{x_t},\overrightarrow{h_{t-1}})
\end{equation}

A the simple version of the RNN is defined by \cref{eq:srnn}.

\begin{equation}\label{eq:srnn}
  \overrightarrow{h_t}=tanh([\overrightarrow{x_t},\overrightarrow{h_{t-1}}]\cdot w + \overrightarrow{b})
\end{equation}

Where ($w$,$\overrightarrow{b}$) are the pair of weights matrix and bias vectors. $\overrightarrow{x_t}$ is the input vector and $\overrightarrow{h_t}$ is the hidden state vector.

The graphical representation of the simple RNN cell is shown in \cref{fig:rnnCell}.

\begin{figure}[htbp]
  \centering
  \begin{minipage}{\columnwidth}
    \subfloat[Simple \acs{RNN} cell\label{fig:rnnCell}]{\includesvg[width=\columnwidth]{rnn/rnnCell.svg}}
  \end{minipage}
  \begin{minipage}{\columnwidth}
    \subfloat[Legend\label{leg:cells}]{\includesvg[width=\columnwidth,pretex=\footnotesize]{cellsLegend.svg}}
  \end{minipage}
  \caption{}
\end{figure}

\subsection{Long Short Term Memory (LSTM)}

LSTMs are a type of RNN used to solve the vanishing gradiant problem \cite{firstLSTM}. It was improved a few times before becoming the modern LSTM \cite{improvLSTM}. The LSTM differs from a simple RNN because of its second feedback variable being the cell state ($\overrightarrow{c_t}$).

The LSTM contains four activated gates, that each serve its own purpose. The input gate \cref{eq:inputG} controls whether the cell state is updated, the forget gate \cref{eq:forgetG} that determines how the current cell state is affected by the old cell state, the output gate \cref{eq:outputG} that controls how much the hidden state is affected by the cell state. The candidate cell state gate \cref{eq:candCell} computes the change in the future cell state.

\begin{equation}\label{eq:inputG}
  \overrightarrow{i_t}=\sigma ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_i + \overrightarrow{b_i})
\end{equation}
\begin{equation}\label{eq:forgetG}
  \overrightarrow{f_t}=\sigma ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_f + \overrightarrow{b_f})
\end{equation}
\begin{equation}\label{eq:outputG}
  \overrightarrow{o_t}=\sigma ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_o + \overrightarrow{b_o})
\end{equation}
\begin{equation}\label{eq:candCell}
  \overrightarrow{\tilde{c}_t}=tanh([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_c+ \overrightarrow{b_c})
\end{equation}

The next part of the LSTM consist of computing the new cell state in the following equation :

\begin{equation}\label{eq:cellS}
  \overrightarrow{c_t}=\overrightarrow{f_t}\odot \overrightarrow{c_{t-1}} + \overrightarrow{i_t} \odot \overrightarrow{\tilde{c}_t}
\end{equation}

The hidden state is then determined using the current cell state \cref{eq:hiddenS}.

\begin{equation}\label{eq:hiddenS}
  \overrightarrow{h_t}=\overrightarrow{o_t}\odot tanh(\overrightarrow{c_t})
\end{equation}

Where ($w_i$,$\overrightarrow{b_i}$), ($w_f$,$\overrightarrow{b_f}$), ($w_o$,$\overrightarrow{b_o}$) and ($w_c$,$\overrightarrow{b_c}$) are the pair of weights matrixes and bias vectors for the input, forget, output and candidate cell gates respectively. $\overrightarrow{x_t}$ is the input vector and $\overrightarrow{h_t}$ is the hidden state vector.

The graphical representation of an LSTM cell is found in \cref{fig:lstmCell}.

\begin{figure}[htbp]
  \centering
  \includesvg[width=\columnwidth]{lstm/lstmCell.svg}
  \caption{\acs{LSTM} cell, adapted from \cite{wikiLSTM}\label{fig:lstmCell}}
\end{figure}

\subsection{Gated Recurrent Units (GRU)}

GRUs are another type of RNN. It is also known to reduce the effect of the vanishing gradient problem. It was first introduced to improve translation techniques \cite{gru}.

The \ac{GRU} is very often compared to the \ac{LSTM}, being sometimes assimilated as a type of \ac{LSTM} \cite{nbLSTM}. Their performance was found to be very similar in most situations \cite{gruVSlstm}, making those two types of \acp{RNN} coexistant in the modern machine learning world.

There are two versions of the \ac{GRU}, both are found on the internet, they are known as the encoder and decoder version \cite{gru}. They were originally designed to encode the message to translate and then decode in the translation. PyTorch only supports the decoder version \cite{gruPyTorch}, while the Keras library supports both \cite{gruKeras} chosen by changing an argument. This work only uses the encoder version, for that reason, it will be the only one described.

It contains an update gate \cref{eq:updateG}, a reset gate \cref{eq:resetG}, a candidate activation gate \cref{eq:candActivG}. The hidden state is then computed \cref{eq:gruHidG} using the previous results.

\begin{equation}\label{eq:updateG}
  \overrightarrow{z_t}=\sigma ([\overrightarrow{x_t},\overrightarrow{h_{t-1}}] \cdot w_z + \overrightarrow{b_z})
\end{equation}
\begin{equation}\label{eq:resetG}
  \overrightarrow{r_t}=\sigma ([\overrightarrow{x_t},\overrightarrow{h_{t-1}}] \cdot w_r + \overrightarrow{b_r})
\end{equation}
\begin{equation}\label{eq:candActivG}
  \overrightarrow{\hat{h_t}}=tanh(\overrightarrow{x_t}\cdot w_{hx}+(\overrightarrow{r_t}\odot\overrightarrow{h_{t-1}}) \cdot w_{hh} + \overrightarrow{b_h})
\end{equation}
\begin{equation}\label{eq:gruHidG}
  \overrightarrow{h_t}=(\overrightarrow{1}-\overrightarrow{z_t})\odot \overrightarrow{h_{t-1}} + \overrightarrow{z_t}\odot \overrightarrow{\hat{h_t}}
\end{equation}

Where ($w_z$,$\overrightarrow{b_z}$), ($w_r$,$\overrightarrow{b_r}$),($w_{hx}$,$w_{hh}$,$\overrightarrow{b_h}$) are the weights matrixes and bias vectors for the update, reset and candidate activation gates respectively.

A visual representation of the encoder \ac{GRU} cell is available in \cref{fig:encoderGruCell}.

\begin{figure}[htbp]
  \centering
  \includesvg[width=\columnwidth]{gru/encoderCell.svg}
  \caption{Encoder \acs{GRU} cell, legend in \cref{leg:cells}\label{fig:encoderGruCell}}
\end{figure}

\subsection{Memristors}

Memristors are the lesser known fourth fundamental passive component of electronics, among resistors, capacitors and inductor.
It was first theorized in 1971 as a missing fundamental component in \cite{TheoMemristor}. The name comes from the blend of \textit{memory} and \textit{resistance}.
The missing component linking the four fundamental circuit variables, voltage ($v$), charge ($q$), current ($i$) and flux ($\phi$). \Cref{fig:fundComp} shows the four fundamental variables are on each side of the square, with the ones on opposite sides being linked by the following equations :

\begin{equation}
  d\phi = v\cdot dt
\end{equation}

\begin{equation}
  dq = i\cdot dt
\end{equation}

Resistors, capacitors and inductors were already very established and well known components, so it was theorized that a fourth device should then exist to physically link flux ($\phi$) and charge ($q$).  The flux in this case is not a magnetic flux and is defined as such : $ d\phi=v\cdot dt \Rightarrow \phi =  \int v \,dt  $.

The component stayed theoretical until 2008 when it was implemented in a physical device for the first time \cite{memristorFab}. It took 37 years to have an actual working device.

An extension to the memristor, reffered to as the memristive device, was theorized in 1976 \cite{memrestiveDev}. The difference between the memristor and the memristive devices is its internal behavior. Memristive devices are commonly referred to as memristors as well.

\begin{figure}[htbp]
  \centering
  \includesvg[width=0.65\columnwidth,pretex=\tiny]{memristor/memristor}
  \caption{Fundamental passive components, adapted from \cite{memWiki}}
  \label{fig:fundComp}
\end{figure}

\subsection{Memristors Crossbar Arrays}

Setting memristors in a crossbar array allows to perform analog Vector Matrix Multiplication (VMM), also called Multiply and Accumulate. \Cref{fig:crossbar} shows what a typical crossbar array looks like.

\begin{figure}[htbp]
  \centering
  \includesvg[width=.6\columnwidth,pretex=\small]{crossbar/crossbar}
  \caption{Crossbar array schematics, inspired from \cite{xbarFigures}}
  \label{fig:crossbar}
\end{figure}

The circuit uses physical properties of electrical systems to perform analog computation. The following part will focus only on the circuit node in \cref{fig:crossNode}.

\begin{figure}[htbp]
  \centering
  \includesvg[wdith=0.4\columnwidth,pretex=\scriptsize]{crossbar/node}
  %\def\svgheigth{3.5cm}
  %\input{crossbar/node.pdf_tex}
  \caption{Memristor crossbar node of the $i^{th}$ line and $j^{th}$ column}
  \label{fig:crossNode}
\end{figure}

A voltage is applied on the $i^{th}$ line, and because every column is virtually grounded, the voltage applied to the memristor, with resistance $R_i$, is $V_i$. By applying Ohm's law, we know that the current flowing into the memristor ($I_{i}$) is bound by the following equation :

\begin{equation}
  V_i = R\cdot I_{i} \Rightarrow I_{i} = V_i\cdot (\frac{1}{R_i})= V_i\cdot\sigma_i
\end{equation}
With $\sigma_i$ being the conductance of memristor, defined as $\sigma_i=\frac{1}{R_i}$.

This line then joins the column where a current of $I_{i,j-1}$ is flowing, then according to Kirchhoff's current law the resulting current is :
\begin{equation}
  I_{j,i} = I_{j,i-1}+I_{i} = I_{j,i-1} + V_i\cdot\sigma_i
\end{equation}
Unfolding the equations will give the current at the bottom of the column, for example, the current at the bottom of the first column in \cref{fig:crossbar} is :
\begin{equation}
  I_1=\sigma_1\cdot V_1 + \sigma_2\cdot V_2 + \sigma_3\cdot V_3
\end{equation}
With $\sigma_1$, $\sigma_2$ and $\sigma_3$ being the conductance of the 3 memristors in the first column.

### 2023-10-30 paper/tools.tex
Worked on paper+corrected errors

\section{Python tools}

\subsection{Weights generation}

The weigths were trained in python \cite{python} using the tensorflow library \cite{tensorflow}.

The weights need to be trained. The training is performed using the analog activation functions generated by the circuits in \cref{subsec:af}. The circuit is thus trained as if it were trained directly on the chip, circuit's inaccuracies aside.

The weights are constrained to make sure the voltage is within the active voltage range of the circuit. If the voltage gets out of range, the weights will not be correct as the voltage cannot get out of range in the physical circuit.

Once the weights are trained, they are saved in a file, that will later serve as the basis to generate the netlist.

The code used to generate the weights is available on my github repo \cite{lstmWei}.

\subsection{Netlist generation}

In order to be able to run the simulation with different kinds of \ac{NN} architecture of varying sizes. The point of the part is thus to explain how the netlist generator tool works.

The netlist is generated using a python script to generate a SPICE netlist. It can generate the netlist for different kinds of circuits. It can generate both LSTM and GRU circuits, with any size of input, time steps or even number of hidden states. When it is possible the serial size can also be selected.

The weights are stored in a single file. This file contains a brief description of the architecture being used. The first index stores this desciption. The following indexes are for the weights of each layer, in the order given in the description. The weights are stored in a python list and are distributed among the different resistances.

All the code used to generate the netlist can be found on my github page \cite{lstmGen}.

### 2023-10-30 state/crossbar.tex
Worked on paper+corrected errors

@@
 This circuit works great for this use as all the computation is done almost instantly and using physics laws.
-Figure \ref{fig:crossbar} shows what a typical crossbar array looks like.
+\Cref{fig:crossbar} shows what a typical crossbar array looks like.
 
@@
 It uses physical properties of electrical systems to perform analog computing.
-In what follows I will use the circuit node in figure \ref{fig:crossNode} to explain the theory behind the memristor crossbar array.
+In what follows I will use the circuit node in \cref{fig:crossNode} to explain the theory behind the memristor crossbar array.
 \begin{figure}[H]
@@
 First of all, a voltage is applied on the $i^{th}$ line, because every column is virtually grounded, the voltage applied to the memristor, of a set resistance $R_i$, is $V_i$.
-As such and using Ohm's law, we know that the current flowing into the memristor ($I_{mem}$) is bound by the following equation :
+As such and using Ohm's law, we know that the current flowing into the memristor ($I_{i}$) is bound by the following equation :
 
 \begin{equation}
-  V_i = R\cdot I_{mem} \Rightarrow I_{mem} = V_i\cdot (\frac{1}{R})=I_{mem} = V_i\cdot\sigma_i
+  V_i = R\cdot I_{i} \Rightarrow I_{i} = V_i\cdot (\frac{1}{R})= V_i\cdot\sigma_i
 \end{equation}
@@
 \begin{equation}
-  I_{j,i} = I_{j,i-1}+I_{mem} = I_{j,i-1} + V_i\cdot\sigma_i
+  I_{j,i} = I_{j,i-1}+I_{i} = I_{j,i-1} + V_i\cdot\sigma_i
 \end{equation}

### 2023-10-30 state/lstm.tex
Correcting inappropriate raccourcis and correcting design
Removed a todo
Worked on paper+corrected errors

@@
 
-The first \ac{LSTM} was invented in 1997 by Hochreiter and Schmidhuber \cite{firstLSTM}. \acp{LSTM} changed a lot through the years to become what they are now.
+The first \ac{LSTM} was invented in 1997 by Hochreiter and Schmidhuber \cite{firstLSTM}. \acp{LSTM} changed a lot through the years to become what they are now \cite{improvLSTM}.
 
@@
 The most comprehensible example is considering a sentence.
-%TODO : find example)
 
@@
   \includesvg[width=\textwidth,pretex=\large]{lstm/lstmCell.svg}
+  \caption{\acs{LSTM} cell, adapted from \cite{wikiLSTM}}
   \label{fig:lstmCell}
-  \caption{\acs{LSTM} cell, adapted from \cite{wikiLSTM}}
 \end{figure}
@@
 \Cref{fig:lstmCell} shows the complexity of the \ac{LSTM} architecture.
-In an \ac{LSTM}, each gate is a different \ac{NN} and then activated with either a \ac{tanh} or a sigmoid activation function.
+In an \ac{LSTM}, each gate is a different \ac{NN} and then acivated with either a \ac{tanh} or a sigmoid activation function.
 Each input to the cell is a vector.
@@
 
-The equations that haven't been rewritten simply stay the same.
+The equations that have not been rewritten simply stay the same.
 
@@
 
-Once again, the equations that haven't been rewritten are unchanged.
+Once again, the equations that have not been rewritten are unchanged.
 

### 2023-10-30 paper/dataset.tex
New figure
finished dataset

@@
 The dataset then contains twelve years of monthly data so $144$ sample points.
-Although the esults of this problem are not necessarely tailored to the system, in the sense that it is not time sensitive information, it is nonetheless a problem that can, in a straightforward manner, demonstrate the predictive capacity of the proposed model.
-It is extracted from the tutorial available at \cite{airline}.
-The dataset is available at \cite{datasets}, the name of the file containing the dataset is \textit{airline-passengers.csv}.
+Although the results of this problem are not necessarely tailored to the system, in the sense that it is not time sensitive information, it is nonetheless a problem that can, in a straightforward manner, demonstrate the predictive capacity of the proposed model.
+The dataset is graphed in \cref{graph:airline}.
 
@@
 
-\begin{figure}[H]
+\begin{figure}[h]
   \centering
-\includesvg[width=\textwidth]{datasets/airline}
+  \includesvg[width=\columnwidth,pretex=\scriptsize]{datasets/airline}
   \caption{The airline dataset. The vertical lines represent a full year.}
@@
 \end{figure}
+
+The $144$ sample points have been transformed into $142$ data points for training.
+The data is grouped by three, two for the input and one the target output.
+Two third of the dataset is being used for training and the other third is used for validation.
+
+\subsection{C. elegans}
+
+This dataset is far more interesting and complex than the latter.
+This data set aims to use \acp{LSTM} to mimic the behavior of real neurons.
+As explained in \cite{celegans}, Caenorhabditis elegans are simple organisms that are getting very popular for whole brain organization studies.
+The point of this problem is to reproduce the nervous system of the \ac{C.
+elegans}.
+This is done using recorded data of the input of 4 neurons and the output of 4 other neurons.
+
+The dataset is great for our study because :
+\begin{itemize}
+  \item It comes from a very recent paper from the research group in which this work is also being developed.
+  \item It aims at reproducing the behavior of the brain of a simple organism (\ac{C. elegans}). In the (very) long term, a full parts of the human brain could be replaced by a very low powered chip.
+\end{itemize}
+
+The dataset contains a set of fourty set of input/output sequences.
+Each set of input and output set contains $500ms$ of data with a time step of $0.5ms$, making a thousand points for each neurons.
+This data is recorded after applying current to the input neurons (PLML2, PLMR, AVBL, AVBR), and monitoring the output of four neurons (DB1, LUAL, PVR, VB1) that are known to have strong activity in response to the four inputs neurons.
+The simulation is done using a known model of \ac{C.
+elegans} connectome (complete overview of the brains connections) \cite{celegans}.

### 2023-10-30 design/dataset.tex
Possibly finished conc
Worked on paper+corrected errors
corrections

@@
 To make sure that the terminal voltage does not reach the threshold as explained in \cref{sec:xbarCircuit}.
-The voltage threshold ($V_{threshold}$) was chosen to be $V_{threshold}=0.1V$ because the memristors used in this thesis are purely theorical.
+The voltage threshold ($V_{read}$) was chosen to be $V_{read}=0.1V$ because the memristors used in this thesis are purely theorical.
 The dataset needs to be formated not to exceed this value when converted to voltage (\cref{tab:valConv}), the dataset needs to be trained while not exceeding $1$.
 
-\subsection{Airline}
+\subsection{Airline passengers}
 
@@
 The dataset then contains twelve years of monthly data so $144$ sample points.
-Although the esults of this problem are not necessarely tailored to the system, in the sense that it is not time sensitive information, it is nonetheless a problem that can, in a straightforward manner, demonstrate the predictive capacity of the proposed model.
+Although the results of this problem are not necessarely tailored to the system, in the sense that it is not time sensitive information, it is nonetheless a problem that can, in a straightforward manner, demonstrate the predictive capacity of the proposed model.
 It is extracted from the tutorial available at \cite{airline}.
@@
 
-\subsection{\ac{C. elegans}}
+\subsection{\ac{C. elegans}}\label{subsec:celegans}
 

### 2023-10-30 paper/results.tex

\section{Results}

The target for our circuit is to reproduce, as closely as possible the digital prediction.

\subsection{Airline passengers}

The problem is solved using an LSTM or GRU with four hidden state connected to a Dense layer with an output size of one. The weights were trained for three hundreds epochs.

\begin{figure}[h]
  \centering
  \includesvg[width=\columnwidth, pretex=\scriptsize]{results/airline/analog}
  \caption{Analog predictions trained with the analog activation functions.}
  \label{graph:airlineAnalog}
\end{figure}

The analog results obtained in \cref{graph:airlineAnalog} are very close to the digital predictions. This is confirmed by the values in \cref{tab:airlineAnalog}. The results the closest to the target are the ones generated using a serial size of one ($n_s=1$). Indeed, its error ($28.4$) is lower than the error from the digital results ($52.3$) to the original target values, almost half of it. The other ones are still quite close to the targeted curve, but are still lower. The errors are higher when running the system in serial mode. This higher inaccuracy is probably due to the data staying for longer in the memory cells as the time steps get longer.

\begin{table}[h]
  \caption{\acp{RMSE} of each analog prediction to their associated digital prediction}
  \label{tab:airlineAnalog}
  \centering
  \begin{tabular}{|c|c|c|c|c|}
    \cline{2-5}
    \multicolumn{1}{c}{}& \multicolumn{3}{|c|}{Analog prediction} &Target\\
    \cline{1-4}
    Serial size & $n_s=1$ & $n_s=2$ & $n_s=4$ & data\\
    \hline
    Digital prediction & $28.4$ & $92.6$ & $68.5$ & $52.3$\\
    \hline
  \end{tabular}
\end{table}

The GRU circuit not being fully working produces very off predictions. For this reason, the GRU results are not included.

\subsection{C. elegans}

The problem is solved using an LSTM or GRU with eights hidden state connected to a Dense layer with an output size of four. The weights were trained for one thousand epochs.

\begin{figure}[h]
  \centering
  \includesvg[width=\columnwidth,pretex=\scriptsize]{results/celegans/out5}
  \caption{\acs{C. elegans} analog responses with their digital counterparts}
  \label{graph:celegansAnalog0}
\end{figure}

\Cref{graph:celegansAnalog0} combines the digital and analog of all four output neurons. However the analog predictions are quite unstable and barely readable. The origin of this issue being unknown, the output are thus artificially smoothed out using software. The

\begin{figure}[h]
  \centering
  \includesvg[width=\columnwidth,pretex=\scriptsize]{results/celegans/smooth5DB1}
  \caption{Averaged analog prediction forscriptsize DB1 neuron\label{graph:smooth5BD1Celegans}}
\end{figure}
\begin{figure}[h]
  \centering
  \includesvg[width=\columnwidth,pretex=\scriptsize]{results/celegans/smooth5LUAL}
  \caption{Averaged analog prediction forscriptsize LUAL neuron\label{graph:smooth5LUALCelegans}}
\end{figure}
\begin{figure}[h]
  \includesvg[width=\columnwidth,pretex=\scriptsize]{results/celegans/smooth5PVR}
  \caption{Averaged analog prediction forscriptsize PVR neuron\label{graph:smooth5PVRCelegans}}
\end{figure}
\begin{figure}[h]
  \includesvg[width=\columnwidth,pretex=\scriptsize]{results/celegans/smooth5VB1}
  \caption{Averaged analog prediction for the VB1 neuron\label{graph:smooth5VB1Celegans}}
\end{figure}

### 2023-10-30 conclusions/conclusion.tex
Supposedly finsihed sending for review

@@
 The predictions are a bit late and it is not clear what is causing this specific issue.
-A theory is that the memory cells deteriorate its stored value by a very small amount every time step, thus the stored value is affect a lot after a large amount of time steps.
+A theory is that the memory cells deteriorate its stored value by a very small amount every time step, thus the stored value is affected a lot after a large amount of time steps.
 
@@
 
-The airline problem feeding two inputs to the circuit to get a single output, it will have to be run $144$ times to get all the output data.
+The airline problem feeding two inputs to the circuit to get a single output, it will have to be run $142$ times to get all the output data.
 The inference time depends on the serial size being used :

### 2023-10-30 paper/conclusion.tex
Supposedly finsihed sending for review

--- 2023-10-29 conclusions/conclusion.tex
@@
-\section{Conclusion}\label{sec:conc}
+\section{Conclusion}
 
-\subsection{Performances}
-
-\subsubsection{\ac{LSTM}}
+\subsection{\ac{LSTM}}
 
@@
 The best performance is the with the airline dataset, using the circuit with a serial size of one ($n_s=1$) (\cref{tab:airlineAnalog}).
-While the results are not quite right, it is assumed that this issue will be elevated once inSitu training is implemented (\cref{subsec:inSitu}).
+While the results are not quite right, it is assumed that this issue will be elevated once inSitu training is implemented.
 
@@
 
-With a simple input sequence, such as sequence 5 (\cref{graph:io5Celegans}), the resulting predictions are quite good and on par with what is expected.
+With a simple input sequence, such as sequence 5 (\cref{graph:celegansAnalog1}), the resulting predictions are quite good and on par with what is expected.
 The predictions are a bit late and it is not clear what is causing this specific issue.
-A theory is that the memory cells deteriorate its stored value by a very small amount every time step, thus the stored value is affect a lot after a large amount of time steps.
+A theory is that the memory cells deteriorate its stored value by a very small amount every time step, thus the stored value is affectd a lot after a large amount of time steps.
 
 More complex input sequences produce even more inaccuracies.
-Indeed, when working with sequence 15 (\cref{graph:io15Celegans}), the inaccuracies explode and gives out a barely usable prediction.
+Indeed, when working with sequence 15, not shown in this paper, the inaccuracies explode and gives out a barely usable prediction.
 The curve generated does not fit its digital counterpart anymore.
@@
 
-\subsubsection{\ac{GRU}}
+\subsection{\ac{GRU}}
 
-The \ac{GRU} circuit, while being present, is still a work in progress.
+The \ac{GRU} circuit, despite not having result in this paper, is still a work in progress.
 The outputed predictions are scaled down for a still unknown reason, but show the right output shape.
@@
 
-The inference time is quite tricky to deal with.
-While the analog circuit's inference time is set, it gets more complicated when trying to get the digital inference time.
-The digital inference time depends on the computer it is being run on.
-The \acp{CPU} used in embedded sytems, the main application for this work, are usually not very fast as they are intended to be low powered.
-This parts will thus only dicuss of the inference of the analog circuit.
-
-\subsubsection{Airline inference}
-
-The airline problem feeding two inputs to the circuit to get a single output, it will have to be run $144$ times to get all the output data.
-The inference time depends on the serial size being used :
-
-\begin{itemize}
-  \item $n_s=1$ : The output can be read as a voltage on the output net from $t_0=18\mu s$ to $t_1=22\mu s$.
-  \item $n_s=2$ : The output can be read as a voltage on the output net from $t_0=34\mu s$ to $t_1=38\mu s$.
-  \item $n_s=4$ : The output can be read as a voltage on the output net from $t_0=66\mu s$ to $t_1=70\mu s$.
-\end{itemize}
-
-Those are the times it takes to compute one of the outputs. It takes $144$ times longer to get all of the results.
-
-\subsubsection{\ac{C. elegans} inference}
-
-This sequence is a bit more complicated, for every input sequence with a thousand time steps, there are a thousand output vectors being read. Here, the inference time also depends on the serial size.
-
-\begin{table}[H]
-  \centering
-  \begin{tabular}{|c|c|c|c|c|}
-    \hline
-    \rowcolor{gray}
-    Serial size & $n_s=1$ & $n_s=2$ & $n_s=4$ & $n_s=8$\\
-    \hline
-    Start of first output & $9\mu s$ & $17\mu s$ & $33\mu s$ & $65\mu s$\\
-    \hline
-    End of first output & $13\mu s$ & $21\mu s$ & $37\mu s$ & $69\mu s$\\
-    \hline
-    Output period & $9\mu s$ & $17\mu s$ & $33\mu s$ & $65\mu s$\\
-    \hline
-    Start of last output & $9ms$ & $17ms$ & $33ms$ & $65ms$\\
-    \hline
-    End of last output & $9ms +4\mu s$ & $17ms+4\mu s$ & $33ms+4\mu s$ & $65ms+4\mu s$\\
-    \hline
-  \end{tabular}
-  \caption{First and last read times for the different serial sizes}
-  \label{tab:readTimesCelegans}
-\end{table}
-
-\Cref{tab:readTimesCelegans} shows the first and last output time range with the period.
-All of the output times can be determined using the first time range and using the period.
-A sequence consists of $500ms$ as explained in \cref{subsec:celegans}, the time steps are physically $500\mu s$ appart, in the case where the circuit would be used to do real time inference, the circuit would have to be slowed down to be on sync with the inputs.
-The circuit could could be used in a real time use for this time sensitive problem.
-
-\subsection{onChip area}\label{subsec:area}
-
-The \ac{LSTM} and \ac{GRU} blocks' onChip area are impossible to estimate, as parts of the circuits are made of verilog-A models.
-However, to get a general idea of the area of the chip, it can be assumed that the area of the circuit mainly depends on the the area of a memrisor, because the area of a memristor is much greater than the ones of the other components.
-Since the number of memristors is the number of weights in the circuit, the minimum area of any \ac{NN} can be determined.
-The area for a memristor was determined using the feature size of the memristors that can be fabricated at \ac{INESC}.
-The typical memristor that can be fabricated at \ac{INESC} is $3\mu m$, which would make the approximate area for a memristor $A_{memristor}=9\mu m^2=9\cdot 10^{-12} m^2$
-
-\Cref{tab:areas} contains several \acp{NN} models with their number of parameters, a minimum onChip area and an estimated feature size.
-None of those values have any scientific ground, they are just here to give general idea of the kind of circuit thta could be fabricated.
-Those areas only considers the area of the memristors, using a two memristor per synapse architecture like the one used in the thesis.
-
-The feature size represents the length of the side of the chip if the chip was manufactured in a square.
-
-The values of \acp{LLM}, such as \ac{LLaMA}-2 \textbf{Meta}'s \ac{LLM} and the two last versions of \acp{GPT} the \ac{LLM} that powers \textit{ChatGPT}, were included to show how the area of the chip scales up.
-Those models do not use \acp{RNN}, and futhermore, give out ridiculous areas.
-Fabricating such cicuits in analog is not feasible witht the current technologies.
-
-\begin{table}[H]
-  \centering
-  \begin{tabular}{|c|c|c|c|}
-    \hline
-    \rowcolor{gray}
-    Model & Parameters & Minimum area & Approximate feature size \\
-    \hline
-    Airline \ac{GRU} $n_h=4$ & $77$ & $1386\mu m^2$ & $37.2\mu m$\\
-    \hline
-    Airline \ac{LSTM} $n_h=4$ & $101$ & $1818\mu m^2$ & $42.6\mu m$\\
-    \hline
-    \ac{C. elegans} \ac{GRU} $n_h=8$ & $348$ & $6264\mu m^2$ & $79.1\mu m$\\
-    \hline
-    \ac{C. elegans} \ac{LSTM} $n_h=8$ & $452$ & $8136\mu m^2$ & $90.2\mu m$\\
-    \hline
-    \multirow{3}{*}{\acs{LLaMA}-2} & $7\cdot 10^9$ & $1260 cm^2$ & $35.5 cm$ \\
-    \cline{2-4}
-    & $13\cdot 10^9$ & $2340 cm^2$ & $48.4 cm$ \\
-    \cline{2-4}
-    & $70\cdot 10^9$ & $1.26 m^2$ & $1.12 m$ \\
-    \hline
-    \acs{GPT}-3 & $175\cdot 10^9$ & $3.14 m^2$ & $1.77m$\\
-    \hline
-    \acs{GPT}-4 & $1\cdot 10^{12}$ & $18.0 m^2$ & $4.24m$\\
-    \hline
-  \end{tabular}
-  \caption{Estimated onChip area for different models of \acp{NN}}
-  \label{tab:areas}
-\end{table}
-
-Nevertheless, it is interresting to look at those approximation.
-It shows that the simple models used in this work would take a very small onChip area, even though this a very raw number.
-It also shows the current limit with state of the art \acp{NN} at the current time.
-
-The onChip area of small \acp{NN} such as the ones used in this work are very reasonable and could easily be fabricated for embedded use.
+The execution time is only interresting when dealing with time sensistive data, like with the C.
+elegans problem.
+The problem gets input every $500\mu s$, however, the circuit takes at most $65\mu s$ to compute when using eight hidden states and a serial size of eight.
+This execution time would even allow to take in more frequent inputs, like the $100\mu s$ sequences dealt with in \cite{celegans}.

### 2023-10-30 paper/results.tex

@@
 \Cref{graph:celegansAnalog0} combines the digital and analog of all four output neurons.
-However the analog predictions are quite unstable and barely readable.
+However the analog predictions are quite unstable.
 The origin of this issue being unknown, the output are thus artificially smoothed out using software.
-The
+The graph with the smoothed out curves is \cref{graph:celegansAnalog1}.
 
@@
   \centering
-  \includesvg[width=\columnwidth,pretex=\scriptsize]{results/celegans/smooth5DB1}
-  \caption{Averaged analog prediction forscriptsize DB1 neuron\label{graph:smooth5BD1Celegans}}
+  \includesvg[width=\columnwidth,pretex=\tiny]{results/celegans/smooth5}
+  \caption{\acs{C. elegans} analog responses with their digital counterparts}
+  \label{graph:celegansAnalog1}
 \end{figure}
-\begin{figure}[h]
-  \centering
-  \includesvg[width=\columnwidth,pretex=\scriptsize]{results/celegans/smooth5LUAL}
-  \caption{Averaged analog prediction forscriptsize LUAL neuron\label{graph:smooth5LUALCelegans}}
-\end{figure}
-\begin{figure}[h]
-  \includesvg[width=\columnwidth,pretex=\scriptsize]{results/celegans/smooth5PVR}
-  \caption{Averaged analog prediction forscriptsize PVR neuron\label{graph:smooth5PVRCelegans}}
-\end{figure}
-\begin{figure}[h]
-  \includesvg[width=\columnwidth,pretex=\scriptsize]{results/celegans/smooth5VB1}
-  \caption{Averaged analog prediction for the VB1 neuron\label{graph:smooth5VB1Celegans}}
-\end{figure}
+
+The output neurons predictions are quite good, expect for the unstable analog predictions. The results are very promising.
+
+The other sequences of inputs were ran but connot all be shown here. More details on the C . elegans analog predictions are available in the thesis this paper goes with.

### 2023-10-31 paper/circuit.tex
Rearranged figures
Removed color
Fixed tables
added lstm symbol

@@
-\section{Design}
+\section{Circuits}
 
@@
 
-\begin{table}[h]
+\begin{table}[b]
+  \caption{Real/Voltage Conversion Table.}
+  \label{tab:valConv}
   \centering
@@
     \hline
-    \rowcolor{gray}
-    Real value & Voltage \\
+    \textbf{Real value} & \textbf{Voltage} \\
     \hline
@@
   \end{tabular}
-  \caption{Real/Voltage Conversion Table.}
-  \label{tab:valConv}
 \end{table}
@@
 
-\begin{figure}[htbp]
+\begin{figure}[t]
   \centering
@@
 
-\begin{table}[htbp]
+\begin{table}[b]
   \caption{Circuits parameters}
@@
     \hline
-    \rowcolor{gray}
-    Parameter & Sigmoid & \ac{tanh} \\
+    \textbf{Parameter} & \textbf{Sigmoid} & \textbf{\ac{tanh}} \\
     \hline
@@
 
-\begin{figure}[h]
+\begin{figure}[t]
   \centering
@@
 
-\begin{figure}[htbp]
+\begin{figure}[b]
   \centering
@@
 
-\begin{figure}[h]
+\begin{figure}[b]
   \centering
@@
 
-\begin{figure*}[h]
+\begin{figure*}[t]
   \centering
@@
 
+The symbol of the \ac{LSTM} circuit can be found in \cref{sym:lstm}.
+It depends on several parameters, the number of inputs ($n_i$), the number of hidden states ($n_h$), the serial size of the system ($n_s$) and the number of time steps ($n_{ts}$).
+
+\begin{figure}[b]
+  \centering
+  \includesvg[height=1cm,pretex=\footnotesize]{lstm/lstmSymbol}
+  \caption{Symbol used for the \ac{LSTM} circuit}
+  \label{sym:lstm}
+\end{figure}
+
 \subsection{GRU analog implementation}
@@
 For that reason it has a similar circuit to LSTM circuit (\cref{circt:lstm}).
-The system is built, once again, using crossbar array (\cref{sec:xbarCircuit}) with ($n_i+n_h+1$,$n_h$, $1$) as parameters.
+The system is built, once again, using crossbar array with ($n_i+n_h+1$,$n_h$, $1$) as parameters.
 

### 2023-10-31 paper/dataset.tex
Rearranged figures
Raccourci

@@
 This data set aims to use \acp{LSTM} to mimic the behavior of real neurons.
-As explained in \cite{celegans}, Caenorhabditis elegans are simple organisms that are getting very popular for whole brain organization studies.
+As explained in \cite{celegans}, Caenorhabditis elegans (C.
+elegans)  are simple organisms that are getting very popular for whole brain organization studies.
 The point of this problem is to reproduce the nervous system of the \ac{C.
@@
 
+An example of an input/output sequence is shown in \cref{graph:celegansAnalog1} along with the digital and analog predictions.
+
 The dataset contains a set of fourty set of input/output sequences.

### 2023-10-31 paper/conclusion.tex
Fixed one mistake

@@
 The predictions are a bit late and it is not clear what is causing this specific issue.
-A theory is that the memory cells deteriorate its stored value by a very small amount every time step, thus the stored value is affectd a lot after a large amount of time steps.
+A theory is that the memory cells deteriorate its stored value by a very small amount every time step, thus the stored value is affected a lot after a large amount of time steps.
 

### 2023-10-31 design/models.tex
Wrong links

@@
 
-This is taken care of using in reality two parts, the actual voltage multiplier (\cref{eq:voltMult}) and a non inverting amplifier (\cref{eq:invAmp}), the circuit of which is available in \cref{apsec:invAmp}.
+This is taken care of using in reality two parts, the actual voltage multiplier (\cref{eq:voltMult}) and a non inverting amplifier (\cref{eq:invAmp}), the circuit of which is available in \cref{apsec:sumInv}.
 
@@
 
-Where $V_{out}$ is the output voltage the inverting amplifier (\cref{apsec:invAmp}), $V_{voltMult}$ is the out voltage of the voltage multiplier itself and $V_{in_1}$ and $V_{in_2}$ are the input voltages.
+Where $V_{out}$ is the output voltage the inverting amplifier (\cref{apsec:sumInv}), $V_{voltMult}$ is the out voltage of the voltage multiplier itself and $V_{in_1}$ and $V_{in_2}$ are the input voltages.
 

### 2023-11-02 conclusions/imag.tex
Impact environmental imag de merde

\section{Environmental and societal impact}

\subsection{Personnal environmental impact}

\subsection{Global impact}

\subsubsection{Environmental impact}

\subsubsection{Societal impact}

\subsection{Host structure's policy}

### 2023-11-27 design/crossbar.tex
corrected a few things

@@
 This design is the one that is used in \cite{doubleMem}.
-Let's assume that a given memristor has a resistance range of $R\in[R_{min},R_{max}]$, that means its conductance range is $\sigma \in [\sigma_{min},\sigma_{max}]$ (with $\sigma_{min}= \frac{1}{R_{max}}$ and $\sigma_{max}= \frac{1}{R_{min}}$).
+Let's assume that a given memristor has a resistance range of $R\in[R_{min},R_{max}]$, that means its conductance range is $ G \in [ G_{min}, G_{max}]$ (with $ G_{min}= \frac{1}{R_{max}}$ and $ G_{max}= \frac{1}{R_{min}}$).
 This design works using two \ac{opAmp} connected to $V_{cm}$ with the positive pin and the negative pin to the output of the crossbar array.
@@
   \label{eq:doubleMem2}
-  V_{opAmp_1}=y_0=-R_f\cdot(i_--i_+)=R_f\cdot(i_+-i_-)=R_f\cdot\sum_{k=0}^n(\sigma_{k+}-\sigma_{k-})\cdot x_k
+  V_{opAmp_1}=y_0=-R_f\cdot(i_--i_+)=R_f\cdot(i_+-i_-)=R_f\cdot\sum_{k=0}^n( G_{k+}- G_{k-})\cdot x_k
 \end{equation}
-With $i_+=\sum_{k=0}^n\sigma_{k+}\cdot x_k$ and $i_-=\sum_{k=0}^n\sigma_{k-}\cdot x_k$.
+With $i_+=\sum_{k=0}^n G_{k+}\cdot x_k$ and $i_-=\sum_{k=0}^n G_{k-}\cdot x_k$.
 
@@
     \hline
-    Maximum weight & $\sigma_{max}-\sigma_{min}$ & $\sigma_{max} -\overline{\sigma}$\\
+    Maximum weight & $ G_{max}- G_{min}$ & $ G_{max} -\overline{ G}$\\
     \hline
-    Minimum weight & $\sigma_{min}-\sigma_{max}$ & $\sigma_{min} -\overline{\sigma}$\\
+    Minimum weight & $ G_{min}- G_{max}$ & $ G_{min} -\overline{ G}$\\
     \hline
-    Range & $2\cdot(\sigma_{max}-\sigma_{min})$&$\sigma_{max}-\sigma_{min}$\\
+    Range & $2\cdot( G_{max}- G_{min})$&$ G_{max}- G_{min}$\\
     \hline

### 2023-11-27 design/memcell.tex
Renamed files and folder
area stuff
Worked on paper+corrected errors
corrected a few things

--- 2023-10-02 3.Design/memcell.tex
@@
 The circuit is shown in \cref{circt:memcell}. The value/voltage is trapped in the capacitor using \ac{CMOS} switches.
+
+The \ac{CMOS} switches found in the circuit have width of $w=200nm$ and a length of $l=60nm$.
 
@@
 
-For this circuit the approximate onChip area :
-
-\begin{equation}
-  A_{memcell}=8\cdot A_{CMOS}+2\cdot A_{inv}+A_{opAmp}+A_{cap}
-\end{equation}
-
-With $A_{CMOS} = w\cdot l = 2\cdot 10^{-7} \cdot 6 \cdot 10^{-8} = 1.2 \cdot 10^{-14} m^2$
-
 \subsection{Symbol}
@@
   \includesvg[height=2.5cm]{memcell/memcellSymbol}
-  \caption{Memory cell symbol with the input enable pin (top) and the output enable pin (bottom). The left and right pins are interchangebly the input and ouput for the sake of readability}
-  \label{circt:memcell}
+  \caption{Memory cell symbol with the input enable pin (top) and the output enable pin (bottom). The left and right pins can be either input or output. The component is non-linear.}
+  \label{sym:memcell}
 \end{figure}
+
+For this circuit the approximate onChip area is :
+
+\begin{equation}
+  A_{memcell}=8\cdot A_{CMOS}+2\cdot A_{inv}+A_{opAmp}+A_{cap}
+\end{equation}
+
+With $A_{CMOS} = w\cdot l$
 

### 2023-11-27 design/run.tex
corrected a few things

@@
 The time step ($T$) value was chosen to be $T=8\mu s$, the pause between two \ac{LSTM} steps is set to be $\frac{T}{8}=1\mu s$.
-The value for the maximum and minimum resistances value of the memristor used ($R_{max}$ and $R_{min}$, respectively) are set to be in acordance with the \textbf{KNOWN}'s memristors.
+The value for the maximum and minimum resistances value of the memristor used ($R_{max}$ and $R_{min}$, respectively) are set to be in acordance with the \textbf{KNOWM}\textregistered{}'s memristors \cite{Knowm}.
 The resistance values used for the simulation are thus $R_{max}=10^6 \Omega=1M\Omega$ and $R_{min}=10^4 \Omega = 10k\Omega$.

### 2023-11-27 state/crossbar.tex
corrected a few things

@@
 \begin{equation}
-  V_i = R\cdot I_{i} \Rightarrow I_{i} = V_i\cdot (\frac{1}{R})= V_i\cdot\sigma_i
+  V_i = R\cdot I_{i} \Rightarrow I_{i} = V_i\cdot (\frac{1}{R})= V_i\cdot G_i
 \end{equation}
-With $\sigma_i$ being the conductance of memristor, defined as $\sigma_i=\frac{1}{R_i}$.
+With $G_i$ being the conductance of memristor, defined as $ G_i=\frac{1}{R_i}$.
 
@@
 \begin{equation}
-  I_{j,i} = I_{j,i-1}+I_{i} = I_{j,i-1} + V_i\cdot\sigma_i
+  I_{j,i} = I_{j,i-1}+I_{i} = I_{j,i-1} + V_i\cdot G_i
 \end{equation}
@@
 \begin{equation}
-  I_1=\sigma_1\cdot V_1 + \sigma_2\cdot V_2 + \sigma_3\cdot V_3
+  I_1= G_1\cdot V_1 +  G_2\cdot V_2 +  G_3\cdot V_3
 \end{equation}
-With $\sigma_1$, $\sigma_2$ and $\sigma_3$ being the conductance of the 3 memristors in the first column.
+With $G_1$, $G_2$ and $G_3$ being the conductance of the 3 memristors in the first column.
 

### 2023-11-27 paper/circuit.tex
Reorganize graphs
Corrections to paper

@@
 
-\begin{table}[b]
+\begin{table}[t]
   \caption{Circuits parameters}
@@
 
-\begin{figure}[t]
+\begin{figure}[b]
   \centering
@@
 
-\begin{figure}[b]
+\begin{figure}[t]
   \centering
@@
     V_{opAmp_1}=y_0=-R_f\cdot(i_--i_+)=R_f\cdot(i_+-i_-)\\
-    =R_f\cdot\sum_{k=0}^n(\sigma_{k+}-\sigma_{k-})\cdot x_k
+    =R_f\cdot\sum_{k=0}^n( G_{k+}- G_{k-})\cdot x_k
   \end{array}
 \end{equation}
-With $i_+=\sum_{k=0}^n\sigma_{k+}\cdot x_k$ and $i_-=\sum_{k=0}^n\sigma_{k-}\cdot x_k$.
+With $i_+=\sum_{k=0}^n G_{k+}\cdot x_k$ and $i_-=\sum_{k=0}^n G_{k-}\cdot x_k$.
 

### 2023-11-27 paper/state.tex
corrected mistake one
Memristor desc
Corrections to paper

@@
 
-A the simple version of the RNN is defined by \cref{eq:srnn}.
+The simple version of the RNN is defined by \cref{eq:srnn}.
 
@@
 \begin{equation}\label{eq:inputG}
-  \overrightarrow{i_t}=\sigma ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_i + \overrightarrow{b_i})
+  \overrightarrow{i_t}= G ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_i + \overrightarrow{b_i})
 \end{equation}
 \begin{equation}\label{eq:forgetG}
-  \overrightarrow{f_t}=\sigma ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_f + \overrightarrow{b_f})
+  \overrightarrow{f_t}= G ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_f + \overrightarrow{b_f})
 \end{equation}
 \begin{equation}\label{eq:outputG}
-  \overrightarrow{o_t}=\sigma ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_o + \overrightarrow{b_o})
+  \overrightarrow{o_t}= G ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_o + \overrightarrow{b_o})
 \end{equation}
@@
 \begin{equation}\label{eq:updateG}
-  \overrightarrow{z_t}=\sigma ([\overrightarrow{x_t},\overrightarrow{h_{t-1}}] \cdot w_z + \overrightarrow{b_z})
+  \overrightarrow{z_t}= G ([\overrightarrow{x_t},\overrightarrow{h_{t-1}}] \cdot w_z + \overrightarrow{b_z})
 \end{equation}
 \begin{equation}\label{eq:resetG}
-  \overrightarrow{r_t}=\sigma ([\overrightarrow{x_t},\overrightarrow{h_{t-1}}] \cdot w_r + \overrightarrow{b_r})
+  \overrightarrow{r_t}= G ([\overrightarrow{x_t},\overrightarrow{h_{t-1}}] \cdot w_r + \overrightarrow{b_r})
 \end{equation}
@@
 
+A memristor is used for its ability to change its internal resistance based on the current that flowed through it.
+
 \subsection{Memristors Crossbar Arrays}
@@
 \begin{equation}
-  V_i = R\cdot I_{i} \Rightarrow I_{i} = V_i\cdot (\frac{1}{R_i})= V_i\cdot\sigma_i
+  V_i = R\cdot I_{i} \Rightarrow I_{i} = V_i\cdot (\frac{1}{R_i})= V_i\cdot G_i
 \end{equation}
-With $\sigma_i$ being the conductance of memristor, defined as $\sigma_i=\frac{1}{R_i}$.
+With $ G_i$ being the conductance of memristor, defined as $ G_i=\frac{1}{R_i}$.
 
@@
 \begin{equation}
-  I_1=\sigma_1\cdot V_1 + \sigma_2\cdot V_2 + \sigma_3\cdot V_3
+  I_1= G_1\cdot V_1 +  G_2\cdot V_2 +  G_3\cdot V_3
 \end{equation}
-With $\sigma_1$, $\sigma_2$ and $\sigma_3$ being the conductance of the 3 memristors in the first column.
+With $ G_1$, $ G_2$ and $ G_3$ being the conductance of the 3 memristors in the first column.

### 2023-11-29 appendices/circuits.tex
Reading to make sure nothing is missing
Changed a bunch of files

@@
-\fancychapter{Less important circuits}
+\chapter{Circuits}
 \label{ap:circuit}
+\clearpage
 
@@
 
-\section{Summing inverter}\label{apsec:sumInv}
+\section{Linear combination of inputs}\label{apsec:sumInv}
 
-The summing amplifier in a very general circuit that is used in a plethera of different situations. A modulable circuit with a variable number of input ($n$) in available in \cref{circt:sumInv}.
+The linear combination of inputs is a very general circuit that is used in a plethera of different situations.
+A modulable circuit with a variable number of input ($n+1$) in available in \cref{circt:sumInv}.
 
@@
 \begin{itemize}
-  \item As a voltage inverter around $V_{cm}$, using a single input ($n=1$) and using $R_0=R$. This is used in \cref{sec:af}.
+  \item As a voltage mirror around $V_{cm}$, using a single input ($n=1$) and using $R_0=R$. This is used in \cref{sec:af}.
   \item As a inverting amplifier after a voltage multiplier (\cref{subsec:voltmult}). Using one input ($n=1$) and $R>R_0$. It can be found in \cref{sec:lstmCircuit,sec:gruCircuit}.

### 2023-11-29 conclusions/further.tex
Wrong links
Changed a bunch of files

@@
 
-\subsubsection{Only change the value of one of the two memristor}
-
 This is a potential idea on how to improve the time it takes to compute a single epoch.
@@
 This solution only works in fully parallel mode.
-  \item For the cell state memory cells, adding a capacitor and a buffer (\cref{ap:buffer}) after the output of the memory cell group.
+  \item For the cell state memory cells, adding a capacitor and a buffer (\cref{apsec:buffer}) after the output of the memory cell group.
 This capacitor is here to store the voltage after the output enable flag goes to low.

### 2023-11-29 design/dataset.tex
Changed a bunch of files

@@
 
-This is a regression problem, the role of the \ac{LSTM} is to predict the number of passenger flying the next month being given previous months' passenger count.
+This is a forecasting problem, the role of the \ac{LSTM} is to predict the number of passenger flying the next month being given previous months' passenger count.
 
@@
 
-\subsubsection{Data format}
-
+The data in the dataset has to be formatted to be used with the \ac{LSTM}.
 The dataset contains $144$ sample points.
@@
 
-\subsection{\ac{C. elegans}}\label{subsec:celegans}
+\subsection{\acl{C. elegans}}\label{subsec:celegans}
 
@@
 
-\subsubsection{Dataset format}
-
-The dataset contains a set of fourty set of input/output.
+The dataset is already formatted in a time depedent manner and contains a set of fourty set of input/output.
 Each input/output set contains $500ms$ of data with a time step of $0.5ms$, making a thousand points for each neurons.

### 2023-11-29 results/airline.tex
corrections
Changed a bunch of files

@@
 
-\subsubsection{Timing}
-
-The timing of the flags are quite hard to keep track of, for this reason all of them are going to be shown for the full duration of the system execution to get one output.
+The timing of the flags can be quite hard to keep track of, for this reason all of them are going to be shown for the full duration of the system execution to get one output.
 
@@
 \Cref{tab:airlineAnalog} contains the errors of the analog predictions.
-It shows that, as can be observed, the results the closest to the target are the ones generated using a serial size of one ($n_s=1$).
+It can be observed that the results the closest to the target are the ones generated using a serial size of one ($n_s=1$).
 Indeed, its error ($28.4$) is lower than the error from the digital results ($52.3$) to the original target values, almost half of it.

### 2023-11-29 results/celegans.tex
Fixing not number of epochs
Changed a bunch of files

@@
 
-\subsubsection{Timing}
+The \ac{NN} was trained for $1000$ epochs.
 
-The timing of the flags are quite hard to keep track of, for this reason all of them are going to be shown for the full duration of the system execution to get one output.
+The timing of the flags can be quite hard to keep track of, for this reason all of them are going to be shown for the full duration of the system execution to get one output.
 

### 2023-11-29 state/lstm.tex
Worked on diapo + changed state of the art
Changed a bunch of files

@@
-\section{\acs{LSTM}}\label{sec:lstm}
+\section{\acl{LSTM}}\label{sec:lstm}
 \acfp{LSTM} are a type of \ac{RNN} used to analyze sequences of data. They are capable of predicting data based on previous data points.
@@
 \begin{equation}\label{eq:inputG}
-  \overrightarrow{i_t}=\sigma ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_i + \overrightarrow{b_i})
+  \overrightarrow{i_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_i + \overrightarrow{b_i})
 \end{equation}
 \begin{equation}\label{eq:forgetG}
-  \overrightarrow{f_t}=\sigma ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_f + \overrightarrow{b_f})
+  \overrightarrow{f_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_f + \overrightarrow{b_f})
 \end{equation}
 \begin{equation}\label{eq:outputG}
-  \overrightarrow{o_t}=\sigma ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_o + \overrightarrow{b_o})
+  \overrightarrow{o_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_o + \overrightarrow{b_o})
 \end{equation}
@@
 \begin{equation}\label{eq:candCell}
-  \overrightarrow{\tilde{c}_t}=tanh([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_c+ \overrightarrow{b_c})
+  \overrightarrow{\tilde{c}_t}=tanh([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_c+ \overrightarrow{b_c})
 \end{equation}
@@
 
-\subsubsection{Vanilla \ac{LSTM}}
+\subsubsection{Vanilla \acs{LSTM}}
 This variant of the \ac{LSTM} was originally the only \ac{LSTM} network, hence its name.
@@
 \begin{equation}\label{eq:inputGVanilla}
-  \overrightarrow{i_t}=\sigma ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_i + \overrightarrow{b_i} +\overrightarrow{c_{t-1}}\odot \overrightarrow{p_f})
+  \overrightarrow{i_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_i + \overrightarrow{b_i} +\overrightarrow{c_{t-1}}\odot \overrightarrow{p_f})
 \end{equation}
 \begin{equation}\label{eq:forgetGVanilla}
-  \overrightarrow{f_t}=\sigma ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_f + \overrightarrow{b_f}+\overrightarrow{c_{t-1}}\odot \overrightarrow{p_i})
+  \overrightarrow{f_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_f + \overrightarrow{b_f}+\overrightarrow{c_{t-1}}\odot \overrightarrow{p_i})
 \end{equation}
 \begin{equation}\label{eq:ouputGVanilla}
-  \overrightarrow{o_t}=\sigma ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_o + \overrightarrow{b_o}+\overrightarrow{c_{t}}\odot \overrightarrow{p_o})
+  \overrightarrow{o_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_o + \overrightarrow{b_o}+\overrightarrow{c_{t}}\odot \overrightarrow{p_o})
 \end{equation}
@@
 
-\subsubsection{\acf{FGR} \ac{LSTM}}
-This variant of the \ac{LSTM} is the most complex of all.
+\subsubsection{\acl{FGR} \ac{LSTM}}
+The \ac{FGR} \ac{LSTM} is one of the most complex variant of the \ac{LSTM}.
 This is due to the amount of feedback connections.
@@
 \begin{equation}\label{eq:inputGFGR}
-  \overrightarrow{i_t}=\sigma ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_i  + \overrightarrow{i_{t-1}}\cdot R_{ii} + \overrightarrow{f_{t-1}}\cdot R_{if} + \overrightarrow{o_{t-1}}\cdot R_{io} + \overrightarrow{b_i} +\overrightarrow{c_{t-1}}\odot \overrightarrow{p_i})
+  \overrightarrow{i_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_i  + \overrightarrow{i_{t-1}}\cdot R_{ii} + \overrightarrow{f_{t-1}}\cdot R_{if} + \overrightarrow{o_{t-1}}\cdot R_{io} + \overrightarrow{b_i} +\overrightarrow{c_{t-1}}\odot \overrightarrow{p_i})
 \end{equation}
 \begin{equation}\label{eq:forgetGFGR}
-  \overrightarrow{f_t}=\sigma ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_f  + \overrightarrow{i_{t-1}}\cdot R_{fi} + \overrightarrow{f_{t-1}}\cdot R_{ff} + \overrightarrow{o_{t-1}}\cdot R_{fo} + \overrightarrow{b_f} +\overrightarrow{c_{t-1}}\odot \overrightarrow{p_f})
+  \overrightarrow{f_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_f  + \overrightarrow{i_{t-1}}\cdot R_{fi} + \overrightarrow{f_{t-1}}\cdot R_{ff} + \overrightarrow{o_{t-1}}\cdot R_{fo} + \overrightarrow{b_f} +\overrightarrow{c_{t-1}}\odot \overrightarrow{p_f})
 \end{equation}
 \begin{equation}\label{eq:ouputGFGR}
-  \overrightarrow{o_t}=\sigma ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_o + \overrightarrow{i_{t-1}}\cdot R_{oi} + \overrightarrow{f_{t-1}}\cdot R_{of} + \overrightarrow{o_{t-1}}\cdot R_{oo} + \overrightarrow{b_o} +\overrightarrow{c_{t-1}}\odot \overrightarrow{p_o})
+  \overrightarrow{o_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_o + \overrightarrow{i_{t-1}}\cdot R_{oi} + \overrightarrow{f_{t-1}}\cdot R_{of} + \overrightarrow{o_{t-1}}\cdot R_{oo} + \overrightarrow{b_o} +\overrightarrow{c_{t-1}}\odot \overrightarrow{p_o})
 \end{equation}

### 2023-11-30 0.Intro/resumo.tex
Changed a bunch of files
Added almost all symbols

--- 2023-10-19 0.Intro/abstract.tex
@@
-\begin{abstract}
+\begin{resumo}
 
-  \acp{AI} are one of the main research domain in computer sciences and is used in other scientific domains.
-One great application of \ac{AI}, is in embedded systems.
-Lots of small electronic devices around us could benefit us.
-\textbf{Google} and \textbf{Apple} are pushing \ac{AI} in phones with their in house designed ARM processors having, very advertised, tensor cores.
-  Being able to run low powered \acp{AI} is thus one of the biggest computer objective of the deceny.
-  Analog computers are known to be a low powered technology that most importantly give almost instant results.
-An analog \ac{ASIC} could then be fabricated for specific embedded applications.
-This would definitely improve digital technologies in terms of time, and maybe in power consumption as well.
-Which are two prerequisite for embedded system use.
-  For this reason this work aims at creating a functioning software simulation of analog circuit capable of running \ac{AI}.
-The thesis will focus on the software simulation, the first step of the work in order to have a fully working chip able to run \ac{AI} algorithms.
+  Neste trabalho, propomos uma estrutura anal\'{o}gica para m\'{o}dulos recorrentes baseados em memristors para computa\c{c}\~{a}o neurom\'{o}rfica.
+O sistema \'{e} totalmente anal\'{o}gico e implementa um de circuito \acs{LSTM} funcional bem como um circuito GRU, ainda em desenvolvimento.
+Ambos os blocos cont\^{e}m memristors para serem usados como pesos em um circuito VMM anal\'{o}gico capaz.
+Esses blocos permitem a execu\c{c}\~{a}o de c\'{a}lculos necess\'{a}rios para Redes Neurais Recorrentes de dimens\~{a}o arbitr\'{a}ria, e de forma rapida e eficiente, em circuito integrado.
+Como parte dos blocos \ac{LSTM} e \ac{GRU}, foi projetado um circuito anal\'{o}gico que implementa fun\c{c}\~{o}es de ativa\c{c}\~{a}o.
+Este circuito espec\'{i}fico \'{e} capaz de reproduzir fun\c{c}\~{o}es semelhantes \`{a}s sigmoid e tanh, com formas e intervalos de valor similares.
+O trabalho tamb\'{e}m inclui a implementa\c{c}\~{a}o de uma c\'{e}lula de mem\'{o}ria usada para armazenar um valor anal\'{o}gico por um curto per\'{i}odo de tempo.
+Os c\'{a}lculos do bloco LSTM podem ser executado em s\'{e}rie ou paralelo, com um grau de serializa\c{c}\~{a}o arbitr\'{a}rio.
+A serializa\c{c}\~{a}o do sistema permite economizar \'{a}rea no chip, a custo do tempo de execu\c{c}\~{a}o.
+Adicionalmente, at\'{e} \`{a} data de publica\c{c}\~{a}o esta \'{e} a primeira implementa\c{c}\~{a}o anal\'{o}gica do comportamento do \ac{C.
+elegans} usando o bloco \ac{LSTM} em circuito integrado.
+Um sistema anal\'{o}gico deste tipo estabelece a base para a implementa\c{c}\~{a}o em tempo real de sistemas nervosos.
 
-\end{abstract}
+\end{resumo}

### 2023-11-30 design/crossbar.tex
Added almost all symbols

@@
 This design is the one that is used in \cite{doubleMem}.
-Let's assume that a given memristor has a resistance range of $R\in[R_{min},R_{max}]$, that means its conductance range is $ G \in [ G_{min}, G_{max}]$ (with $ G_{min}= \frac{1}{R_{max}}$ and $ G_{max}= \frac{1}{R_{min}}$).
-This design works using two \ac{opAmp} connected to $V_{cm}$ with the positive pin and the negative pin to the output of the crossbar array.
+Let's assume that a given memristor has a resistance range of $\symR\in[\symR_{min},\symR_{max}]$, that means its conductance range is $ \symG \in [ \symG_{min}, \symG_{max}]$ (with $ \symG_{min}= \frac{1}{\symR_{max}}$ and $ \symG_{max}= \frac{1}{\symR_{min}}$).
+This design works using two \ac{opAmp} connected to $\symv_{cm}$ with the positive pin and the negative pin to the output of the crossbar array.
 \Cref{eq:doubleMem0,eq:doubleMem1,eq:doubleMem2} are describing how this architecture works.
@@
 
-Using $x_k$ as the voltage for the input line $k$. The highest \ac{opAmp} is identified as $opAmp_0$ and the lowest $opAmp_1$.
+Using $\symv_k$ as the voltage for the input line $k$. The highest \ac{opAmp} is identified as $opAmp_0$ and the lowest $opAmp_1$.
 
-For the sake of simplicity, the following equations considers the ground to be $V_{cm}$.
+For the sake of simplicity, the following equations considers the ground to be $\symv_{cm}$.
 
@@
   \label{eq:doubleMem0}
-  V_{opAmp_0}=-R_r\cdot i_+
+  \symv_{opAmp_0}=-\symR_r\cdot \symi_+
 \end{equation}
@@
   \label{eq:doubleMem1}
-  i_{R_f}=i_-+\frac{V_{opAmp_0}}{R_r}=i_--i_+
+  \symi_{\symR_f}=\symi_-+\frac{\symv_{opAmp_0}}{\symR_r}=\symi_--\symi_+
 \end{equation}
@@
   \label{eq:doubleMem2}
-  V_{opAmp_1}=y_0=-R_f\cdot(i_--i_+)=R_f\cdot(i_+-i_-)=R_f\cdot\sum_{k=0}^n( G_{k+}- G_{k-})\cdot x_k
+  \symv_{opAmp_1}=\symv_{out}=-\symR_f\cdot(\symi_--\symi_+)=\symR_f\cdot(\symi_+-\symi_-)=\symR_f\cdot\sum_{k=0}^n( \symG_{k+}- \symG_{k-})\cdot \symv_k
 \end{equation}
-With $i_+=\sum_{k=0}^n G_{k+}\cdot x_k$ and $i_-=\sum_{k=0}^n G_{k-}\cdot x_k$.
+With $\symi_+=\sum_{k=0}^n \symG_{k+}\cdot \symv_k$ and $\symi_-=\sum_{k=0}^n \symG_{k-}\cdot \symv_k$.
 
@@
     \hline
-    Maximum weight & $ G_{max}- G_{min}$ & $ G_{max} -\overline{ G}$\\
+    Maximum weight & $ \symG_{max}- \symG_{min}$ & $ \symG_{max} -\overline{ \symG}$\\
     \hline
-    Minimum weight & $ G_{min}- G_{max}$ & $ G_{min} -\overline{ G}$\\
+    Minimum weight & $ \symG_{min}- \symG_{max}$ & $ \symG_{min} -\overline{ \symG}$\\
     \hline
-    Range & $2\cdot( G_{max}- G_{min})$&$ G_{max}- G_{min}$\\
+    Range & $2\cdot( \symG_{max}- \symG_{min})$&$ \symG_{max}- \symG_{min}$\\
     \hline
@@
 \begin{equation}
-  A_{xbar}(n_i,n_o,n_s)=
+  \symA_{xbar}(n_i,n_o,n_s)=
   \begin{cases}
-    2\cdot n_i\cdot n_o \cdot A_{memristor}+n_o\cdot(2\cdot[A_{opAmp}+A_{R_r}]+A_{R_f})& n_s=1\\
+    2\cdot n_i\cdot n_o \cdot \symA_{memristor}+n_o\cdot(2\cdot[\symA_{opAmp}+\symA_{R_r}]+\symA_{R_f})& n_s=1\\
     \begin{array}{2}
-      2\cdot n_i\cdot n_o \cdot n_s\cdot A_{memristor}+2\cdot A_{CMOS}\cdot n_o\cdot n_s\\
-      +n_o\cdot(2\cdot[A_{opAmp}+A_{R_r}]+A_{R_f})
+      2\cdot n_i\cdot n_o \cdot n_s\cdot \symA_{memristor}+2\cdot \symA_{CMOS}\cdot n_o\cdot n_s\\
+      +n_o\cdot(2\cdot[\symA_{opAmp}+\symA_{R_r}]+\symA_{R_f})
     \end{array}
@@
 
-With $A_{CMOS}=w\cdot l$.
+With $\symA_{CMOS}=w\cdot l$.
 
@@
 \begin{equation}\label{eq:memThres}
-  |V_{mem}|= |x_i-V_{cm}|\le V_{read}
+  |\symv_{mem}|= |\symv_i-\symv_{cm}|\le \symv_{read}
 \end{equation}
 
-Where $V_{mem}$ is the terminal voltage of a memristor.
+Where $\symv_{mem}$ is the terminal voltage of a memristor.
 

### 2023-11-30 design/genwei.tex
Added almost all symbols

@@
 
-The weights need to be contrained during the training to make sure that the voltage of the analog circuit stays within its operating voltage range ($[0,V_{dd}]$).
+The weights need to be contrained during the training to make sure that the voltage of the analog circuit stays within its operating voltage range ($[0,\symv_{dd}]$).
 To avoid any unwanted behavior when the system supposedly goes out of range, the weights in the \ac{LSTM} are contrained.
@@
 
-The worst case scenario for an \ac{LSTM} is when every value reaches the voltage threshold ($V_{read}$) (\cref{sec:xbarCircuit}) and every weight is maximized ($w_{max}$).
+The worst case scenario for an \ac{LSTM} is when every value reaches the voltage threshold ($\symv_{read}$) (\cref{sec:xbarCircuit}) and every weight is maximized ($\symw_{max}$).
 When that is the case the output of the \acp{VMM} is found in \cref{eq:weiCons}.
@@
 \begin{equation}\label{eq:weiCons}
-  V_{read}\cdot w_{max} \cdot(n_i+n_h)+w_{max}+V_{cm}= V_{dd}
+  \symv_{read}\cdot \symw_{max} \cdot(n_i+n_h)+\symw_{max}+\symv_{cm}= \symv_{dd}
 \end{equation}
 
-We can then determine the maximum acceptable value for the weights ($w_{max}$), this value has thus been determined to be the one in \cref{eq:weiConsRes}.
+We can then determine the maximum acceptable value for the weights ($\symw_{max}$), this value has thus been determined to be the one in \cref{eq:weiConsRes}.
 
 \begin{equation}\label{eq:weiConsRes}
-  w_{max}=\frac{V_{dd}-V_{cm}}{V_{read}\cdot(n_i+n_h)+1}
+  \symw_{max}=\frac{\symv_{dd}-\symv_{cm}}{\symv_{read}\cdot(n_i+n_h)+1}
 \end{equation}
 
-The analog system being completly centered and symetrical around $V_{cm}$ the value for the minimum weight is the opposite to the maximum one (\cref{eq:weiConsRes1})
+The analog system being completly centered and symetrical around $\symv_{cm}$ the value for the minimum weight is the opposite to the maximum one (\cref{eq:weiConsRes1})
 
 \begin{equation}\label{eq:weiConsRes1}
-  w_{min}=-w_{max}=-\frac{V_{dd}-V_{cm}}{V_{read}\cdot(n_i+n_h)+1}
+  \symw_{min}=-\symw_{max}=-\frac{\symv_{dd}-\symv_{cm}}{\symv_{read}\cdot(n_i+n_h)+1}
 \end{equation}

### 2023-11-30 design/gru.tex
Changed a bunch of files
Added almost all symbols

@@
   \item $e_{next}$ is the enable flag on in between two time steps.
-  \item $R_{amp0}$ and $R_{amp1}$ are the two resistances used to amplify the output voltage of the voltage multipliers.
-The ratio of the resistances value must of $\frac{R_{amp1}}{R_{amp0}}=10$.
+  \item $\symR_{amp0}$ and $\symR_{amp1}$ are the two resistances used to amplify the output voltage of the voltage multipliers.
+The ratio of the resistances value must of $\frac{\symR_{amp1}}{\symR_{amp0}}=10$.
 The resistances must stay arround the values of the resistances used around the circuit, especially those of the memristors.
-  \item $V_{inv}$ is the voltage used to perform the $1-x$ operation.
+  \item $\symv_{inv}$ is the voltage used to perform the $1-x$ operation.
 \end{itemize}
@@
 
-\subsubsection{The $1-x$ operation}
+\subsection{The $1-x$ operation}
 
@@
 \begin{equation}\label{eq:1minus1}
-  f_v(v)=-(v-V_{inv})+V_{inv}=2\cdot V_{inv} -v
+  f_\symv(\symv)=-(\symv-\symv_{inv})+\symv_{inv}=2\cdot \symv_{inv} -\symv
 \end{equation}
@@
 \begin{equation}\label{eq:1minus2}
-  f_v(v)=(1-V_{cm})-(v-V_{cm})+V_{cm}= 0.1-v+2\cdot V_{cm}
+  f_\symv(\symv)=(1-\symv_{cm})-(\symv-\symv_{cm})+\symv_{cm}= 0.1-\symv+2\cdot \symv_{cm}
 \end{equation}
 
-It can thus be found that the value for $V_{inv}$ is $V_{inv}=V_{cm}+0.05$
+It can thus be found that the value for $\symv_{inv}$ is $\symv_{inv}=\symv_{cm}+0.05$
 
@@
 \begin{equation}\label{eq:serProb}
-  \overrightarrow{\hat{h_t}}=tanh(\overrightarrow{x_t}\cdot w_{hx}+(\overrightarrow{r_t}\odot\overrightarrow{h_{t-1}}) \cdot w_{hh} + \overrightarrow{b_h})
+  \symvhh_t=tanh(\symvx_t\cdot \symmw_{hx}+(\symvr_t\odot\symvh_{t-1}) \cdot \symmw_{hh} + \symvb_h)
 \end{equation}
@@
   \begin{array}{3}
-    A_{gru}(n_i,n_h)=4\cdot A_{xbar}(n_i,n_h,1)\\+ \frac{n_h}{n_s}\cdot (5\cdot A_{af}+3\cdot A_{voltMult}+3\cdot A_{R_{amp0}}+2\cdot A_{R_{amp1}}+2\cdot A_{opAmp})\\+5\cdot n_h\cdot A_{memcell}
+    \symA_{gru}(n_i,n_h)=4\cdot \symA_{xbar}(n_i,n_h,1)\\+ \frac{n_h}{n_s}\cdot (5\cdot \symA_{af}+3\cdot \symA_{voltMult}+3\cdot \symA_{\symR_{amp0}}+2\cdot \symA_{\symR_{amp1}}+2\cdot \symA_{opAmp})\\+5\cdot n_h\cdot \symA_{memcell}
   \end{array}

### 2023-11-30 design/lstm.tex
Worked on paper+corrected errors
Added almost all symbols

@@
   \item $e_{next}$ is the enable flag on in between two time steps.
-  \item $R_{amp0}$ and $R_{amp1}$ are the two resistances used to amplify the output voltage of the voltage multipliers.
-The ratio of the resistances value must of $\frac{R_{amp1}}{R_{amp0}}=10$.
+  \item $\symR_{amp0}$ and $\symR_{amp1}$ are the two resistances used to amplify the output voltage of the voltage multipliers.
+The ratio of the resistances value must of $\frac{\symR_{amp1}}{\symR_{amp0}}=10$.
 The resistances must stay arround the values of the resistances used around the circuit, especially those of the memristors.
@@
   \begin{array}{3}
-    A_{lstm}(n_i,n_h,n_s)=4\cdot A_{xbar}(n_i,\frac{n_h}{n_s},n_s)\\+ \frac{n_h}{n_s}\cdot (5\cdot A_{af}+3\cdot A_{voltMult}+3\cdot A_{R_{amp0}}+2\cdot A_{R_{amp1}}+2\cdot A_{opAmp})\\+5\cdot n_h\cdot A_{memcell}
+    \symA_{lstm}(n_i,n_h,n_s)=4\cdot \symA_{xbar}(n_i,\frac{n_h}{n_s},n_s)\\+ \frac{n_h}{n_s}\cdot (5\cdot \symA_{af}+3\cdot \symA_{voltMult}+3\cdot \symA_{\symR_{amp0}}+2\cdot \symA_{\symR_{amp1}}+2\cdot \symA_{opAmp})\\+5\cdot n_h\cdot \symA_{memcell}
   \end{array}

### 2023-11-30 design/main.tex
Correcting inappropriate raccourcis and correcting design
Added almost all symbols

@@
 
-The system will be working with a $V_{dd}$ of $1.8V$. Such a value was chosen because this is a low power system.
+The system will be working with a $\symv_{dd}$ of $1.8V$. Such a value was chosen because this is a low power system.
 
 The way values are encoded in the analog system will be descibed here as it serves for the entire thesis.
-In order for the system to support negative numbers we're going to use a $V_{cm}$ set to $\frac{V_{dd}}{2}$.
-That means that $V_{cm}=\frac{V_{dd}}{2}=0.9V$.
-This $V_{cm}$ will then describe a zero.
+In order for the system to support negative numbers we're going to use a $\symv_{cm}$ set to $\frac{\symv_{dd}}{2}$.
+That means that $\symv_{cm}=\frac{\symv_{dd}}{2}=0.9V$.
+This $\symv_{cm}$ will then describe a zero.
 A step of one was chosen to be $0.1V$ in the analog circuit.
-\Cref{tab:valConv} shows the conversions from a real number to it's voltage equivalent.
+\Cref{tab:valConv} shows the conversions from a real number to its voltage equivalent.
 
@@
     \hline
-    $x$ & $V(x)=\frac{x}{10}+V_{cm}$\\
+    $x$ & $\symv(x)=\frac{x}{10}+\symv_{cm}$\\
     \hline
-    $real(v)=(v-0.9)\cdot 10$ & $v$\\
+    $real(\symv)=(\symv-\symv_{cm})\cdot 10$ & \symv\\
     \hline
@@
 
-Since the system cannot reach voltage outside of the operating range with the intended behavior, the voltage is then restricted to $V\in [0,1.8]$.
+Since the system cannot reach voltage outside of the operating range with the intended behavior, the voltage is then restricted to $\symv\in [0,1.8]$.
 This means that the range of real value that the systems can handle is $x\in [-9,9]$.

### 2023-11-30 design/memcell.tex
Added almost all symbols

@@
 \begin{equation}
-  A_{memcell}=8\cdot A_{CMOS}+2\cdot A_{inv}+A_{opAmp}+A_{cap}
+  \symA_{memcell}=8\cdot \symA_{CMOS}+2\cdot \symA_{inv}+\symA_{opAmp}+\symA_{cap}
 \end{equation}
 
-With $A_{CMOS} = w\cdot l$
+With $\symA_{CMOS} = w\cdot l$
 
@@
 
-This circuit is pretty straight forward, when the input enable pin is high (up to $V_{dd}$), then the first switch is open and the capacitor is being charged, if it is low, then the switch is closed and the capacitor keeps the voltage and thus the analog data.
+This circuit is pretty straight forward, when the input enable pin is high (up to $\symv_{dd}$), then the first switch is open and the capacitor is being charged, if it is low, then the switch is closed and the capacitor keeps the voltage and thus the analog data.
 When the output enable is high, the buffer (\cref{apsec:buffer}) forwards the voltage stored in the capactor to the output pin without emptying the capacitor.

### 2023-11-30 design/models.tex
Added almost all symbols

@@
   \label{eq:opAmp}
-  V_{out}=\mu \cdot (V_+-V_-)
+  \symv_{out}=\mu \cdot (\symv_+-\symv_-)
 \end{equation}
@@
 Indeed, because of the voltage conversion (\cref{tab:valConv}), the equation needs to take the conversion into consideration.
-The final equation needs to be \cref{eq:finalVoltMult} in order to have $1\cdot 1=1$ ($V_{in_1}\cdot V_{in_2}=1V$, with $V_{in_1}=V_{in_2}=1V$)
+The final equation needs to be \cref{eq:finalVoltMult} in order to have $1\cdot 1=1$ ($\symv_{in_1}\cdot \symv_{in_2}=1V$, with $\symv_{in_1}=\symv_{in_2}=1V$)
 
 \begin{equation}\label{eq:finalVoltMult}
-  V_{out}=10\cdot(V_{in_1}-V_{cm})\cdot (V_{in_2}-V_{cm}) + V_{cm}
+  \symv_{out}=10\cdot(\symv_{in_1}-\symv_{cm})\cdot (\symv_{in_2}-\symv_{cm}) + \symv_{cm}
 \end{equation}
@@
 \begin{equation}\label{eq:voltMult}
-  V_{voltMult}=-(V_{in_1}-V_{cm})\cdot (V_{in_2}-V_{cm}) + V_{cm}
+  \symv_{voltMult}=-(\symv_{in_1}-\symv_{cm})\cdot (\symv_{in_2}-\symv_{cm}) + \symv_{cm}
 \end{equation}
 \begin{equation}\label{eq:invAmp}
-  V_{out}=-(V_{voltMult}-V_{cm})\cdot10+V_{cm}
+  \symv_{out}=-(\symv_{voltMult}-\symv_{cm})\cdot10+\symv_{cm}
 \end{equation}
 
-Where $V_{out}$ is the output voltage the inverting amplifier (\cref{apsec:sumInv}), $V_{voltMult}$ is the out voltage of the voltage multiplier itself and $V_{in_1}$ and $V_{in_2}$ are the input voltages.
+Where $\symv_{out}$ is the output voltage the inverting amplifier (\cref{apsec:sumInv}), $\symv_{voltMult}$ is the out voltage of the voltage multiplier itself and $\symv_{in_1}$ and $\symv_{in_2}$ are the input voltages.
 

### 2023-11-30 design/run.tex
Added almost all symbols

--- 2023-10-28 design/run.tex
@@
 The time step ($T$) value was chosen to be $T=8\mu s$, the pause between two \ac{LSTM} steps is set to be $\frac{T}{8}=1\mu s$.
-The value for the maximum and minimum resistances value of the memristor used ($R_{max}$ and $R_{min}$, respectively) are set to be in acordance with the \textbf{KNOWN}'s memristors.
-The resistance values used for the simulation are thus $R_{max}=10^6 \Omega=1M\Omega$ and $R_{min}=10^4 \Omega = 10k\Omega$.
+The value for the maximum and minimum resistances value of the memristor used ($\symR_{max}$ and $\symR_{min}$, respectively) are set to be in acordance with the \textbf{KNOWM}\textregistered{}'s memristors \cite{Knowm}.
+The resistance values used for the simulation are thus $\symR_{max}=10^6 \Omega=1M\Omega$ and $\symR_{min}=10^4 \Omega = 10k\Omega$.
 

### 2023-11-30 design/wei2res.tex
Worked on appendix
corrected a few things
Added almost all symbols

@@
   \label{eq:wei2res0}
-  w=R_f\cdot(\sigma_+-\sigma_-)=\frac{R_f}{R_+}-\frac{R_f}{R_-}
+  \symw=\symR_f\cdot( \symG_+- \symG_-)=\frac{\symR_f}{\symR_+}-\frac{\symR_f}{\symR_-}
 \end{equation}
@@
 \begin{itemize}
-  \item $w_{max}=R_f\cdot(\sigma_{max}-\sigma_{min})$
-  \item $w_{min}=-w_{max}=R_f\cdot(\sigma_{min}-\sigma_{max})$
+  \item $\symw_{max}=\symR_f\cdot( \symG_{max}- \symG_{min})$
+  \item $\symw_{min}=-\symw_{max}=\symR_f\cdot( \symG_{min}- \symG_{max})$
 \end{itemize}
 
-For the rest of this chapter we consider that $R_f$ is set to the middle point of $R_{max}$ and $R_{min}$ meaning that $R_f=\frac{R_{min}+R_{max}}{2}$.
+For the rest of this chapter we consider that $\symR_f$ is set to the middle point of $\symR_{max}$ and $\symR_{min}$ meaning that $\symR_f=\frac{\symR_{min}+\symR_{max}}{2}$.
 
-Since we only have one equation (\cref{eq:wei2res0}) for 2 unknowns ($R_+$ and $R_-$), we need to set a second equation.
-This is done by centering the resistances around $R_f$, this means that \cref{eq:wei2res1} is the second equation, that makes our problem now solvable.
+Since we only have one equation (\cref{eq:wei2res0}) for 2 unknowns ($\symR_+$ and $\symR_-$), we need to set a second equation.
+This is done by centering the resistances around $\symR_f$, this means that \cref{eq:wei2res1} is the second equation, that makes our problem now solvable.
 
@@
   \label{eq:wei2res1}
-  R_f=\frac{R_-+R_+}{2}
+  \symR_f=\frac{\symR_-+\symR_+}{2}
 \end{equation}
@@
   \begin{cases}
-    w=\frac{R_f}{R_+}-\frac{R_f}{R_-}\\
-    R_f=\frac{R_-+R_+}{2}
+    \symw=\frac{\symR_f}{\symR_+}-\frac{\symR_f}{\symR_-}\\
+    \symR_f=\frac{\symR_-+\symR_+}{2}
   \end{cases}
@@
 
-By solving \cref{eq:wei2res01}, we find the \cref{eq:wei2res2} that gives the values for $R_-$ and $R_+$. All the steps for solving the equations can be found in \cref{apsec:wei2res}.
+By solving \cref{eq:wei2res01}, we find the \cref{eq:wei2res2} that gives the values for $\symR_-$ and $\symR_+$. All the steps for solving the equations can be found in \cref{apsec:wei2res}.
 
 The real value to voltage conversion (\cref{tab:valConv}) does not affect this part of the system as the crossbar array already works in voltages.
-Replacing the resistor variable in \cref{eq:doubleMem2} by $w$ using \cref{eq:wei2res0} would not change the results.
+Replacing the resistor variable in \cref{eq:doubleMem2} by \symw using \cref{eq:wei2res0} would not change the results.
 
@@
   \begin{cases}
-    R_+= (w+1-\sqrt{w^2+1})\cdot\frac{R_f}{w}\\
-    R_-=2\cdot R_f -R_+
+    \symR_+= (\symw+1-\sqrt{\symw^2+1})\cdot\frac{\symR_f}{\symw}\\
+    \symR_-=2\cdot \symR_f -\symR_+
   \end{cases}
@@
 
-While solving the system (\cref{apsec:wei2res}), we get two potential equations for $R_+$.
-Graphing them shows that one of them is outside of the memristor's resistance range ($[R_{min},R_{max}]$).
+While solving the system (\cref{apsec:wei2res}), we get two potential equations for $\symR_+$.
+Graphing them shows that one of them is outside of the memristor's resistance range ($[\symR_{min},\symR_{max}]$).
 Leaving us with only one equation because the other one is physically unreachable.
@@
 The resolution of the memristor (the precision of the resistances at which the memristor can be set) is simulated by limiting the generated resistances values to having only two significant figures.
+
+The python implementation is available in \cref{apsec:wei2res}.

### 2023-11-30 state/crossbar.tex
Added almost all symbols

@@
   %\input{crossbar/node.pdf_tex}
-  \caption{Memristor crossbar node of the $i^{th}$ line and $j^{th}$ column}
+  \caption{Memristor crossbar node of the $k^{th}$ line and $j^{th}$ column}
   \label{fig:crossNode}
@@
 
-First of all, a voltage is applied on the $i^{th}$ line, because every column is virtually grounded, the voltage applied to the memristor, of a set resistance $R_i$, is $V_i$.
-As such and using Ohm's law, we know that the current flowing into the memristor ($I_{i}$) is bound by the following equation :
+First of all, a voltage is applied on the $k^{th}$ line, because every column is virtually grounded, the voltage applied to the memristor, of a set resistance $\symR_k$, is $\symv_k$.
+As such and using Ohm's law, we know that the current flowing into the memristor ($\symi_{k}$) is bound by the following equation :
 
 \begin{equation}
-  V_i = R\cdot I_{i} \Rightarrow I_{i} = V_i\cdot (\frac{1}{R})= V_i\cdot G_i
+  \symv_k = \symR_k\cdot \symi_{k} \Rightarrow \symi_{k} = \symv_k\cdot (\frac{1}{\symR_k})= \symv_k\cdot \symG_k
 \end{equation}
-With $G_i$ being the conductance of memristor, defined as $ G_i=\frac{1}{R_i}$.
+With $\symG_k$ being the conductance of memristor, defined as $\symG_k=\frac{1}{\symR_k}$.
 
-This line then joins the column where a current of $I_{i,j-1}$ is flowing, then according to Kirchhoff's current law the resulting current is :
+This line then joins the column where a current of $\symi_{j,k-1}$ is flowing, then according to Kirchhoff's current law the resulting current is :
 \begin{equation}
-  I_{j,i} = I_{j,i-1}+I_{i} = I_{j,i-1} + V_i\cdot G_i
+  \symi_{j,k} = \symi_{j,k-1}+\symi_{k} = \symi_{j,k-1} + \symv_k\cdot \symG_k
 \end{equation}
@@
 \begin{equation}
-  I_1= G_1\cdot V_1 +  G_2\cdot V_2 +  G_3\cdot V_3
+  \symi_1= \symG_1\cdot \symv_1 +  \symG_2\cdot \symv_2 +  \symG_3\cdot \symv_3
 \end{equation}
-With $G_1$, $G_2$ and $G_3$ being the conductance of the 3 memristors in the first column.
+With $\symG_1$, $\symG_2$ and $\symG_3$ being the conductance of the 3 memristors in the first column.
 

### 2023-11-30 state/gru.tex
Correctiion diogo state
Changed a bunch of files
Added almost all symbols

@@
-\section{\acs{GRU}}\label{sec:gru}
+\section{\acl{GRU}}\label{sec:gru}
 The \acf{GRU} is another type of \ac{RNN}. It is also known to reduce the effect of the vanishing gradient problem. It was first introduced to improve translation techniques \cite{gru}.
@@
 \begin{equation}\label{eq:updateG}
-  \overrightarrow{z_t}=\sigma ([\overrightarrow{x_t},\overrightarrow{h_{t-1}}] \cdot w_z + \overrightarrow{b_z})
+  \symvz_t=\sigma ([\symvx_t,\symvh_{t-1}] \cdot \symmw_z + \symvb_z)
 \end{equation}
 \begin{equation}\label{eq:resetG}
-  \overrightarrow{r_t}=\sigma ([\overrightarrow{x_t},\overrightarrow{h_{t-1}}] \cdot w_r + \overrightarrow{b_r})
+  \symvr_t=\sigma ([\symvx_t,\symvh_{t-1}] \cdot \symmw_r + \symvb_r)
 \end{equation}
 \begin{equation}\label{eq:candActivG}
-  \overrightarrow{\hat{h_t}}=tanh(\overrightarrow{x_t}\cdot w_{hx}+(\overrightarrow{r_t}\odot\overrightarrow{h_{t-1}}) \cdot w_{hh} + \overrightarrow{b_h})
+  \symvhh_t=tanh(\symvx_t\cdot \symmw_{hx}+(\symvr_t\odot\symvh_{t-1}) \cdot \symvw_{hh} + \symvb_h)
 \end{equation}
 \begin{equation}\label{eq:gruHidG}
-  \overrightarrow{h_t}=(\overrightarrow{1}-\overrightarrow{z_t})\odot \overrightarrow{h_{t-1}} + \overrightarrow{z_t}\odot \overrightarrow{\hat{h_t}}
+  \symvh_t=(\vunit-\symvz_t)\odot \symvh_{t-1} + \symvz_t\odot \symvhh_t
 \end{equation}
 
-Where ($w_z$,$\overrightarrow{b_z}$), ($w_r$,$\overrightarrow{b_r}$),($w_{hx}$,$w_{hh}$,$\overrightarrow{b_h}$) are the weights matrixes and bias vectors for the update, reset and candidate activation gates respectively.
+Where ($\symmw_z$,$\symvb_z$), ($\symmw_r$,$\symvb_r$),($\symmw_{hx}$,$\symmw_{hh}$,$\symvb_h$) are the weights matrixes and bias vectors for the update, reset and candidate activation gates respectively.
 
@@
   \includesvg[width=\textwidth,pretex=\large]{gru/encoderCell.svg}
+  \caption{Encoder \acs{GRU} cell, legend in \cref{leg:cells}}
   \label{fig:encoderGruCell}
-  \caption{Encoder \acs{GRU} cell, legend in \cref{leg:cells}}
 \end{figure}
@@
 \begin{equation}\label{eq:gruHidG1}
-  \overrightarrow{h_t}=\overrightarrow{z_t}\odot \overrightarrow{h_{t-1}} + (\overrightarrow{1}-\overrightarrow{z_t})\odot \overrightarrow{\hat{h_t}}
+  \symvh_t=\symvz_t\odot \symvh_{t-1} + (\vunit-\symvz_t)\odot \symvh_t
 \end{equation}
@@
 \begin{equation}\label{eq:candActivG1}
-  \overrightarrow{\hat{h_t}}=tanh(\overrightarrow{x_t}\cdot w_{hx}+ \overrightarrow{b_{hx}}+\overrightarrow{r_t}\odot[\overrightarrow{h_{t-1}} \cdot w_{hh} + \overrightarrow{b_{hh}}])
+  \symvhh_t=tanh(\symvx_t\cdot \symmw_{hx}+ \symvb_{hx}+\symvr_t\odot[\symvh_{t-1} \cdot \symmw_{hh} + \symvb_{hh}])
 \end{equation}
@@
   \includesvg[width=\textwidth,pretex=\large]{gru/decoderCell.svg}
+  \caption{Decoder \acs{GRU} cell, legend in \cref{leg:cells}}
   \label{fig:decoderGruCell}
-  \caption{Decoder \acs{GRU} cell, legend in \cref{leg:cells}}
 \end{figure}
@@
 \begin{equation}\label{eq:lstmGru0}
-  \overrightarrow{f_t}=\sigma ([\overrightarrow{x_t},\overrightarrow{h_{t-1}}] \cdot w_f + \overrightarrow{b_f})
+  \symvf_t=\sigma ([\symvx_t,\symvh_{t-1}] \cdot \symmw_f + \symvb_f)
 \end{equation}
 \begin{equation}\label{eq:lstmGru1}
-  \overrightarrow{o_t}=\sigma ([\overrightarrow{x_t},\overrightarrow{h_{t-1}}] \cdot w_o + \overrightarrow{b_o})
+  \symvo_t=\sigma ([\symvx_t,\symvh_{t-1}] \cdot \symmw_o + \symvb_o)
 \end{equation}
 \begin{equation}\label{eq:lstmGru2}
-  \overrightarrow{i_t}=\overrightarrow{1}-\overrightarrow{f_t}
+  \symvi_t=\vunit-\symvf_t
 \end{equation}
 \begin{equation}\label{eq:lstmGru3}
-  \overrightarrow{\tilde{c_t}}=tanh(\overrightarrow{x_t}\cdot w_{cx}+(\overrightarrow{o_t}\odot\overrightarrow{h_{t-1}}) \cdot w_{ch} + \overrightarrow{b_c})
+  \symvct_t=tanh(\symvx_t\cdot \symmw_{cx}+(\symvo_t\odot\symvh_{t-1}) \cdot \symmw_{ch} + \symvb_c)
 \end{equation}
 \begin{equation}\label{eq:lstmGru4}
-  \overrightarrow{h_t}=(\overrightarrow{i_t})\odot \overrightarrow{h_{t-1}} + \overrightarrow{f_t}\odot \overrightarrow{\tilde{h_t}}
+  \symvh_t=(\symvi_t)\odot \symvh_{t-1} + \symvf_t\odot \symvct_t
 \end{equation}

### 2023-11-30 state/lstm.tex
Added almost all symbols

@@
 \begin{equation}\label{eq:inputG}
-  \overrightarrow{i_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_i + \overrightarrow{b_i})
+  \symvi_t=\sigma ([\symvx_{t},\symvh_{t-1}]\cdot \symmw_i + \symvb_i)
 \end{equation}
 \begin{equation}\label{eq:forgetG}
-  \overrightarrow{f_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_f + \overrightarrow{b_f})
+  \symvf_t=\sigma ([\symvx_{t},\symvh_{t-1}]\cdot \symmw_f + \symvb_f)
 \end{equation}
 \begin{equation}\label{eq:outputG}
-  \overrightarrow{o_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_o + \overrightarrow{b_o})
+  \symvo_t=\sigma ([\symvx_{t},\symvh_{t-1}]\cdot \symmw_o + \symvb_o)
 \end{equation}
 
-Where ($w_i$,$\overrightarrow{b_i}$), ($w_f$,$\overrightarrow{b_f}$) and ($w_o$,$\overrightarrow{b_o}$) are the pair of weights matrixes and bias vectors for the input, forget and output gates respectively.
-$\overrightarrow{x_t}$ is the input vector and $\overrightarrow{h_t}$ is the hidden state vector.
+Where ($\symmw_i$,$\symvb_i$), ($\symmw_f$,$\symvb_f$) and ($\symmw_o$,$\symvb_o$) are the pair of weights matrixes and bias vectors for the input, forget and output gates respectively.
+$\symvx_t$ is the input vector and $\symvh_t$ is the hidden state vector.
 
@@
 \begin{equation}\label{eq:candCell}
-  \overrightarrow{\tilde{c}_t}=tanh([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_c+ \overrightarrow{b_c})
+  \symvct_t=tanh([\symvx_{t},\symvh_{t-1}]\cdot \symmw_c+ \symvb_c)
 \end{equation}
 \begin{equation}\label{eq:cellS}
-  \overrightarrow{c_t}=\overrightarrow{f_t}\odot \overrightarrow{c_{t-1}} + \overrightarrow{i_t} \odot \overrightarrow{\tilde{c}_t}
+  \symvc_t=\symvf_t\odot \symvc_{t-1} + \symvi_t \odot \symvct_t
 \end{equation}
 
-Where $w_c$ and $\overrightarrow{b_c}$ are the weights matrix and bias vector for the candidate cell state.
+Where $\symmw_c$ and $\symvb_c$ are the weights matrix and bias vector for the candidate cell state.
 
@@
 \begin{equation}\label{eq:hiddenS}
-  \overrightarrow{h_t}=\overrightarrow{o_t}\odot tanh(\overrightarrow{c_t})
+  \symvh_t=\symvo_t\odot tanh(\symvc_t)
 \end{equation}
 
-We set $x_1$ as the first input and define $\overrightarrow{h_0}$ as a zero only vector.
+We set $x_1$ as the first input and define $\symvh_0$ as a zero only vector.
 
@@
 \begin{equation}\label{eq:inputGVanilla}
-  \overrightarrow{i_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_i + \overrightarrow{b_i} +\overrightarrow{c_{t-1}}\odot \overrightarrow{p_f})
+  \symvi_t=\sigma ([\symvx_{t},\symvh_{t-1}]\cdot \symmw_i + \symvb_i +\symvc_{t-1}\odot \symvp_f)
 \end{equation}
 \begin{equation}\label{eq:forgetGVanilla}
-  \overrightarrow{f_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_f + \overrightarrow{b_f}+\overrightarrow{c_{t-1}}\odot \overrightarrow{p_i})
+  \symvf_t=\sigma ([\symvx_{t},\symvh_{t-1}]\cdot \symmw_f + \symvb_f+\symvc_{t-1}\odot \symvp_i)
 \end{equation}
 \begin{equation}\label{eq:ouputGVanilla}
-  \overrightarrow{o_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_o + \overrightarrow{b_o}+\overrightarrow{c_{t}}\odot \overrightarrow{p_o})
+  \symvo_t=\sigma ([\symvx_{t},\symvh_{t-1}]\cdot \symmw_o + \symvb_o+\symvc_{t}\odot \symvp_o)
 \end{equation}
 
-With $\overrightarrow{p_f}$, $\overrightarrow{p_i}$ and $\overrightarrow{p_o}$ being the peephole weights vectors.
-Their size is the same as the size of the hidden state vector ($\overrightarrow{h_t}$).
-$\odot$ is the pointwise multiplication operator.
-
+With $\symvp_f$, $\symvp_i$ and $\symvp_o$ being the peephole weights vectors. Their size is the same as the size of the hidden state vector ($\symvh_t$).
 The equations that have not been rewritten simply stay the same.
@@
 \begin{equation}\label{eq:inputGFGR}
-  \overrightarrow{i_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_i  + \overrightarrow{i_{t-1}}\cdot R_{ii} + \overrightarrow{f_{t-1}}\cdot R_{if} + \overrightarrow{o_{t-1}}\cdot R_{io} + \overrightarrow{b_i} +\overrightarrow{c_{t-1}}\odot \overrightarrow{p_i})
+  \symvi_t=\sigma ([\symvx_{t},\symvh_{t-1}]\cdot \symmw_i  + \symvi_{t-1}\cdot \symmw_{ii} + \symvf_{t-1}\cdot \symmw_{if} + \symvo_{t-1}\cdot \symmw_{io} + \symvb_i +\symvc_{t-1}\odot \symvp_i)
 \end{equation}
 \begin{equation}\label{eq:forgetGFGR}
-  \overrightarrow{f_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_f  + \overrightarrow{i_{t-1}}\cdot R_{fi} + \overrightarrow{f_{t-1}}\cdot R_{ff} + \overrightarrow{o_{t-1}}\cdot R_{fo} + \overrightarrow{b_f} +\overrightarrow{c_{t-1}}\odot \overrightarrow{p_f})
+  \symvf_t=\sigma ([\symvx_{t},\symvh_{t-1}]\cdot \symmw_f  + \symvi_{t-1}\cdot \symmw_{fi} + \symvf_{t-1}\cdot \symmw_{ff} + \symvo_{t-1}\cdot \symmw_{fo} + \symvb_f +\symvc_{t-1}\odot \symvp_f)
 \end{equation}
 \begin{equation}\label{eq:ouputGFGR}
-  \overrightarrow{o_t}=\sigma ([\overrightarrow{x_{t}},\overrightarrow{h_{t-1}}]\cdot w_o + \overrightarrow{i_{t-1}}\cdot R_{oi} + \overrightarrow{f_{t-1}}\cdot R_{of} + \overrightarrow{o_{t-1}}\cdot R_{oo} + \overrightarrow{b_o} +\overrightarrow{c_{t-1}}\odot \overrightarrow{p_o})
+  \symvo_t=\sigma ([i\symvx_{t},\symvh_{t-1}]\cdot \symmw_o + \symvi_{t-1}\cdot \symmw_{oi} + \symvf_{t-1}\cdot \symmw_{of} + \symvo_{t-1}\cdot \symmw_{oo} + \symvb_o +\symvc_{t-1}\odot \symvp_o)
 \end{equation}
 
-With $R_{fi}$ being the weight matrix of the input feedback connection to compute the forget gate, and the same goes for the other weight matrixes.
+With $\symmw_{fi}$ being the weight matrix of the input feedback connection to compute the forget gate, and the same goes for the other weight matrixes.
 
@@
     \begin{itemize}
-      \item the candidate cell and thus becomes $ \overrightarrow{\tilde{c}_t}=(w_c\cdot[\overrightarrow{x_{t-1}},\overrightarrow{h_{t-1}}] + \overrightarrow{b_c}) $, this is called the \ac{NIAF}
-      \item the output of the previous cell state and thus becomes $ \overrightarrow{h_t}=\overrightarrow{o_t}\odot \overrightarrow{c_t} $, and is called \ac{NOAF}
+      \item the candidate cell and thus becomes $ \symvct_t=([\symvx_{t-1},\symvh_{t-1}]\cdot\symmw_c + \symvb_c) $, this is called the \ac{NIAF}
+      \item the output of the previous cell state and thus becomes $ \symvh_t=\symvo_t\odot \symvc_t $, and is called \ac{NOAF}
     \end{itemize}
@@
     \begin{itemize}
-      \item \ac{NIG} : $\overrightarrow{i_t}=1$
-      \item \ac{NOG} : $\overrightarrow{o_t}=1$
-      \item \ac{NFG} : $\overrightarrow{f_t}=1$
+      \item \ac{NIG} : $\symvi_t=1$
+      \item \ac{NOG} : $\symvo_t=1$
+      \item \ac{NFG} : $\symvf_t=1$
     \end{itemize}

### 2023-11-30 state/memristors.tex
Correctiion diogo state
Added almost all symbols

@@
 It was first theorized in 1971 by L. Chua from UC Berkeley, in \cite{TheoMemristor}. The name comes from the blend of \textit{memory} and \textit{resistance}.
-The theory behind this component was extracted from a missing component to link the four fundamental circuit variables, voltage ($v$), charge ($q$), current ($i$) and flux ($\phi$).
+The theory behind this component was extracted from a missing component to link the four fundamental circuit variables, voltage (\symv), charge (\symq), current (\symq) and flux (\symphi).
 \Cref{fig:fundComp} shows the four fundamental variables are on each side of the square, with the ones on opposite sides being linked by the following equations :
@@
 \begin{equation}
-  d\phi = v\cdot dt
+  d\symphi = \symv\cdot d\symt
 \end{equation}
@@
 \begin{equation}
-  dq = i\cdot dt
+  d\symq = \symi\cdot d\symt
 \end{equation}
 
-Resistors, capacitors and inductors were already very established and well known components, so it was theorized that a fourth device should then exist to physically link flux ($\phi$) and charge ($q$).
-The flux in this case is not a magnetic flux and is defined as such : $ d\phi=v\cdot dt \Rightarrow \phi =  \int v \,dt  $.
+Resistors, capacitors and inductors were already very established and well known components, so it was theorized that a fourth device should then exist to physically link flux (\symphi) and charge (\symq).
+The flux in this case is not a magnetic flux and is defined as such : $d\symphi=\symv\cdot d\symt \Rightarrow \symphi =  \int \symv \,d\symt$.
 
@@
   \includesvg[width=0.55\textwidth]{memristor/memristor}
-  \caption{Fundamental passive components}
+  \caption{Fundamental passive components, adapted from \cite{memWiki}}
   \label{fig:fundComp}
@@
 \subsection{Equations}
-A memristor links the flux ($\phi$) and charge ($q$) by the memristance.
+A memristor links the flux (\symphi) and charge (\symq) by the memristance.
 This memristance is defined with the following equation :
 \begin{equation}
-  M(q)=\frac{d\phi}{dq}
+  \symM(\symq)=\frac{d\symphi}{d\symq}
 \end{equation}
-It can be compared with the other fundamental components like resistor ($R(i)=\frac{dv}{di}$), capacitor ($\frac{1}{C(q)}=\frac{dv}{dq}$) and inductor ($L(i)=\frac{d\phi}{di}$).
+It can be compared with the other fundamental components like resistor ($\symR(\symi)=\frac{d\symv}{d\symi}$), capacitor ($\frac{1}{\symC(\symq)}=\frac{d\symv}{d\symq}$) and inductor ($\symL(\symi)=\frac{d\symphi}{d\symi}$).
 We can then extract a more useful equation in an actual circuit :
 \begin{equation}
-  v(t)=M(q(t))\cdot i(t)
+  \symv(\symt)=\symM(\symq(\symt))\cdot \symi(\symt)
 \end{equation}
@@
 \begin{equation}
-  W(\phi)=\frac{dq}{d\phi}
+  \symW(\symphi)=\frac{d\symq}{d\symphi}
 \end{equation}
@@
 \begin{equation}\label{eq:memristiveDev}
-  v(t)=M(x,i)\cdot i(t)
+  \symv(\symt)=\symM(x,\symi)\cdot \symi(\symt)
 \end{equation}
@@
 \begin{equation}
-  i(t)=W(x,v)\cdot v(t)
+  \symi(\symt)=\symW(x,\symv)\cdot \symv(\symt)
 \end{equation}
-In all of the previous equations, $v$ is the voltage in Volt ($V$), $i$ is the current in Ampere ($A$), $\phi$ is the flux in Weber ($Wb$), $q$ is the charge in Coulomb ($C$), $M$ is the memristance in Ohm ($\Omega$) and $W$ is the memductance in Siemens ($S$ or $\Omega^{-1}$).
+In all of the previous equations, \symv is the voltage in Volt ($V$), \symi is the current in Ampere ($A$), \symphi is the flux in Weber ($Wb$), \symq is the charge in Coulomb ($C$), \symM is the memristance in Ohm ($\Omega$) and \symW is the memductance in Siemens ($S$ or $\Omega^{-1}$).
 

### 2023-11-30 state/model.tex
Removing space
Changed a bunch of files
Added almost all symbols

@@
 \begin{equation}
-  \frac{dx}{dt}=f(x,v)
+  \frac{dx}{d\symt}=f(x,\symv)
 \end{equation}
 
-Where $v$ is the voltage, $t$ is time and $f$ is a function.
+Where \symv is the voltage, \symt is time.
 
@@
 \begin{equation}\label{eq:funcDesc}
-  f(x,v)=g(x)\cdot s(v)
+  f(x,\symv)=g(x)\cdot s(\symv)
 \end{equation}
@@
 \begin{equation}\label{eq:memModel0}
-  i(R,v)=
+  \symi(\symR,\symv)=
   \begin{cases}
-    \frac{a_p}{R}\cdot sinh(b_p\cdot v) & v\ge 0\\
-    \frac{a_n}{R}\cdot sinh(b_n\cdot v) & v<0\\
+    \frac{a_p}{\symR}\cdot sinh(b_p\cdot \symv) & \symv\ge 0\\
+    \frac{a_n}{\symR}\cdot sinh(b_n\cdot \symv) & \symv<0\\
   \end{cases}
@@
 \begin{equation}\label{eq:memModel1}
-  \frac{dR}{dt}=f(R,v)=s(v)\cdot g(R,v)
+  \frac{d\symR}{d\symt}=f(\symR,\symv)=s(\symv)\cdot g(\symR,\symv)
 \end{equation}
@@
 \begin{equation}\label{eq:memModel2}
-  s(v)=
+  s(\symv)=
   \begin{cases}
@@
 \begin{equation}\label{eq:memModel3}
-  g(R,v)=
+  g(\symR,\symv)=
   \begin{cases}
-    (R-r_p(v))^2& v>0\\
-    (R-r_n(v))^2& v<0\\
-    0 &  v=0\\
+    (\symR-r_p(\symv))^2& \symv>0\\
+    (\symR-r_n(\symv))^2& \symv<0\\
+    0 &  \symv=0\\
   \end{cases}
@@
   \begin{cases}
-    r_p(v)= r_{p,0} + r_{p,1}\cdot v& v>0\\
-    r_n(v)= r_{r,0} + r_{n,1}\cdot v& v\le 0\\
+    r_p(\symv)= r_{p,0} + r_{p,1}\cdot \symv& \symv>0\\
+    r_n(\symv)= r_{r,0} + r_{n,1}\cdot \symv& \symv\le 0\\
   \end{cases}
@@
 
-The changes of resistive states from the initial resistance ($R_0$) are computed by \cref{eq:memModel5} extracted from \cite{memCadenceModel}.
-It uses the bias voltage ($V_b$) of the pulse applied to the memristor.
+The changes of resistive states from the initial resistance ($\symR_0$) are computed by \cref{eq:memModel5} extracted from \cite{memCadenceModel}.
+It uses the bias voltage ($\symv_b$) of the pulse applied to the memristor.
 The pulse has a duration of $t$.
@@
 
-The final Verilog-A code is available at \cref{apsec:memModel}.
+The final Verilog-A code is available in \cite{memCadenceModel}.
 
@@
 The model implements, as with physical memristors, a threshold voltage value from which its resistive states starts changing.
-This threshold value is $V_{read}$.
-For this specific model the threshold voltage is set to $V_{read}=500mV$.
+This threshold value is $\symv_{read}$.
+For this specific model the threshold voltage is set to $\symv_{read}=500mV$.
 
@@
 \begin{itemize}
-  \item The reading of the resistive state, by applying a triangular wave that is bound by the threshold value $V_{read}$.
-The triangular wave starts from $0V$, goes to $V_{read}$ then to $-V_{read}$ before going back to $0V$.
-  \item The writing to the memristor. During this part the resistive state will change. This is done by applying a square wave of variable width ($t_{\Delta}$) with a set bias voltage ($V_{bias}$).
+  \item The reading of the resistive state, by applying a triangular wave that is bound by the threshold value $\symv_{read}$.
+The triangular wave starts from $0V$, goes to $\symv_{read}$ then to $-\symv_{read}$ before going back to $0V$.
+  \item The writing to the memristor.
+During this part the resistive state will change.
+This is done by applying a square wave of variable width ($\symt_{\Delta}$) with a set bias voltage ($\symv_{bias}$).
 \end{itemize}
@@
 
-The resistive state of the memristor changes depending on the pulse width ($t_{\Delta}$) and the bias voltage ($V_{bias}$) used for each pulse.
+The resistive state of the memristor changes depending on the pulse width ($\symt_{\Delta}$) and the bias voltage ($\symv_{bias}$) used for each pulse.
 
@@
   \includesvg[width=\textwidth]{memristor/memPulseChange.svg}
-  \caption{Memristor's resistive state under different pulses width (Fixed bias voltage $V_{bias}=1.8V$)}
+  \caption{Memristor's resistive state under different pulses width (Fixed bias voltage $\symv_{bias}=\symv_{dd}=1.8V$)}
   \label{graph:pulseChange}
@@
 Indeed, the larger the pulse widths, the faster the internal resistance reaches the bounding resistance (\cref{eq:memModel4}).
-The bounding resistance being the same as all curves use the same bias voltage ($V_{bias}=1.8V$), the curves will all eventually reach the max resistance.
+The bounding resistance being the same as all curves use the same bias voltage ($\symv_{bias}=\symv_{dd}=1.8V$), the curves will all eventually reach the max resistance.
 
@@
 Applying higher voltage also allows to reach a set resistance faster as the slopes are bigger the higher the bias voltage gets.
-
 The resistive state can also be lowered by applying negative voltages.%TODO

### 2023-11-30 state/rnn.tex
Worked on paper+corrected errors
Changed a bunch of files
Added almost all symbols

--- 2023-10-25 state/rnn.tex
@@
-\section{\acs{RNN}}\label{sec:rnn}
+\section{\acl{RNN}}\label{sec:rnn}
 
@@
 \begin{equation}\label{eq:rnn}
-  \overrightarrow{h_t}=f(\overrightarrow{x_t},\overrightarrow{h_{t-1}})
+  \symvh_t=f(\symvx_t,\symvh_{t-1})
 \end{equation}
 
-The hidden state ($\overrightarrow{h_t}$) carries all the past informations for the next time step. It also serves as the output of the \ac{RNN}.
+The hidden state ($\symvh_t$) carries all the past informations for the next time step. It also serves as the output of the \ac{RNN}.
 
-They are trained the same way \ac{NN} are, measuring the error, backpropagate and adjust the weights accordingly.
+They are trained the same way \ac{NN} are, measuring the error, backpropagating and adjusting the weights accordingly.
 
-\subsection{Simple \ac{RNN}}
+\subsection{Simple \acl{RNN}}
 
@@
 \begin{equation}\label{eq:srnn}
-  \overrightarrow{h_t}=tanh([\overrightarrow{x_t},\overrightarrow{h_{t-1}}]\cdot w + b)
+  \symvh_t=tanh([\symvx_t,\symvh_{t-1}]\cdot \symmw + \symvb)
 \end{equation}
 
-Where $t\in\mathbb{N}^*$ is the time index, $w$ the weight matrix, $b$ the bias matrix, $\overrightarrow{x_t}$ the input vector and $\overrightarrow{h_{t}}$ the hidden state of the \ac{RNN}.
+Where $t\in\mathbb{N}^*$ is the time index, \symmw the weight matrix, \symvb the bias vector, \symvx the input vector and \symvh the hidden state of the \ac{RNN}.
 

### 2023-11-30 paper/circuit.tex
fixed paper

@@
 
-The system will be working with a $V_{dd}$ of $1.8V$. Such a value was chosen because this is a low power system.
+The system will be working with a $\symv_{dd}$ of $1.8V$. Such a value was chosen because this is a low power system.
 
 The way values are encoded in the analog system will be descibed here as it serves for the entire thesis.
-In order for the system to support negative numbers we're going to use a $V_{cm}$ set to $\frac{V_{dd}}{2}$.
-That means that $V_{cm}=\frac{V_{dd}}{2}=0.9V$.
-This $V_{cm}$ will then describe a zero.
+In order for the system to support negative numbers we're going to use a $\symv_{cm}$ set to $\frac{\symv_{dd}}{2}$.
+That means that $\symv_{cm}=\frac{\symv_{dd}}{2}=0.9V$.
+This $\symv_{cm}$ will then describe a zero.
 A step of one was chosen to be $0.1V$ in the analog circuit.
@@
     \hline
-    $x$ & $V(x)=\frac{x}{10}+V_{cm}$\\
+    $x$ & $\symv(x)=\frac{x}{10}+\sym_{cm}$\\
     \hline
-    $real(v)=(v-0.9)\cdot 10$ & $v$\\
+    $real(\symv)=(\symv-\symv_{cm})\cdot 10$ & $\symv$\\
     \hline
@@
     \hline
-    $V_1$ & \multicolumn{2}{c|}{$1.1V$}\\
+    $\symv_1$ & \multicolumn{2}{c|}{$1.1V$}\\
     \hline
-    $V_2$ & \multicolumn{2}{c|}{$635mV$}\\
+    $\symv_2$ & \multicolumn{2}{c|}{$635mV$}\\
     \hline
-    $V_3$ & $0.8V$ & $550mV$\\
+    $\symv_3$ & $0.8V$ & $550mV$\\
     \hline
-    $i_{dc}$ & \multicolumn{2}{c|}{$150uA$}\\
+    $\symi_{dc}$ & \multicolumn{2}{c|}{$150uA$}\\
     \hline
@@
     \hline
-    $R_1$ & \multicolumn{2}{c|}{$5k\Omega$}\\
+    $\symR_1$ & \multicolumn{2}{c|}{$5k\Omega$}\\
     \hline
-    $R_2$ & \multicolumn{2}{c|}{$10k\Omega$}\\
+    $\symR_2$ & \multicolumn{2}{c|}{$10k\Omega$}\\
     \hline
-    $R_3$ & $2k\Omega$ & $4k\Omega$\\
+    $\symR_3$ & $2k\Omega$ & $4k\Omega$\\
     \hline
@@
 Using two memristor per synapse doubles the area but doubles the weight range \cite{doubleMem} and allows to easily use negavite weights.
-The output voltage will be centered around $V_{cm}$ and be compliant with the standard set in \cref{tab:valConv}.
+The output voltage will be centered around $\symv_{cm}$ and be compliant with the standard set in \cref{tab:valConv}.
 
 This design is the one that is used in \cite{doubleMem}.
-Let's assume that a given memristor has a resistance range of $R\in[R_{min},R_{max}]$, that means its conductance range is $ G \in [ G_{min}, G_{max}]$ (with $ G_{min}= \frac{1}{R_{max}}$ and $ G_{max}= \frac{1}{R_{min}}$).
-This design works using two \ac{opAmp} connected to $V_{cm}$ with the positive pin and the negative pin to the output of the crossbar array.
+Let's assume that a given memristor has a resistance range of $\symR\in[\symR_{min},\symR_{max}]$, that means its conductance range is $ \symG \in [ \symG_{min}, \symG_{max}]$ (with $ \symG_{min}= \frac{1}{\symR_{max}}$ and $ \symG_{max}= \frac{1}{\symR_{min}}$).
+This design works using two \ac{opAmp} connected to $\symv_{cm}$ with the positive pin and the negative pin to the output of the crossbar array.
 \Cref{eq:doubleMem0,eq:doubleMem1,eq:doubleMem2} are describing how this architecture works.
@@
   \begin{array}{2}
-    V_{opAmp_1}=y_0=-R_f\cdot(i_--i_+)=R_f\cdot(i_+-i_-)\\
-    =R_f\cdot\sum_{k=0}^n( G_{k+}- G_{k-})\cdot x_k
+    \symv_{opAmp_1}=\symv_{out}=-\symR_f\cdot(\symi_--\symi_+)=\symR_f\cdot(\symi_+-\symi_-)\\
+    =\symR_f\cdot\sum_{k=0}^n( \symG_{k+}- \symG_{k-})\cdot \symv_k
   \end{array}
 \end{equation}
-With $i_+=\sum_{k=0}^n G_{k+}\cdot x_k$ and $i_-=\sum_{k=0}^n G_{k-}\cdot x_k$.
+With $\symi_+=\sum_{k=0}^n \symG_{k+}\cdot \symv_k$ and $\symi_-=\sum_{k=0}^n \symG_{k-}\cdot \symv_k$.
 
@@
 \begin{equation}\label{eq:finalVoltMult}
-  V_{out}=10\cdot(V_{in_1}-V_{cm})\cdot (V_{in_2}-V_{cm}) + V_{cm}
+  \symv_{out}=10\cdot(\symv_{in_1}-\symv_{cm})\cdot (\symv_{in_2}-\symv_{cm}) + \symv_{cm}
 \end{equation}
@@
 \begin{equation}\label{eq:voltMult}
-  V_{voltMult}=-(V_{in_1}-V_{cm})\cdot (V_{in_2}-V_{cm}) + V_{cm}
+  \symv_{voltMult}=-(\symv_{in_1}-\symv_{cm})\cdot (\symv_{in_2}-\symv_{cm}) + \symv_{cm}
 \end{equation}
 \begin{equation}\label{eq:invAmp}
-  V_{out}=-(V_{voltMult}-V_{cm})\cdot10+V_{cm}
+  \symv_{out}=-(\symv_{voltMult}-\symv_{cm})\cdot10+\symv_{cm}
 \end{equation}
 
-Where $V_{out}$ is the output voltage the inverting amplifier, $V_{voltMult}$ is the out voltage of the voltage multiplier itself and $V_{in_1}$ and $V_{in_2}$ are the input voltages.
+Where $\symv_{out}$ is the output voltage the inverting amplifier, $\symv_{voltMult}$ is the out voltage of the voltage multiplier itself and $\symv_{in_1}$ and $\symv_{in_2}$ are the input voltages.
 
@@
   \item $e_{next}$ is the enable flag on in between two time steps.
-  \item $R_{amp0}$ and $R_{amp1}$ are the two resistances used to amplify the output voltage of the voltage multipliers. Their value is set so that $\frac{R_{amp1}}{R_{amp0}}=10$.
+  \item $\symR_{amp0}$ and $\symR_{amp1}$ are the two resistances used to amplify the output voltage of the voltage multipliers. Their value is set so that $\frac{\symR_{amp1}}{\symR_{amp0}}=10$.
 \end{itemize}
@@
 \begin{equation}\label{eq:1minus1}
-  f_v(v)=-(v-V_{inv})+V_{inv}=2\cdot V_{inv} -v
+  f_v(\symv)=-(\symv-\symv_{inv})+\symv_{inv}=2\cdot \symv_{inv} -\symv
 \end{equation}
-With $V_{inv}=V_{cm}+0.05$.
+With $\symv_{inv}=\symv_{cm}+0.05$.
 

### 2023-11-30 paper/state.tex
fixed paper

@@
 
-The feedback connection is caracterized by the hidden state vector ($\overrightarrow{h_t}$) which serves as the output and as part of the input for the next time step, a RNN is defined as \cref{eq:rnn}.
+The feedback connection is caracterized by the hidden state vector ($\symvh_t$) which serves as the output and as part of the input for the next time step, a RNN is defined as \cref{eq:rnn}.
 
 \begin{equation}\label{eq:rnn}
-  \overrightarrow{h_t}=f(\overrightarrow{x_t},\overrightarrow{h_{t-1}})
+  \symvh_t=f(\symvx_t,\symvh_{t-1})
 \end{equation}
@@
 \begin{equation}\label{eq:srnn}
-  \overrightarrow{h_t}=tanh([\overrightarrow{x_t},\overrightarrow{h_{t-1}}]\cdot w + \overrightarrow{b})
+  \symvh_t=tanh([\symvx_t,\symvh_{t-1}]\cdot \symmw + \symvb)
 \end{equation}
 
-Where ($w$,$\overrightarrow{b}$) are the pair of weights matrix and bias vectors. $\overrightarrow{x_t}$ is the input vector and $\overrightarrow{h_t}$ is the hidden state vector.
+Where (\symmw,\symvb) are the pair of weights matrix and bias vectors. $\symvx_t$ is the input vector and $\symvh_t$ is the hidden state vector.
 
@@
 It was improved a few times before becoming the modern LSTM \cite{improvLSTM}.
-The LSTM differs from a simple RNN because of its second feedback variable being the cell state ($\overrightarrow{c_t}$).
+The LSTM differs from a simple RNN because of its second feedback variable being the cell state ($\symvc_t$).
 
@@
 \begin{equation}\label{eq:inputG}
-  \overrightarrow{i_t}= G ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_i + \overrightarrow{b_i})
+  \symvi_t= \sigma([\symvx_{t_1},\symvh_{t-1}]\cdot \symmw_i + \symvb_i)
 \end{equation}
 \begin{equation}\label{eq:forgetG}
-  \overrightarrow{f_t}= G ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_f + \overrightarrow{b_f})
+  \symvf_t= \sigma([\symvx_{t_1},\symvh_{t-1}]\cdot \symmw_f + \symvb_f)
 \end{equation}
 \begin{equation}\label{eq:outputG}
-  \overrightarrow{o_t}= G ([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_o + \overrightarrow{b_o})
+  \symvo_t= \sigma([\symvx_{t_1},\symvh_{t-1}]\cdot \symmw_o + \symvb_o)
 \end{equation}
 \begin{equation}\label{eq:candCell}
-  \overrightarrow{\tilde{c}_t}=tanh([\overrightarrow{x_{t_1}},\overrightarrow{h_{t-1}}]\cdot w_c+ \overrightarrow{b_c})
+  \symvct_t=tanh([\symvx_{t_1},\symvh_{t-1}]\cdot \symmw_c+ \symvb_c)
 \end{equation}
@@
 \begin{equation}\label{eq:cellS}
-  \overrightarrow{c_t}=\overrightarrow{f_t}\odot \overrightarrow{c_{t-1}} + \overrightarrow{i_t} \odot \overrightarrow{\tilde{c}_t}
+  \symvc_t=\symvf_t\odot \symvc_{t-1} + \symvi_t \odot \symvct_t
 \end{equation}
@@
 \begin{equation}\label{eq:hiddenS}
-  \overrightarrow{h_t}=\overrightarrow{o_t}\odot tanh(\overrightarrow{c_t})
+  \symvh_t=\symvo_t\odot tanh(\symvc_t)
 \end{equation}
 
-Where ($w_i$,$\overrightarrow{b_i}$), ($w_f$,$\overrightarrow{b_f}$), ($w_o$,$\overrightarrow{b_o}$) and ($w_c$,$\overrightarrow{b_c}$) are the pair of weights matrixes and bias vectors for the input, forget, output and candidate cell gates respectively.
-$\overrightarrow{x_t}$ is the input vector and $\overrightarrow{h_t}$ is the hidden state vector.
+Where ($\symmw_i$,$\symvb_i$), ($\symmw_f$,$\symvb_f$), ($\symmw_o$,$\symvb_o$) and ($\symmw_c$,$\symvb_c$) are the pair of weights matrixes and bias vectors for the input, forget, output and candidate cell gates respectively.
+$\symvx_t$ is the input vector and $\symvh_t$ is the hidden state vector.
 
@@
 \begin{equation}\label{eq:updateG}
-  \overrightarrow{z_t}= G ([\overrightarrow{x_t},\overrightarrow{h_{t-1}}] \cdot w_z + \overrightarrow{b_z})
+  \symvz_t= \sigma([\symvx_t,\symvh_{t-1}] \cdot \symmw_z + \symvb_z)
 \end{equation}
 \begin{equation}\label{eq:resetG}
-  \overrightarrow{r_t}= G ([\overrightarrow{x_t},\overrightarrow{h_{t-1}}] \cdot w_r + \overrightarrow{b_r})
+  \symvr_t= \sigma([\symvx_t,\symvh_{t-1}] \cdot \symmw_r + \symvb_r)
 \end{equation}
 \begin{equation}\label{eq:candActivG}
-  \overrightarrow{\hat{h_t}}=tanh(\overrightarrow{x_t}\cdot w_{hx}+(\overrightarrow{r_t}\odot\overrightarrow{h_{t-1}}) \cdot w_{hh} + \overrightarrow{b_h})
+  \symvhh_t=tanh(\symvx_t\cdot \symmw_{hx}+(\symvr_t\odot\symvh_{t-1}) \cdot \symmw_{hh} + \symvb_h)
 \end{equation}
 \begin{equation}\label{eq:gruHidG}
-  \overrightarrow{h_t}=(\overrightarrow{1}-\overrightarrow{z_t})\odot \overrightarrow{h_{t-1}} + \overrightarrow{z_t}\odot \overrightarrow{\hat{h_t}}
+  \symvh_t=(\vunit-\symvz_t)\odot \symvh_{t-1} + \symvz_t\odot \symvhh_t
 \end{equation}
 
-Where ($w_z$,$\overrightarrow{b_z}$), ($w_r$,$\overrightarrow{b_r}$),($w_{hx}$,$w_{hh}$,$\overrightarrow{b_h}$) are the weights matrixes and bias vectors for the update, reset and candidate activation gates respectively.
+Where ($\symmw_z$,$\symvb_z$), ($\symmw_r$,$\symvb_r$),($\symmw_{hx}$,$\symmw_{hh}$,$\symvb_h$) are the weights matrixes and bias vectors for the update, reset and candidate activation gates respectively.
 
@@
 It was first theorized in 1971 as a missing fundamental component in \cite{TheoMemristor}. The name comes from the blend of \textit{memory} and \textit{resistance}.
-The missing component linking the four fundamental circuit variables, voltage ($v$), charge ($q$), current ($i$) and flux ($\phi$).
+The missing component linking the four fundamental circuit variables, voltage (\symv), charge (\symq), current (\symi) and flux (\symphi).
 \Cref{fig:fundComp} shows the four fundamental variables are on each side of the square, with the ones on opposite sides being linked by the following equations :
@@
 \begin{equation}
-  dq = i\cdot dt
+  d\symq = \symi\cdot d\symt
 \end{equation}
 
-Resistors, capacitors and inductors were already very established and well known components, so it was theorized that a fourth device should then exist to physically link flux ($\phi$) and charge ($q$).
-The flux in this case is not a magnetic flux and is defined as such : $ d\phi=v\cdot dt \Rightarrow \phi =  \int v \,dt  $.
+Resistors, capacitors and inductors were already very established and well known components, so it was theorized that a fourth device should then exist to physically link flux (\symphi) and charge (\symq).
+The flux in this case is not a magnetic flux and is defined as such : $ d\symphi=\symv\cdot d\symt \Rightarrow \symphi =  \int \symv \,d\symt  $.
 
@@
 
-A voltage is applied on the $i^{th}$ line, and because every column is virtually grounded, the voltage applied to the memristor, with resistance $R_i$, is $V_i$.
-By applying Ohm's law, we know that the current flowing into the memristor ($I_{i}$) is bound by the following equation :
+A voltage is applied on the $k^{th}$ line, and because every column is virtually grounded, the voltage applied to the memristor, with resistance $R_i$, is $V_i$.
+By applying Ohm's law, we know that the current flowing into the memristor ($\symi_{k}$) is bound by the following equation :
 
 \begin{equation}
-  V_i = R\cdot I_{i} \Rightarrow I_{i} = V_i\cdot (\frac{1}{R_i})= V_i\cdot G_i
+  \symv_k = \symR\cdot \symi_{k} \Rightarrow \symi_{k} = \symv_k\cdot (\frac{1}{\symR_k})= \symv_k\cdot \symG_k
 \end{equation}
-With $ G_i$ being the conductance of memristor, defined as $ G_i=\frac{1}{R_i}$.
+With $ \symG_k$ being the conductance of memristor, defined as $ \symG_k=\frac{1}{\symR_k}$.
 
-This line then joins the column where a current of $I_{i,j-1}$ is flowing, then according to Kirchhoff's current law the resulting current is :
+This line then joins the column where a current of $\symi_{j,k-1}$ is flowing, then according to Kirchhoff's current law the resulting current is :
 \begin{equation}
-  I_{j,i} = I_{j,i-1}+I_{i} = I_{j,i-1} + V_i\cdot G_i
+  \symi_{j,k} = \symi_{j,k-1}+\symi_{k} = \symi_{j,k-1} + \symv_k\cdot \symG_k
 \end{equation}
@@
 \begin{equation}
-  I_1= G_1\cdot V_1 +  G_2\cdot V_2 +  G_3\cdot V_3
+  \symi_1= \symG_1\cdot \symv_1 +  \symG_2\cdot \symv_2 +  \symG_3\cdot \symv_3
 \end{equation}
-With $ G_1$, $ G_2$ and $ G_3$ being the conductance of the 3 memristors in the first column.
+With $\symG_1$, $\symG_2$ and $\symG_3$ being the conductance of the 3 memristors in the first column.

### 2023-12-13 conclusions/imag.tex
Started imag envirnoment

@@
 
+The subject of this thesis being purely theoretical, running simulations only, the environmental and societal impact will focus on the system if it were to become real.
+This assumes that the entire system is working and is being fabricated for embedded use.
+The final product can be used in both comsumer and business oriented product.
+
 \subsubsection{Environmental impact}

### 2023-12-13 state/main.tex
Correctiion diogo state

@@
 \label{cap:state}
-
-This chapter will go over the several subjects that need to be known in order to understand the thesis.
-
-Math is the best way to universally describe all scientific models. For this reason, this chapter, and the thesis as whole, will contains equations.
-
-It seems only logical to define all the tools that are going to be used in the thesis.
-
-\begin{itemize}
-  \item Vectors are displayed as $\overrightarrow{v}$ and are considered to be vertical matrixes, matrixes of size ($1\times n$) for vector of size $n$.
-  \item $\overrightarrow{1}$ represent a vector that only contains ones of variable size.
-  \item $\odot$ is used for the Hadamart product, that is the point wise multiplication of each element of the matrixes
-  \item $\cdot$ is used in two cases :
-    \begin{itemize}
-      \item When used with scalars, it is the product of those scalars.
-      \item When used with matrixes, it is the matrix product
-    \end{itemize}
-  \item $+$ is also used in two cases :
-    \begin{itemize}
-      \item When used with scalars, it is the sum of those scalars.
-      \item When used with vectors (vertical matrixes), it is the point wise sum, basically adding each element to each other. The result is a vector of the same size.
-    \end{itemize}
-\end{itemize}
 

### 2023-12-15 conclusions/imag.tex
Worked quite a bit on environmental impact

@@
 
+The system would be found in the form of a chip, they would need to be fabricated using a very precise method.
+Chip fabrication doesn't come cheap for the environment.
+Indeed, the manufacturing process of computer chips, while integral to technological advancements, poses significant environmental challenges.
+Chip manufacturing involves several stages, including silicon wafer production, lithography, etching, and assembly.
+These processes consume vast amounts of energy and water while generating substantial waste.
+Semiconductor fabrication demands clean rooms with controlled environments, necessitating extensive energy consumption for air conditioning and filtration systems.
+Additionally, the use of hazardous chemicals, such as solvents and acids, in chip etching and cleaning stages contributes to chemical waste and potential environmental pollution if not managed and disposed of properly.
+Moreover, the disposal of silicon scraps and other manufacturing byproducts requires careful handling to prevent environmental contamination.
+The machines used to fabricate the chip are not taken into account here as their impact is negligible next to the fabrication itself.
+For example, TSMC's electricity consumption in 2021 accounted for 5 \% of Taiwan total electricity consumption \cite{chipEnv}.
+Furthermore, in a report from greenpeace, the projected emissions in 2030 of the chip manufacturing industry is of 86 million metric tons of carbon dioxide equivalent, which is more than the total emissions of Portugal over 2021 and 2022 \cite{greenReport,worldEmissions}.
+
+When the product reaches end of life, it needs to be disposed of properly to avoid soil and water contamination due to the release of hazardous substances.
+Recycling computer chips is a critical facet of sustainable electronic waste management, aiming to mitigate the environmental repercussions of discarded electronic devices.
+The recycling process typically involves several stages, including collection, sorting, dismantling, and extraction of valuable materials like silicon, copper, gold, and other metals.
+However, chip recycling poses challenges due to the complex composition of electronic components and the intricate structures of integrated circuits.
+These challenges often necessitate advanced techniques such as shredding, mechanical separation, and chemical treatments to extract and purify the reusable materials.
+Although chip recycling reduces the need for raw materials and minimizes landfill accumulation, the process itself demands significant energy and resources.
+Therefore, while chip recycling offers a sustainable approach to managing electronic waste, continuous advancements in recycling technologies and effective waste management policies are imperative to diminish the environmental impact associated with both chip manufacturing and disposal.
+
+Subtantial gain in power consumption.
+
 \subsubsection{Societal impact}

### 2023-12-16 conclusions/imag.tex
a few lines

@@
 The system would be found in the form of a chip, they would need to be fabricated using a very precise method.
-Chip fabrication doesn't come cheap for the environment.
+Chip fabrication does not come cheap for the environment.
 Indeed, the manufacturing process of computer chips, while integral to technological advancements, poses significant environmental challenges.
@@
 
+The potential chip that would be created has been analyzed, there are also other aspects to take into account.
+The system being designed to run \acp{NN}, the training of the \ac{NN} also consumes energy
+
 Subtantial gain in power consumption.

### 2023-12-19 conclusions/imag.tex
Few more lines, plus déblocage de moi

@@
 The potential chip that would be created has been analyzed, there are also other aspects to take into account.
-The system being designed to run \acp{NN}, the training of the \ac{NN} also consumes energy
+The system being a way to run a wide variety of \acp{NN}, depending on the scale of the project the cip is used in, the impact of training is different.
+However, the main application being for embedded systems, training the system for a few hours is negligible when the final circuit is used thousands to millions times in its life cycle.
+The inference of the \ac{NN} also has an environmental impact due to the power it uses.
+But the goal of building an analog \ac{NN} is to reduce both execution time and energy consumption, and while the energy comsumption wasn't evaluated in this thesis
 

### 2023-12-20 conclusions/imag.tex
New text, missing societal

@@
 The inference of the \ac{NN} also has an environmental impact due to the power it uses.
-But the goal of building an analog \ac{NN} is to reduce both execution time and energy consumption, and while the energy comsumption wasn't evaluated in this thesis
+But the goal of building an analog \ac{NN} is to reduce both execution time and energy consumption, and while the energy comsumption wasn't evaluated in this thesis, it can be assumed, from the usual trend of analog circuits, a much lower energy consumption than with digital chips.
+This, scaled with thousands of units, offers a subtantial gain in overall power consumption.
 
-Subtantial gain in power consumption.
+Once again, the system having a wide spectrum of applications, the rebound effect can't be determined, it all depends on the kind of \ac{NN} that will be running on the chip.
+Nonetheless, the system being designed around embedded systems, the improvements made to the power consumption will affect the battery life of the product using the chip.
 
@@
 
+The societal impacts are hard t
+
 \subsection{Host structure's policy}

### 2023-12-21 conclusions/imag.tex
Final version of my thesis

@@
 \subsection{Personnal environmental impact}
+
+My personnal environmental impact was overall very low.
+
+I was walking from home to the lab everyday, which almost has no carbon footprint, and determining such a number is impossible.
+
+I was already in Portugal when my thesis started so there wasn't any trips to join there.
+Nevertheless, for the sake of having things to talk about, I'm going to take into account the personnal flights home.
+One way between Toulouse and Lisbon is estimated to have a carbon footprint of $234 kg\, eq.\,CO_2$.
+Having made a total of 7 trips, my total carbon footprint for traveling during this period is of about $1638 kg\, eq.\, CO_2$.
+
+The computers I was working on were my owns, they both have an average power consumption of about $25W$ or $50W$ total.
+In Portugal, the average carbon footprint of $1kW\cdot h^{-1}$ of electricity is of about $60g\,eq.\,CO_2$ \cite{portugalElec}.
+I used my computers an average of 7 hours a day for an estimated 175 days of work.
+This accounts for a total of $6.125kW\cdot h^{-1}$ which makes a carbon footprint of $3 675g\,eq.\,CO_2$.
+
+All this data added together makes a total of $1642kg\,eq.\,CO_2$ personnal carbon footprint over my thesis in Portugal.
+This number is quite high, but this is due to the plane trips, the computer consumption is negligeable in the end.
 
@@
 
-\subsubsection{Environmental impact}
+\subsubsection{Environmental impact}\label{subsubsec:envImp}
 
@@
 
-The societal impacts are hard t
+The societal of such a projects are quite hard to determines since it all depends on the ways it is used, which can be for basically anything that a \ac{NN} can do.
+Nevertheless, RGPD isn't a concern for the chip itself, the chip being completely isolated from the internet or any other connection as of itself.
+
+Using this chip in embedded systems might increase the digital divide, as they would push more state of the art hardware to market and thus rendering old technologies less relevant.
+However, the examples in \cref{sec:motivation} are more seemless uses.
+The gait detection \cite{gaitDS,gaitDig,gait} could be used in a security camera, without the user having to do anything different.
+The camera stabilization \cite{videoStab} if integrated in a chip within a camera would record a better stabilized video.
+
+The good or bad impacts of the chip depends on its use. For a quite farfetch analogy, nuclear power has been used to both make environment friendly fuel and bombs.
 
 \subsection{Host structure's policy}
+
+The lab that hosted me during my master thesis doesn't have a policy concerning the envirnoment.
+\ac{INESC}-MN being a micro and nanofabrication lab, they own and run most of the machines required to fabricate chips.
+The machines are not state of the art and mostly used for prototyping, for the lab itself or for contracted companies.
+As explained in \cref{subsubsec:envImp}, those machines are very power hungry and thus explains why any other measures would be rendered useless next to this.

### 2024-01-08 imag.tex
New version for imag

--- 2023-12-21 conclusions/imag.tex
@@
-\section{Environmental and societal impact}
+\fancychapter{Environmental and societal impact}\label{cap:rse}
 
-\subsection{Personnal environmental impact}
+\section{Personnal environmental impact}
 
@@
 
-\subsection{Global impact}
+\section{Global impact}
 
@@
 
-\subsubsection{Environmental impact}\label{subsubsec:envImp}
+\subsection{Environmental impact}\label{subsec:envImp}
 
@@
 The potential chip that would be created has been analyzed, there are also other aspects to take into account.
-The system being a way to run a wide variety of \acp{NN}, depending on the scale of the project the cip is used in, the impact of training is different.
+The system being a way to run a wide variety of \acp{NN}, depending on the scale of the project the chip is used in, the impact of training is different.
 However, the main application being for embedded systems, training the system for a few hours is negligible when the final circuit is used thousands to millions times in its life cycle.
@@
 
-\subsubsection{Societal impact}
+\subsection{Societal impact}
 
@@
 
-\subsection{Host structure's policy}
+\section{Host structure's policy}
 
 The lab that hosted me during my master thesis doesn't have a policy concerning the envirnoment.
-\ac{INESC}-MN being a micro and nanofabrication lab, they own and run most of the machines required to fabricate chips.
+\acs{INESC}-MN being a micro and nanofabrication lab, they own and run most of the machines required to fabricate chips.
 The machines are not state of the art and mostly used for prototyping, for the lab itself or for contracted companies.
-As explained in \cref{subsubsec:envImp}, those machines are very power hungry and thus explains why any other measures would be rendered useless next to this.
+As explained in \cref{subsec:envImp}, those machines are very power hungry and thus explains why any other measures would be rendered useless next to this.
